### 2018-09-25 article.tex
First draft

\documentclass{article}

% Target journal: Molecular Phylogenetics and Evolution
%
% From author guidelines:
%
% Short communications of approximately 3000 words are also accepted. 
% These papers should contain no more than two figures, two tables, and thirty references. 
% A short abstract of fewer than 200 words is acceptable.

% Annotation/feedback commands
\newcommand*\rampal[1]{\textcolor{red}{\textbf{[RSE: #1]}}}
\newcommand*\richel[1]{\textcolor{orange}{\textbf{[RJCB: #1]}}}
\newcommand*\gio[1]{\textcolor{green}{\textbf{[RJCB: #1]}}}

% Bibliography
\usepackage{natbib}
\bibpunct{(}{)}{;}{a}{}{;}

\usepackage[english]{babel}

% Use 'It was found that something is something (Name 1234)' style
\setcitestyle{authoryear,open={},close={}}

% Affiliations
\usepackage{authblk}
\title{The error in Bayesian phylogenetic reconstruction when speciation co-occurs}

\author[1]{Giovanni Laudanno}
\author[1]{Rich\`el J.C. Bilderbeek}
\author[1]{Rampal S. Etienne}
\affil[1]{Groningen Institute for Evolutionary Life Sciences, University of Groningen, Groningen, The Netherlands}

% Use double spacing
\usepackage{setspace}
\doublespacing

\usepackage{pgf}
\usepackage{hyperref}
\usepackage{verbatim}

% Adds numbered lines
\usepackage{lineno}
\linenumbers

\hyphenation{
  BEAST
  Pa-ra-me-ter
  Drum-mond 
  Bayes-ian 
  Mr-Bayes 
  ap-proach-es 
  Rev-Bayes 
  cre-ate
  spe-ci-a-tion-com-ple-tion
  pro-trac-ted
 }

\begin{document}

\maketitle

\begin{abstract}

  % From 'How to construct a Nature summary paragraph'

  % A short abstract of fewer than 200 words is acceptable.

  % One or two sentences providing a basic
  % introduction to the field,
  % comprehensible to a scientist in any discipline.
  The tools for reconstructing phylogenetic relationships between taxonomic 
  units (e.g. species) have become very advanced in the last three decades.

  % Two to three sentences of
  % more detailed background, comprehensible to
  % scientists in related disciplines.
  Among the most popular tools are Bayesian approaches, 
  such as \\ BEAST, MrBayes and RevBayes, 
  that use efficient tree sampling routines to create a posterior probability distribution 
  of the phylogenetic tree. 
  A feature of these approaches is the possibility to incorporate known 
  or hypothesized structure of the phylogenetic tree through the tree prior. 
  It has been shown that the effect of the prior on the posterior distribution 
  of trees can be substantial.

  % One sentence clearly stating the general
  % problem being addressed by this particular
  % study.
  Currently implemented tree priors assume that speciation events are independent,
  where we know that speciation can coincide, for example, when trigger by a
  larger geographic change.

  % One sentence summarising the main
  % result (with the words “here we show”
  % their equivalent).
  Here we explore the effects of ignoring 
  speciation co-occurence with an extensive simulation study.

  % Two or three sentences explaining what
  % the main result reveals in direct
  % comparison to what was thought to be the case
  % previously, or how the main result adds to
  % previous knowledge.

  % One or two sentences to put the results into a
  % more general context.

  % Two or three sentences to provide a
  % broader perspective, readily comprehensible
  % to a scientist in any discipline, may be included
  % in the first paragraph if the editor considers that the accessibility of the paper is significantly enhanced
  % by their inclusion. Under these circumstances, the length of the paragraph can be up to 300 words.
  % (The above example is 190 words without the final section, and 250 words with it).

  We compare the inferred tree to the simulated tree, and find that ....

\end{abstract}

{\bf Keywords:} computational biology, evolution, phylogenetics, Bayesian analysis, tree prior

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The computational tools that are currently available 
to the phylogeneticists go beyond the wildest 
imagination of those living four decades ago.
Advances in computational power allowed the first cladograms to be inferred 
from DNA alignments in 1981 (\cite{felsenstein1981}), and  
the first Bayesian tools emerged in 1996 (\cite{rannala1996}),
providing unprecedented flexibility in the setup of a phylogenetic model.

Currently, the most popular Bayesian phylogenetics tools are \\ 
BEAST (\cite{beast}) and its offshoot BEAST2 (\cite{beast2}), 
MrBayes (\cite{mrbayes}) and RevBayes (\cite{revbayes}). 
They allow to incorporate known or hypothesized structure of a phylogenetic 
tree-to-be-inferred through model priors. 
With these priors and an alignment of DNA, RNA or protein sequences, 
they create a sample of the posterior distribution
of phylogenies and parameter estimates (of the models used as a prior), 
in which more probable combinations are represented more often.
Each of these tools use efficient tree sampling routines to rapidly create an 
informative posterior.

The model priors in Bayesian phylogenetic reconstruction 
can be grouped into three categories: (1) site model, specifying 
nucleotide substitutions, (2) clock model, specifying
the rate of mutation per lineage in time, and (3) tree model, 
constituting the speciation model underlying branching events (speciation) 
and branch termination (extinction).
The choice of site model (\cite{posada_and_buckley_2004}), 
clock model (\cite{baele_et_al_2012}) 
or tree prior (\cite{moller2018, yang_and_ranalla_2005}) is known to affect
the posterior.

Current phylogenetic tools use tree priors 
that assume speciation events are independent, 
whilst we know multiple examples of adaptive radiations 
in multiple species families can co-occur, 
for example \richel{@gio: please add examples}.

The (constant-rate) birth-death (BD) model is a commonly 
used tree prior, but it ignores the co-occurence of speciation.
It even assume that two speciation events at exactly the same time
has zero likelihood!
The multiple birth-death (MBD) model, an extension of 
the BD model, does incorporate the idea that speciation can co-occur.

\richel{explain model here, example is below in the comments}
% In this model, a branching event does not give rise to a new species, but to
% a new species-to-be, called an incipient species. Such an incipient
% species may go extinct, finish its speciation to become a good species, or give
% rise to new incipient species. Protracted speciation may explain observed 
% declines in lineage accumulation (\cite{etienne_and_rosindell_2012}).

Unfortunately, a tree prior according to this model, 
providing the probability of a species tree under the MBD model, 
is unavailable in current Bayesian phylogenetic tools. 
Whilst a likelihood equation has been derived (\richel{cite yourself here}),
it has not been implemented as tree prior yet. 
There are various reasons for this.
First, the computation of the MBD likelihood involves solving a set of 
non-linear differential equations, and while this computation is quite fast, 
it still takes much more time than the corresponding probability 
of the BD model which is a simple analytical formula. 
In a Bayesian MCMC chain, the tree prior probability must be calculated many times, 
and hence the total computation will take considerably longer with a PBD tree prior.

Here we aim to explore the effect of using the
BD prior on MBD simulated phylogenies.
In brief, we simulate phylogenies with co-occuring speciation events using the MBD process. 
Given this species tree, we simulate a DNA sequence alignment. Then, we use BEAST2 on these alignments
to infer a posterior of phylogenies, using a BD prior. We quantify the difference
between the (BD) posterior phylogenies and the simulated (MBD) species tree.
Furthermore, while we evidently know the clock and site models used in the simulation, 
using a different clock and/or site model prior in inference 
may compensate or increase this difference between inferred and simulated tree. 
To study this, we also explore the effect of 
a different clock and site model prior in inference.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Methods (but we are not allowed to use this header)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The MBD model has \richel{@gio: how many?} parameters, depicted in table \ref{table:parameters}.

\richel{@gio: describe parameter values used here, example is below}
% The per species speciation completion rates $\lambda$ we use are $0.1$, $0.3$, $1.0$ and $10^9$. This means that there is a $0.15 dt$ ($0.35 dt$, $1.0 dt$, $10^9 dt$) probability of speciation completion occurring in an infinitesimal time $dt$. 
% We use per species extinction rates of $\mu = \mu_g = \mu_i$ $0.0$, $0.1$ and $0.2$.
% For each combination of $\lambda$, $\mu$ and $n$, we use a speciation initiation rate $b = b_i = b_g$ that gives a given value of the expected number of species. We use expected mean tree sizes $n$ of 50, 100 and 200 good taxa and the corresponding speciation initiation rates for all parameter combinations are shown in table \ref{table:pbd_parameters}.
% Our parameters are inspired by existing work on protracted speciation \cite{etienne_and_rosindell_2012}\cite{etienne2014}.

We use \richel{@gio: parameter setting here} as our control for which the MBD model reduces to the BD model.

We simulate protracted birth-death trees, using the MBD package (\cite{pbd}) in the R programming language (\cite{r}).
The first tree has a random number generator seed of 1, which is incremented by 1 for each simulated tree.
For each combination of \richel{@gio: parameter values here}, 
we generate incipient species trees with a crown age of 15 million years.
Only trees with the desired number of good taxa are kept.

We create one data set to explore parameter space, 
All the trees with the correct number of good species are kept.
Based on the species tree, we simulate a DNA alignment that has the same history
as this species tree, using the \verb;phangorn; package (\cite{phangorn}). 
We set the nucleotides of the DNA alignment to follow a Jukes-Cantor (\cite{jc69})
nucleotide substitution model, in which all nucleotide-to-nucleotide transitions
are equally likely. 
The DNA sequence of the root ancestor consists of four equally sized single-nucleotide 
blocks of adenine, cytosine, guanine and thymine respectively. 
For example, for a DNA sequence length of 12, this would be AAACCCGGGTTT. 
The order of nucletides does not matter in this study, 
because we do not consider several partitions of the sequence with their own parameters. 
Only the frequency of occurrence matters.
In our Bayesian inference (see below) we use the same site model as the (obviously correct) site model prior,
but we also explore the effect of assuming a more complex site model prior.
We predict with the more complex substitution model, 
that there will be more noise and hence our inference error will increase.
On the other hand, we dare not rule out that the inference error will decrease,
due to more flexibility in the more complex prior.
We set the mutation rate in such a way to maximize the information contained in the alignment.
To do so, we set the mutation rate such that we expect on average one (possibly silent) mutation per nucleotide
between crown age and present, which equates to $\frac{1}{15}$ mutations
per million years.
The DNA sequence length is chosen to provide a
resolution of $10^3$ years, 
% If sims take too long, we will decrease this resolution
that is, to have one expected nucleotide change 
per $10^3$ years per lineage on average. As one nucleotide is expected 
to have on average one (possibly silent) mutation per 15 million years, $15 \cdot 10^3$
nucleotides result in 1 mutation per alignment per $10^3$ years (which is
coincidentally the same as \cite{moller2018}). 
The simulation of these DNA aligments follows a strict clock model, 
which we will specify as one of the two clock models assumed in the Bayesian inference (see below).

\richel{must rewrite, use pirouette as a starting point}
From an alignment, we run a Bayesian analysis and 
create a posterior distribution of trees and parameters
using the \verb;pirouette; (\cite{pirouette}) package
that sets the input parameters similar to BEAUti 2 and then runs BEAST2. 
For our site model, we assume either a Jukes-Cantor or GTR nucleotide substitution model.
The Jukes-Cantor model is the correct one, as it is used for simulating that alignment,
where the GTR model is the site model that is picked as a default by most users.
For our clock model, we assume either a strict or relaxed log-normal 
clock model. 
Also here, the strict clock model is the correct one, as it is used for simulating the alignment,
but the relaxed log-normal clock model is the one most commonly used.
We set the BD model as a tree prior, 
as gauging the effect of this incorrect assumption is the goal of this study. 
We assume an MRCA prior with a tight normal distribution
around the crown age, by choosing the crown age as mean, and a standard deviation 
of $0.5 \cdot 10^{-3}$ time units,
resulting in 95\% of the crown ages inferred have the same resolution (of $10^{-3}$ time 
units) as the alignment. 
We ran the MCMC chain to generate 1111 states,
of which we remove the first 10\% (also called the 'burn-in'). 
Of the remaining
1000 MCMC states, the effective sample size (ESS) of the posterior 
must at least be 200
for a strong enough inference (\cite{beastbook}). An ESS can be increased by increasing
the number of samples or decreasing the autocorrelation between samples. 
If the ESS is less than 200, we decrease autocorrelation by doubling 
the MCMC sampling interval of that simulation, until the ESS exceeds 200.

We compare each posterior phylogeny to the (sampled) species tree
using the nLTT statistic (\cite{janzen2015}), from the \verb;nLTT; package (\cite{nltt}). 
The nLTT statistic equals the area between the normalized
lineages-through-time-plots of two phylogenies, which has a range 
from zero (for identical phylogenies) to one. We use inference error 
and nLTT statistic interchangeably. Comparing the simulated species tree
with each of the posterior species trees yields a distribution of nLTT statistics.

The input trees generated with a \richel{@gio: parameter that is set to reduce the MBD model to BD} 
allow us to measure the noise of the experiment.

We produce one data set as a comma-separated file.
The general data set has ?144 \richel{recalc} different combinations
of biological parameter combinations, site and clock models.
The data set to investigate sampling has ?552 \richel{recalc} different combinations
of biological parameter combinations, site models, clock models 
and sampling methods. The experiment is computationally intensive:
pilot experiments show that the experiment takes roughly 100 days
of CPU time and 20 days of wall clock time (which includes the queued 
waiting for computational resources) per replicate. 
Due to this, we choose to perform ten replicates, so that the complete
experiment will take an acceptable time of roughly seven months.

For both data sets, we display the nLTT statistics distribution per
biological parameter combination as a violin plot.
We only show the nLTT distributions
that were generated under the (correct) assumptions of a Jukes-Cantor site model
and a strict clock model, separated per sampling method used. 
We display the nLTT statistic distributions separated per site or clock model 
in the supplementary information.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{fig_general.png}
  \caption{
    nLTT statistic distribution per biological parameter set, using the
    general data set, 
    under the (correct) assumptions of a strict clock and Jukes-Cantor site model.
  }
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Glossary}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Please supply, as a separate list, the definitions of field-specific terms used in your article.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
  \centering 
  \begin{tabular}{l p{0.65\textwidth}}
    \hline
    Term                  & Definition \\
    \hline
    \hline
    Phylogenetics         & The inference of evolutionary relationships of groups of organisms using genetics \\
    Model prior           & Knowledge or assumptions about the ontogeny of evolutionary histories \\
    Posterior             & A collection of phylogenies and parameter estimates, in which more probable combinations (determined by the data and the model prior) are presented more frequently \\
    \hline
  \end{tabular}
  \caption{
    Glossary
  }
  \label{table:glossary}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Bibliography
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MEE style
\bibliographystyle{mee}
\bibliography{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{fig_clock_model.png}
  \caption{
    nLTT statistic distribution per biological parameter set per clock model,
    using the general data set, 
    under the (correct) assumption of a Jukes-Cantor site model.
  }
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{fig_site_model.png}
  \caption{
    nLTT statistic distribution per biological parameter set per site model,
    using the general data set, 
    under the (correct) assumption of a strict clock model.
  }
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
  \centering 
  \begin{tabular}{p{0.05\textwidth} p{0.6\textwidth} p{0.2\textwidth}}
    \hline
                          & Description & Values \\
    \hline
    \hline
    $\lambda$             & \richel{@gio: MBD params here} & 0.1, 0.3, 1.0, $10^9$ \\
    $\lambda$             & Speciation rate & 0.1, 0.3, 1.0, $10^9$ \\
    $\mu$                 & Extinction rate & 0.0, 0.1, 0.2 \\
    \hline
    $n$                   & Number of good taxa & 50, 100, 200 \\
    $t_c$                 & Crown age & 15 \\
    $\sigma_c$            & Standard deviation around crown age & 0.001 \\
    $M_s$                 & Sampling method & S, L, R \\
    $M_c$                 & Clock model & S, RLN \\
    $M_t$                 & Site model & JC69, GTR \\
    $r$                   & Mutation rate & $\frac{1}{15}$ \\
    $l_a$                 & DNA alignment length & $15K$ \\
    $f_i$                 & MCMC sampling interval & 1K or more \\
    $R_i$                 & RNG seed incipient tree and randomly sampled species tree & 1, 2, etc. \\
    $R_a$                 & RNG seed alignment simulation & $R_i$ \\
    $R_b$                 & RNG seed BEAST2 & $R_i$ \\
    \hline
  \end{tabular}
  \caption{
    Overview of the simulation parameters. Above the horizontal line is 
    the biological parameter set. 
    The RNG seed $R_i$ is 1 for the first simulation, 2 for the next,
    and so on.  
    The clock models are abbreviated as 'S' for a strict and 'RLN' for a relaxed log-normal model.
    The site models are abbreviated as 'JC69' for Jukes-Cantor (\cite{jc69}) and 'GTR' for the generalized 
    time-reversible model (\cite{gtr}).
  }
  \label{table:parameters}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
  \centering 
  \begin{tabular}{l l}
    \hline
    $n$ & Description \\
    \hline
    \hline
    $12$ \richel{recalc} & simulation parameters, see table \ref{table:parameters} \\
    $1000$ & nLTT statistic values \\
    $11$ & ESSes of all parameters estimated by BEAST2 (see specs below) \\
    \hline
  \end{tabular}
  \caption{
    Specification of the data sets. Each row will contain one experiment,
    where the columns contain parameters, measurements and diagnostics.
    This table displays the content of the columns. 
    $n$ denotes the number
    of columns a certain item will occupy, resulting in a table of 
    1023 \richel{recalc} columns and 20K rows.
  }
  \label{table:specs}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
  \centering 
  \begin{tabular}{l l}
    \hline
    \# & Description \\
    \hline
    \hline
    1 & posterior \\
    2 & likelihood \\
    3 & prior \\
    4 & treeLikelihood \\
    5 & TreeHeight \\
    6 & BirthDeath \\
    7 & BDBirthRate \\
    8 & BDDeathRate \\
    9 & logP.mrca \\
    10 & mrcatime \\
    11 & clockRate \\
    \hline
  \end{tabular}
  \caption{
    Overview of the 11 parameters estimated by BEAST2
  }
  \label{table:estimated_parameters}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgements}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\richel{put this section here, as the journal does not request for this}
We would like to thank the Center for Information Technology of the University of Groningen for their support
and for providing access to the Peregrine high performance computing cluster.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Authors' contributions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\richel{put this section here, as the journal does not request for this}
RSE \richel{@gio: I assume this this is true?} conceived the idea for this experiment.
GL created and tested the MBD package.
RJCB created and tested the experiment.
GL and RJCB wrote the first draft of the manuscript. 
RSE contributed substantially to revisions.

\end{document}

### 2018-09-26 article.tex
Add @Giappo comments

@@
 
-Current phylogenetic tools use tree priors 
-that assume speciation events are independent, 
-whilst we know multiple examples of adaptive radiations 
-in multiple species families can co-occur, 
-for example \richel{@gio: please add examples}.
+%Current phylogenetic tools use tree priors 
+%that assume speciation events are independent, 
+%whilst we know multiple examples of adaptive radiations 
+%in multiple species families can co-occur, 
+%for example 
+\richel{@gio: please add examples}.\gio{I've taken the freedom to change your proposed structure.
+Feel free to change whatever you don't like it.
+I might be not completely accurate on the biological background.}
+
+Current phylogenetic tools use tree priors that assume that only a single speciation event can occur at the same time.
+While this assumption has been proved to be useful to construct a wide variety of successful models, MBD model relaxes this
+hypothesis allowing for a description of a different kind of events, where the environmental changes can act as species pump.
+This kind of feature can be particularly efficient to describe those kinds of diversification process 
+characterized by a tremendous tempo, 
+where a great number of species is produced in relatively short time intervals. 
+A prime example of that could be given by Cichlid fish diversification in the African Great Lakes: Malawi, Tanganyika and Victoria.
 
@@
 \richel{explain model here, example is below in the comments}
+\gio{If I described the process in the same way you report in the example I would probably end up writing the same things that we say a few lines below, where we describe the parameters.
+Don't you think?}
 % In this model, a branching event does not give rise to a new species, but to
@@
 First, the computation of the MBD likelihood involves solving a set of 
-non-linear differential equations, and while this computation is quite fast, 
+non-linear differential equations \gio{@richel: are they actually non-linear?}, and while this computation is quite fast, 
 it still takes much more time than the corresponding probability 
@@
 
-The MBD model has \richel{@gio: how many?} parameters, depicted in table \ref{table:parameters}.
+The MBD model has 4 parameters, depicted in table \ref{table:parameters}.
+Parameters $\lambda$ and $\mu$ correspond, respectively, to the usual per-species speciation and extinction rates.
+The model also introduces two new parameters: $\nu$ is the total rate for an environmental change to trigger, which leads to a potentially multiple speciation event.
+If such event triggers each species present at that moment in time can undergo a speciation event with probability $q$.
 
-\richel{@gio: describe parameter values used here, example is below}
+\richel{@gio: describe parameter values used here, example is below}\gio{@richel: I described the meaning of each parameter.
+I guess it is fine in this way.
+For the setting I think we have to wait to decide which values are actually worth a try}
 % The per species speciation completion rates $\lambda$ we use are $0.1$, $0.3$, $1.0$ and $10^9$.
@@
 The first tree has a random number generator seed of 1, which is incremented by 1 for each simulated tree.
-For each combination of \richel{@gio: parameter values here}, 
-we generate incipient species trees with a crown age of 15 million years.
+For each combination of $(\lambda, \mu, \nu, q) $\richel{@gio: parameter values here}, 
+we generate incipient species trees with a crown age of 15 million years \gio{are we sure we wanna try 15 million years.
+so far i've been trying only with 10 million years, which most of the time is working really well}.
 Only trees with the desired number of good taxa are kept.

### 2018-09-26 article.tex
Giappo's comments start with 'GL:'
Wrote v0.3, fix #8

@@
 \newcommand*\richel[1]{\textcolor{orange}{\textbf{[RJCB: #1]}}}
-\newcommand*\gio[1]{\textcolor{green}{\textbf{[RJCB: #1]}}}
+\newcommand*\gio[1]{\textcolor{green}{\textbf{[GL: #1]}}}
 
@@
 
-Currently, the most popular Bayesian phylogenetics tools are \\ 
+Currently, the most popular Bayesian phylogenetics tools are
 BEAST (\cite{beast}) and its offshoot BEAST2 (\cite{beast2}), 
@@
 
-%Current phylogenetic tools use tree priors 
-%that assume speciation events are independent, 
-%whilst we know multiple examples of adaptive radiations 
-%in multiple species families can co-occur, 
-%for example 
-\richel{@gio: please add examples}.\gio{I've taken the freedom to change your proposed structure.
-Feel free to change whatever you don't like it.
-I might be not completely accurate on the biological background.}
+\gio{I've taken the freedom to change your proposed structure. Feel free to change whatever you don't like it. I might be not completely accurate on the biological background.}
+\richel{It's just fine :-)}
 
-Current phylogenetic tools use tree priors that assume that only a single speciation event can occur at the same time.
-While this assumption has been proved to be useful to construct a wide variety of successful models, MBD model relaxes this
-hypothesis allowing for a description of a different kind of events, where the environmental changes can act as species pump.
-This kind of feature can be particularly efficient to describe those kinds of diversification process 
-characterized by a tremendous tempo, 
-where a great number of species is produced in relatively short time intervals. 
-A prime example of that could be given by Cichlid fish diversification in the African Great Lakes: Malawi, Tanganyika and Victoria.
+Current phylogenetic tools assume that only a single speciation event can occur at the same time.
+While this assumption is useful to construct a wide variety of successful models \richel{@gio: citation here},
+they disallow for environmental changes that trigger speciation on a large scale,
+for example, the cichlid fish diversification in the 
+African Great Lakes: Malawi, Tanganyika and Victoria \richel{@gio: citation here}.
 
-The (constant-rate) birth-death (BD) model is a commonly 
-used tree prior, but it ignores the co-occurence of speciation.
-It even assume that two speciation events at exactly the same time
-has zero likelihood!
-The multiple birth-death (MBD) model, an extension of 
-the BD model, does incorporate the idea that speciation can co-occur.
+The (constant-rate) birth-death (BD) model embodies the common assumption that 
+only a single speciation event can occur at the same time.
+The MBD model relaxes this assumption, allowing events in which 
+large-scale environmental changes lead to a great number of species 
+in relatively short time intervals.
 
-\richel{explain model here, example is below in the comments}
 \gio{If I described the process in the same way you report in the example I would probably end up writing the same things that we say a few lines below, where we describe the parameters.
 Don't you think?}
-% In this model, a branching event does not give rise to a new species, but to
-% a new species-to-be, called an incipient species. Such an incipient
-% species may go extinct, finish its speciation to become a good species, or give
-% rise to new incipient species. Protracted speciation may explain observed 
-% declines in lineage accumulation (\cite{etienne_and_rosindell_2012}).
+\richel{You described the model in the Methods. I moved it to here}
+In the MBD model, parameters $\lambda$ and $\mu$ correspond, respectively, 
+to the usual per-species speciation and extinction rates. 
+The model also introduces two new parameters: 
+$\nu$ is the total rate for an environmental change to trigger, 
+which leads to a potentially multiple speciation event 
+\richel{How exactly? When the environment change is triggered, is the speciation rate $q$ or $\lambda + q$? What ends the change?}.
+If such event triggers each species present at that moment in time can undergo a speciation event with probability $q$.
 
@@
 The MBD model has 4 parameters, depicted in table \ref{table:parameters}.
-Parameters $\lambda$ and $\mu$ correspond, respectively, to the usual per-species speciation and extinction rates.
-The model also introduces two new parameters: $\nu$ is the total rate for an environmental change to trigger, which leads to a potentially multiple speciation event.
-If such event triggers each species present at that moment in time can undergo a speciation event with probability $q$.
 
-\richel{@gio: describe parameter values used here, example is below}\gio{@richel: I described the meaning of each parameter.
-I guess it is fine in this way.
-For the setting I think we have to wait to decide which values are actually worth a try}
+\richel{describe parameter values used here, example is below}
+\gio{@richel: I described the meaning of each parameter. I guess it is fine in this way. For the setting I think we have to wait to decide which values are actually worth a try}
+\richel{I assume you have some idea of parameter values? For example, I think you will not use a $\lambda$ of -3 or 1234. Please put some reasonable values here, like I did in Figure 1 for you}
 % The per species speciation completion rates $\lambda$ we use are $0.1$, $0.3$, $1.0$ and $10^9$.
@@
 % Our parameters are inspired by existing work on protracted speciation \cite{etienne_and_rosindell_2012}\cite{etienne2014}.
-
-We use \richel{@gio: parameter setting here} as our control for which the MBD model reduces to the BD model.
+We use \richel{please check me} $\nu = 0$ as our control for which the MBD model reduces to the BD model.
 
@@
 The first tree has a random number generator seed of 1, which is incremented by 1 for each simulated tree.
-For each combination of $(\lambda, \mu, \nu, q) $\richel{@gio: parameter values here}, 
-we generate incipient species trees with a crown age of 15 million years \gio{are we sure we wanna try 15 million years.
-so far i've been trying only with 10 million years, which most of the time is working really well}.
+For each combination of $\lambda, \mu, \nu$ and $q$, 
+we generate species trees with a crown age of 15 million years 
+\gio{are we sure we wanna try 15 million years. so far i've been trying only with 10 million years, which most of the time is working really well}
+\richel{I suggest yes. Most papers associated with Rampal use 15 million years, including my raket paper. Just use lower speciation rates}.
 Only trees with the desired number of good taxa are kept.

### 2018-09-27 article.tex
Updates from ShareLaTeX

@@
 
-\gio{I've taken the freedom to change your proposed structure. Feel free to change whatever you don't like it. I might be not completely accurate on the biological background.}
-\richel{It's just fine :-)}
-
 Current phylogenetic tools assume that only a single speciation event can occur at the same time.
-While this assumption is useful to construct a wide variety of successful models \richel{@gio: citation here},
+While this assumption is useful to construct a wide variety of successful models \richel{@gio: citation here} \gio{@richel: basically all the models we know are based on this assumpion: DDD, PBD, BISSE, MUSSE, SECSSE, any other SSE, DAISIE etc.
+etc.
+It's a very very general feature.
+Maybe being specific could lead the reader to consider things that are, in the end, not essential to the story we want to tell here.
+Do you still think we need it?},
 they disallow for environmental changes that trigger speciation on a large scale,
@@
 \richel{How exactly? When the environment change is triggered, is the speciation rate $q$ or $\lambda + q$? What ends the change?}.
+\gio{If the environmental change triggers only $q$ is taken in consideration.
+$\lambda + q$ is not even a possibility, as $\lambda$ is a rate while $q$ is a probability.
+Do you think we need to be clearer on that?}
 If such event triggers each species present at that moment in time can undergo a speciation event with probability $q$.
@@
 \gio{are we sure we wanna try 15 million years. so far i've been trying only with 10 million years, which most of the time is working really well}
-\richel{I suggest yes. Most papers associated with Rampal use 15 million years, including my raket paper. Just use lower speciation rates}.
+\richel{I suggest yes. Most papers associated with Rampal use 15 million years, including my raket paper. Just use lower speciation rates}
+\gio{In general it is ok for me.
+Keep in mind, though, that allowing multiple speciations may lead to an explosion in the number of species.
+Increasing the time by a factor of n usually means increasing the expected number of species at the present by a factor proportional to $e^n$}.
 Only trees with the desired number of good taxa are kept.
@@
 Of the remaining
-1000 MCMC states, the effective sample size (ESS) of the posterior 
+1000 MCMC states, the effective sample size (ESS) \gio{@richel: could it be nicer, when citing the "effective sample size" for the first time (in order to use the acronym thereafter), to put all in capital letters "Effective Sample Size"} of the posterior
 must at least be 200
@@
 
-The input trees generated with a \richel{@gio: parameter that is set to reduce the MBD model to BD} 
+The input trees generated with a \richel{@gio: parameter that is set to reduce the MBD model to BD}\gio{@richel: setting $nu = 0$ denies any chance for multiple speciations to occur.
+If so, $q$ setting becomes unimportant.
+As a convention is possible to set $q=0$}
 allow us to measure the noise of the experiment.
@@
 % Bibliography
+\gio{bibliography is missing. The only bib file present does not correspond to the bibliography showed in the pdf file.}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     \hline
-    $\lambda$             & \richel{@gio: MBD params here} & 0.1, 0.3, 1.0, $10^9$ \\
-    $\lambda$             & Speciation rate & 0.1, 0.3, 1.0, $10^9$ \\
-    $\mu$                 & Extinction rate & 0.0, 0.1, 0.2 \\
+    $\lambda$             & \gio{@richel: I insert some params that can make sense. Let's keep en eye on this anyways} & 0.1, 0.3, 1.0, $10^9$ \\
+    $\lambda$             & Per-species speciation rate & 0.2 \\
+    $\mu$                 & Per-species extinction rate & 0.0, 0.15 \\
+    $\nu$                 & Multiple speciation trigger rate & 1.5, 2.0, 2.5 \\
+    $q$                   & Per-species probability of multiple speciation  & 0.1, 0.15 \\
+    \hline
     \hline
@@
 \richel{put this section here, as the journal does not request for this}
-RSE \richel{@gio: I assume this this is true?} conceived the idea for this experiment.
+RSE \richel{@gio: I assume this this is true?} \gio{@richel: it is true} conceived the idea for this experiment.
 GL created and tested the MBD package.

### 2018-09-27 article.tex
Write v0.5, fix #11

--- 2018-09-26 article.tex
@@
 
-\gio{I've taken the freedom to change your proposed structure. Feel free to change whatever you don't like it. I might be not completely accurate on the biological background.}
-\richel{It's just fine :-)}
-
 Current phylogenetic tools assume that only a single speciation event can occur at the same time.
-While this assumption is useful to construct a wide variety of successful models \richel{@gio: citation here},
+While this assumption is useful to construct a wide variety of successful models \richel{@gio: citation here} \gio{@richel: basically all the models we know are based on this assumpion: DDD, PBD, BISSE, MUSSE, SECSSE, any other SSE, DAISIE etc.
+etc.
+It's a very very general feature.
+Maybe being specific could lead the reader to consider things that are, in the end, not essential to the story we want to tell here.
+Do you still think we need it?},
+\richel{Yes, I think here would be a fine spot to cite some of those models, 
+I think BiSSE, DAISIE, DDD and PBD would be appropriate} 
 they disallow for environmental changes that trigger speciation on a large scale,
@@
 to the usual per-species speciation and extinction rates. 
-The model also introduces two new parameters: 
-$\nu$ is the total rate for an environmental change to trigger, 
-which leads to a potentially multiple speciation event 
-\richel{How exactly? When the environment change is triggered, is the speciation rate $q$ or $\lambda + q$? What ends the change?}.
-If such event triggers each species present at that moment in time can undergo a speciation event with probability $q$.
+Additionally, $\nu$ is the rate at which an environmental change is triggered.
+When that event is triggered, all species at that moment
+have a probability $q$ to speciatiate (independent on $\lambda$).
+The number of species that speciate due to this can also be zero. 
+\richel{Is this correct?}.
 
@@
 The MBD model has 4 parameters, depicted in table \ref{table:parameters}. 
+We pick values of $\nu$ in such a way we expect a multiple speciation
+event to be triggered zero ($\nu = 0$), once, twice, four and eight times 
+\richel{I assume you can calculate the correct $\nu$}.
+We pick values of $q$ that are 0.0 (a speciation barrier at the
+triggered event), 0.25, 0.5 and 1.0.
+We set our extinction rate $\mu$ to 0.1 in all simulation.
+As we select our phylogenies on their number of lineages,
+we calculate $\lambda$ in a such a way that the mean expected number
+of lineages equals the desired numbers of taxa of 50, 100 and 200. 
+For $\nu = 0$, the model falls back to a standard BD model.
+Note that the $\lambda$ and $q$ have different units
+and it is a misconception to think that for $\lambda = q$ (already
+impossible due to their units) the MBD model would reduce to a BD model.
 
-\richel{describe parameter values used here, example is below}
-\gio{@richel: I described the meaning of each parameter. I guess it is fine in this way. For the setting I think we have to wait to decide which values are actually worth a try}
-\richel{I assume you have some idea of parameter values? For example, I think you will not use a $\lambda$ of -3 or 1234. Please put some reasonable values here, like I did in Figure 1 for you}
-% The per species speciation completion rates $\lambda$ we use are $0.1$, $0.3$, $1.0$ and $10^9$.
-This means that there is a $0.15 dt$ ($0.35 dt$, $1.0 dt$, $10^9 dt$) probability of speciation completion occurring in an infinitesimal time $dt$.
-% We use per species extinction rates of $\mu = \mu_g = \mu_i$ $0.0$, $0.1$ and $0.2$.
-% For each combination of $\lambda$, $\mu$ and $n$, we use a speciation initiation rate $b = b_i = b_g$ that gives a given value of the expected number of species.
-We use expected mean tree sizes $n$ of 50, 100 and 200 good taxa and the corresponding speciation initiation rates for all parameter combinations are shown in table \ref{table:pbd_parameters}.
-% Our parameters are inspired by existing work on protracted speciation \cite{etienne_and_rosindell_2012}\cite{etienne2014}.
-We use \richel{please check me} $\nu = 0$ as our control for which the MBD model reduces to the BD model.
-
-We simulate protracted birth-death trees, using the MBD package (\cite{pbd}) in the R programming language (\cite{r}).
+We simulate protracted birth-death trees, using the MBD package (\cite{mbd}) in the R programming language (\cite{r}).
 The first tree has a random number generator seed of 1, which is incremented by 1 for each simulated tree.
@@
 we generate species trees with a crown age of 15 million years 
-\gio{are we sure we wanna try 15 million years. so far i've been trying only with 10 million years, which most of the time is working really well}
-\richel{I suggest yes. Most papers associated with Rampal use 15 million years, including my raket paper. Just use lower speciation rates}.
+\gio{In general [15 million years] is ok for me.
+Keep in mind, though, that allowing multiple speciations may lead to an explosion in the number of species.
+Increasing the time by a factor of n usually means increasing the expected number of species at the present by a factor proportional to $e^n$}
+\richel{I know}.
 Only trees with the desired number of good taxa are kept.
 
-We create one data set to explore parameter space, 
-All the trees with the correct number of good species are kept.
-Based on the species tree, we simulate a DNA alignment that has the same history
-as this species tree, using the \verb;phangorn; package (\cite{phangorn}). 
-We set the nucleotides of the DNA alignment to follow a Jukes-Cantor (\cite{jc69})
-nucleotide substitution model, in which all nucleotide-to-nucleotide transitions
-are equally likely. 
+From an (MBD) species tree, we create a BEAST2 posterior using
+the 'pirouette' (\cite{pirouette}) R package: 
+'pirouette' first simulates a DNA alignment that has the same history
+as the species tree, using the \verb;phangorn; package (\cite{phangorn}).
 The DNA sequence of the root ancestor consists of four equally sized single-nucleotide 
-blocks of adenine, cytosine, guanine and thymine respectively. 
-For example, for a DNA sequence length of 12, this would be AAACCCGGGTTT. 
-The order of nucletides does not matter in this study, 
-because we do not consider several partitions of the sequence with their own parameters. 
-Only the frequency of occurrence matters.
+blocks of adenine, cytosine, guanine and thymine 
+respectively (for example, for a DNA sequence length of 12, this would 
+be AAACCCGGGTTT). Throughout evolutionary time, we use equal 
+mutation rates between the four DNA nucleotides, 
+also called the Jukes-Cantor (\cite{jc69}) nucleotide substitution model.
+The neat seperation of the nucleotides is for visualization and debugging
+purposes and has no effect in any other way. 
+The equal amount of nucleotides does matter, assuring any nucleotide
+mutation is equally likely to be observed.
+
 In our Bayesian inference (see below) we use the same site model as the (obviously correct) site model prior,
@@
 
-\richel{must rewrite, use pirouette as a starting point}
-From an alignment, we run a Bayesian analysis and 
-create a posterior distribution of trees and parameters
-using the \verb;pirouette; (\cite{pirouette}) package
-that sets the input parameters similar to BEAUti 2 and then runs BEAST2. 
-For our site model, we assume either a Jukes-Cantor or GTR nucleotide substitution model.
+From here, the 'babette' R package (\cite{babette})
+takes over and converts the DNA alignment to a BEAST2 posterior.
+We set up the BEAST2 analysis to assume either a Jukes-Cantor or GTR nucleotide substitution model.
 The Jukes-Cantor model is the correct one, as it is used for simulating that alignment,
@@
 Of the remaining
-1000 MCMC states, the effective sample size (ESS) of the posterior 
+1000 MCMC states, the Effective Sample Size (ESS) of the posterior 
 must at least be 200
@@
 
-The input trees generated with a \richel{@gio: parameter that is set to reduce the MBD model to BD} 
+The input trees generated with a $\nu = 0$, 
+in which all BEAST2's
+assumptions are met,
 allow us to measure the noise of the experiment.
@@
 
-For both data sets, we display the nLTT statistics distribution per
-biological parameter combination as a violin plot.
+We display the data set as an nLTT statistics distribution per
+parameter combination as a violin plot.
 We only show the nLTT distributions
@@
     \hline
-    $\lambda$             & \richel{@gio: MBD params here} & 0.1, 0.3, 1.0, $10^9$ \\
-    $\lambda$             & Speciation rate & 0.1, 0.3, 1.0, $10^9$ \\
-    $\mu$                 & Extinction rate & 0.0, 0.1, 0.2 \\
+    $\lambda$             & Per-species speciation rate & calculated \\
+    $\mu$                 & Per-species extinction rate & 0.0, 0.1 \\
+    $\nu$                 & Multiple speciation trigger rate & occurs never, once, twice, four and eight times \\
+    $q$                   & Per-species probability of multiple speciation & 0, 0.25, 0.5, 1.0 \\
+    \hline
     \hline
@@
     $\sigma_c$            & Standard deviation around crown age & 0.001 \\
-    $M_s$                 & Sampling method & S, L, R \\
     $M_c$                 & Clock model & S, RLN \\
@@
     $f_i$                 & MCMC sampling interval & 1K or more \\
-    $R_i$                 & RNG seed incipient tree and randomly sampled species tree & 1, 2, etc. \\
+    $R_i$                 & RNG seed MBD tree generation & 1, 2, etc. \\
     $R_a$                 & RNG seed alignment simulation & $R_i$ \\
@@
   \caption{
-    Overview of the simulation parameters. Above the horizontal line is 
-    the biological parameter set. 
+    Overview of the simulation parameters. Above the horizontal line are 
+    the MBD model's parameters. 
     The RNG seed $R_i$ is 1 for the first simulation, 2 for the next,
@@
 \richel{put this section here, as the journal does not request for this}
-RSE \richel{@gio: I assume this this is true?} conceived the idea for this experiment.
+RSE conceived the idea for this experiment.
 GL created and tested the MBD package.

### 2018-09-27 article.tex
Condition on the number of triggered events

@@
 We pick values of $\nu$ in such a way we expect a multiple speciation
-event to be triggered zero ($\nu = 0$), once, twice, four and eight times 
-\richel{I assume you can calculate the correct $\nu$}.
+event to be triggered zero ($\nu = 0$), once, twice , four and eight times 
+\richel{I assume you can calculate the correct $\nu$}. For each
+expected number of triggered events, we only keep those phylogenies
+that actually realized the expected number of triggered events.
 We pick values of $q$ that are 0.0 (a speciation barrier at the
@@
 The general data set has ?144 \richel{recalc} different combinations
-of biological parameter combinations, site and clock models.
-The data set to investigate sampling has ?552 \richel{recalc} different combinations
-of biological parameter combinations, site models, clock models 
-and sampling methods. The experiment is computationally intensive:
+of parameter combinations.
+The experiment is computationally intensive:
 pilot experiments show that the experiment takes roughly 100 days
@@
 We display the data set as an nLTT statistics distribution per
-parameter combination as a violin plot.
+parameter combination as a faceted violin plot, showing
+the effect of the number of species (a proxy for the amount of 
+information), the number of triggered events and the intensity of such
+a triggered event.
 We only show the nLTT distributions
@@
   \caption{
-    nLTT statistic distribution per biological parameter set, using the
-    general data set, 
+    nLTT statistic distribution per setup,
     under the (correct) assumptions of a strict clock and Jukes-Cantor site model.
@@
 \gio{bibliography is missing. The only bib file present does not correspond to the bibliography showed in the pdf file.}
+\richel{Weird. I see it both locally and on Overleaf. Sent email with my screenshots, show me yours to see if I can help}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2018-10-01 article.tex
Changes @Giappo

@@
 
-Current phylogenetic tools assume that only a single speciation event can occur at the same time.
+Current phylogenetic tools assume that only a single speciation event can occur at any given time.
 While this assumption is useful to construct a wide variety of successful models \richel{@gio: citation here} \gio{@richel: basically all the models we know are based on this assumpion: DDD, PBD, BISSE, MUSSE, SECSSE, any other SSE, DAISIE etc.
@@
 \richel{Yes, I think here would be a fine spot to cite some of those models, 
-I think BiSSE, DAISIE, DDD and PBD would be appropriate} 
-they disallow for environmental changes that trigger speciation on a large scale,
-for example, the cichlid fish diversification in the 
-African Great Lakes: Malawi, Tanganyika and Victoria \richel{@gio: citation here}.
+I think BiSSE, DAISIE, DDD and PBD would be appropriate}\gio{CLOSED} 
+(for example as in \cite{Maddison2007biSSE}, \cite{Valente2015}, \cite{etienne2012diversity}, \cite{etienne2014estimating})
+they disallow for environmental changes that trigger speciation on a large scale.
 
 The (constant-rate) birth-death (BD) model embodies the common assumption that 
-only a single speciation event can occur at the same time.
+only a single speciation event can occur at any given time.
 The MBD model relaxes this assumption, allowing events in which 
 large-scale environmental changes lead to a great number of species 
-in relatively short time intervals.
+in relatively short time intervals. Such hypothesis can be useful to describe, for example, 
+systems like cichlid fish diversification in the 
+African Great Lakes: Malawi, Tanganyika and Victoria (\cite{janzen2016}, \cite{janzen2017})\richel{@gio: citation here}\gio{CLOSED}.
 
@@
 Don't you think?}
-\richel{You described the model in the Methods. I moved it to here}
+\richel{You described the model in the Methods. I moved it to here}\gio{CLOSED}
 In the MBD model, parameters $\lambda$ and $\mu$ correspond, respectively, 
@@
 Additionally, $\nu$ is the rate at which an environmental change is triggered.
-When that event is triggered, all species at that moment
-have a probability $q$ to speciatiate (independent on $\lambda$).
+When such event is triggered, all species present in the phylogeny at that moment
+have a probability $q$ to speciate (independent on $\lambda$).
 The number of species that speciate due to this can also be zero. 
-\richel{Is this correct?}.
+\richel{Is this correct?}\gio{this is the code used for the simulation: "stats::rbinom(n = 1, size = N, prob = q)". It can actually yield zero. Now I am wondering if it is correct or not...}.
 
@@
 is unavailable in current Bayesian phylogenetic tools. 
-Whilst a likelihood equation has been derived (\richel{cite yourself here}),
+Whilst a likelihood equation has been derived (\richel{cite yourself here}\gio{CLOSED})(\cite{mbd}),
 it has not been implemented as tree prior yet. 
@@
 event to be triggered zero ($\nu = 0$), once, twice , four and eight times 
-\richel{I assume you can calculate the correct $\nu$}. For each
+\richel{I assume you can calculate the correct $\nu$} \gio{I dunno if it's possible to do so.
+This kind of calculation, especially for $\mu>0$ might be very cumbersome.
+One thing Rampal and I discussed to do was to use an equivalent $\lambda$ in BD to have a similar amount of mutations for both kind of simulated trees.
+In the code there should be already something like that (it might be not completely accurate though).
+I explain myself better: a) you simulate with some MBD par setup; b) you count the amount of mutations you had; c) from this number you decide what $lambda$ to use to simulate the BD process to have the (expected) same amount of mutations.
+It's probably much much easier than going from BD to MBD.}.
+For each
 expected number of triggered events, we only keep those phylogenies
@@
 \gio{bibliography is missing. The only bib file present does not correspond to the bibliography showed in the pdf file.}
-\richel{Weird. I see it both locally and on Overleaf. Sent email with my screenshots, show me yours to see if I can help}
+\richel{Weird. I see it both locally and on Overleaf. Sent email with my screenshots, show me yours to see if I can help}\gio{CLOSED}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2018-10-01 article.tex
Write v0.7, fix #18

@@
 Current phylogenetic tools assume that only a single speciation event can occur at any given time.
-While this assumption is useful to construct a wide variety of successful models \richel{@gio: citation here} \gio{@richel: basically all the models we know are based on this assumpion: DDD, PBD, BISSE, MUSSE, SECSSE, any other SSE, DAISIE etc.
-etc.
-It's a very very general feature.
-Maybe being specific could lead the reader to consider things that are, in the end, not essential to the story we want to tell here.
-Do you still think we need it?},
-\richel{Yes, I think here would be a fine spot to cite some of those models, 
-I think BiSSE, DAISIE, DDD and PBD would be appropriate}\gio{CLOSED} 
-(for example as in \cite{Maddison2007biSSE}, \cite{Valente2015}, \cite{etienne2012diversity}, \cite{etienne2014estimating})
-they disallow for environmental changes that trigger speciation on a large scale.
+While this assumption is useful to construct a wide variety of successful 
+models (for example: \cite{Maddison2007biSSE}, \cite{Valente2015}, \cite{etienne2012diversity}, \cite{etienne2014estimating}),
+they disallow for environmental changes that trigger speciations in multiple clades at a same point in time.
 
@@
 systems like cichlid fish diversification in the 
-African Great Lakes: Malawi, Tanganyika and Victoria (\cite{janzen2016}, \cite{janzen2017})\richel{@gio: citation here}\gio{CLOSED}.
+African Great Lakes: Malawi, Tanganyika and Victoria (\cite{janzen2016}, \cite{janzen2017}).
 
-\gio{If I described the process in the same way you report in the example I would probably end up writing the same things that we say a few lines below, where we describe the parameters.
-Don't you think?}
-\richel{You described the model in the Methods. I moved it to here}\gio{CLOSED}
 In the MBD model, parameters $\lambda$ and $\mu$ correspond, respectively, 
@@
 The number of species that speciate due to this can also be zero. 
-\richel{Is this correct?}\gio{this is the code used for the simulation: "stats::rbinom(n = 1, size = N, prob = q)". It can actually yield zero. Now I am wondering if it is correct or not...}.
+\richel{Is this correct?}
+\gio{
+  this is the code used for the simulation: 
+  "stats::rbinom(n = 1, size = N, prob = q)". 
+  It can actually yield zero. 
+  Now I am wondering if it is correct or not...
+}
+\richel{
+  I read in the PDF describing the likelihood 
+  derivation 'The $L^\lambda_m$ term takes into account, being in 
+  the state $Q^k_m(t)$, all the possible ways to have at least one birth event from whichever pool'.
+  I would suggested changing the code to
+  "1 + stats::rbinom(n = 1, size = N, prob = q)"
+  and test it somehow to be sure the sim matches the likelihood
+}	.
 
@@
 is unavailable in current Bayesian phylogenetic tools. 
-Whilst a likelihood equation has been derived (\richel{cite yourself here}\gio{CLOSED})(\cite{mbd}),
+Whilst a likelihood equation has been derived (\cite{mbd}),
 it has not been implemented as tree prior yet. 
@@
 First, the computation of the MBD likelihood involves solving a set of 
-non-linear differential equations \gio{@richel: are they actually non-linear?}, and while this computation is quite fast, 
+non-linear differential equations 
+\gio{are they actually non-linear?}
+\richel{Definitely, I see an exponentiation and combinatorial term, both are non-linear}, 
+and while this computation is quite fast, 
 it still takes much more time than the corresponding probability 
@@
 
+\richel{This setup is wrong, discuss described setup by Gio today}
 The MBD model has 4 parameters, depicted in table \ref{table:parameters}. 
@@
 event to be triggered zero ($\nu = 0$), once, twice , four and eight times 
-\richel{I assume you can calculate the correct $\nu$} \gio{I dunno if it's possible to do so.
-This kind of calculation, especially for $\mu>0$ might be very cumbersome.
-One thing Rampal and I discussed to do was to use an equivalent $\lambda$ in BD to have a similar amount of mutations for both kind of simulated trees.
-In the code there should be already something like that (it might be not completely accurate though).
-I explain myself better: a) you simulate with some MBD par setup; b) you count the amount of mutations you had; c) from this number you decide what $lambda$ to use to simulate the BD process to have the (expected) same amount of mutations.
-It's probably much much easier than going from BD to MBD.}.
-For each
+\gio{
+  One thing Rampal and I discussed to do was to use an 
+  equivalent $\lambda$ in BD to have a similar amount of mutations for both kind of simulated trees. 
+  In the code there should be already something like 
+  that (it might be not completely accurate though). I explain myself 
+  better: 
+  a) you simulate with some MBD par setup; 
+  b) you count the amount of mutations you had; 
+  c) from this number you decide what $lambda$ to use to simulate the BD process 
+  to have the (expected) same amount of mutations. It's probably much much easier 
+  than going from BD to MBD.
+}
+. For each
 expected number of triggered events, we only keep those phylogenies
@@
 we generate species trees with a crown age of 15 million years 
-\gio{In general [15 million years] is ok for me.
-Keep in mind, though, that allowing multiple speciations may lead to an explosion in the number of species.
-Increasing the time by a factor of n usually means increasing the expected number of species at the present by a factor proportional to $e^n$}
-\richel{I know}.
 Only trees with the desired number of good taxa are kept.
@@
 % Bibliography
-\gio{bibliography is missing. The only bib file present does not correspond to the bibliography showed in the pdf file.}
-\richel{Weird. I see it both locally and on Overleaf. Sent email with my screenshots, show me yours to see if I can help}\gio{CLOSED}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2018-10-04 article.tex
Updates from ShareLaTeX

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{abstract}
@@
   % comprehensible to a scientist in any discipline.
-  The tools for reconstructing phylogenetic relationships between taxonomic 
-  units (e.g. species) have become very advanced in the last three decades.
 
@@
   % scientists in related disciplines.
-  Among the most popular tools are Bayesian approaches, 
-  such as \\ BEAST, MrBayes and RevBayes, 
-  that use efficient tree sampling routines to create a posterior probability distribution 
-  of the phylogenetic tree. 
-  A feature of these approaches is the possibility to incorporate known 
-  or hypothesized structure of the phylogenetic tree through the tree prior. 
-  It has been shown that the effect of the prior on the posterior distribution 
-  of trees can be substantial.
 
@@
   % study.
-  Currently implemented tree priors assume that speciation events are independent,
-  where we know that speciation can coincide, for example, when trigger by a
-  larger geographic change.
 
@@
   % their equivalent).
-  Here we explore the effects of ignoring 
-  speciation co-occurence with an extensive simulation study.
 
@@
 
-  We compare the inferred tree to the simulated tree, and find that ....
-
 \end{abstract}
@@
 {\bf Keywords:} computational biology, evolution, phylogenetics, Bayesian analysis, tree prior
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\gio{@richel: A) I am throwing down some main points to follow to develop a convincing story that follows a clear logic.
+It might happen that some of these points are repeated and/or in the wrong order.
+We need to check that.}
+\gio{@richel: B) According to my fine graining approach we should at each step deepen every small section.
+At a certain level I think we can start to re-coarse-grain what we wrote to create the abstract.}
 
@@
 \section{Introduction}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{itemize}
 
-The computational tools that are currently available 
-to the phylogeneticists go beyond the wildest 
-imagination of those living four decades ago.
-Advances in computational power allowed the first cladograms to be inferred 
-from DNA alignments in 1981 (\cite{felsenstein1981}), and  
-the first Bayesian tools emerged in 1996 (\cite{rannala1996}),
-providing unprecedented flexibility in the setup of a phylogenetic model.
+\item Tools like BEAST provide the possibility to infer a phylogeny from genetic data (DNA, RNA, proteins).
 
-Currently, the most popular Bayesian phylogenetics tools are
-BEAST (\cite{beast}) and its offshoot BEAST2 (\cite{beast2}), 
-MrBayes (\cite{mrbayes}) and RevBayes (\cite{revbayes}). 
-They allow to incorporate known or hypothesized structure of a phylogenetic 
-tree-to-be-inferred through model priors. 
-With these priors and an alignment of DNA, RNA or protein sequences, 
-they create a sample of the posterior distribution
-of phylogenies and parameter estimates (of the models used as a prior), 
-in which more probable combinations are represented more often.
-Each of these tools use efficient tree sampling routines to rapidly create an 
-informative posterior.
+\item There are a number of options that can be set up while using BEAST.
 
-The model priors in Bayesian phylogenetic reconstruction 
-can be grouped into three categories: (1) site model, specifying 
-nucleotide substitutions, (2) clock model, specifying
-the rate of mutation per lineage in time, and (3) tree model, 
-constituting the speciation model underlying branching events (speciation) 
-and branch termination (extinction).
-The choice of site model (\cite{posada_and_buckley_2004}), 
-clock model (\cite{baele_et_al_2012}) 
-or tree prior (\cite{moller2018, yang_and_ranalla_2005}) is known to affect
-the posterior.
+\item Current limits in current tools.
 
-Current phylogenetic tools assume that only a single speciation event can occur at any given time.
-While this assumption is useful to construct a wide variety of successful 
-models (for example: \cite{Maddison2007biSSE}, \cite{Valente2015}, \cite{etienne2012diversity}, \cite{etienne2014estimating}),
-they disallow for environmental changes that trigger speciations in multiple clades at a same point in time.
+\item BEAST2 gives us the possibility to introduce new tree priors to infer phylogenies based on different assumptions.
 
+\item One of these is the multiple birth hypothesis.
+
+\item Multiple birth hypothesis can be useful to explain a phenomenon that have always puzzled evolutionary biologists: what are the drivers of the diversification processes for those phylogenies that show an impressive amount of speciation events in relatively short times.
 The (constant-rate) birth-death (BD) model embodies the common assumption that 
@@
 
-In the MBD model, parameters $\lambda$ and $\mu$ correspond, respectively, 
+\item However it can be argued that current tree priors could be good enough to detect such events involving, at the same time, a lower level of complexity.
+If this is the case one should always be more keen to adopt the simplest model.
+
+\item Here we present our study with the aim of proving that there is indeed the necessity of a new tree prior for such cases.
+
+\end{itemize}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Methods}
+
+\subsection{Model}
+\begin{itemize}
+
+\item Current phylogenetic tools assume that only a single speciation event can occur at any given time.
+While this assumption is useful to construct a wide variety of successful 
+models (for example: \cite{Maddison2007biSSE}, \cite{Valente2015}, \cite{etienne2012diversity}, \cite{etienne2014estimating}),
+they disallow for environmental changes that trigger speciations in multiple clades at a same point in time.
+
+\item The (constant-rate) birth-death (BD) model embodies the common assumption that 
+only a single speciation event can occur at any given time.
+The MBD model relaxes this assumption, allowing events in which 
+large-scale environmental changes lead to a great number of species 
+in relatively short time intervals. Such hypothesis can be useful to describe, for example, 
+systems like cichlid fish diversification in the 
+African Great Lakes: Malawi, Tanganyika and Victoria (\cite{janzen2016}, \cite{janzen2017}).
+
+\item In the MBD model, parameters $\lambda$ and $\mu$ correspond, respectively, 
 to the usual per-species speciation and extinction rates. 
@@
 The number of species that speciate due to this can also be zero.
-\richel{Is this correct?}
-\gio{
-  this is the code used for the simulation: 
-  "stats::rbinom(n = 1, size = N, prob = q)". 
-  It can actually yield zero. 
-  Now I am wondering if it is correct or not...
+
+\item It is also possible to write down a likelihood function for such processes as in \cite{mbd}.
+
+\end{itemize}
+
+\subsection{Simulations}
+\begin{itemize}
+
+\item To prove our hypothesis we simulate two twin datasets. All the simulations are produced in continuous time, using the Doob-Gillespie algorithm.
+
+\item We start simulating a number (100?) of trees for the MBD model.
+We then use, for each tree, the 'pirouette' package (\cite{pirouette}) to obtain the sequence alignments corresponding to the given tree.
+'pirouette' then calls BEAST to use these alignments to infer a posterior distribution of phylogenies.
+To perform this operation BEAST will use a BD prior.
+
+\item For each tree 
+\gio{QUESTION 1: Do we convert each single tree or the entire distribution?} 
+\richel{SUGGEST 1: Each tree, as (1) it's easier (2) I so no benefit (nor difference) with the second option} 
+generated under the MBD model we aim to generate a "twin" tree under the BD model in order to perform the same analysis and compare the results.
+To do so we need to be sure that the comparison is fair.
+The method we adopt to achieve this goal is to impose that the average number of substitution will be the same
+\gio{QUESTION 2: What criterion are we going to use? Same amount of information? Same number of tips? Something else? The answer has to be given mostly according to biological reasons, I think.}.
+\richel{SUGGEST 2: I would prefer the same amount of tips and mutations. See figure.}
+\item Here we explain how the conversion is done, with formulas if necessary.
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{mbd.jpg}
+  \caption{
+    How to create twin trees and alignments. From a focal MBD tree, a twin tree is produced as 
+    such: (1) estimate the $\lambda_{BD}$ to get the same expected number of tips, (2) simulate a BD tree with that amount of tips (discard trees with different number of tips), (3) estimate a mutation rate to get an alignment with the same expected number of mutations, (4) simulate alignments with that amount of mutations (discard those that don't, the picture shows an alignment that should be discarded)
   }
-\richel{
-  I read in the PDF describing the likelihood 
-  derivation 'The $L^\lambda_m$ term takes into account, being in 
-  the state $Q^k_m(t)$, all the possible ways to have at least one birth event from whichever pool'.
-  I would suggested changing the code to
-  "1 + stats::rbinom(n = 1, size = N, prob = q)"
-  and test it somehow to be sure the sim matches the likelihood
-}	.
+\end{figure}
 
-Unfortunately, a tree prior according to this model, 
-providing the probability of a species tree under the MBD model, 
-is unavailable in current Bayesian phylogenetic tools. 
-Whilst a likelihood equation has been derived (\cite{mbd}),
-it has not been implemented as tree prior yet. 
-There are various reasons for this.
-First, the computation of the MBD likelihood involves solving a set of 
-non-linear differential equations 
-\gio{are they actually non-linear?}
-\richel{Definitely, I see an exponentiation and combinatorial term, both are non-linear}, 
-and while this computation is quite fast, 
-it still takes much more time than the corresponding probability 
-of the BD model which is a simple analytical formula. 
-In a Bayesian MCMC chain, the tree prior probability must be calculated many times, 
-and hence the total computation will take considerably longer with a PBD tree prior.
+\item We explained how we set the parameters for each twin BD tree.
+Using this rules we generate a BD dateset.
+We repeat the raket analysis, producing alignments for each tree and subsequently using BEAST to produce a posterior for each of them.
 
-Here we aim to explore the effect of using the
-BD prior on MBD simulated phylogenies.
-In brief, we simulate phylogenies with co-occuring speciation events using the MBD process. 
-Given this species tree, we simulate a DNA sequence alignment. Then, we use BEAST2 on these alignments
-to infer a posterior of phylogenies, using a BD prior. We quantify the difference
-between the (BD) posterior phylogenies and the simulated (MBD) species tree.
-Furthermore, while we evidently know the clock and site models used in the simulation, 
-using a different clock and/or site model prior in inference 
-may compensate or increase this difference between inferred and simulated tree. 
-To study this, we also explore the effect of 
-a different clock and site model prior in inference.
+\item Now we have to datasets of posteriors to compare, one for the BD model and one for the MBD model.
 
+\item To compare the results for the two models we measure the inference error using the nLTT statistic between known/true tree and posterior/inferred trees.
+\richel{Rewrote this a bit.
+I suggest to drop the Bayes factor for now.
+It can be used if we would let BEAST2 infer with different nucleotide substitution models (a.k.a.
+site models), to answer the question: which nucleotide substitution model should be used in inference?
+Using different site models was Rampal's idea and I hope we can cite the raket paper to demonstrate it does not matter}
+
+\item The BF method allows us to compare the two marginal distributions. \gio{insert mathematical formula}
+
+\end{itemize}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Methods (but we are not allowed to use this header)
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\richel{This setup is wrong, discuss described setup by Gio today}
-The MBD model has 4 parameters, depicted in table \ref{table:parameters}. 
-We pick values of $\nu$ in such a way we expect a multiple speciation
-event to be triggered zero ($\nu = 0$), once, twice , four and eight times 
-\gio{
-  One thing Rampal and I discussed to do was to use an 
-  equivalent $\lambda$ in BD to have a similar amount of mutations for both kind of simulated trees. 
-  In the code there should be already something like 
-  that (it might be not completely accurate though). I explain myself 
-  better: 
-  a) you simulate with some MBD par setup; 
-  b) you count the amount of mutations you had; 
-  c) from this number you decide what $lambda$ to use to simulate the BD process 
-  to have the (expected) same amount of mutations. It's probably much much easier 
-  than going from BD to MBD.
-}
-. For each
-expected number of triggered events, we only keep those phylogenies
-that actually realized the expected number of triggered events.
-We pick values of $q$ that are 0.0 (a speciation barrier at the
-triggered event), 0.25, 0.5 and 1.0.
-We set our extinction rate $\mu$ to 0.1 in all simulation.
-As we select our phylogenies on their number of lineages,
-we calculate $\lambda$ in a such a way that the mean expected number
-of lineages equals the desired numbers of taxa of 50, 100 and 200. 
-For $\nu = 0$, the model falls back to a standard BD model.
-Note that the $\lambda$ and $q$ have different units
-and it is a misconception to think that for $\lambda = q$ (already
-impossible due to their units) the MBD model would reduce to a BD model.
-
-We simulate protracted birth-death trees, using the MBD package (\cite{mbd}) in the R programming language (\cite{r}).
-The first tree has a random number generator seed of 1, which is incremented by 1 for each simulated tree.
-For each combination of $\lambda, \mu, \nu$ and $q$, 
-we generate species trees with a crown age of 15 million years 
-Only trees with the desired number of good taxa are kept.
-
-From an (MBD) species tree, we create a BEAST2 posterior using
-the 'pirouette' (\cite{pirouette}) R package: 
-'pirouette' first simulates a DNA alignment that has the same history
-as the species tree, using the \verb;phangorn; package (\cite{phangorn}).
-The DNA sequence of the root ancestor consists of four equally sized single-nucleotide 
-blocks of adenine, cytosine, guanine and thymine 
-respectively (for example, for a DNA sequence length of 12, this would 
-be AAACCCGGGTTT). Throughout evolutionary time, we use equal 
-mutation rates between the four DNA nucleotides, 
-also called the Jukes-Cantor (\cite{jc69}) nucleotide substitution model.
-The neat seperation of the nucleotides is for visualization and debugging
-purposes and has no effect in any other way. 
-The equal amount of nucleotides does matter, assuring any nucleotide
-mutation is equally likely to be observed.
-
-In our Bayesian inference (see below) we use the same site model as the (obviously correct) site model prior,
-but we also explore the effect of assuming a more complex site model prior.
-We predict with the more complex substitution model, 
-that there will be more noise and hence our inference error will increase.
-On the other hand, we dare not rule out that the inference error will decrease,
-due to more flexibility in the more complex prior.
-We set the mutation rate in such a way to maximize the information contained in the alignment.
-To do so, we set the mutation rate such that we expect on average one (possibly silent) mutation per nucleotide
-between crown age and present, which equates to $\frac{1}{15}$ mutations
-per million years.
-The DNA sequence length is chosen to provide a
-resolution of $10^3$ years, 
-% If sims take too long, we will decrease this resolution
-that is, to have one expected nucleotide change 
-per $10^3$ years per lineage on average. As one nucleotide is expected 
-to have on average one (possibly silent) mutation per 15 million years, $15 \cdot 10^3$
-nucleotides result in 1 mutation per alignment per $10^3$ years (which is
-coincidentally the same as \cite{moller2018}). 
-The simulation of these DNA aligments follows a strict clock model, 
-which we will specify as one of the two clock models assumed in the Bayesian inference (see below).
-
-From here, the 'babette' R package (\cite{babette})
-takes over and converts the DNA alignment to a BEAST2 posterior.
-We set up the BEAST2 analysis to assume either a Jukes-Cantor or GTR nucleotide substitution model.
-The Jukes-Cantor model is the correct one, as it is used for simulating that alignment,
-where the GTR model is the site model that is picked as a default by most users.
-For our clock model, we assume either a strict or relaxed log-normal 
-clock model. 
-Also here, the strict clock model is the correct one, as it is used for simulating the alignment,
-but the relaxed log-normal clock model is the one most commonly used.
-We set the BD model as a tree prior, 
-as gauging the effect of this incorrect assumption is the goal of this study. 
-We assume an MRCA prior with a tight normal distribution
-around the crown age, by choosing the crown age as mean, and a standard deviation 
-of $0.5 \cdot 10^{-3}$ time units,
-resulting in 95\% of the crown ages inferred have the same resolution (of $10^{-3}$ time 
-units) as the alignment. 
-We ran the MCMC chain to generate 1111 states,
-of which we remove the first 10\% (also called the 'burn-in'). 
-Of the remaining
-1000 MCMC states, the Effective Sample Size (ESS) of the posterior 
-must at least be 200
-for a strong enough inference (\cite{beastbook}). An ESS can be increased by increasing
-the number of samples or decreasing the autocorrelation between samples. 
-If the ESS is less than 200, we decrease autocorrelation by doubling 
-the MCMC sampling interval of that simulation, until the ESS exceeds 200.
-
-We compare each posterior phylogeny to the (sampled) species tree
-using the nLTT statistic (\cite{janzen2015}), from the \verb;nLTT; package (\cite{nltt}). 
-The nLTT statistic equals the area between the normalized
-lineages-through-time-plots of two phylogenies, which has a range 
-from zero (for identical phylogenies) to one. We use inference error 
-and nLTT statistic interchangeably. Comparing the simulated species tree
-with each of the posterior species trees yields a distribution of nLTT statistics.
-
-The input trees generated with a $\nu = 0$, 
-in which all BEAST2's
-assumptions are met,
-allow us to measure the noise of the experiment.
-
-We produce one data set as a comma-separated file.
-The general data set has ?144 \richel{recalc} different combinations
-of parameter combinations.
-The experiment is computationally intensive:
-pilot experiments show that the experiment takes roughly 100 days
-of CPU time and 20 days of wall clock time (which includes the queued 
-waiting for computational resources) per replicate. 
-Due to this, we choose to perform ten replicates, so that the complete
-experiment will take an acceptable time of roughly seven months.
-
-We display the data set as an nLTT statistics distribution per
-parameter combination as a faceted violin plot, showing
-the effect of the number of species (a proxy for the amount of 
-information), the number of triggered events and the intensity of such
-a triggered event.
-We only show the nLTT distributions
-that were generated under the (correct) assumptions of a Jukes-Cantor site model
-and a strict clock model, separated per sampling method used. 
-We display the nLTT statistic distributions separated per site or clock model 
-in the supplementary information.
 
@@
 \section{Results}
+\begin{itemize}
+
+\item
+
+\item
+
+\end{itemize}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_general.png}
-  \caption{
-    nLTT statistic distribution per setup,
-    under the (correct) assumptions of a strict clock and Jukes-Cantor site model.
-  }
-\end{figure}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Glossary}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Please supply, as a separate list, the definitions of field-specific terms used in your article.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table}
-  \centering 
-  \begin{tabular}{l p{0.65\textwidth}}
-    \hline
-    Term                  & Definition \\
-    \hline
-    \hline
-    Phylogenetics         & The inference of evolutionary relationships of groups of organisms using genetics \\
-    Model prior           & Knowledge or assumptions about the ontogeny of evolutionary histories \\
-    Posterior             & A collection of phylogenies and parameter estimates, in which more probable combinations (determined by the data and the model prior) are presented more frequently \\
-    \hline
-  \end{tabular}
-  \caption{
-    Glossary
-  }
-  \label{table:glossary}
-\end{table}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% Bibliography
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% MEE style
+% Bibliography % MEE style
 \bibliographystyle{mee}
@@
 \appendix
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_clock_model.png}
-  \caption{
-    nLTT statistic distribution per biological parameter set per clock model,
-    using the general data set, 
-    under the (correct) assumption of a Jukes-Cantor site model.
-  }
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_site_model.png}
-  \caption{
-    nLTT statistic distribution per biological parameter set per site model,
-    using the general data set, 
-    under the (correct) assumption of a strict clock model.
-  }
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table}
-  \centering 
-  \begin{tabular}{p{0.05\textwidth} p{0.6\textwidth} p{0.2\textwidth}}
-    \hline
-                          & Description & Values \\
-    \hline
-    \hline
-    $\lambda$             & Per-species speciation rate & calculated \\
-    $\mu$                 & Per-species extinction rate & 0.0, 0.1 \\
-    $\nu$                 & Multiple speciation trigger rate & occurs never, once, twice, four and eight times \\
-    $q$                   & Per-species probability of multiple speciation & 0, 0.25, 0.5, 1.0 \\
-    \hline
-    \hline
-    $n$                   & Number of good taxa & 50, 100, 200 \\
-    $t_c$                 & Crown age & 15 \\
-    $\sigma_c$            & Standard deviation around crown age & 0.001 \\
-    $M_c$                 & Clock model & S, RLN \\
-    $M_t$                 & Site model & JC69, GTR \\
-    $r$                   & Mutation rate & $\frac{1}{15}$ \\
-    $l_a$                 & DNA alignment length & $15K$ \\
-    $f_i$                 & MCMC sampling interval & 1K or more \\
-    $R_i$                 & RNG seed MBD tree generation & 1, 2, etc. \\
-    $R_a$                 & RNG seed alignment simulation & $R_i$ \\
-    $R_b$                 & RNG seed BEAST2 & $R_i$ \\
-    \hline
-  \end{tabular}
-  \caption{
-    Overview of the simulation parameters. Above the horizontal line are 
-    the MBD model's parameters. 
-    The RNG seed $R_i$ is 1 for the first simulation, 2 for the next,
-    and so on.  
-    The clock models are abbreviated as 'S' for a strict and 'RLN' for a relaxed log-normal model.
-    The site models are abbreviated as 'JC69' for Jukes-Cantor (\cite{jc69}) and 'GTR' for the generalized 
-    time-reversible model (\cite{gtr}).
-  }
-  \label{table:parameters}
-\end{table}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table}
-  \centering 
-  \begin{tabular}{l l}
-    \hline
-    $n$ & Description \\
-    \hline
-    \hline
-    $12$ \richel{recalc} & simulation parameters, see table \ref{table:parameters} \\
-    $1000$ & nLTT statistic values \\
-    $11$ & ESSes of all parameters estimated by BEAST2 (see specs below) \\
-    \hline
-  \end{tabular}
-  \caption{
-    Specification of the data sets. Each row will contain one experiment,
-    where the columns contain parameters, measurements and diagnostics.
-    This table displays the content of the columns. 
-    $n$ denotes the number
-    of columns a certain item will occupy, resulting in a table of 
-    1023 \richel{recalc} columns and 20K rows.
-  }
-  \label{table:specs}
-\end{table}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table}
-  \centering 
-  \begin{tabular}{l l}
-    \hline
-    \# & Description \\
-    \hline
-    \hline
-    1 & posterior \\
-    2 & likelihood \\
-    3 & prior \\
-    4 & treeLikelihood \\
-    5 & TreeHeight \\
-    6 & BirthDeath \\
-    7 & BDBirthRate \\
-    8 & BDDeathRate \\
-    9 & logP.mrca \\
-    10 & mrcatime \\
-    11 & clockRate \\
-    \hline
-  \end{tabular}
-  \caption{
-    Overview of the 11 parameters estimated by BEAST2
-  }
-  \label{table:estimated_parameters}
-\end{table}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Acknowledgements}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\richel{put this section here, as the journal does not request for this}
-We would like to thank the Center for Information Technology of the University of Groningen for their support
-and for providing access to the Peregrine high performance computing cluster.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Authors' contributions}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\richel{put this section here, as the journal does not request for this}
-RSE conceived the idea for this experiment.
-GL created and tested the MBD package.
-RJCB created and tested the experiment.
-GL and RJCB wrote the first draft of the manuscript. 
-RSE contributed substantially to revisions.
-
 \end{document}

### 2018-10-04 article.tex
Updates from ShareLaTeX

@@
 At a certain level I think we can start to re-coarse-grain what we wrote to create the abstract.}
+\richel{I enjoy this approach! And I know Rampal does not :-)}
+\richel{Have you already looked up for a target journal? I know how a journal's constraints have helped me 
+in writing an article}
 
@@
 only a single speciation event can occur at any given time.
-The MBD model relaxes this assumption, allowing events in which 
+The Multiple-Birth-Death (MBD) model \richel{I feel MBSD (Multiple Birth Single Death) may be a better name: extinctions are still one at a time} relaxes this assumption, allowing events in which 
 large-scale environmental changes lead to a great number of species 
@@
 have a probability $q$ to speciate (independent on $\lambda$).
-The number of species that speciate due to this can also be zero.
+The number of species that speciate due to this can also be zero .
 

### 2018-10-04 article.tex
New version, fix #21

--- 2018-10-04 article.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\gio{@richel: A) I am throwing down some main points to follow to develop a convincing story that follows a clear logic.
-It might happen that some of these points are repeated and/or in the wrong order.
-We need to check that.}
-\gio{@richel: B) According to my fine graining approach we should at each step deepen every small section.
-At a certain level I think we can start to re-coarse-grain what we wrote to create the abstract.}
+\gio{According to my fine graining approach we should at each step deepen every small section. At a certain level I think we can start to re-coarse-grain what we wrote to create the abstract.}
+\richel{I enjoy this approach! Did some minor fine-graining}
+\richel{Have you already looked up for a target journal? I know how a journal's constraints have helped me 
+in writing an article, for example, by having a maximum number of pictures}
 
@@
 
-\item Tools like BEAST provide the possibility to infer a phylogeny from genetic data (DNA, RNA, proteins).
+\item There are many contemporary tools that provide the possibility to infer a phylogeny from genetic data (DNA, RNA, proteins). A popular Bayesian
+phylogenetic tool is called BEAST and its cousin BEAST2.
 
-\item There are a number of options that can be set up while using BEAST.
+\item BEAST is very flexible in setting up all possible phylogenetic priors (e.g. site/clock/speciation model).
 
@@
 
-\item BEAST2 gives us the possibility to introduce new tree priors to infer phylogenies based on different assumptions.
+\item BEAST2 gives us the possibility to introduce new tree priors to infer phylogenies based on different assumptions on how the speciation process
+takes place.
 
-\item One of these is the multiple birth hypothesis.
+\item One of such speciation processes is the multiple birth hypothesis,
+a new model (described below) and thus absent in BEAST.
 
-\item Multiple birth hypothesis can be useful to explain a phenomenon that have always puzzled evolutionary biologists: what are the drivers of the diversification processes for those phylogenies that show an impressive amount of speciation events in relatively short times.
+\item The Multiple birth hypothesis can be useful to explain a phenomenon that has always puzzled evolutionary biologists: what are the drivers of the diversification processes for those phylogenies that show an impressive amount of speciation events in relatively short times?
 The (constant-rate) birth-death (BD) model embodies the common assumption that 
 only a single speciation event can occur at any given time.
-The MBD model relaxes this assumption, allowing events in which 
+The multiple-birth-death (MBD) model \richel{I feel MBSD (Multiple Birth Single Death) may be a better name: extinctions are still one at a time} relaxes this assumption, allowing events in which 
 large-scale environmental changes lead to a great number of species 
-in relatively short time intervals. Such hypothesis can be useful to describe, for example, 
-systems like cichlid fish diversification in the 
+in relatively short time intervals. Such a hypothesis may be a better fit to describe the burst in cichlid fish diversification in systems like in the 
 African Great Lakes: Malawi, Tanganyika and Victoria (\cite{janzen2016}, \cite{janzen2017}).
 
-\item However it can be argued that current tree priors could be good enough to detect such events involving, at the same time, a lower level of complexity.
+\item However, it may be that current BD tree priors are good enough at detecting such events, with a (preferred) lower level of complexity.
 If this is the case one should always be more keen to adopt the simplest model.
 
-\item Here we present our study with the aim of proving that there is indeed the necessity of a new tree prior for such cases.
+\item Here we present our study with the aim of exploring when using a more complex MBD tree prior is warranted.
 
@@
 only a single speciation event can occur at any given time.
-The MBD model relaxes this assumption, allowing events in which 
+The multiple-birth-death (MBD) model relaxes this assumption, allowing events in which 
 large-scale environmental changes lead to a great number of species 
@@
 \item In the MBD model, parameters $\lambda$ and $\mu$ correspond, respectively, 
-to the usual per-species speciation and extinction rates. 
+to the common per-species speciation and extinction rates. 
 Additionally, $\nu$ is the rate at which an environmental change is triggered.
 When such event is triggered, all species present in the phylogeny at that moment
-have a probability $q$ to speciate (independent on $\lambda$).
-The number of species that speciate due to this can also be zero.
+have a probability $q$ to speciate at that time, which is 
+independent on $\lambda$.
 
@@
 
-\item We start simulating a number (100?) of trees for the MBD model.
-We then use, for each tree, the 'pirouette' package (\cite{pirouette}) to obtain the sequence alignments corresponding to the given tree.
-'pirouette' then calls BEAST to use these alignments to infer a posterior distribution of phylogenies.
-To perform this operation BEAST will use a BD prior.
+\item We start simulating 1000 \richel{I will measure the number of trees we'll be able to simulate within a short enough time, when the experiment is set up} MBD trees.
+From each MBD tree, a a DNA sequence alignment is simulated, after
+which that alignment starts a Bayesian analysis. We use 
+the 'pirouette' package (\cite{pirouette}) to call the BEAST2 tool 
+suite from R. We let the Bayesian analysis assume a BD prior, to investigate
+the error this inference makes due to this.
 
 \item For each tree 
-\gio{QUESTION 1: Do we convert each single tree or the entire distribution?} 
-\richel{SUGGEST 1: Each tree, as (1) it's easier (2) I so no benefit (nor difference) with the second option} 
 generated under the MBD model we aim to generate a "twin" tree under the BD model in order to perform the same analysis and compare the results.
 To do so we need to be sure that the comparison is fair.
-The method we adopt to achieve this goal is to impose that the average number of substitution will be the same
-\gio{QUESTION 2: What criterion are we going to use? Same amount of information? Same number of tips? Something else? The answer has to be given mostly according to biological reasons, I think.}.
-\richel{SUGGEST 2: I would prefer the same amount of tips and mutations. See figure.}
-\item Here we explain how the conversion is done, with formulas if necessary.
+The method we adopt to achieve this goal is to impose that the amount of information used in the Bayesian information will be the same.
+These are the same number of taxa and the same (expected)
+number of DNA mutations. See appendix for the exact procedure.
 
@@
 Using this rules we generate a BD dateset.
-We repeat the raket analysis, producing alignments for each tree and subsequently using BEAST to produce a posterior for each of them.
+We repeat the analysis, producing alignments for each tree and subsequently using BEAST to produce a posterior for each of them.
 
-\item Now we have to datasets of posteriors to compare, one for the BD model and one for the MBD model.
+\item Now we have two datasets of posteriors to compare, one for the BD model and one for the MBD model.
 
 \item To compare the results for the two models we measure the inference error using the nLTT statistic between known/true tree and posterior/inferred trees.
-\richel{Rewrote this a bit.
-I suggest to drop the Bayes factor for now.
-It can be used if we would let BEAST2 infer with different nucleotide substitution models (a.k.a.
-site models), to answer the question: which nucleotide substitution model should be used in inference?
-Using different site models was Rampal's idea and I hope we can cite the raket paper to demonstrate it does not matter}
+\richel{
+  I would love to describe this more concrete. For example, when do
+  we say something has an effect? If we avoid making such judgements,
+  how will we visualize?
+}
 
-\item The BF method allows us to compare the two marginal distributions. \gio{insert mathematical formula}
+\richel{I removed the Bayes factor text. It is useful when letting BEAST2
+pick more/overly complex models and see if that more complex model fits the
+data better (penalized by its increased complexity, similar to the AIC). It has
+its uses, but I am unsure if we already want to discuss this now
+or first focus on the proper tree twinning}
 
@@
 \item
+\richel{
+  I guess you know I am a fan of the Open Science Framework,
+  in which you first register you work before you do the experiment
+  (note: I will do some small pilots to estimate the complete time
+  of the experiment). I think it is the proper and superior science,
+  which helps us against writing down bullshit stories after having
+  obtained the results (e.g. 'We expected A and indeed found it!').
+  It also helps me structure my work: first think
+  deeply about the experiment, then do it (instead of the mixing
+  up the two phases). What are your thoughts on that?
+}
 
@@
 
+\section{Creating twin trees}
+
+\richel{I put your work here and LaTeXed it. Maybe this moves
+to the Methods, but this seemed a good place to start}
+
+The expected number of mutations, $m$ of a phylogeny 
+with crown age $T$,
+depends on the number of DNA nucleotides $L$,
+per-site mutation rate $\mu$ and the number of species $n$
+of time as such:
+
+\begin{equation}
+m = L \cdot \mu \cdot \int_{t=0}^{T} n_t\ dt
+\end{equation}
+
+\richel{I suggest to use a different symbol for mutation rate, 
+  for example $\alpha$ or $\rho$. Using lambda here as well feels
+  needlessly confusing.
+}
+
+It can be easily seen considering that $\mu$ is the number of mutations
+per unit time per site. For this reason it's needed to multiply by
+time and number of sites.
+
+To obtain twin trees with an equal amount of information, we impose
+that the expected number of mutations in an MBD tree, $m_{MBD}$ equals
+the expected number of mutations in a BD tree, $m_{BD}$:
+
+\begin{equation}
+m_{MBD} = m_{BD}
+\end{equation}
+
+From which follows:
+
+\begin{equation}
+m_{BD} = L \cdot \mu \cdot \int_{t=0}^{T} n_{BD}(t)\ dt
+\end{equation}
+
+Obviously we cannot know $n_{BD}(t)$ before running the simulations but
+we can replace it with a proxy, for example the average number of
+species in time according to the BD model
+\richel{I suggest $n_{BD} = n_{MBD}$ and only 
+change $\mu_{BD}$ to reach $m_{MBD} = m_{BD}$}
+
+\gio{START}
+
+$$
+<n_{BD}(t)> = n(t=0) * exp[(mu - lambda) * t] (6)
+$$
+
+The relation to use to get the equivalent lambda should therefore be
+
+$$
+m_{MBD} 
+  = L * mu * integral[exp[(mu - lambda) * t] dt] 
+  = (L * mu * exp[(mu - lambda) * t]) / (mu - lambda) (7)
+$$
+
+Here everything is known but lambda. So solve the lambda and use that value.
+
+My doubt is if we need to use $m_{MBD}$ for the single tree or the same
+quantity averaged on the full MBD dataset $<m_{MBD}>$.
+
+Does it make sense to you? Do you think is better to use the
+individual $m_{MBD}$ for each tree or the average across the whole
+dataset?
+
+\gio{END}
+
+\richel{I think I see your point: in single trees stochasticity works per tree twin, in a distribution per distribution. To circumvent this: what if I would be 
+able to measure the actual number of mutations? Then twinning trees by
+having an exact equal amount of mutations would be very clean. Let me know: 
+you think I should investigate this?}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2018-10-05 article.tex
Changes @Giappo

@@
 
+To obtain twin trees with an equal amount of information, we impose
+that the expected number of mutations in an MBD tree, $m_{MBD}$ equals
+the expected number of mutations in a BD tree, $m_{BD}$:
+
+\begin{equation}
+m_{MBD} = m_{BD} \label{m equivalence}
+\end{equation}
+
+We first generate a set of MBD trees. For each of them we can measure the amount of mutations $m_{MBD}$.
+
 \richel{I put your work here and LaTeXed it. Maybe this moves
 to the Methods, but this seemed a good place to start}
+\gio{Thanks Richel, It was actually the next thing that I planned to do. I think this should definitely go to the methods}
 
-The expected number of mutations, $m$ of a phylogeny 
-with crown age $T$,
-depends on the number of DNA nucleotides $L$,
-per-site mutation rate $\mu$ and the number of species $n$
-of time as such:
+The number of mutations $m$ of a phylogeny 
+with crown age $-T$ in fact is given by
 
 \begin{equation}
-m = L \cdot \mu \cdot \int_{t=0}^{T} n_t\ dt
+m = L \cdot \rho \cdot \int_{0}^{T} n(t)\ dt \label{m calculation}
 \end{equation}
+
+where $L$ is the number of DNA nucleotides, 
+$\rho$ is the per-site per-species mutation rate and
+$n(t)$ the number of species at each time.
 
@@
 }
+\gio{ok}
 
-It can be easily seen considering that $\mu$ is the number of mutations
-per unit time per site. For this reason it's needed to multiply by
-time and number of sites.
+%It can be easily seen considering that $\rho$ is the number of mutations
+%per unit time per site. For this reason it's needed to multiply by
+%time and number of sites. 
+\gio{This feels kind of a repetitions of what we wrote before the formula. I comment it. We can think of reinsert it afterwards, if needed (see comment above).}
 
-To obtain twin trees with an equal amount of information, we impose
-that the expected number of mutations in an MBD tree, $m_{MBD}$ equals
-the expected number of mutations in a BD tree, $m_{BD}$:
+Since we cannot know $n_{BD}(t)$ before running simulations
+we need to replace it with a proxy. For this reason we will use the average number of
+species in time according to the BD model. 
+It's well known that this is equal to \gio{insert proper citation}
 
 \begin{equation}
-m_{MBD} = m_{BD}
+    <n_{BD}>(t) = n_{0} \cdot e^{(\mu_{BD} - \lambda_{BD})t} \label{BD average n}
 \end{equation}
 
-From which follows:
+where $n_{0} = n_{BD}(-T) = n_{MBD}(-T)$ is the initial number of species at the crown age.
+From \ref{m equivalence}, \ref{m calculation} and \ref{BD average n} follows:
 
 \begin{equation}
-m_{BD} = L \cdot \mu \cdot \int_{t=0}^{T} n_{BD}(t)\ dt
+m_{MBD} = L \cdot \rho \cdot \int_{0}^{T} <n_{BD}>(t)\ dt \\
+= L \cdot \rho \cdot n_{0} \cdot \left[ \frac{e^{(\mu_{BD} - \lambda_{BD})T} - 1}{\mu_{BD} - \lambda_{BD}} \right]
 \end{equation}
 
-Obviously we cannot know $n_{BD}(t)$ before running the simulations but
-we can replace it with a proxy, for example the average number of
-species in time according to the BD model
+If we set $\mu_{BD} = \mu_{MBD}$ and reverse this relation we can extrapolate the value of $\lambda_{BD}$ to use to generate BD trees.
+
 \richel{I suggest $n_{BD} = n_{MBD}$ and only 
 change $\mu_{BD}$ to reach $m_{MBD} = m_{BD}$}
+\gio{@Richel: Don't you think it might make more sense to set $\mu_{BD} = \mu_{MBD}$?
+What changes in the two model is the way we use to generate new species, not the way to remove them.
+Maybe one thing that's possible to do would be to make $\lambda_{BD}$ a function of time, a bit like Théo is doing in his comparison between DD and TD4 (which, by the way, seem to yield very different results).
+In case you are wondering the theory can be found in Caesar's master thesis.}
 
-\gio{START}
+%$$
+%<n_{BD}(t)> = n(t=0) * exp[(mu - lambda) * t] (6)
+%$$
 
-$$
-<n_{BD}(t)> = n(t=0) * exp[(mu - lambda) * t] (6)
-$$
+%The relation to use to get the equivalent lambda should therefore be
 
-The relation to use to get the equivalent lambda should therefore be
+%$$ 
+%m_{MBD} 
+%  = L * mu * integral[exp[(mu - lambda) * t] dt] 
+%  = (L * mu * exp[(mu - lambda) * t]) / (mu - lambda) (7)
+%$$
 
-$$
-m_{MBD} 
-  = L * mu * integral[exp[(mu - lambda) * t] dt] 
-  = (L * mu * exp[(mu - lambda) * t]) / (mu - lambda) (7)
-$$
+%Here everything is known but lambda. So solve the lambda and use that value.
 
-Here everything is known but lambda. So solve the lambda and use that value.
-
-My doubt is if we need to use $m_{MBD}$ for the single tree or the same
+\gio{My doubt is if we need to use $m_{MBD}$ for the single tree or the same
 quantity averaged on the full MBD dataset $<m_{MBD}>$.
-
-Does it make sense to you? Do you think is better to use the
+Do you think is better to use the
 individual $m_{MBD}$ for each tree or the average across the whole
-dataset?
+dataset?}
 

### 2018-10-05 article.tex
Next iteration

@@
 
-\richel{I put your work here and LaTeXed it. Maybe this moves
-to the Methods, but this seemed a good place to start}
-\gio{Thanks Richel, It was actually the next thing that I planned to do. I think this should definitely go to the methods}
+\gio{I think this should definitely go to the methods}
+\richel{I put it there for use, hopefully at a spot you liked :-)}
 
-The number of mutations $m$ of a phylogeny 
-with crown age $-T$ in fact is given by
+The expected number of mutations $m$ of a phylogeny 
+\richel{
+  I think 'expected number of mutations' would be more correct.
+  Do you agree?
+}
+with crown age $T$ in fact is given by
+\richel{
+  above stood 'crown age -T'. 
+  I feel that a crown age is a positive number,
+  but I know you have had a good reason.
+  Perhaps better would be to write something explicit like:
+  t_now - t_crown = T.
+  Looking forward for a better suggestion than mine :-)
+}
 
@@
 
-\richel{I suggest to use a different symbol for mutation rate, 
-  for example $\alpha$ or $\rho$. Using lambda here as well feels
-  needlessly confusing.
-}
-\gio{ok}
-
 %It can be easily seen considering that $\rho$ is the number of mutations
@@
 \gio{This feels kind of a repetitions of what we wrote before the formula. I comment it. We can think of reinsert it afterwards, if needed (see comment above).}
+\richel{
+  I suggest to remove such commented-out lines. 
+  Although I sometimes get attached to my sentences,
+  they clog up the document by non-info and 
+  I usually delete them anyways in the end.
+  I cannot remember ever regretting this (would I, 
+  I could find it in the git history).
+  Would you agree?
+}
 
 Since we cannot know $n_{BD}(t)$ before running simulations
-we need to replace it with a proxy. For this reason we will use the average number of
+we need to replace it with a proxy. 
+For this reason we will use the average number of
 species in time according to the BD model. 
 It's well known that this is equal to \gio{insert proper citation}
+\richel{
+  I see you use angle bracket as a notation for the expected
+  value. I usually see 'E(x)' as the expected value for 'x',
+  and this is used at the beloved https://en.wikipedia.org/wiki/Expected_value.
+  What are the reasons you prefer the notation with the angle brackets?  
+}
 
@@
 \richel{I suggest $n_{BD} = n_{MBD}$ and only 
-change $\mu_{BD}$ to reach $m_{MBD} = m_{BD}$}
+change $\rho_{BD}$ to reach $<m_{MBD}> = <m_{BD}>$}
 \gio{@Richel: Don't you think it might make more sense to set $\mu_{BD} = \mu_{MBD}$?
@@
 In case you are wondering the theory can be found in Caesar's master thesis.}
+\richel{
+  I fully agree to use the same extinction rates!
+  The 'mu' used in the context of mutations (now 'rho') messed me up.
+  I hope this is clear now. To recap: 
+  (1) calculate the speciation rate of the twin tree as you wrote down excellently,
+  (2) simulate a twin tree with same number of taxa,
+  (3) calculate the mutation rates of the trees, 
+    so their alignments contain as much information
+}
 
@@
 dataset?}
-
-\gio{END}
-
-\richel{I think I see your point: in single trees stochasticity works per tree twin, in a distribution per distribution. To circumvent this: what if I would be 
-able to measure the actual number of mutations? Then twinning trees by
-having an exact equal amount of mutations would be very clean. Let me know: 
-you think I should investigate this?}
+\richel{
+  I think a per-tree calculation of the mutation rates 
+  is the best we can do. As there was some noise between us above,
+  I think the iteration after the next will allow us to get a better
+  idea about this
+}
 

### 2018-10-05 article.tex
Moved to methods, embedded in text

@@
 \item For each tree 
-generated under the MBD model we aim to generate a "twin" tree under the BD model in order to perform the same analysis and compare the results.
-To do so we need to be sure that the comparison is fair.
-The method we adopt to achieve this goal is to impose that the amount of information used in the Bayesian information will be the same.
-These are the same number of taxa and the same (expected)
-number of DNA mutations. See appendix for the exact procedure.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{mbd.jpg}
-  \caption{
-    How to create twin trees and alignments. From a focal MBD tree, a twin tree is produced as 
-    such: (1) estimate the $\lambda_{BD}$ to get the same expected number of tips, (2) simulate a BD tree with that amount of tips (discard trees with different number of tips), (3) estimate a mutation rate to get an alignment with the same expected number of mutations, (4) simulate alignments with that amount of mutations (discard those that don't, the picture shows an alignment that should be discarded)
+generated under the MBD model we aim to generate a "twin" tree under the BD model in order to perform a fair comparison, using trees 
+with the same amount of information, 
+which is the number of taxa 
+and the same (expected) number of DNA mutations.
+To obtain these twin trees, 
+\richel{
+  I suggest to first start with the equal number of taxa,
+  and the calculation of the speciation rates first.
 }
-\end{figure}
-
-\item We explained how we set the parameters for each twin BD tree.
-Using this rules we generate a BD dateset.
-We repeat the analysis, producing alignments for each tree and subsequently using BEAST to produce a posterior for each of them.
-
-\item Now we have two datasets of posteriors to compare, one for the BD model and one for the MBD model.
-
-\item To compare the results for the two models we measure the inference error using the nLTT statistic between known/true tree and posterior/inferred trees.
-\richel{
-  I would love to describe this more concrete. For example, when do
-  we say something has an effect? If we avoid making such judgements,
-  how will we visualize?
-}
-
-\richel{I removed the Bayes factor text. It is useful when letting BEAST2
-pick more/overly complex models and see if that more complex model fits the
-data better (penalized by its increased complexity, similar to the AIC). It has
-its uses, but I am unsure if we already want to discuss this now
-or first focus on the proper tree twinning}
-
-\end{itemize}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
-\begin{itemize}
-
-\item
-\richel{
-  I guess you know I am a fan of the Open Science Framework,
-  in which you first register you work before you do the experiment
-  (note: I will do some small pilots to estimate the complete time
-  of the experiment). I think it is the proper and superior science,
-  which helps us against writing down bullshit stories after having
-  obtained the results (e.g. 'We expected A and indeed found it!').
-  It also helps me structure my work: first think
-  deeply about the experiment, then do it (instead of the mixing
-  up the two phases). What are your thoughts on that?
-}
-
-\item
-
-\end{itemize}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Bibliography % MEE style
-\bibliographystyle{mee}
-\bibliography{article}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\appendix
-
-\section{Creating twin trees}
-
-To obtain twin trees with an equal amount of information, we impose
+we impose
 that the expected number of mutations in an MBD tree, $m_{MBD}$ equals
@@
 
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{mbd.jpg}
+  \caption{
+    How to create twin trees and alignments. From a focal MBD tree, a twin tree is produced as 
+    such: (1) estimate the $\lambda_{BD}$ to get the same expected number of tips, (2) simulate a BD tree with that amount of tips (discard trees with different number of tips), (3) estimate a mutation rate to get an alignment with the same expected number of mutations, (4) simulate alignments with that amount of mutations (discard those that don't, the picture shows an alignment that should be discarded)
+  }
+\end{figure}
+
+\item We explained how we set the parameters for each twin BD tree.
+Using this rules we generate a BD dateset.
+We repeat the analysis, producing alignments for each tree and subsequently using BEAST to produce a posterior for each of them.
+
+\item Now we have two datasets of posteriors to compare, one for the BD model and one for the MBD model.
+
+\item To compare the results for the two models we measure the inference error using the nLTT statistic between known/true tree and posterior/inferred trees.
+\richel{
+  I would love to describe this more concrete. For example, when do
+  we say something has an effect? If we avoid making such judgements,
+  how will we visualize?
+}
+
+\richel{I removed the Bayes factor text. It is useful when letting BEAST2
+pick more/overly complex models and see if that more complex model fits the
+data better (penalized by its increased complexity, similar to the AIC). It has
+its uses, but I am unsure if we already want to discuss this now
+or first focus on the proper tree twinning}
+
+\end{itemize}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Results}
+\begin{itemize}
+
+\item
+\richel{
+  I guess you know I am a fan of the Open Science Framework,
+  in which you first register you work before you do the experiment
+  (note: I will do some small pilots to estimate the complete time
+  of the experiment). I think it is the proper and superior science,
+  which helps us against writing down bullshit stories after having
+  obtained the results (e.g. 'We expected A and indeed found it!').
+  It also helps me structure my work: first think
+  deeply about the experiment, then do it (instead of the mixing
+  up the two phases). What are your thoughts on that?
+}
+
+\item
+
+\end{itemize}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Bibliography % MEE style
+\bibliographystyle{mee}
+\bibliography{article}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\appendix
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2018-10-05 article.tex
Finished next iteration

@@
   Perhaps better would be to write something explicit like:
-  t_now - t_crown = T.
+  t\_now - t\_crown = T.
   Looking forward for a better suggestion than mine :-)
@@
   value. I usually see 'E(x)' as the expected value for 'x',
-  and this is used at the beloved https://en.wikipedia.org/wiki/Expected_value.
+  and this is used at the beloved https://en.wikipedia.org/wiki/Expected\_value.
   What are the reasons you prefer the notation with the angle brackets?  

### 2018-10-08 article.tex
Changes @Giappo

@@
 \richel{Have you already looked up for a target journal? I know how a journal's constraints have helped me 
+\gio{Honestly I have literally no idea how to select a good journal for this kind of article.}
 in writing an article, for example, by having a maximum number of pictures}
@@
 \item There are many contemporary tools that provide the possibility to infer a phylogeny from genetic data (DNA, RNA, proteins). A popular Bayesian
-phylogenetic tool is called BEAST and its cousin BEAST2.
+phylogenetic tool is called BEAST (\cite{beast}) and its cousin BEAST2 (\cite{beast2}).
 
-\item BEAST is very flexible in setting up all possible phylogenetic priors (e.g. site/clock/speciation model).
+\item BEAST is very flexible, providing the user with the option to set up all possible phylogenetic priors (e.g. site/clock/speciation model).
 
-\item Current limits in current tools.
+%Current limits in current tools.
+\item However, currently available priors can be not suitable to analyze some specific datasets. 
+%With this work we aim to test whether or not the implementation of a new prior model is beneficial to study a specific kind of diversification process.
 
@@
 \item One of such speciation processes is the multiple birth hypothesis,
-a new model (described below) and thus absent in BEAST.
+a new model (described below) and thus currently absent in BEAST.
 
@@
 only a single speciation event can occur at any given time.
-The multiple-birth-death (MBD) model \richel{I feel MBSD (Multiple Birth Single Death) may be a better name: extinctions are still one at a time} relaxes this assumption, allowing events in which 
+The multiple-birth-death (MBD) model 
+\richel{I feel MBSD (Multiple Birth Single Death) may be a better name: extinctions are still one at a time}
+\gio{I always had the same doubt. The problem is that a good name should be, at the same time, both descriptive and short. I'll think about that.}
+relaxes this assumption, allowing events in which 
 large-scale environmental changes lead to a great number of species 
-in relatively short time intervals. Such a hypothesis may be a better fit to describe the burst in cichlid fish diversification in systems like in the 
+in relatively short time intervals. Such a hypothesis may be a better fit to describe the burst in in systems like cichlid fish diversification in the 
 African Great Lakes: Malawi, Tanganyika and Victoria (\cite{janzen2016}, \cite{janzen2017}).
@@
 While this assumption is useful to construct a wide variety of successful 
-models (for example: \cite{Maddison2007biSSE}, \cite{Valente2015}, \cite{etienne2012diversity}, \cite{etienne2014estimating}),
+models (e.g \cite{Maddison2007biSSE}, \cite{Valente2015}, \cite{etienne2012diversity}, \cite{etienne2014estimating}),
 they disallow for environmental changes that trigger speciations in multiple clades at a same point in time.
@@
 \item In the MBD model, parameters $\lambda$ and $\mu$ correspond, respectively, 
-to the common per-species speciation and extinction rates. 
-Additionally, $\nu$ is the rate at which an environmental change is triggered.
+to the common per-species speciation and extinction rates present also in the standard BD model. 
+Additionally, MBD relies on two additional parameters. Parameter $\nu$ is the rate at which an environmental change is triggered.
 When such event is triggered, all species present in the phylogeny at that moment
@@
 
-\item We start simulating 1000 \richel{I will measure the number of trees we'll be able to simulate within a short enough time, when the experiment is set up} MBD trees.
-From each MBD tree, a a DNA sequence alignment is simulated, after
-which that alignment starts a Bayesian analysis. We use 
+\item We start simulating $N_{S} = 1000$ \richel{I will measure the number of trees we'll be able to simulate within a short enough time, when the experiment is set up} MBD trees.
+From each MBD tree, a DNA sequence alignment is simulated.
+For each sequence alignment we then perform a Bayesian analysis to recover a posterior distribution of trees, each composed of $N_{P}$ phylogenies.
+Such analysis is performed using
 the 'pirouette' package (\cite{pirouette}) to call the BEAST2 tool 
@@
 \item For each tree 
-generated under the MBD model we aim to generate a "twin" tree under the BD model in order to perform a fair comparison, using trees 
-with the same amount of information, 
-which is the number of taxa 
-and the same (expected) number of DNA mutations.
+generated under the MBD model we aim to generate a "twin" tree under the BD model in order to perform a fair comparison. 
+We want trees from the two models to contain the same amount of information, 
+i.e. the same (expected) number of DNA mutations. %option 1
+%i.e. the same number of taxa at the present. %option 2
 To obtain these twin trees, 
@@
 }
+\gio{@richel: both assumptions seem to make sense to me. If possible I would run both and see which one performs better. For now let's stick to option 1 in the manuscript.}
 we impose
@@
 }
-with crown age $T$ in fact is given by
+\gio{I don't agree. It is expected if you use an expected number of species over time. If you have the correct $n(t)$ in principle you should be able to get the exact number of mutations.}
+with crown age $-T$ (with $T>0$) in fact is given by
 \richel{
@@
 }
+\gio{We can specify that the crown age is $-T$ where $T$ is a positive number (which should be already clear, I guess, if we report it as $-T$).
+In my opinion this makes everything nicer when we have to write down the integral which should, in principle, be $\int_{-T}^{0} f(t-(-T))$ but using integral properties can be rewritten as $\int_{0}^{T} f(t)$.
+Having a definite integral starting from $0$ always looks better ^^.}
 
@@
 }
+\gio{Agree. Remove them. CLOSED}
 
@@
 }
+\gio{Because the capital E is inelegant, plebeian and boor. And we want to be classy, don't we? :D}
 
@@
 
-%$$
-%<n_{BD}(t)> = n(t=0) * exp[(mu - lambda) * t] (6)
-%$$
-
-%The relation to use to get the equivalent lambda should therefore be
-
-%$$ 
-%m_{MBD} 
-%  = L * mu * integral[exp[(mu - lambda) * t] dt] 
-%  = (L * mu * exp[(mu - lambda) * t]) / (mu - lambda) (7)
-%$$
-
-%Here everything is known but lambda. So solve the lambda and use that value.
-
 \gio{My doubt is if we need to use $m_{MBD}$ for the single tree or the same
@@
 }
-
+\gio{I am still very debated about this. I actually would like to ask also Rampal's opinion.
+}
 \begin{figure}[!htbp]
@@
 \item We explained how we set the parameters for each twin BD tree.
-Using this rules we generate a BD dateset.
+Using this rules we generate a BD dataset.
 We repeat the analysis, producing alignments for each tree and subsequently using BEAST to produce a posterior for each of them.
 
-\item Now we have two datasets of posteriors to compare, one for the BD model and one for the MBD model.
+\subsection{Model selection}
 
-\item To compare the results for the two models we measure the inference error using the nLTT statistic between known/true tree and posterior/inferred trees.
+\item So far we have simulated two datasets of trees under the two models: $\{T_{i}^{BD}\}_{i=1}^{N_{S}}$ and $\{T_{i}^{MBD}\}_{i=1}^{N_{S}}$.
+We used them to generate a dataset of alignments for each model: $\{X^{BD}_{i}\}_{i=1}^{N_{S}}$ and $\{X^{MBD}_{i}\}_{i=1}^{N_{S}}$.
+From each dataset we produced a posterior distribution from a BD prior:
+$P_{i}(\theta | X^{BD}_{i}, BD)$ and $P_{i}(\theta | X^{MBD}_{i}, BD)$.
+\gio{1) We might want to rename the models, e.g. BD = (0) and MBD = (1). These names with capital letters are too big and ugly;
+2) Writing down things to tidy up manuscript and thoughts I just realized that actually the "model" is the same for the two posteriors.
+In fact with the word "model" Maturana et al.
+refer to the prior used (see eq.
+1, $\pi(\theta | M)$).
+In such case are we sure we can actually use the Bayes factor for model selection?}
+%Now we have two datasets  of posteriors to compare, one for the BD model and one for the MBD model.
+
+\subsubsection{Option 1: nLTT statistics}
+\item To compare the results for the two models we measure the inference error using the nLTT statistic between known/true tree and posterior/inferred trees (\cite{nltt}).
+To obtain such statistics the procedure is the following:
+
+- From each tree $T_{i,j}^{M}$ (with $j=1,...,N_{S}$) belonging to the posterior $P_{i}(\theta | X^{M}_{i}, BD)$ and relative to the model $M$, we extrapolate the lineage-through-time (LTT), in other words we measure the number of species as a function of time $n_{i,j}(t)$.
+To allow a comparison we normalize dividing by the maximum number of species of each tree, i.e.
+the number of tips at the present $N_{i,j}(t)=\frac{n_{i,j}(t)}{n^{max}_{i,j}}$.
+We then define the nLTT measure as
+
+$nLTT_{i,j} = \int_{0}^{T} \abs{N_{i,j}(t) - N_{T_{i}}} dt$
+
+\gio{I am running out of letters :(}
 \richel{
@@
 
+\subsubsection{Option 2: Bayes Factor (BF)}
+
 \richel{I removed the Bayes factor text. It is useful when letting BEAST2
@@
 or first focus on the proper tree twinning}
+\gio{But after we describe the tree twinning we will have to describe the method we use for model selection anyways. BF and nLTT are mutually exclusive choices, right?}
+\gio{The point is: what method do we choose?
+According to the choice we have to describe that.
+I'd also like to know Rampal's opinion on that.
+In principle I like more the BF method cause it looks more solid while the nLTT looks a bit like the last resort to me.}
 
@@
 }
-
+\gio{I agree on the principle but I don't fully agree on the open science procedure.
+Sometimes you do find something good while you're looking for something else.
+That happened many times in the history of science.
+This doesn't necessarily implies that we will end up p-hacking the results.
+I think we need to be open to new findings that might be not forecasted by our hypothesis.
+I think this is the good way to do science: be critical with our results but not imposing that reality will follow the theory.}
 \item

### 2018-10-08 article.tex
Next iteration

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\gio{According to my fine graining approach we should at each step deepen every small section. At a certain level I think we can start to re-coarse-grain what we wrote to create the abstract.}
-\richel{I enjoy this approach! Did some minor fine-graining}
-\richel{Have you already looked up for a target journal? I know how a journal's constraints have helped me 
+\richel{Have you already looked up for a target journal?} 
 \gio{Honestly I have literally no idea how to select a good journal for this kind of article.}
-in writing an article, for example, by having a maximum number of pictures}
+\richel{I suggest you ask Rampal}
 
@@
 We want trees from the two models to contain the same amount of information, 
-i.e. the same (expected) number of DNA mutations. %option 1
-%i.e. the same number of taxa at the present. %option 2
+i.e. the same (expected) number of DNA mutations and 
+the same number of taxa at the present.
 To obtain these twin trees, 
 \richel{
-  I suggest to first start with the equal number of taxa,
-  and the calculation of the speciation rates first.
+  I suggest to first start describing how to create the twin tree,
+  calculation of the speciation rates, then the twin alignment.
 }
-\gio{@richel: both assumptions seem to make sense to me. If possible I would run both and see which one performs better. For now let's stick to option 1 in the manuscript.}
+\gio{@richel: both assumptions seem to make sense to me. 
+  If possible I would run both and see which one performs better. For now
+  let's stick to option 1 in the manuscript.
+}
+\richel{
+  Sorry for the confusion: the previous comment was about the order
+  of describing out methods in this manuscript. As of the methods,  
+  I really meant to create twin trees with 
+  both equal number of taxa and equal expected number of mutations.
+}
+
 we impose
@@
 
-\gio{I think this should definitely go to the methods}
-\richel{I put it there for use, hopefully at a spot you liked :-)}
-
 The expected number of mutations $m$ of a phylogeny 
@@
 }
-\gio{I don't agree. It is expected if you use an expected number of species over time. If you have the correct $n(t)$ in principle you should be able to get the exact number of mutations.}
+\gio{
+  I don't agree. It is expected if you use an expected number 
+  of species over time. If you have the correct $n(t)$ in principle you 
+  should be able to get the exact number of mutations.
+}
+\richel{
+  If you have the correct $n(t)$ you can indeed calculate exactly what
+  the number of mutations will be, if and only if the mutation rate is
+  zero. Except for this travial case, I am curious about how you
+  calculate this. What is the precise number of mutations on a DNA
+  sequence of 1000 base pairs over one time unit with an 
+  expected average per-nucleotide
+  mutation rate of once per time? I would say one expects on average 1000 mutations
+  to occur. What would your claim be?
+}
+
 with crown age $-T$ (with $T>0$) in fact is given by
@@
 }
-\gio{We can specify that the crown age is $-T$ where $T$ is a positive number (which should be already clear, I guess, if we report it as $-T$).
-In my opinion this makes everything nicer when we have to write down the integral which should, in principle, be $\int_{-T}^{0} f(t-(-T))$ but using integral properties can be rewritten as $\int_{0}^{T} f(t)$.
-Having a definite integral starting from $0$ always looks better ^^.}
+\gio{
+  We can specify that the crown age is $-T$ where $T$ is a positive number 
+  (which should be already clear, I guess, 
+  if we report it as $-T$). In my opinion this makes everything 
+  nicer when we have to write down the integral which should, 
+  in principle, be $\int_{-T}^{0} f(t-(-T))$ but using integral properties 
+  can be rewritten as $\int_{0}^{T} f(t)$. Having a definite integral 
+  starting from $0$ always looks better ^^.
+}
+\richel{
+  I see, in a conversion you would say 'the crown age of the phylogeny is 
+  minus ten million years', where I would the say 'the crown age of the 
+  phylogeny is ten million years' to say exactly the same thing!
+  To try to resolve this, I checked one (actually two) of Rampal's papers (
+    'Diversity-dependence brings molecular phylogenies closer to agreement with the fossil record',
+    by Rampal, Bart Haegeman, Tanja Stadler, Tracy Aze, Paul N. Pearson, Andy Purvis, Albert B. Phillimore
+  ). Rampal avoids this point altogether: '... crown age $t_c = t_p − t_1$'
+  and '... and crown ages (5, 10 and 15 Myr)'. I suggest we do the same in
+  both writing and derivations: he is even more famous than us :-) 
+}
 
@@
 $n(t)$ the number of species at each time.
-
-%It can be easily seen considering that $\rho$ is the number of mutations
-%per unit time per site. For this reason it's needed to multiply by
-%time and number of sites. 
-\gio{This feels kind of a repetitions of what we wrote before the formula. I comment it. We can think of reinsert it afterwards, if needed (see comment above).}
-\richel{
-  I suggest to remove such commented-out lines. 
-  Although I sometimes get attached to my sentences,
-  they clog up the document by non-info and 
-  I usually delete them anyways in the end.
-  I cannot remember ever regretting this (would I, 
-  I could find it in the git history).
-  Would you agree?
-}
-\gio{Agree. Remove them. CLOSED}
 
@@
 }
-\gio{Because the capital E is inelegant, plebeian and boor. And we want to be classy, don't we? :D}
+\gio{
+  Because the capital E is inelegant, plebeian and boor. 
+  And we want to be classy, don't we? :D
+}
+\richel{
+  If the angle brackets are the new hip kids in mathematics town, sure,
+  we'll use those!
+}
 
@@
 }
-\gio{I am still very debated about this. I actually would like to ask also Rampal's opinion.
+\gio{
+  I am still very debated about this. I actually would like to ask also Rampal's opinion.
 }
+\richel{
+  Sounds like a good idea!
+}
+
 \begin{figure}[!htbp]
@@
 
-\item So far we have simulated two datasets of trees under the two models: $\{T_{i}^{BD}\}_{i=1}^{N_{S}}$ and $\{T_{i}^{MBD}\}_{i=1}^{N_{S}}$.
+\item So far we have simulated two datasets of trees under the two models: 
+$\{T_{i}^{BD}\}_{i=1}^{N_{S}}$ and $\{T_{i}^{MBD}\}_{i=1}^{N_{S}}$.
 We used them to generate a dataset of alignments for each model: $\{X^{BD}_{i}\}_{i=1}^{N_{S}}$ and $\{X^{MBD}_{i}\}_{i=1}^{N_{S}}$.
@@
 $P_{i}(\theta | X^{BD}_{i}, BD)$ and $P_{i}(\theta | X^{MBD}_{i}, BD)$.
-\gio{1) We might want to rename the models, e.g. BD = (0) and MBD = (1). These names with capital letters are too big and ugly;
-2) Writing down things to tidy up manuscript and thoughts I just realized that actually the "model" is the same for the two posteriors.
-In fact with the word "model" Maturana et al.
-refer to the prior used (see eq.
-1, $\pi(\theta | M)$).
-In such case are we sure we can actually use the Bayes factor for model selection?}
-%Now we have two datasets  of posteriors to compare, one for the BD model and one for the MBD model.
+\gio{
+  1) We might want to rename the models, e.g. BD = (0) and MBD = (1). These names with capital letters are too big and ugly;
+}
+\richel{
+  I definitely get your point! 
+  I would prefer something expressive over a number, perhaps a fancy letter, like
+  for example $\mathbb{B}$ and $\mathbb{M}$ for a Birth-Death and Multiple-Birth Death. 
+}
+\gio{
+2) Writing down things to tidy up manuscript and thoughts I just realized that 
+  actually the "model" is the same for the two posteriors. 
+  In fact with the word "model" Maturana et al. refer to the prior 
+  used (see eq. 1, $\pi(\theta | M)$). In such case are we sure we can actually 
+  use the Bayes factor for model selection?
+}
+\richel{
+  Correct! The assumed model of the posterior is a Birth-Death model. To do
+  model selection, we could assume a different nucleotide substitution 
+  model (GTR instead of JC69) and/or a different clock model (relaxed clock,
+  instead of a strict clock). I hope this clears things up. I will add
+  'generative model' for the model (BD or MBD) that created a phylogeny
+  and 'inferential model' for the model used in generating a posterior.
+}
 
 \subsubsection{Option 1: nLTT statistics}
-\item To compare the results for the two models we measure the inference error using the nLTT statistic between known/true tree and posterior/inferred trees (\cite{nltt}).
+\item To compare the results for the two models we measure the inference 
+error using the nLTT statistic between known/true tree and 
+posterior/inferred trees (\cite{nltt}). 
 To obtain such statistics the procedure is the following:
 
-- From each tree $T_{i,j}^{M}$ (with $j=1,...,N_{S}$) belonging to the posterior $P_{i}(\theta | X^{M}_{i}, BD)$ and relative to the model $M$, we extrapolate the lineage-through-time (LTT), in other words we measure the number of species as a function of time $n_{i,j}(t)$.
-To allow a comparison we normalize dividing by the maximum number of species of each tree, i.e.
-the number of tips at the present $N_{i,j}(t)=\frac{n_{i,j}(t)}{n^{max}_{i,j}}$.
-We then define the nLTT measure as
+- From each tree $T_{i,j}^{M}$ (with $j=1,...,N_{S}$) 
+  belonging to the posterior $P_{i}(\theta | X^{M}_{i}, BD)$ 
+  and relative to the model $M$, we extrapolate the lineage-through-time (LTT), 
+  in other words we measure the number of species as a function of 
+  time $n_{i,j}(t)$. To allow a comparison we normalize dividing by the 
+  maximum number of species of each tree, i.e. the number of tips at the 
+  present $N_{i,j}(t)=\frac{n_{i,j}(t)}{n^{max}_{i,j}}$. We then define the 
+  nLTT measure as
 
@@
 \gio{I am running out of letters :(}
-\richel{
-  I would love to describe this more concrete. For example, when do
-  we say something has an effect? If we avoid making such judgements,
-  how will we visualize?
+\richel{Haha! I suggest to use the same equation and symbols 
+  as equation 1 in
+  the nLTT article of Janzen, Hoehna and Etienne, 2015:
 }
+
+$$
+\delta nLTT = \int_{0}^{1} \abs{nLTT_1(t) - nLTT_2{t}} dt$
+$$
 
@@
 
-\richel{I removed the Bayes factor text. It is useful when letting BEAST2
+\richel{
+  I removed the Bayes factor text. It is useful when letting BEAST2
   pick more/overly complex models and see if that more complex model fits the
@@
   its uses, but I am unsure if we already want to discuss this now
-or first focus on the proper tree twinning}
-\gio{But after we describe the tree twinning we will have to describe the method we use for model selection anyways. BF and nLTT are mutually exclusive choices, right?}
-\gio{The point is: what method do we choose?
+  or first focus on the proper tree twinning
+}
+\gio{
+  But after we describe the tree twinning we will have to describe 
+  the method we use for model selection anyways. 
+  BF and nLTT are mutually exclusive choices, right?
+}
+\richel{
+  BF and nLTT are not mutually exclusive choices, but do complementary things:
+  nLTT measures the difference between the 'true' tree and the inferred trees,
+  which is the most important thing: we are interested in the error BEAST2
+  makes when using an overly simple (single-)BD tree prior.
+  With the BF we can compare BEAST2 posteriors that use (the same
+  tree prior, but) different site models (e.g. JC69, the simplest, 
+  versus GTR, the most complex). Although we know we generated the DNA
+  alignment with JC69, there are some hints that the GTR model is the
+  superior choice to do inference with.
+}
+\gio{
+  The point is: what method do we choose? 
   According to the choice we have to describe that. 
   I'd also like to know Rampal's opinion on that. 
-In principle I like more the BF method cause it looks more solid while the nLTT looks a bit like the last resort to me.}
+  In principle I like more the BF method cause it looks 
+  more solid while the nLTT looks a bit like the last resort to me.
+}
+\richel{
+  I hope this is now more clear. If not, I'll happily
+  elaborate :-)
+}
 
@@
 I think this is the good way to do science: be critical with our results but not imposing that reality will follow the theory.}
+\richel{
+  Looks like we are on the same page here: we both dislike making up bullshit 
+  stories after the results are known. Great!
+}
 \item

### 2018-10-15 article.tex
Correctly converts to PDF

@@
 
+\usepackage{amssymb}
 \usepackage[english]{babel}
@@
 \usepackage{verbatim}
+\usepackage{comment}
 
@@
 have a probability $q$ to speciate at that time, which is 
-independent on $\lambda$.
+independent on $\lambda$. Polytomies are not allowed in such process as each species can speciate only once at the time.
 
@@
 the 'pirouette' package (\cite{pirouette}) to call the BEAST2 tool 
-suite from R. We let the Bayesian analysis assume a BD prior, to investigate
-the error this inference makes due to this.
+suite from R. We let the Bayesian analysis assume a BD prior in both cases, to investigate
+the extent of the error we make under this assumption.
 
 \item For each tree 
-generated under the MBD model we aim to generate a "twin" tree under the BD model in order to perform a fair comparison. 
+generated under the MBD model we aim to generate a "twin" tree under the BD model.
+With the word "twin" we denote a tree generated starting from the respective MBD tree, in order to perform a fair comparison with it.
+This operation has to be done because we want to compare two trees that are generated by different processes.
+To do so we infer the parameters $\lambda_{BD}$ and $\mu_{BD}$ from the MBD maximizing the likelihood under a BD model.
+To perform this operation we use the function "\texttt{bd\_ML}" from the package "\texttt{DDD}" (\cite{etienne2012diversity}).
+
+\item We then exploit such parameters to generate a BD tree using the function "\texttt{tess.sim.taxa.age}" from the package "\texttt{TESS}" (\cite{Hoehna2013}).
+We simulate the tree in such a way the new tree has the same number of tips and the same crown age as the MBD tree.
+We furthermore require that the BD tree conserve the topology of the MBD tree
+
+\gio{see the razzo pipeline file: I am not sure what is the best way to do so}.
+
+\begin{comment}
 We want trees from the two models to contain the same amount of information, 
@@
 To obtain these twin trees, 
+\end{comment}
 \richel{
@@
 }
+\gio{The procedure is explained in the razzo pipeline markdown file. See there. CLOSED}
 
@@
   the number of mutations will be, if and only if the mutation rate is
-  zero. Except for this travial case, I am curious about how you
+  zero. Except for this trivial case, I am curious about how you
   calculate this. What is the precise number of mutations on a DNA
@@
 }
+\gio{Now I agree. CLOSED}
 
@@
   I would prefer something expressive over a number, perhaps a fancy letter, like
-  for example $\mathbb{B}$ and $\mathbb{M}$ for a Birth-Death and Multiple-Birth Death. 
+  for example $\mathbb{B}$ and $\mathbb{M}$ 
+  for a Birth-Death and Multiple-Birth Death. 
 }
@@
   and 'inferential model' for the model used in generating a posterior.
+}
 }
@@
 
-$nLTT_{i,j} = \int_{0}^{T} \abs{N_{i,j}(t) - N_{T_{i}}} dt$
+$nLTT_{i,j} = \int_{0}^{T} | N_{i,j}(t) - N_{T_{i}} | dt$
 
@@
 $$
-\delta nLTT = \int_{0}^{1} \abs{nLTT_1(t) - nLTT_2{t}} dt$
+\Delta nLTT = \int_{0}^{1} | nLTT_1(t) - nLTT_2{t} | dt
 $$

### 2018-10-15 article.tex
Changes by @richelbilderbeek

@@
 \gio{Honestly I have literally no idea how to select a good journal for this kind of article.}
-\richel{I suggest you ask Rampal}
+\richel{
+  May I suggest we aim for Molecular Phylogenetics and Evolution, the same journal as the raket paper?
+}
 
@@
 \item However, currently available priors can be not suitable to analyze some specific datasets. 
-%With this work we aim to test whether or not the implementation of a new prior model is beneficial to study a specific kind of diversification process.
+  With this work we aim to test whether or not the implementation of a new prior model 
+  is beneficial to study a specific kind of diversification process.
 
@@
 The multiple-birth-death (MBD) model 
-\richel{I feel MBSD (Multiple Birth Single Death) may be a better name: extinctions are still one at a time}
-\gio{I always had the same doubt. The problem is that a good name should be, at the same time, both descriptive and short. I'll think about that.}
 relaxes this assumption, allowing events in which 
@@
 \gio{see the razzo pipeline file: I am not sure what is the best way to do so}.
+\richel{Looks fine to me!}
 
-\begin{comment}
-We want trees from the two models to contain the same amount of information, 
-i.e. the same (expected) number of DNA mutations and 
-the same number of taxa at the present.
-To obtain these twin trees, 
-\end{comment}
-\richel{
-  I suggest to first start describing how to create the twin tree,
-  calculation of the speciation rates, then the twin alignment.
-}
-\gio{@richel: both assumptions seem to make sense to me. 
-  If possible I would run both and see which one performs better. For now
-  let's stick to option 1 in the manuscript.
-}
-\richel{
-  Sorry for the confusion: the previous comment was about the order
-  of describing out methods in this manuscript. As of the methods,  
-  I really meant to create twin trees with 
-  both equal number of taxa and equal expected number of mutations.
-}
-\gio{The procedure is explained in the razzo pipeline markdown file. See there. CLOSED}
-
-we impose
-that the expected number of mutations in an MBD tree, $m_{MBD}$ equals
-the expected number of mutations in a BD tree, $m_{BD}$:
+We want the MBD and twin BD trees to contain the same amount of information, 
+i.e. the same number of DNA mutations and the same number of taxa at the present:
 
@@
 
-We first generate a set of MBD trees. For each of them we can measure the amount of mutations $m_{MBD}$.
-
 The expected number of mutations $m$ of a phylogeny 
-\richel{
-  I think 'expected number of mutations' would be more correct.
-  Do you agree?
-}
-\gio{
-  I don't agree. It is expected if you use an expected number 
-  of species over time. If you have the correct $n(t)$ in principle you 
-  should be able to get the exact number of mutations.
-}
-\richel{
-  If you have the correct $n(t)$ you can indeed calculate exactly what
-  the number of mutations will be, if and only if the mutation rate is
-  zero. Except for this trivial case, I am curious about how you
-  calculate this. What is the precise number of mutations on a DNA
-  sequence of 1000 base pairs over one time unit with an 
-  expected average per-nucleotide
-  mutation rate of once per time? I would say one expects on average 1000 mutations
-  to occur. What would your claim be?
-}
-\gio{Now I agree. CLOSED}
-
 with crown age $-T$ (with $T>0$) in fact is given by
 \richel{
-  above stood 'crown age -T'. 
-  I feel that a crown age is a positive number,
-  but I know you have had a good reason.
-  Perhaps better would be to write something explicit like:
-  t\_now - t\_crown = T.
-  Looking forward for a better suggestion than mine :-)
-}
-\gio{
-  We can specify that the crown age is $-T$ where $T$ is a positive number 
-  (which should be already clear, I guess, 
-  if we report it as $-T$). In my opinion this makes everything 
-  nicer when we have to write down the integral which should, 
-  in principle, be $\int_{-T}^{0} f(t-(-T))$ but using integral properties 
-  can be rewritten as $\int_{0}^{T} f(t)$. Having a definite integral 
-  starting from $0$ always looks better ^^.
-}
-\richel{
-  I see, in a conversion you would say 'the crown age of the phylogeny is 
-  minus ten million years', where I would the say 'the crown age of the 
-  phylogeny is ten million years' to say exactly the same thing!
-  To try to resolve this, I checked one (actually two) of Rampal's papers (
-    'Diversity-dependence brings molecular phylogenies closer to agreement with the fossil record',
-    by Rampal, Bart Haegeman, Tanja Stadler, Tracy Aze, Paul N. Pearson, Andy Purvis, Albert B. Phillimore
-  ). Rampal avoids this point altogether: '... crown age $t_c = t_p − t_1$'
-  and '... and crown ages (5, 10 and 15 Myr)'. I suggest we do the same in
-  both writing and derivations: he is even more famous than us :-) 
+  So one of use likes '-T', the other likes 'T'. How to resolve this?
 }
@@
 
+The parameter we'll tune is $\rho$ ... \richel{elaborate here :-)}
+
 Since we cannot know $n_{BD}(t)$ before running simulations
@@
 It's well known that this is equal to \gio{insert proper citation}
-\richel{
-  I see you use angle bracket as a notation for the expected
-  value. I usually see 'E(x)' as the expected value for 'x',
-  and this is used at the beloved https://en.wikipedia.org/wiki/Expected\_value.
-  What are the reasons you prefer the notation with the angle brackets?  
-}
-\gio{
-  Because the capital E is inelegant, plebeian and boor. 
-  And we want to be classy, don't we? :D
-}
-\richel{
-  If the angle brackets are the new hip kids in mathematics town, sure,
-  we'll use those!
-}
 
@@
 If we set $\mu_{BD} = \mu_{MBD}$ and reverse this relation we can extrapolate the value of $\lambda_{BD}$ to use to generate BD trees.
-
-\richel{I suggest $n_{BD} = n_{MBD}$ and only 
-change $\rho_{BD}$ to reach $<m_{MBD}> = <m_{BD}>$}
-\gio{@Richel: Don't you think it might make more sense to set $\mu_{BD} = \mu_{MBD}$?
-What changes in the two model is the way we use to generate new species, not the way to remove them.
-Maybe one thing that's possible to do would be to make $\lambda_{BD}$ a function of time, a bit like Théo is doing in his comparison between DD and TD4 (which, by the way, seem to yield very different results).
-In case you are wondering the theory can be found in Caesar's master thesis.}
-\richel{
-  I fully agree to use the same extinction rates!
-  The 'mu' used in the context of mutations (now 'rho') messed me up.
-  I hope this is clear now. To recap: 
-  (1) calculate the speciation rate of the twin tree as you wrote down excellently,
-  (2) simulate a twin tree with same number of taxa,
-  (3) calculate the mutation rates of the trees, 
-    so their alignments contain as much information
-}
-
-\gio{My doubt is if we need to use $m_{MBD}$ for the single tree or the same
-quantity averaged on the full MBD dataset $<m_{MBD}>$.
-Do you think is better to use the
-individual $m_{MBD}$ for each tree or the average across the whole
-dataset?}
-\richel{
-  I think a per-tree calculation of the mutation rates 
-  is the best we can do. As there was some noise between us above,
-  I think the iteration after the next will allow us to get a better
-  idea about this
-}
-\gio{
-  I am still very debated about this. I actually would like to ask also Rampal's opinion.
-}
-\richel{
-  Sounds like a good idea!
-}
 
@@
 }
-\richel{
-  I definitely get your point! 
-  I would prefer something expressive over a number, perhaps a fancy letter, like
-  for example $\mathbb{B}$ and $\mathbb{M}$ 
-  for a Birth-Death and Multiple-Birth Death. 
-}
-\gio{
-2) Writing down things to tidy up manuscript and thoughts I just realized that 
-  actually the "model" is the same for the two posteriors. 
-  In fact with the word "model" Maturana et al. refer to the prior 
-  used (see eq. 1, $\pi(\theta | M)$). In such case are we sure we can actually 
-  use the Bayes factor for model selection?
-}
-\richel{
-  Correct! The assumed model of the posterior is a Birth-Death model. To do
-  model selection, we could assume a different nucleotide substitution 
-  model (GTR instead of JC69) and/or a different clock model (relaxed clock,
-  instead of a strict clock). I hope this clears things up. I will add
-  'generative model' for the model (BD or MBD) that created a phylogeny
-  and 'inferential model' for the model used in generating a posterior.
-}
-}
 
-\subsubsection{Option 1: nLTT statistics}
 \item To compare the results for the two models we measure the inference 
@@
 
-\subsubsection{Option 2: Bayes Factor (BF)}
-
-\richel{
-  I removed the Bayes factor text. It is useful when letting BEAST2
-  pick more/overly complex models and see if that more complex model fits the
-  data better (penalized by its increased complexity, similar to the AIC). It has
-  its uses, but I am unsure if we already want to discuss this now
-  or first focus on the proper tree twinning
-}
-\gio{
-  But after we describe the tree twinning we will have to describe 
-  the method we use for model selection anyways. 
-  BF and nLTT are mutually exclusive choices, right?
-}
-\richel{
-  BF and nLTT are not mutually exclusive choices, but do complementary things:
-  nLTT measures the difference between the 'true' tree and the inferred trees,
-  which is the most important thing: we are interested in the error BEAST2
-  makes when using an overly simple (single-)BD tree prior.
-  With the BF we can compare BEAST2 posteriors that use (the same
-  tree prior, but) different site models (e.g. JC69, the simplest, 
-  versus GTR, the most complex). Although we know we generated the DNA
-  alignment with JC69, there are some hints that the GTR model is the
-  superior choice to do inference with.
-}
-\gio{
-  The point is: what method do we choose? 
-  According to the choice we have to describe that. 
-  I'd also like to know Rampal's opinion on that. 
-  In principle I like more the BF method cause it looks 
-  more solid while the nLTT looks a bit like the last resort to me.
-}
-\richel{
-  I hope this is now more clear. If not, I'll happily
-  elaborate :-)
-}
-
 \end{itemize}
@@
 \item
-\richel{
-  I guess you know I am a fan of the Open Science Framework,
-  in which you first register you work before you do the experiment
-  (note: I will do some small pilots to estimate the complete time
-  of the experiment). I think it is the proper and superior science,
-  which helps us against writing down bullshit stories after having
-  obtained the results (e.g. 'We expected A and indeed found it!').
-  It also helps me structure my work: first think
-  deeply about the experiment, then do it (instead of the mixing
-  up the two phases). What are your thoughts on that?
-}
-\gio{I agree on the principle but I don't fully agree on the open science procedure.
-Sometimes you do find something good while you're looking for something else.
-That happened many times in the history of science.
-This doesn't necessarily implies that we will end up p-hacking the results.
-I think we need to be open to new findings that might be not forecasted by our hypothesis.
-I think this is the good way to do science: be critical with our results but not imposing that reality will follow the theory.}
-\richel{
-  Looks like we are on the same page here: we both dislike making up bullshit 
-  stories after the results are known. Great!
-}
+
 \item

### 2018-11-26 article.tex
Section seperating comment lines are 80 chars
Limit text lines to 80 chars

@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{abstract}
@@
 {\bf Keywords:} computational biology, evolution, phylogenetics, Bayesian analysis, tree prior
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \richel{Have you already looked up for a target journal?} 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
@@
 
-\item There are many contemporary tools that provide the possibility to infer a phylogeny from genetic data (DNA, RNA, proteins). A popular Bayesian
-phylogenetic tool is called BEAST (\cite{beast}) and its cousin BEAST2 (\cite{beast2}).
+\item There are many contemporary tools that provide the possibility 
+to infer a phylogeny from genetic data (DNA, RNA, proteins). 
+A popular Bayesian phylogenetic tool is called BEAST (\cite{beast}) 
+and its cousin BEAST2 (\cite{beast2}).
 
-\item BEAST is very flexible, providing the user with the option to set up all possible phylogenetic priors (e.g. site/clock/speciation model).
+\item BEAST is very flexible, providing the user with the option 
+to set up all possible phylogenetic priors (e.g. site/clock/speciation model).
 
 %Current limits in current tools.
-\item However, currently available priors can be not suitable to analyze some specific datasets. 
-  With this work we aim to test whether or not the implementation of a new prior model 
+\item However, currently available priors can be not suitable 
+to analyze some specific datasets. 
+With this work we aim to test whether or not 
+the implementation of a new prior model 
 is beneficial to study a specific kind of diversification process.
 
-\item BEAST2 gives us the possibility to introduce new tree priors to infer phylogenies based on different assumptions on how the speciation process
-takes place.
+\item BEAST2 gives us the possibility to introduce new tree priors 
+to infer phylogenies based on different assumptions 
+on how the speciation process takes place.
 
@@
 
-\item The Multiple birth hypothesis can be useful to explain a phenomenon that has always puzzled evolutionary biologists: what are the drivers of the diversification processes for those phylogenies that show an impressive amount of speciation events in relatively short times?
+\item The Multiple birth hypothesis can be useful to explain a phenomenon 
+that has always puzzled evolutionary biologists: 
+what are the drivers of the diversification processes 
+for those phylogenies that show an impressive amount of speciation events 
+in relatively short times? 
 The (constant-rate) birth-death (BD) model embodies the common assumption that 
@@
 large-scale environmental changes lead to a great number of species 
-in relatively short time intervals. Such a hypothesis may be a better fit to describe the burst in in systems like cichlid fish diversification in the 
-African Great Lakes: Malawi, Tanganyika and Victoria (\cite{janzen2016}, \cite{janzen2017}).
+in relatively short time intervals. 
+Such a hypothesis may be a better fit to describe the burst in systems 
+like cichlid fish diversification in the 
+African Great Lakes: Malawi, Tanganyika and Victoria 
+(\cite{janzen2016}, \cite{janzen2017}).
 
-\item However, it may be that current BD tree priors are good enough at detecting such events, with a (preferred) lower level of complexity.
+\item However, it may be that current BD tree priors are good enough 
+at detecting such events, with a (preferred) lower level of complexity. 
 If this is the case one should always be more keen to adopt the simplest model.
 
-\item Here we present our study with the aim of exploring when using a more complex MBD tree prior is warranted.
+\item Here we present our study with the aim of exploring 
+when using a more complex MBD tree prior is warranted.
 
 \end{itemize}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Methods}
@@
 
-\item Current phylogenetic tools assume that only a single speciation event can occur at any given time.
+\item Current phylogenetic tools assume that only a single speciation event 
+can occur at any given time.
 While this assumption is useful to construct a wide variety of successful 
-models (e.g \cite{Maddison2007biSSE}, \cite{Valente2015}, \cite{etienne2012diversity}, \cite{etienne2014estimating}),
-they disallow for environmental changes that trigger speciations in multiple clades at a same point in time.
+models (e.g \cite{Maddison2007biSSE}, \cite{Valente2015}, 
+\cite{etienne2012diversity}, \cite{etienne2014estimating}),
+they disallow for environmental changes that trigger speciations 
+in multiple clades at a same point in time.
 
-\item The (constant-rate) birth-death (BD) model embodies the common assumption that 
+\item The (constant-rate) birth-death (BD) model 
+embodies the common assumption that 
 only a single speciation event can occur at any given time.
-The multiple-birth-death (MBD) model relaxes this assumption, allowing events in which 
+The multiple-birth-death (MBD) model relaxes this assumption, 
+allowing events in which 
 large-scale environmental changes lead to a great number of species 
-in relatively short time intervals. Such hypothesis can be useful to describe, for example, 
+in relatively short time intervals. 
+Such hypothesis can be useful to describe, for example, 
 systems like cichlid fish diversification in the 
-African Great Lakes: Malawi, Tanganyika and Victoria (\cite{janzen2016}, \cite{janzen2017}).
+African Great Lakes: Malawi, Tanganyika and Victoria 
+(\cite{janzen2016}, \cite{janzen2017}).
 
 \item In the MBD model, parameters $\lambda$ and $\mu$ correspond, respectively, 
-to the common per-species speciation and extinction rates present also in the standard BD model. 
-Additionally, MBD relies on two additional parameters. Parameter $\nu$ is the rate at which an environmental change is triggered.
-When such event is triggered, all species present in the phylogeny at that moment
+to the common per-species speciation and extinction rates present 
+also in the standard BD model. 
+Additionally, MBD relies on two additional parameters. 
+Parameter $\nu$ is the rate at which an environmental change is triggered.
+When such event is triggered, 
+all species present in the phylogeny at that moment
 have a probability $q$ to speciate at that time, which is 
-independent on $\lambda$. Polytomies are not allowed in such process as each species can speciate only once at the time.
+independent on $\lambda$. 
+Polytomies are not allowed in such process 
+as each species can speciate only once at the time.
 
-\item It is also possible to write down a likelihood function for such processes as in \cite{mbd}.
+\item It is also possible to write down a likelihood function 
+for such processes as in \cite{mbd}.
 
@@
 
-\item To prove our hypothesis we simulate two twin datasets. All the simulations are produced in continuous time, using the Doob-Gillespie algorithm.
+\item To prove our hypothesis we simulate two twin datasets. 
+All the simulations are produced in continuous time, 
+using the Doob-Gillespie algorithm.
 
-\item We start simulating $N_{S} = 1000$ \richel{I will measure the number of trees we'll be able to simulate within a short enough time, when the experiment is set up} MBD trees.
+\item We start simulating $N_{S} = 1000$ MBD trees. 
 From each MBD tree, a DNA sequence alignment is simulated. 
-For each sequence alignment we then perform a Bayesian analysis to recover a posterior distribution of trees, each composed of $N_{P}$ phylogenies.
+For each sequence alignment we then perform a Bayesian analysis 
+to recover a posterior distribution of trees, 
+each composed of $N_{P}$ phylogenies. 
 Such analysis is performed using 
 the 'pirouette' package (\cite{pirouette}) to call the BEAST2 tool 
-suite from R. We let the Bayesian analysis assume a BD prior in both cases, to investigate
-the extent of the error we make under this assumption.
+suite from R. 
+We let the Bayesian analysis assume a BD prior in both cases, 
+to investigate the extent of the error we make under this assumption.
 
-\item For each tree 
-generated under the MBD model we aim to generate a "twin" tree under the BD model.
-With the word "twin" we denote a tree generated starting from the respective MBD tree, in order to perform a fair comparison with it.
-This operation has to be done because we want to compare two trees that are generated by different processes.
-To do so we infer the parameters $\lambda_{BD}$ and $\mu_{BD}$ from the MBD maximizing the likelihood under a BD model.
-To perform this operation we use the function "\texttt{bd\_ML}" from the package "\texttt{DDD}" (\cite{etienne2012diversity}).
+\item For each tree generated under the MBD model 
+we aim to generate a "twin" tree under the BD model. 
+With the word "twin" 
+we denote a tree generated starting from the respective MBD tree, 
+in order to perform a fair comparison with it. 
+This operation has to be done, 
+because we want to compare two trees 
+that are generated by different processes. 
+To do so we infer the parameters $\lambda_{BD}$ and $\mu_{BD}$ 
+from the MBD maximizing the likelihood under a BD model. 
+To perform this operation we use the function "\texttt{bd\_ML}" 
+from the package "\texttt{DDD}" (\cite{etienne2012diversity}).
 
-\item We then exploit such parameters to generate a BD tree using the function "\texttt{tess.sim.taxa.age}" from the package "\texttt{TESS}" (\cite{Hoehna2013}).
-We simulate the tree in such a way the new tree has the same number of tips and the same crown age as the MBD tree.
-We furthermore require that the BD tree conserve the topology of the MBD tree
-
-\gio{see the razzo pipeline file: I am not sure what is the best way to do so}.
-\richel{Looks fine to me!}
+\item We then exploit such parameters to generate a BD tree 
+using the function "\texttt{tess.sim.taxa.age}" 
+from the package "\texttt{TESS}" (\cite{Hoehna2013}). 
+We simulate the tree in such a way the new tree 
+has the same number of tips and the same crown age as the MBD tree. 
+We furthermore require that the BD tree conserve the topology of the MBD tree.
 
@@
 
-where $n_{0} = n_{BD}(-T) = n_{MBD}(-T)$ is the initial number of species at the crown age.
+where $n_{0} = n_{BD}(-T) = n_{MBD}(-T)$ is the initial number of species 
+at the crown age.
 From \ref{m equivalence}, \ref{m calculation} and \ref{BD average n} follows:
@@
 
-If we set $\mu_{BD} = \mu_{MBD}$ and reverse this relation we can extrapolate the value of $\lambda_{BD}$ to use to generate BD trees.
+If we set $\mu_{BD} = \mu_{MBD}$ and reverse this relation 
+we can extrapolate the value of $\lambda_{BD}$ to use to generate BD trees.
 
@@
   \caption{
-    How to create twin trees and alignments. From a focal MBD tree, a twin tree is produced as 
-    such: (1) estimate the $\lambda_{BD}$ to get the same expected number of tips, (2) simulate a BD tree with that amount of tips (discard trees with different number of tips), (3) estimate a mutation rate to get an alignment with the same expected number of mutations, (4) simulate alignments with that amount of mutations (discard those that don't, the picture shows an alignment that should be discarded)
+    How to create twin trees and alignments. 
+    From a focal MBD tree, a twin tree is produced as 
+    such: (1) estimate the $\lambda_{BD}$ to get 
+    the same expected number of tips, (2) simulate a BD tree 
+    with that amount of tips (discard trees with different number of tips), 
+    (3) estimate a mutation rate to get an alignment 
+    with the same expected number of mutations, (4) simulate alignments 
+    with that amount of mutations (discard those that don't, 
+    the picture shows an alignment that should be discarded) 
   }
@@
 Using this rules we generate a BD dataset. 
-We repeat the analysis, producing alignments for each tree and subsequently using BEAST to produce a posterior for each of them.
+We repeat the analysis, producing alignments for each tree 
+and subsequently using BEAST to produce a posterior for each of them.
 
@@
 \end{itemize}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Results}
@@
 \end{itemize}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Bibliography % MEE style
@@
 \bibliography{article}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \appendix
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2018-11-27 article.tex
@richelbilderbeek suggests to use MBD and BD, instead of 0 and 1
Add hypthesis, progress https://github.com/richelbilderbeek/razzo/issues/143

@@
 
+% Hypothesis: general
+\item We hypothesize that the error made today, using BD tree priors,
+increases with an increased number or stronger effect of multiple birth events.
+This is straightforward: without multiple birth events or such event having
+no effect, the MBD model falls back to a BD model. We expect larger
+errors when we deviate more from the BD model's assumptions.
+
+\item Additionally, we hypothesize MBD having a stronger effect if the normal
+speciation process is less pronounced. The more speciations are caused by the
+BD process, there are relatively less multiple-birth events.
+
+\item We expect the effect of extinction rates to be neutral, as extinctions will
+hit lineages created by both speciation processes equally.
+
+\item Summarized, we expect the error made be correlated to the number 
+of species created by the multiple-birth process over the total number 
+of species created:
+
+\begin{equation}
+<e> = f(\frac{n_{taxa}^{MBD}}{n_{taxa}^{BD} + n_{taxa}^{MBD}})
+\end{equation}
+
+Where $<e>$ denotes the expected error, $f$ is a monotonously increasing 
+function of unknown shape, $n_{taxa}^{MBD}$ is the number of taxa created
+in multiple-birth events and $n_{taxa}^{BD}$ is the number of taxa created
+by the standard BD specition process.
+
 \end{itemize}
@@
 
-\subsection{Model selection}
+\subsection{Measuring the inference error}
 
@@
 \gio{
-  1) We might want to rename the models, e.g. BD = (0) and MBD = (1). These names with capital letters are too big and ugly;
+  1) We might want to rename the models, e.g. BD = (0) and MBD = (1). 
+  These names with capital letters are too big and ugly;
+}
+\richel{
+  I would strongly prefer MBD and BD, as I feel replacing the big ugly 
+  capital letters by short pretty numbers hurts readability even more 
 }
@@
 
+\subsection{Model selection}
+
 \end{itemize}

### 2018-11-27 article.tex
Expect timing to have no effect, progress https://github.com/richelbilderbeek/razzo/issues/143

@@
 
-\item Summarized, we expect the error made be correlated to the number 
+\item To put into an explicit equation, 
+we expect the error made be correlated to the number 
 of species created by the multiple-birth process over the total number 
@@
 by the standard BD specition process.
+
+\item Due to the proportionality of the term within $f$, 
+we expect the timing of a multiple births event (be it close to the
+crown age or close to the present) to have no effect. Compared to a late
+multiple birth event, an early multiple birth event may have a 
+longer-lasting effect (as the next speciation event will be later), but it
+will create less new species, as there are still fewer taxa.
 

### 2018-11-27 article.tex
Describe marginal likelihood estimation, progress in https://github.com/richelbilderbeek/razzo/issues/136 and https://github.com/richelbilderbeek/razzo/issues/137

@@
 
+We simulate alignments using the simplest nucleotide substitution model (JC69),
+the simplest clock model (strict). It is thus imminent to assume these
+models in our Bayesian inference. Nevertheless, the phylogeny the alignment
+was based on, could have followed either an MBD or BD tree model, 
+where we in both cases assume a BD tree model. This will have 
+an unknown effect on our inference: it may theoretically be that an MBD model
+generates (a tree that generates) an alignment in which a different site 
+and/or clock model is favored.
+
+We investigate this by measuring if the generative model (with the simplest
+nucleotide substitution and simplest clock model) is indeed selected 
+to be the best fitting model. 
+To be precise, we look at the model 
+with the highest marginal likelihood 
+(also called evidence \cite{mackay2003information}),
+$f(D|M)$, which is the probability of the data D given model M.
+In the context of this research, D consists of the DNA alignment,
+and M is the combination of site, clock and tree models.
+
+To estimate the marginal likelihood, 
+we use an algorithm named nested sampling \cite{skilling2006nested}.
+Nested sampling is attractive to use
+in a phylogentic context, as it gives a good estimation,
+requires little tuning \cite{maturana2018}.
+Nested sampling is available as a BEAST2 package
+and can be used by babette \cite{babette}.
+
+The nested sampling algorithm stops its run 
+when the marginal likelihood estimation error 
+reaches below a certain tolerance.  
+Similar to \cite{maturana2018},
+we use a (relative) error tolerance $\epsilon$ of $10^{-13}$,
+1 particle to explore the parameter space
+and 100 active points. 
+To achieve the latter, we use the MCMC chain length $L_c$ of 1M 
+(as also used in the parameter estimates),
+and a sub-chain length $L_{sc}$ of 10K.
+
+The models we use in our model comparison are the four combinations
+of two site models and two clock models. We use the JC69 site model, which
+is the (generative and) simplest model and GTR, the site model with most
+degrees of freedom. For the clock models, we use the strict clock model,
+which is the (generative and) simplest clock model, and the RLN clock model.
+
+From these four marginal likelihood estimates, we calculate the weight of
+the generative model and plot this in figure 2. We do this for both the 
+alignments derived from the MBD tree and the BD twin tree. We expect that
+the generative model has the heighest weight in both the MBD and BD alignments.
+We expect this weight to be higher in the BD alignments.
+
 \end{itemize}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{table}
+  \centering 
+  \begin{tabular}{p{0.05\textwidth} p{0.6\textwidth} p{0.2\textwidth}}
+    \hline
+                          & Description & Value(s) \\
+    \hline
+    \hline
+    $L_c$                 & MCMC chain length & $10^6$ \\
+    $L_{sc}$              & MCMC sub-chain length & $10^4$ \\
+    $\epsilon$            & relative error tolerance in marginal likelihood estimation & $10^{-13}$ \\
+    \hline
+  \end{tabular}
+  \caption{
+    Overview of the simulation parameters. 
+  }
+  \label{table:parameters}
+\end{table}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2018-12-03 article.tex
Put hypotheses in figures

@@
 errors when we deviate more from the BD model's assumptions. 
-
-\item Additionally, we hypothesize MBD having a stronger effect if the normal
+Additionally, we hypothesize MBD having a stronger effect if the normal
 speciation process is less pronounced. The more speciations are caused by the
 BD process, there are relatively less multiple-birth events.
-
-\item We expect the effect of extinction rates to be neutral, as extinctions will
-hit lineages created by both speciation processes equally.
-
-\item To put into an explicit equation, 
+To put this hypothesis, H1, into an explicit equation, 
 we expect the error made be correlated to the number 
@@
 
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{fig_h_1.png}
+  \caption{
+    Hypothesis 1: we expect the error made be correlated to the number 
+    of species created by the multiple-birth process over the total number 
+    of species created
+  }
+  \label{fig_h_1}
+\end{figure}
+
+\item We have the hypothesis, H2, 
+that the effect of extinction rates is neutral, 
+as extinctions will hit lineages created by both speciation processes equally.
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{fig_h_2.png}
+  \caption{
+    Hypothesis 2: the effect of extinction rates is neutral
+  }
+  \label{fig_h_2}
+\end{figure}
+
 \item Due to the proportionality of the term within $f$, 
-we expect the timing of a multiple births event (be it close to the
+we have the hypothesis, H3, 
+that the timing of a multiple birth event (be it close to the
 crown age or close to the present) to have no effect. Compared to a late
@@
 will create less new species, as there are still fewer taxa.
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{fig_h_3.png}
+  \caption{
+    Hypothesis 3: the timing of a multiple birth event has no effect.
+  }
+  \label{fig_h_3}
+\end{figure}
 
@@
 
-\item We start simulating $N_{S} = 1000$ MBD trees. 
-From each MBD tree, a DNA sequence alignment is simulated. 
+\item We start simulating $N_{S} = 1000$ MBD trees,
+with either 50, 100 and 200 taxa.
+We have the hypothesis H5, that the number of taxa does not
+have an effect on the error being made, as there is no
+diversity dependency in any of the processes.
+We have the hypothesis H6, that for a higher number of taxa 
+the variance in the error decreases, as more information is present
+in the simulated phylogenies.
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{fig_h_5.png}
+  \caption{
+    Hypothesis 5: the number of taxa does not
+    have an effect on the error being made
+  }
+  \label{fig_h_5}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{fig_h_6.png}
+  \caption{
+    Hypothesis 6: for a higher number of taxa 
+    the variance in the error decreases
+  }
+  \label{fig_h_6}
+\end{figure}
+
+\item From each MBD tree, a DNA sequence alignment is simulated. 
 For each sequence alignment we then perform a Bayesian analysis 
@@
 We furthermore require that the BD tree conserve the topology of the MBD tree.
+We have hypothesis H4 that, compared to the MBD trees, 
+the error will be less in the BD twin tree.
+The difference between the errors made in MBD and twin BD trees indicates
+the impact the MBD process has on the error we make in inference using a
+contemporary BD prior.
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{fig_h_4.png}
+  \caption{
+    Hypothesis 4: compared to the MBD trees, 
+    the error will be less in the BD twin tree
+  }
+  \label{fig_h_4}
+\end{figure}
 

### 2018-12-03 article.tex
Add hypothesis 7

@@
   \label{fig_h_1}
+\end{figure}
+
+\item We have the hypothesis, H7, 
+that there is no difference in inference errors if the MBD process
+generates many species in many modest MB events or few intense MB events.
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{fig_h_7.png}
+  \caption{
+    Hypothesis 7: there is no difference in inference errors if the MBD process
+    generates many species in many modest MB events or few intense MB events
+  }
+  \label{fig_h_7}
 \end{figure}

### 2018-12-03 article.tex
Could use all 8 combinations of site and clock models
Add MBD params in table

@@
 which is the (generative and) simplest clock model, and the RLN clock model.
+\richel{Could also just be all site models and clock models = 8 models}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{table}
+  \centering 
+  \begin{tabular}{ l | c | c | c | c   c   c | c }
+    \hline
+idx & $n_{taxa}$ & $f_{taxa}^{MB}$ & $\mu$ & MB regime & $n_{\nu}$ evts & $q$ & $\lambda$ \\
+    \hline
+    \hline
+1 & 50 & 0 & 0 & many modest  & 8 & 1 / 8 & derive \\
+2 & 50 & 0 & 0 & intermediate & 4 & 1 / 4 & derive \\
+3 & 50 & 0 & 0 & few intense & 2 & 1 / 2 & derive \\
+4 & 50 & 0 & 0.1 & many modest  & 8 & 1 / 8 & derive \\
+5 & 50 & 0 & 0.1 & intermediate & 4 & 1 / 4 & derive \\
+6 & 50 & 0 & 0.1 & few intense & 2 & 1 / 2 & derive \\
+7 & 50 & 0 & 0.2 & many modest  & 8 & 1 / 8 & derive \\
+8 & 50 & 0 & 0.2 & intermediate & 4 & 1 / 4 & derive \\
+9 & 50 & 0 & 0.2 & few intense & 2 & 1 / 2 & derive \\
+10 & 50 & 0 & 0.2 & many modest  & 8 & 1 / 8 & derive \\
+11 & 50 & 0.25 & 0 & intermediate & 4 & 1 / 4 & derive \\
+12 & 50 & 0.25 & 0 & few intense & 2 & 1 / 2 & derive \\
+13 & 50 & 0.25 & 0 & many modest  & 8 & 1 / 8 & derive \\
+14 & 50 & 0.25 & 0.1 & intermediate & 4 & 1 / 4 & derive \\
+15 & 50 & 0.25 & 0.1 & few intense & 2 & 1 / 2 & derive \\
+16 & 50 & 0.25 & 0.1 & many modest  & 8 & 1 / 8 & derive \\
+17 & 50 & 0.25 & 0.2 & intermediate & 4 & 1 / 4 & derive \\
+18 & 50 & 0.25 & 0.2 & few intense & 2 & 1 / 2 & derive \\
+19 & 50 & 0.25 & 0.2 & many modest  & 8 & 1 / 8 & derive \\
+20 & 50 & 0.25 & 0.2 & intermediate & 4 & 1 / 4 & derive \\
+21 & 50 & 0.5 & 0 & few intense & 2 & 1 / 2 & derive \\
+22 & 50 & 0.5 & 0 & many modest  & 8 & 1 / 8 & derive \\
+23 & 50 & 0.5 & 0 & intermediate & 4 & 1 / 4 & derive \\
+24 & 50 & 0.5 & 0.1 & few intense & 2 & 1 / 2 & derive \\
+25 & 50 & 0.5 & 0.1 & many modest  & 8 & 1 / 8 & derive \\
+26 & 50 & 0.5 & 0.1 & intermediate & 4 & 1 / 4 & derive \\
+27 & 50 & 0.5 & 0.2 & few intense & 2 & 1 / 2 & derive \\
+28 & 50 & 0.5 & 0.2 & many modest  & 8 & 1 / 8 & derive \\
+29 & 50 & 0.5 & 0.2 & intermediate & 4 & 1 / 4 & derive \\
+30 & 50 & 0.5 & 0.2 & few intense & 2 & 1 / 2 & derive \\
+31 & 50 & 0.75 & 0 & many modest  & 8 & 1 / 8 & derive \\
+32 & 50 & 0.75 & 0 & intermediate & 4 & 1 / 4 & derive \\
+33 & 50 & 0.75 & 0 & few intense & 2 & 1 / 2 & derive \\
+34 & 50 & 0.75 & 0.1 & many modest  & 8 & 1 / 8 & derive \\
+35 & 50 & 0.75 & 0.1 & intermediate & 4 & 1 / 4 & derive \\
+36 & 50 & 0.75 & 0.1 & few intense & 2 & 1 / 2 & derive \\
+37 & 50 & 0.75 & 0.2 & many modest  & 8 & 1 / 8 & derive \\
+38 & 50 & 0.75 & 0.2 & intermediate & 4 & 1 / 4 & derive \\
+39 & 50 & 0.75 & 0.2 & few intense & 2 & 1 / 2 & derive \\
+40 & 50 & 0.75 & 0.2 & many modest  & 8 & 1 / 8 & derive \\
+41 & 50 & 1 & 0 & intermediate & 4 & 1 / 4 & 0 \\
+42 & 50 & 1 & 0 & few intense & 2 & 1 / 2 & 0 \\
+43 & 50 & 1 & 0 & many modest  & 8 & 1 / 8 & 0 \\
+44 & 50 & 1 & 0.1 & intermediate & 4 & 1 / 4 & 0 \\
+45 & 50 & 1 & 0.1 & few intense & 2 & 1 / 2 & 0 \\
+46 & 50 & 1 & 0.1 & many modest  & 8 & 1 / 8 & 0 \\
+47 & 50 & 1 & 0.2 & intermediate & 4 & 1 / 4 & 0 \\
+48 & 50 & 1 & 0.2 & few intense & 2 & 1 / 2 & 0 \\
+49 & 50 & 1 & 0.2 & many modest  & 8 & 1 / 8 & 0 \\
+50 & 50 & 1 & 0.2 & intermediate & 4 & 1 / 4 & 0 \\
+    \hline
+  \end{tabular}
+  \caption{
+    Overview of the MBD parameters 1/3. 
+  }
+  \label{table:mbd_parameters_1}
+\end{table}
+
+\begin{table}
+  \centering 
+  \begin{tabular}{ l | c | c | c | c   c   c | c }
+    \hline
+idx & $n_{taxa}$ & $f_{taxa}^{MB}$ & $\mu$ & MB regime & $n_{\nu}$ evts & $q$ & $\lambda$ \\
+    \hline
+    \hline
+51 & 100 & 0 & 0 & few intense & 2 & 1 / 2 & derive \\
+52 & 100 & 0 & 0 & many modest  & 8 & 1 / 8 & derive \\
+53 & 100 & 0 & 0 & intermediate & 4 & 1 / 4 & derive \\
+54 & 100 & 0 & 0.1 & few intense & 2 & 1 / 2 & derive \\
+55 & 100 & 0 & 0.1 & many modest  & 8 & 1 / 8 & derive \\
+56 & 100 & 0 & 0.1 & intermediate & 4 & 1 / 4 & derive \\
+57 & 100 & 0 & 0.2 & few intense & 2 & 1 / 2 & derive \\
+58 & 100 & 0 & 0.2 & many modest  & 8 & 1 / 8 & derive \\
+59 & 100 & 0 & 0.2 & intermediate & 4 & 1 / 4 & derive \\
+60 & 100 & 0 & 0.2 & few intense & 2 & 1 / 2 & derive \\
+61 & 100 & 0.25 & 0 & many modest  & 8 & 1 / 8 & derive \\
+62 & 100 & 0.25 & 0 & intermediate & 4 & 1 / 4 & derive \\
+63 & 100 & 0.25 & 0 & few intense & 2 & 1 / 2 & derive \\
+64 & 100 & 0.25 & 0.1 & many modest  & 8 & 1 / 8 & derive \\
+65 & 100 & 0.25 & 0.1 & intermediate & 4 & 1 / 4 & derive \\
+66 & 100 & 0.25 & 0.1 & few intense & 2 & 1 / 2 & derive \\
+67 & 100 & 0.25 & 0.2 & many modest  & 8 & 1 / 8 & derive \\
+68 & 100 & 0.25 & 0.2 & intermediate & 4 & 1 / 4 & derive \\
+69 & 100 & 0.25 & 0.2 & few intense & 2 & 1 / 2 & derive \\
+70 & 100 & 0.25 & 0.2 & many modest  & 8 & 1 / 8 & derive \\
+71 & 100 & 0.5 & 0 & intermediate & 4 & 1 / 4 & derive \\
+72 & 100 & 0.5 & 0 & few intense & 2 & 1 / 2 & derive \\
+73 & 100 & 0.5 & 0 & many modest  & 8 & 1 / 8 & derive \\
+74 & 100 & 0.5 & 0.1 & intermediate & 4 & 1 / 4 & derive \\
+75 & 100 & 0.5 & 0.1 & few intense & 2 & 1 / 2 & derive \\
+76 & 100 & 0.5 & 0.1 & many modest  & 8 & 1 / 8 & derive \\
+77 & 100 & 0.5 & 0.2 & intermediate & 4 & 1 / 4 & derive \\
+78 & 100 & 0.5 & 0.2 & few intense & 2 & 1 / 2 & derive \\
+79 & 100 & 0.5 & 0.2 & many modest  & 8 & 1 / 8 & derive \\
+80 & 100 & 0.5 & 0.2 & intermediate & 4 & 1 / 4 & derive \\
+81 & 100 & 0.75 & 0 & few intense & 2 & 1 / 2 & derive \\
+82 & 100 & 0.75 & 0 & many modest  & 8 & 1 / 8 & derive \\
+83 & 100 & 0.75 & 0 & intermediate & 4 & 1 / 4 & derive \\
+84 & 100 & 0.75 & 0.1 & few intense & 2 & 1 / 2 & derive \\
+85 & 100 & 0.75 & 0.1 & many modest  & 8 & 1 / 8 & derive \\
+86 & 100 & 0.75 & 0.1 & intermediate & 4 & 1 / 4 & derive \\
+87 & 100 & 0.75 & 0.2 & few intense & 2 & 1 / 2 & derive \\
+88 & 100 & 0.75 & 0.2 & many modest  & 8 & 1 / 8 & derive \\
+89 & 100 & 0.75 & 0.2 & intermediate & 4 & 1 / 4 & derive \\
+90 & 100 & 0.75 & 0.2 & few intense & 2 & 1 / 2 & derive \\
+91 & 100 & 1 & 0 & many modest  & 8 & 1 / 8 & 0 \\
+92 & 100 & 1 & 0 & intermediate & 4 & 1 / 4 & 0 \\
+93 & 100 & 1 & 0 & few intense & 2 & 1 / 2 & 0 \\
+94 & 100 & 1 & 0.1 & many modest  & 8 & 1 / 8 & 0 \\
+95 & 100 & 1 & 0.1 & intermediate & 4 & 1 / 4 & 0 \\
+96 & 100 & 1 & 0.1 & few intense & 2 & 1 / 2 & 0 \\
+97 & 100 & 1 & 0.2 & many modest  & 8 & 1 / 8 & 0 \\
+98 & 100 & 1 & 0.2 & intermediate & 4 & 1 / 4 & 0 \\
+99 & 100 & 1 & 0.2 & few intense & 2 & 1 / 2 & 0 \\
+100 & 200 & 1 & 0.2 & many modest  & 8 & 1 / 8 & 0 \\
+    \hline
+  \end{tabular}
+  \caption{
+    Overview of the MBD parameters 2/3. 
+  }
+  \label{table:mbd_parameters_2}
+\end{table}
+
+\begin{table}
+  \centering
+  \begin{tabular}{ l | c | c | c | c   c   c | c }
+    \hline
+idx & $n_{taxa}$ & $f_{taxa}^{MB}$ & $\mu$ & MB regime & $n_{\nu}$ evts & $q$ & $\lambda$ \\
+    \hline
+    \hline
+101 & 200 & 0 & 0 & intermediate & 4 & 1 / 4 & derive \\
+102 & 200 & 0 & 0 & few intense & 2 & 1 / 2 & derive \\
+103 & 200 & 0 & 0 & many modest  & 8 & 1 / 8 & derive \\
+104 & 200 & 0 & 0.1 & intermediate & 4 & 1 / 4 & derive \\
+105 & 200 & 0 & 0.1 & few intense & 2 & 1 / 2 & derive \\
+106 & 200 & 0 & 0.1 & many modest  & 8 & 1 / 8 & derive \\
+107 & 200 & 0 & 0.2 & intermediate & 4 & 1 / 4 & derive \\
+108 & 200 & 0 & 0.2 & few intense & 2 & 1 / 2 & derive \\
+109 & 200 & 0 & 0.2 & many modest  & 8 & 1 / 8 & derive \\
+110 & 200 & 0 & 0.2 & intermediate & 4 & 1 / 4 & derive \\
+111 & 200 & 0.25 & 0 & few intense & 2 & 1 / 2 & derive \\
+112 & 200 & 0.25 & 0 & many modest  & 8 & 1 / 8 & derive \\
+113 & 200 & 0.25 & 0 & intermediate & 4 & 1 / 4 & derive \\
+114 & 200 & 0.25 & 0.1 & few intense & 2 & 1 / 2 & derive \\
+115 & 200 & 0.25 & 0.1 & many modest  & 8 & 1 / 8 & derive \\
+116 & 200 & 0.25 & 0.1 & intermediate & 4 & 1 / 4 & derive \\
+117 & 200 & 0.25 & 0.2 & few intense & 2 & 1 / 2 & derive \\
+118 & 200 & 0.25 & 0.2 & many modest  & 8 & 1 / 8 & derive \\
+119 & 200 & 0.25 & 0.2 & intermediate & 4 & 1 / 4 & derive \\
+120 & 200 & 0.25 & 0.2 & few intense & 2 & 1 / 2 & derive \\
+121 & 200 & 0.5 & 0 & many modest  & 8 & 1 / 8 & derive \\
+122 & 200 & 0.5 & 0 & intermediate & 4 & 1 / 4 & derive \\
+123 & 200 & 0.5 & 0 & few intense & 2 & 1 / 2 & derive \\
+124 & 200 & 0.5 & 0.1 & many modest  & 8 & 1 / 8 & derive \\
+125 & 200 & 0.5 & 0.1 & intermediate & 4 & 1 / 4 & derive \\
+126 & 200 & 0.5 & 0.1 & few intense & 2 & 1 / 2 & derive \\
+127 & 200 & 0.5 & 0.2 & many modest  & 8 & 1 / 8 & derive \\
+128 & 200 & 0.5 & 0.2 & intermediate & 4 & 1 / 4 & derive \\
+129 & 200 & 0.5 & 0.2 & few intense & 2 & 1 / 2 & derive \\
+130 & 200 & 0.5 & 0.2 & many modest  & 8 & 1 / 8 & derive \\
+131 & 200 & 0.75 & 0 & intermediate & 4 & 1 / 4 & derive \\
+132 & 200 & 0.75 & 0 & few intense & 2 & 1 / 2 & derive \\
+133 & 200 & 0.75 & 0 & many modest  & 8 & 1 / 8 & derive \\
+134 & 200 & 0.75 & 0.1 & intermediate & 4 & 1 / 4 & derive \\
+135 & 200 & 0.75 & 0.1 & few intense & 2 & 1 / 2 & derive \\
+136 & 200 & 0.75 & 0.1 & many modest  & 8 & 1 / 8 & derive \\
+137 & 200 & 0.75 & 0.2 & intermediate & 4 & 1 / 4 & derive \\
+138 & 200 & 0.75 & 0.2 & few intense & 2 & 1 / 2 & derive \\
+139 & 200 & 0.75 & 0.2 & many modest  & 8 & 1 / 8 & derive \\
+140 & 200 & 0.75 & 0.2 & intermediate & 4 & 1 / 4 & derive \\
+141 & 200 & 1 & 0 & few intense & 2 & 1 / 2 & 0 \\
+142 & 200 & 1 & 0 & many modest  & 8 & 1 / 8 & 0 \\
+143 & 200 & 1 & 0 & intermediate & 4 & 1 / 4 & 0 \\
+144 & 200 & 1 & 0.1 & few intense & 2 & 1 / 2 & 0 \\
+145 & 200 & 1 & 0.1 & many modest  & 8 & 1 / 8 & 0 \\
+146 & 200 & 1 & 0.1 & intermediate & 4 & 1 / 4 & 0 \\
+147 & 200 & 1 & 0.2 & few intense & 2 & 1 / 2 & 0 \\
+148 & 200 & 1 & 0.2 & many modest  & 8 & 1 / 8 & 0 \\
+149 & 200 & 1 & 0.2 & intermediate & 4 & 1 / 4 & 0 \\
+150 & 200 & 1 & 0.2 & few intense & 2 & 1 / 2 & 0 \\
+    \hline
+  \end{tabular}
+  \caption{
+    Overview of the MBD parameters 3/3. 
+  }
+  \label{table:mbd_parameters_3}
+\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2018-12-04 article.tex
EOD
Aim for 'Molecular Phylogenetics and Evolution'

@@
   % comprehensible to a scientist in any discipline.
+  Phylogenetics is the field in which phylogenies are
+  created from DNA, RNA or protein alignments.
+  The evolutionary relationships between species or
+  individuals
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\richel{Have you already looked up for a target journal?} 
-\gio{Honestly I have literally no idea how to select a good journal for this kind of article.}
-\richel{
-  May I suggest we aim for Molecular Phylogenetics and Evolution, the same journal as the raket paper?
-}
 

### 2018-12-04 article.tex
First sketch of abstract

@@
   % comprehensible to a scientist in any discipline.
-  Phylogenetics is the field in which phylogenies are
-  created from DNA, RNA or protein alignments.
-  The evolutionary relationships between species or
-  individuals
+  There exist millions of species on Earth,
+  all originating from a common ancestor billions
+  of years ago.
+  The field of phylogenetics uses heritable material
+  to determine which species are closest related and what are
+  the mathematics that shape speciation.
 
@@
   % scientists in related disciplines.
+  In Bayesian phylogenetics, a DNA/RNA/protein alignment
+  is used to infer a distribution of phylogenies and parameter estimates.
+  To do so, we use assumptions that may be biologically unrealistic,
+  but may give tolerable errors.
 
@@
   % study.
+  Contemporary inference assumes that speciation
+  never co-occurs.
 
@@
   % their equivalent).
+  Here we show the error we make in our inference,
+  when nature has varying degrees of co-occuring speciation.
 
@@
 
+\richel{duplicate paragraph!}
 \item The (constant-rate) birth-death (BD) model 

### 2018-12-04 article.tex
@Giappo will love my pick for the MBD and BD symbol :-)

@@
 \usepackage[english]{babel}
+
+\usepackage{bbm}
+\usepackage{dsfont}
 
@@
 when using a more complex MBD tree prior is warranted.
+To do so, we simulate phylogenies using the MBD process, with
+varying degrees of that process. To be explicit, we define
+that degree, $s$, as the number of extinct and extant species created
+during a co-occuring speciation event, 
+$n_{\mathbb{M}}$, from the total number of extinct and extant species:
+
+\begin{equation}
+s = \frac{n_{\mathbb{M}}}{n_{\mathbb{M}} + n_{\mathbb{B}}}
+\end{equation}
+
+Here, $n_{\mathbb{B}}$ is the number of extinct and extant species created
+during a default single-birth speciation event.
 

### 2018-12-04 article.tex
Put symbols and parameters in different tables

@@
 
+From such phylogenies with different degrees of $s$, we measure the
+inference error we make today, would nature follow such a phylogeny.
+The inference error we make today is caused by the assumption of a BD process
+and by inherent noise in this inference.
+
+We expect
 % Hypothesis: general
@@
 \begin{equation}
-<e> = f(\frac{n_{taxa}^{MBD}}{n_{taxa}^{BD} + n_{taxa}^{MBD}})
+<e> = f(s)
 \end{equation}
@@
     \hline
-                          & Description & Value(s) \\
+                          & Description\\
     \hline
     \hline
-    $L_c$                 & MCMC chain length & $10^6$ \\
-    $L_{sc}$              & MCMC sub-chain length & $10^4$ \\
-    $\epsilon$            & relative error tolerance in marginal likelihood estimation & $10^{-13}$ \\
+    $n_{\mathbb{M}}$      & Number of extinct and extant taxa created by a multiple-birth event \\
+    $n_{\mathbb{B}}$      & Number of extinct and extant taxa created by a single-birth event \\
+    $s$                   & Extent/strength of the multiple birth process \\
+    $L_c$                 & MCMC chain length \\
+    $L_{sc}$              & MCMC sub-chain length \\
+    $\epsilon$            & relative error tolerance in marginal likelihood estimation \\
+    \hline
+  \end{tabular}
+  \caption{
+    Overview of the symbols. 
+  }
+  \label{table:symbols}
+\end{table}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{table}
+  \centering 
+  \begin{tabular}{r l}
+    \hline
+                          & Value(s) \\
+    \hline
+    \hline
+    $L_c$                 & $10^6$ \\
+    $L_{sc}$              & $10^4$ \\
+    $\epsilon$            & $10^{-13}$ \\
     \hline

### 2018-12-04 article.tex
Put symbols and parameters in different tables with a nice layout
Cluster hypotheses as far as this follows naturally

@@
 
-We expect
-% Hypothesis: general
-\item We hypothesize that the error made today, using BD tree priors,
-increases with an increased number or stronger effect of multiple birth events.
-This is straightforward: without multiple birth events or such event having
-no effect, the MBD model falls back to a BD model. We expect larger
-errors when we deviate more from the BD model's assumptions. 
-Additionally, we hypothesize MBD having a stronger effect if the normal
-speciation process is less pronounced. The more speciations are caused by the
-BD process, there are relatively less multiple-birth events.
-To put this hypothesis, H1, into an explicit equation, 
-we expect the error made be correlated to the number 
-of species created by the multiple-birth process over the total number 
-of species created:
+We have the hypothesis $\mathcal{H}_1$ that, for a higher $s$, 
+the inference error $e$ will increase:
 
 \begin{equation}
-<e> = f(s)
+e = f(s)
 \end{equation}
 
-Where $<e>$ denotes the expected error, $f$ is a monotonously increasing 
-function of unknown shape, $n_{taxa}^{MBD}$ is the number of taxa created
-in multiple-birth events and $n_{taxa}^{BD}$ is the number of taxa created
-by the standard BD specition process.
+Where $f$ is a monotonously increasing function of unknown shape.
 
@@
 
-\item We have the hypothesis, H7, 
-that there is no difference in inference errors if the MBD process
-generates many species in many modest MB events or few intense MB events.
+The MBD process has multiple components, that may cause 
+different values of $e$ for identical values of $s$.
+
+One MBD component that may cause 
+different values of $e$ for identical values of $s$
+is the MBD regime: which can be many modest 
+multiple-speciation events, or few intense ones. 
+As we have no prior expectations, we have the (null) hypothesis, 
+$\mathcal{H}_7$, that the MBD regime has no effect on $e$.
 
@@
 
-\item We have the hypothesis, H2, 
-that the effect of extinction rates is neutral, 
-as extinctions will hit lineages created by both speciation processes equally.
+Another MBD component that may cause 
+different values of $e$ for identical values of $s$
+is the effect of extinction. 
+As extinctions will hit lineages created by both speciation processes equally,
+and we have no additionaly prior expectations,
+we have the (null) hypothesis, 
+$\mathcal{H}_2$, that extinction has no effect on $e$.
 
@@
 
-\item Due to the proportionality of the term within $f$, 
-we have the hypothesis, H3, 
-that the timing of a multiple birth event (be it close to the
-crown age or close to the present) to have no effect. Compared to a late
-multiple birth event, an early multiple birth event may have a 
+Another MBD component that may cause 
+different values of $e$ for identical values of $s$,
+is the timing of a multiple-birth event: be it close to the
+crown age or close to the present. 
+Compared to a late multiple birth event, an early multiple birth event may have a 
 longer-lasting effect (as the next speciation event will be later), but it
 will create less new species, as there are still fewer taxa.
+As we have no prior expectations,
+we have the (null) hypothesis, 
+$\mathcal{H}_3$, that the timing of multiple-birth events has no effect on $e$.
 
@@
   \label{fig_h_3}
+\end{figure}
+
+Another MBD component that may cause 
+different values of $e$ for identical values of $s$,
+is the number of taxa in a phylogeny,
+As there is no diversity dependency in any of the processes
+and we have no further prior expectations, we have
+the (null) hypothesis $\mathcal{H}_5$, that the number of
+taxa has no effect on $e$.
+As a higher number of taxa increases the information content in a
+phylogeny, we have hypothesis $\mathcal{H}_6$ that the
+variance in $e$ decreases.
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{fig_h_5.png}
+  \caption{
+    Hypothesis 5: the number of taxa does not
+    have an effect on the error being made
+  }
+  \label{fig_h_5}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{fig_h_6.png}
+  \caption{
+    Hypothesis 6: for a higher number of taxa 
+    the variance in the error decreases
+  }
+  \label{fig_h_6}
 \end{figure}
@@
 with either 50, 100 and 200 taxa.
-We have the hypothesis H5, that the number of taxa does not
-have an effect on the error being made, as there is no
-diversity dependency in any of the processes.
-We have the hypothesis H6, that for a higher number of taxa 
-the variance in the error decreases, as more information is present
-in the simulated phylogenies.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_h_5.png}
-  \caption{
-    Hypothesis 5: the number of taxa does not
-    have an effect on the error being made
-  }
-  \label{fig_h_5}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_h_6.png}
-  \caption{
-    Hypothesis 6: for a higher number of taxa 
-    the variance in the error decreases
-  }
-  \label{fig_h_6}
-\end{figure}
 
@@
   \centering 
-  \begin{tabular}{p{0.05\textwidth} p{0.6\textwidth} p{0.2\textwidth}}
+  \begin{tabular}{r l}
     \hline
@@
     \hline
+    $e$                   & Inference error \\
     $n_{\mathbb{M}}$      & Number of extinct and extant taxa created by a multiple-birth event \\

### 2018-12-04 article.tex
List prior hypotheses

@@
   \centering 
+  \begin{tabular}{r c l}
+    \hline
+    Symbol                & Expectation & Figure \\
+    \hline
+    \hline
+    $\mathcal{H}_1$       & For a higher $s$, $e$ will increase monotonously & 1 \\
+    $\mathcal{H}_2$       & Extinction rates have no effect on $e$ & 2 \\
+    $\mathcal{H}_3$       & The timing of multiple-birth events has no effect on $e$ & 3 \\
+    $\mathcal{H}_4$       & An MBD tree will have a higher $e$ than its twin BD counterpart & 4 \\
+    $\mathcal{H}_5$       & The number of taxa has no effect on $e$ & 5 \\
+    $\mathcal{H}_6$       & A higher number of taxa decreases the variance in $e$ & 6 \\
+    $\mathcal{H}_7$       & The MBD regime has no effect on $e$ & 7 \\
+    \hline
+  \end{tabular}
+  \caption{
+    Overview of the symbols. 
+  }
+  \label{table:hypotheses}
+\end{table}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{table}
+  \centering 
   \begin{tabular}{r l}

### 2018-12-04 article.tex
Remove duplicate paragraph about the African Big Lakes

@@
 in multiple clades at a same point in time.
-
-\richel{duplicate paragraph!}
-\item The (constant-rate) birth-death (BD) model 
-embodies the common assumption that 
-only a single speciation event can occur at any given time.
-The multiple-birth-death (MBD) model relaxes this assumption, 
-allowing events in which 
-large-scale environmental changes lead to a great number of species 
-in relatively short time intervals. 
-Such hypothesis can be useful to describe, for example, 
-systems like cichlid fish diversification in the 
-African Great Lakes: Malawi, Tanganyika and Victoria 
-(\cite{janzen2016}, \cite{janzen2017}).
 

### 2018-12-04 article.tex
Differentiate between number of taxa that are extant (n) and are extant or extinct (N)

@@
 during a co-occuring speciation event, 
-$n_{\mathbb{M}}$, from the total number of extinct and extant species:
+$N_{\mathbb{M}}$, from the total number of extinct and extant species:
 
 \begin{equation}
-s = \frac{n_{\mathbb{M}}}{n_{\mathbb{M}} + n_{\mathbb{B}}}
+s = \frac{N_{\mathbb{M}}}{N_{\mathbb{M}} + N_{\mathbb{B}}}
 \end{equation}
 
-Here, $n_{\mathbb{B}}$ is the number of extinct and extant species created
+Here, $N_{\mathbb{B}}$ is the number of extinct and extant species created
 during a default single-birth speciation event.
@@
 
-\item Current phylogenetic tools assume that only a single speciation event 
+\item 
+\richel{TODO: move to Introduction}
+Current phylogenetic tools assume that only a single speciation event 
 can occur at any given time.
@@
 
-\item In the MBD model, parameters $\lambda$ and $\mu$ correspond, respectively, 
+\item 
+\richel{TODO: move to Introduction}
+In the MBD model, parameters $\lambda$ and $\mu$ correspond, respectively, 
 to the common per-species speciation and extinction rates present 
@@
 
-\item To prove our hypothesis we simulate two twin datasets. 
-All the simulations are produced in continuous time, 
-using the Doob-Gillespie algorithm.
+\item To investigate the effect of $s$ on $e$, we simulate phylogenies
+for different values of $s$ spread equally from zero (no multiple-birth
+event, thus a BD model) to one (species are created only from multiple-birth
+events) and three intermediate values of 0.25, 0.5 and 0.75.
 
@@
     $e$                   & Inference error \\
-    $n_{\mathbb{M}}$      & Number of extinct and extant taxa created by a multiple-birth event \\
-    $n_{\mathbb{B}}$      & Number of extinct and extant taxa created by a single-birth event \\
+    $n_{\mathbb{M}}$      & Number of extant taxa created by a multiple-birth event \\
+    $n_{\mathbb{B}}$      & Number of extant taxa created by a single-birth event \\
+    $N_{\mathbb{M}}$      & Number of extinct and extant taxa created by a multiple-birth event \\
+    $N_{\mathbb{B}}$      & Number of extinct and extant taxa created by a single-birth event \\
     $s$                   & Extent/strength of the multiple birth process \\
@@
     \hline
+    $s$                   & 0.0, 0.25, 0.5, 0.75, 1.0 \\
     $L_c$                 & $10^6$ \\

### 2018-12-04 article.tex
Add fig 1
Add figure references

@@
   }
-  \label{fig_h_1}
+  \label{fig:h_1}
 \end{figure}
@@
   }
-  \label{fig_h_7}
+  \label{fig:h_7}
 \end{figure}
@@
   }
-  \label{fig_h_2}
+  \label{fig:h_2}
 \end{figure}
@@
   }
-  \label{fig_h_3}
+  \label{fig:h_3}
 \end{figure}
@@
   }
-  \label{fig_h_5}
+  \label{fig:h_5}
 \end{figure}
@@
   }
-  \label{fig_h_6}
+  \label{fig:h_6}
 \end{figure}
@@
 event, thus a BD model) to one (species are created only from multiple-birth
-events) and three intermediate values of 0.25, 0.5 and 0.75.
+events) and three intermediate values of 0.25, 0.5 and 0.75. Aggregating all
+components of $s$, we show the effect of $s$ on $e$ in figure 1.
 
@@
   }
-  \label{fig_h_4}
+  \label{fig:h_4}
 \end{figure}
@@
 
-\item
+\item We expected the effect of $s$ on $e$ to be monotonously increasing
+function (figure \ref{fig:h_1}). Figure \ref{fig:1} shows the effect 
+measured in this experiment.
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{fig_1.png}
+  \caption{
+    The effect of s on e
+  }
+  \label{fig:1}
+\end{figure}
 
@@
   \centering 
-  \begin{tabular}{r c l}
+  \begin{tabular}{r c c c}
     \hline
-    Symbol                & Expectation & Figure \\
+    Symbol                & Expectation & Expected figure & Measured figure \\
     \hline
     \hline
-    $\mathcal{H}_1$       & For a higher $s$, $e$ will increase monotonously & 1 \\
-    $\mathcal{H}_2$       & Extinction rates have no effect on $e$ & 2 \\
-    $\mathcal{H}_3$       & The timing of multiple-birth events has no effect on $e$ & 3 \\
-    $\mathcal{H}_4$       & An MBD tree will have a higher $e$ than its twin BD counterpart & 4 \\
-    $\mathcal{H}_5$       & The number of taxa has no effect on $e$ & 5 \\
-    $\mathcal{H}_6$       & A higher number of taxa decreases the variance in $e$ & 6 \\
-    $\mathcal{H}_7$       & The MBD regime has no effect on $e$ & 7 \\
+    $\mathcal{H}_1$       & For a higher $s$, $e$ will increase monotonously & \ref{fig:h_1} & \ref{fig:1} \\
+    $\mathcal{H}_2$       & Extinction rates have no effect on $e$ & \ref{fig:h_2} & Not yet \\
+    $\mathcal{H}_3$       & The timing of multiple-birth events has no effect on $e$ & \ref{fig:h_3} & Not yet \\
+    $\mathcal{H}_4$       & An MBD tree will have a higher $e$ than its twin BD counterpart & \ref{fig:h_4} & Not yet \\
+    $\mathcal{H}_5$       & The number of taxa has no effect on $e$ & \ref{fig:h_5} & Not yet \\
+    $\mathcal{H}_6$       & A higher number of taxa decreases the variance in $e$ & \ref{fig:h_6} & Not yet \\
+    $\mathcal{H}_7$       & The MBD regime has no effect on $e$ & \ref{fig:h_7} & Not yet \\
     \hline

### 2018-12-04 article.tex
Move hypothesis to start of appendix to not seperate symbols and parameter values
Add MBD symbols lambda, mu, nu, q

@@
 events) and three intermediate values of 0.25, 0.5 and 0.75. Aggregating all
-components of $s$, we show the effect of $s$ on $e$ in figure 1.
+components of $s$, we show the effect of $s$ on $e$ in figure \ref{fig:1}.
 
@@
   \centering 
-  \begin{tabular}{r l}
-    \hline
-                          & Description\\
-    \hline
-    \hline
-    $e$                   & Inference error \\
-    $n_{\mathbb{M}}$      & Number of extant taxa created by a multiple-birth event \\
-    $n_{\mathbb{B}}$      & Number of extant taxa created by a single-birth event \\
-    $N_{\mathbb{M}}$      & Number of extinct and extant taxa created by a multiple-birth event \\
-    $N_{\mathbb{B}}$      & Number of extinct and extant taxa created by a single-birth event \\
-    $s$                   & Extent/strength of the multiple birth process \\
-    $L_c$                 & MCMC chain length \\
-    $L_{sc}$              & MCMC sub-chain length \\
-    $\epsilon$            & relative error tolerance in marginal likelihood estimation \\
-    \hline
-  \end{tabular}
-  \caption{
-    Overview of the symbols. 
-  }
-  \label{table:symbols}
-\end{table}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table}
-  \centering 
   \begin{tabular}{r c c c}
@@
     \hline
+                          & Description\\
+    \hline
+    \hline
+    $\lambda$             & Speciation rate from default single-birth process \\
+    $\mu$                 & Extinction rate \\
+    $\nu$                 & Multiple-birth event trigger rate \\
+    $q$                   & Proportion of taxa that speciate during a multiple-birth event \\
+    $e$                   & Inference error \\
+    $n_{\mathbb{M}}$      & Number of extant taxa created by a multiple-birth event \\
+    $n_{\mathbb{B}}$      & Number of extant taxa created by a single-birth event \\
+    $N_{\mathbb{M}}$      & Number of extinct and extant taxa created by a multiple-birth event \\
+    $N_{\mathbb{B}}$      & Number of extinct and extant taxa created by a single-birth event \\
+    $s$                   & Extent/strength of the multiple birth process \\
+    $L_c$                 & MCMC chain length \\
+    $L_{sc}$              & MCMC sub-chain length \\
+    $\epsilon$            & relative error tolerance in marginal likelihood estimation \\
+    \hline
+  \end{tabular}
+  \caption{
+    Overview of the symbols. 
+  }
+  \label{table:symbols}
+\end{table}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{table}
+  \centering 
+  \begin{tabular}{r l}
+    \hline
                           & Value(s) \\
@@
     $s$                   & 0.0, 0.25, 0.5, 0.75, 1.0 \\
+    $\mu$                 & 0.0, 0.1, 0.2 [TODO: can be more extreme] \\
     $L_c$                 & $10^6$ \\

### 2018-12-04 article.tex
Add effect of taxa

@@
 
+\item To see the effect of the number of taxa on the relation 
+between $s$ and $e$, we simulate phylogenies
+for different number of extant taxa $n$. We use values of $n$
+of 50, 100 and 200 extant taxa, as this will result in phylogenies
+that are big enough to be useful, yet small of enough to
+be computationally feasible. For the different values of $n$, we show
+the relation between $s$ and $e$ in figure \ref{fig:5} and its effect
+on the variance of $e$ in figure \ref{fig:6}.
+
+$s$ spread equally from zero (no multiple-birth
+event, thus a BD model) to one (species are created only from multiple-birth
+events) and three intermediate values of 0.25, 0.5 and 0.75. Aggregating all
+components of $s$, we show the effect of $s$ on $e$ in figure \ref{fig:1}.
+
 \item We start simulating $N_{S} = 1000$ MBD trees,
@@
   \label{fig:1}
+\end{figure}
+
+\item We expected the effect of the number of taxa on the relation
+between $s$ and $e$ to be neutral (figure \ref{fig:h_5}). 
+Figure \ref{fig:5} shows the effect 
+measured in this experiment.
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{fig_5.png}
+  \caption{
+    The effect of number of taxa on the relation between s on e
+  }
+  \label{fig:5}
+\end{figure}
+
+\item We expected the effect of the number of taxa on the variance
+of $e$ for similar values of $s$ to decrease (figure \ref{fig:h_6}). 
+Figure \ref{fig:6} shows the effect 
+measured in this experiment.
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{fig_6.png}
+  \caption{
+    The effect of number of taxa on the variance in e
+    for similar values of s
+  }
+  \label{fig:6}
 \end{figure}

### 2018-12-04 article.tex
Remove duplicate from MC
Effect of extinction

@@
 
-$s$ spread equally from zero (no multiple-birth
-event, thus a BD model) to one (species are created only from multiple-birth
-events) and three intermediate values of 0.25, 0.5 and 0.75. Aggregating all
-components of $s$, we show the effect of $s$ on $e$ in figure \ref{fig:1}.
+\item To see the effect of the extinction rate on the relation 
+between $s$ and $e$, we simulate phylogenies
+with different extinction rates $\mu$. We use values of $\mu$
+of 0.0, 0.1 and 0.2. An extinction rate of zero has two features:
+(1) the model falls back to a pure multiple-birth model, 
+(2) all multiple-birth events are observed.
+For the different values of $\mu$, we show
+the relation between $s$ and $e$ in figure \ref{fig:2}.
 
@@
 
+\item We expected the effect of extinction on the relation
+between $s$ and $e$ to be neutral (figure \ref{fig:h_2}). 
+Figure \ref{fig:2} shows the effect 
+measured in this experiment.
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{fig_2.png}
+  \caption{
+    The effect of extinction on the relation between s on e
+  }
+  \label{fig:2}
+\end{figure}
+
 \item
@@
     $\mathcal{H}_1$       & For a higher $s$, $e$ will increase monotonously & \ref{fig:h_1} & \ref{fig:1} \\
-    $\mathcal{H}_2$       & Extinction rates have no effect on $e$ & \ref{fig:h_2} & Not yet \\
+    $\mathcal{H}_2$       & Extinction rates have no effect on $e$ & \ref{fig:h_2} & \ref{fig:2} \\
     $\mathcal{H}_3$       & The timing of multiple-birth events has no effect on $e$ & \ref{fig:h_3} & Not yet \\
     $\mathcal{H}_4$       & An MBD tree will have a higher $e$ than its twin BD counterpart & \ref{fig:h_4} & Not yet \\
-    $\mathcal{H}_5$       & The number of taxa has no effect on $e$ & \ref{fig:h_5} & Not yet \\
-    $\mathcal{H}_6$       & A higher number of taxa decreases the variance in $e$ & \ref{fig:h_6} & Not yet \\
+    $\mathcal{H}_5$       & The number of taxa has no effect on $e$ & \ref{fig:h_5} & \ref{fig:5} \\
+    $\mathcal{H}_6$       & A higher number of taxa decreases the variance in $e$ & \ref{fig:h_6} & \ref{fig:6} \\
     $\mathcal{H}_7$       & The MBD regime has no effect on $e$ & \ref{fig:h_7} & Not yet \\

### 2019-05-07 article.tex
EOD
Split up in article.tex (big picture), content.tex (the content) and (supplement.tex) supplementary materials

--- 2018-12-04 article.tex
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{abstract}
+\include{content}
 
-  % From 'How to construct a Nature summary paragraph'
-
-  % A short abstract of fewer than 200 words is acceptable.
-
-  % One or two sentences providing a basic
-  % introduction to the field,
-  % comprehensible to a scientist in any discipline.
-  There exist millions of species on Earth,
-  all originating from a common ancestor billions
-  of years ago.
-  The field of phylogenetics uses heritable material
-  to determine which species are closest related and what are
-  the mathematics that shape speciation.
-
-  % Two to three sentences of
-  % more detailed background, comprehensible to
-  % scientists in related disciplines.
-  In Bayesian phylogenetics, a DNA/RNA/protein alignment
-  is used to infer a distribution of phylogenies and parameter estimates.
-  To do so, we use assumptions that may be biologically unrealistic,
-  but may give tolerable errors.
-
-  % One sentence clearly stating the general
-  % problem being addressed by this particular
-  % study.
-  Contemporary inference assumes that speciation
-  never co-occurs.
-
-  % One sentence summarising the main
-  % result (with the words “here we show”
-  % their equivalent).
-  Here we show the error we make in our inference,
-  when nature has varying degrees of co-occuring speciation.
-
-  % Two or three sentences explaining what
-  % the main result reveals in direct
-  % comparison to what was thought to be the case
-  % previously, or how the main result adds to
-  % previous knowledge.
-
-  % One or two sentences to put the results into a
-  % more general context.
-
-  % Two or three sentences to provide a
-  % broader perspective, readily comprehensible
-  % to a scientist in any discipline, may be included
-  % in the first paragraph if the editor considers that the accessibility of the paper is significantly enhanced
-  % by their inclusion. Under these circumstances, the length of the paragraph can be up to 300 words.
-  % (The above example is 190 words without the final section, and 250 words with it).
-
-\end{abstract}
-
-{\bf Keywords:} computational biology, evolution, phylogenetics, Bayesian analysis, tree prior
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-\begin{itemize}
-
-\item There are many contemporary tools that provide the possibility 
-to infer a phylogeny from genetic data (DNA, RNA, proteins). 
-A popular Bayesian phylogenetic tool is called BEAST (\cite{beast}) 
-and its cousin BEAST2 (\cite{beast2}).
-
-\item BEAST is very flexible, providing the user with the option 
-to set up all possible phylogenetic priors (e.g. site/clock/speciation model).
-
-%Current limits in current tools.
-\item However, currently available priors can be not suitable 
-to analyze some specific datasets. 
-With this work we aim to test whether or not 
-the implementation of a new prior model 
-is beneficial to study a specific kind of diversification process.
-
-\item BEAST2 gives us the possibility to introduce new tree priors 
-to infer phylogenies based on different assumptions 
-on how the speciation process takes place.
-
-\item One of such speciation processes is the multiple birth hypothesis,
-a new model (described below) and thus currently absent in BEAST.
-
-\item The Multiple birth hypothesis can be useful to explain a phenomenon 
-that has always puzzled evolutionary biologists: 
-what are the drivers of the diversification processes 
-for those phylogenies that show an impressive amount of speciation events 
-in relatively short times? 
-The (constant-rate) birth-death (BD) model embodies the common assumption that 
-only a single speciation event can occur at any given time.
-The multiple-birth-death (MBD) model 
-relaxes this assumption, allowing events in which 
-large-scale environmental changes lead to a great number of species 
-in relatively short time intervals. 
-Such a hypothesis may be a better fit to describe the burst in systems 
-like cichlid fish diversification in the 
-African Great Lakes: Malawi, Tanganyika and Victoria 
-(\cite{janzen2016}, \cite{janzen2017}).
-
-\item However, it may be that current BD tree priors are good enough 
-at detecting such events, with a (preferred) lower level of complexity. 
-If this is the case one should always be more keen to adopt the simplest model.
-
-\item Here we present our study with the aim of exploring 
-when using a more complex MBD tree prior is warranted.
-To do so, we simulate phylogenies using the MBD process, with
-varying degrees of that process. To be explicit, we define
-that degree, $s$, as the number of extinct and extant species created
-during a co-occuring speciation event, 
-$N_{\mathbb{M}}$, from the total number of extinct and extant species:
-
-\begin{equation}
-s = \frac{N_{\mathbb{M}}}{N_{\mathbb{M}} + N_{\mathbb{B}}}
-\end{equation}
-
-Here, $N_{\mathbb{B}}$ is the number of extinct and extant species created
-during a default single-birth speciation event.
-
-From such phylogenies with different degrees of $s$, we measure the
-inference error we make today, would nature follow such a phylogeny.
-The inference error we make today is caused by the assumption of a BD process
-and by inherent noise in this inference.
-
-We have the hypothesis $\mathcal{H}_1$ that, for a higher $s$, 
-the inference error $e$ will increase:
-
-\begin{equation}
-e = f(s)
-\end{equation}
-
-Where $f$ is a monotonously increasing function of unknown shape.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_h_1.png}
-  \caption{
-    Hypothesis 1: we expect the error made be correlated to the number 
-    of species created by the multiple-birth process over the total number 
-    of species created
-  }
-  \label{fig:h_1}
-\end{figure}
-
-The MBD process has multiple components, that may cause 
-different values of $e$ for identical values of $s$.
-
-One MBD component that may cause 
-different values of $e$ for identical values of $s$
-is the MBD regime: which can be many modest 
-multiple-speciation events, or few intense ones. 
-As we have no prior expectations, we have the (null) hypothesis, 
-$\mathcal{H}_7$, that the MBD regime has no effect on $e$.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_h_7.png}
-  \caption{
-    Hypothesis 7: there is no difference in inference errors if the MBD process
-    generates many species in many modest MB events or few intense MB events
-  }
-  \label{fig:h_7}
-\end{figure}
-
-Another MBD component that may cause 
-different values of $e$ for identical values of $s$
-is the effect of extinction. 
-As extinctions will hit lineages created by both speciation processes equally,
-and we have no additionaly prior expectations,
-we have the (null) hypothesis, 
-$\mathcal{H}_2$, that extinction has no effect on $e$.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_h_2.png}
-  \caption{
-    Hypothesis 2: the effect of extinction rates is neutral
-  }
-  \label{fig:h_2}
-\end{figure}
-
-Another MBD component that may cause 
-different values of $e$ for identical values of $s$,
-is the timing of a multiple-birth event: be it close to the
-crown age or close to the present. 
-Compared to a late multiple birth event, an early multiple birth event may have a 
-longer-lasting effect (as the next speciation event will be later), but it
-will create less new species, as there are still fewer taxa.
-As we have no prior expectations,
-we have the (null) hypothesis, 
-$\mathcal{H}_3$, that the timing of multiple-birth events has no effect on $e$.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_h_3.png}
-  \caption{
-    Hypothesis 3: the timing of a multiple birth event has no effect.
-  }
-  \label{fig:h_3}
-\end{figure}
-
-Another MBD component that may cause 
-different values of $e$ for identical values of $s$,
-is the number of taxa in a phylogeny,
-As there is no diversity dependency in any of the processes
-and we have no further prior expectations, we have
-the (null) hypothesis $\mathcal{H}_5$, that the number of
-taxa has no effect on $e$.
-As a higher number of taxa increases the information content in a
-phylogeny, we have hypothesis $\mathcal{H}_6$ that the
-variance in $e$ decreases.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_h_5.png}
-  \caption{
-    Hypothesis 5: the number of taxa does not
-    have an effect on the error being made
-  }
-  \label{fig:h_5}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_h_6.png}
-  \caption{
-    Hypothesis 6: for a higher number of taxa 
-    the variance in the error decreases
-  }
-  \label{fig:h_6}
-\end{figure}
-
-\end{itemize}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Methods}
-
-\subsection{Model}
-\begin{itemize}
-
-\item 
-\richel{TODO: move to Introduction}
-Current phylogenetic tools assume that only a single speciation event 
-can occur at any given time.
-While this assumption is useful to construct a wide variety of successful 
-models (e.g \cite{Maddison2007biSSE}, \cite{Valente2015}, 
-\cite{etienne2012diversity}, \cite{etienne2014estimating}),
-they disallow for environmental changes that trigger speciations 
-in multiple clades at a same point in time.
-
-\item 
-\richel{TODO: move to Introduction}
-In the MBD model, parameters $\lambda$ and $\mu$ correspond, respectively, 
-to the common per-species speciation and extinction rates present 
-also in the standard BD model. 
-Additionally, MBD relies on two additional parameters. 
-Parameter $\nu$ is the rate at which an environmental change is triggered.
-When such event is triggered, 
-all species present in the phylogeny at that moment
-have a probability $q$ to speciate at that time, which is 
-independent on $\lambda$. 
-Polytomies are not allowed in such process 
-as each species can speciate only once at the time.
-
-\item It is also possible to write down a likelihood function 
-for such processes as in \cite{mbd}.
-
-\end{itemize}
-
-\subsection{Simulations}
-\begin{itemize}
-
-\item To investigate the effect of $s$ on $e$, we simulate phylogenies
-for different values of $s$ spread equally from zero (no multiple-birth
-event, thus a BD model) to one (species are created only from multiple-birth
-events) and three intermediate values of 0.25, 0.5 and 0.75. Aggregating all
-components of $s$, we show the effect of $s$ on $e$ in figure \ref{fig:1}.
-
-\item We start simulating $N_{S} = 1000$ MBD trees,
-with either 50, 100 and 200 taxa.
-
-\item From each MBD tree, a DNA sequence alignment is simulated. 
-For each sequence alignment we then perform a Bayesian analysis 
-to recover a posterior distribution of trees, 
-each composed of $N_{P}$ phylogenies. 
-Such analysis is performed using 
-the 'pirouette' package (\cite{pirouette}) to call the BEAST2 tool 
-suite from R. 
-We let the Bayesian analysis assume a BD prior in both cases, 
-to investigate the extent of the error we make under this assumption.
-
-\item For each tree generated under the MBD model 
-we aim to generate a "twin" tree under the BD model. 
-With the word "twin" 
-we denote a tree generated starting from the respective MBD tree, 
-in order to perform a fair comparison with it. 
-This operation has to be done, 
-because we want to compare two trees 
-that are generated by different processes. 
-To do so we infer the parameters $\lambda_{BD}$ and $\mu_{BD}$ 
-from the MBD maximizing the likelihood under a BD model. 
-To perform this operation we use the function "\texttt{bd\_ML}" 
-from the package "\texttt{DDD}" (\cite{etienne2012diversity}).
-
-\item We then exploit such parameters to generate a BD tree 
-using the function "\texttt{tess.sim.taxa.age}" 
-from the package "\texttt{TESS}" (\cite{Hoehna2013}). 
-We simulate the tree in such a way the new tree 
-has the same number of tips and the same crown age as the MBD tree. 
-We furthermore require that the BD tree conserve the topology of the MBD tree.
-We have hypothesis H4 that, compared to the MBD trees, 
-the error will be less in the BD twin tree.
-The difference between the errors made in MBD and twin BD trees indicates
-the impact the MBD process has on the error we make in inference using a
-contemporary BD prior.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_h_4.png}
-  \caption{
-    Hypothesis 4: compared to the MBD trees, 
-    the error will be less in the BD twin tree
-  }
-  \label{fig:h_4}
-\end{figure}
-
-We want the MBD and twin BD trees to contain the same amount of information, 
-i.e. the same number of DNA mutations and the same number of taxa at the present:
-
-\begin{equation}
-m_{MBD} = m_{BD} \label{m equivalence}
-\end{equation}
-
-The expected number of mutations $m$ of a phylogeny 
-with crown age $-T$ (with $T>0$) in fact is given by
-\richel{
-  So one of use likes '-T', the other likes 'T'. How to resolve this?
-}
-
-\begin{equation}
-m = L \cdot \rho \cdot \int_{0}^{T} n(t)\ dt \label{m calculation}
-\end{equation}
-
-where $L$ is the number of DNA nucleotides, 
-$\rho$ is the per-site per-species mutation rate and
-$n(t)$ the number of species at each time.
-
-The parameter we'll tune is $\rho$ ... \richel{elaborate here :-)}
-
-Since we cannot know $n_{BD}(t)$ before running simulations
-we need to replace it with a proxy. 
-For this reason we will use the average number of
-species in time according to the BD model. 
-It's well known that this is equal to \gio{insert proper citation}
-
-\begin{equation}
-    <n_{BD}>(t) = n_{0} \cdot e^{(\mu_{BD} - \lambda_{BD})t} \label{BD average n}
-\end{equation}
-
-where $n_{0} = n_{BD}(-T) = n_{MBD}(-T)$ is the initial number of species 
-at the crown age.
-From \ref{m equivalence}, \ref{m calculation} and \ref{BD average n} follows:
-
-\begin{equation}
-m_{MBD} = L \cdot \rho \cdot \int_{0}^{T} <n_{BD}>(t)\ dt \\
-= L \cdot \rho \cdot n_{0} \cdot \left[ \frac{e^{(\mu_{BD} - \lambda_{BD})T} - 1}{\mu_{BD} - \lambda_{BD}} \right]
-\end{equation}
-
-If we set $\mu_{BD} = \mu_{MBD}$ and reverse this relation 
-we can extrapolate the value of $\lambda_{BD}$ to use to generate BD trees.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{mbd.jpg}
-  \caption{
-    How to create twin trees and alignments. 
-    From a focal MBD tree, a twin tree is produced as 
-    such: (1) estimate the $\lambda_{BD}$ to get 
-    the same expected number of tips, (2) simulate a BD tree 
-    with that amount of tips (discard trees with different number of tips), 
-    (3) estimate a mutation rate to get an alignment 
-    with the same expected number of mutations, (4) simulate alignments 
-    with that amount of mutations (discard those that don't, 
-    the picture shows an alignment that should be discarded) 
-  }
-\end{figure}
-
-\item We explained how we set the parameters for each twin BD tree. 
-Using this rules we generate a BD dataset. 
-We repeat the analysis, producing alignments for each tree 
-and subsequently using BEAST to produce a posterior for each of them.
-
-\subsection{Measuring the inference error}
-
-\item So far we have simulated two datasets of trees under the two models: 
-$\{T_{i}^{BD}\}_{i=1}^{N_{S}}$ and $\{T_{i}^{MBD}\}_{i=1}^{N_{S}}$.
-We used them to generate a dataset of alignments for each model: $\{X^{BD}_{i}\}_{i=1}^{N_{S}}$ and $\{X^{MBD}_{i}\}_{i=1}^{N_{S}}$.
-From each dataset we produced a posterior distribution from a BD prior:
-$P_{i}(\theta | X^{BD}_{i}, BD)$ and $P_{i}(\theta | X^{MBD}_{i}, BD)$.
-\gio{
-  1) We might want to rename the models, e.g. BD = (0) and MBD = (1). 
-  These names with capital letters are too big and ugly;
-}
-\richel{
-  I would strongly prefer MBD and BD, as I feel replacing the big ugly 
-  capital letters by short pretty numbers hurts readability even more 
-}
-
-\item To compare the results for the two models we measure the inference 
-error using the nLTT statistic between known/true tree and 
-posterior/inferred trees (\cite{nltt}). 
-To obtain such statistics the procedure is the following:
-
-- From each tree $T_{i,j}^{M}$ (with $j=1,...,N_{S}$) 
-  belonging to the posterior $P_{i}(\theta | X^{M}_{i}, BD)$ 
-  and relative to the model $M$, we extrapolate the lineage-through-time (LTT), 
-  in other words we measure the number of species as a function of 
-  time $n_{i,j}(t)$. To allow a comparison we normalize dividing by the 
-  maximum number of species of each tree, i.e. the number of tips at the 
-  present $N_{i,j}(t)=\frac{n_{i,j}(t)}{n^{max}_{i,j}}$. We then define the 
-  nLTT measure as
-
-$nLTT_{i,j} = \int_{0}^{T} | N_{i,j}(t) - N_{T_{i}} | dt$
-
-\gio{I am running out of letters :(}
-\richel{Haha! I suggest to use the same equation and symbols 
-  as equation 1 in
-  the nLTT article of Janzen, Hoehna and Etienne, 2015:
-}
-
-$$
-\Delta nLTT = \int_{0}^{1} | nLTT_1(t) - nLTT_2{t} | dt
-$$
-
-\subsection{Model selection}
-
-We simulate alignments using the simplest nucleotide substitution model (JC69),
-the simplest clock model (strict). It is thus imminent to assume these
-models in our Bayesian inference. Nevertheless, the phylogeny the alignment
-was based on, could have followed either an MBD or BD tree model, 
-where we in both cases assume a BD tree model. This will have 
-an unknown effect on our inference: it may theoretically be that an MBD model
-generates (a tree that generates) an alignment in which a different site 
-and/or clock model is favored.
-
-We investigate this by measuring if the generative model (with the simplest
-nucleotide substitution and simplest clock model) is indeed selected 
-to be the best fitting model. 
-To be precise, we look at the model 
-with the highest marginal likelihood 
-(also called evidence \cite{mackay2003information}),
-$f(D|M)$, which is the probability of the data D given model M.
-In the context of this research, D consists of the DNA alignment,
-and M is the combination of site, clock and tree models.
-
-To estimate the marginal likelihood, 
-we use an algorithm named nested sampling \cite{skilling2006nested}.
-Nested sampling is attractive to use
-in a phylogentic context, as it gives a good estimation,
-requires little tuning \cite{maturana2018}.
-Nested sampling is available as a BEAST2 package
-and can be used by babette \cite{babette}.
-
-The nested sampling algorithm stops its run 
-when the marginal likelihood estimation error 
-reaches below a certain tolerance.  
-Similar to \cite{maturana2018},
-we use a (relative) error tolerance $\epsilon$ of $10^{-13}$,
-1 particle to explore the parameter space
-and 100 active points. 
-To achieve the latter, we use the MCMC chain length $L_c$ of 1M 
-(as also used in the parameter estimates),
-and a sub-chain length $L_{sc}$ of 10K.
-
-The models we use in our model comparison are the four combinations
-of two site models and two clock models. We use the JC69 site model, which
-is the (generative and) simplest model and GTR, the site model with most
-degrees of freedom. For the clock models, we use the strict clock model,
-which is the (generative and) simplest clock model, and the RLN clock model.
-\richel{Could also just be all site models and clock models = 8 models}
-
-From these four marginal likelihood estimates, we calculate the weight of
-the generative model and plot this in figure 2. We do this for both the 
-alignments derived from the MBD tree and the BD twin tree. We expect that
-the generative model has the heighest weight in both the MBD and BD alignments.
-We expect this weight to be higher in the BD alignments.
-
-\end{itemize}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
-\begin{itemize}
-
-\item We expected the effect of $s$ on $e$ to be monotonously increasing
-function (figure \ref{fig:h_1}). Figure \ref{fig:1} shows the effect 
-measured in this experiment.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_1.png}
-  \caption{
-    The effect of s on e
-  }
-  \label{fig:1}
-\end{figure}
-
-\item
-
-\end{itemize}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Bibliography % MEE style
-\bibliographystyle{mee}
-\bibliography{article}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\appendix
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table}
-  \centering 
-  \begin{tabular}{r c c c}
-    \hline
-    Symbol                & Expectation & Expected figure & Measured figure \\
-    \hline
-    \hline
-    $\mathcal{H}_1$       & For a higher $s$, $e$ will increase monotonously & \ref{fig:h_1} & \ref{fig:1} \\
-    $\mathcal{H}_2$       & Extinction rates have no effect on $e$ & \ref{fig:h_2} & Not yet \\
-    $\mathcal{H}_3$       & The timing of multiple-birth events has no effect on $e$ & \ref{fig:h_3} & Not yet \\
-    $\mathcal{H}_4$       & An MBD tree will have a higher $e$ than its twin BD counterpart & \ref{fig:h_4} & Not yet \\
-    $\mathcal{H}_5$       & The number of taxa has no effect on $e$ & \ref{fig:h_5} & Not yet \\
-    $\mathcal{H}_6$       & A higher number of taxa decreases the variance in $e$ & \ref{fig:h_6} & Not yet \\
-    $\mathcal{H}_7$       & The MBD regime has no effect on $e$ & \ref{fig:h_7} & Not yet \\
-    \hline
-  \end{tabular}
-  \caption{
-    Overview of the symbols. 
-  }
-  \label{table:hypotheses}
-\end{table}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table}
-  \centering 
-  \begin{tabular}{r l}
-    \hline
-                          & Description\\
-    \hline
-    \hline
-    $\lambda$             & Speciation rate from default single-birth process \\
-    $\mu$                 & Extinction rate \\
-    $\nu$                 & Multiple-birth event trigger rate \\
-    $q$                   & Proportion of taxa that speciate during a multiple-birth event \\
-    $e$                   & Inference error \\
-    $n_{\mathbb{M}}$      & Number of extant taxa created by a multiple-birth event \\
-    $n_{\mathbb{B}}$      & Number of extant taxa created by a single-birth event \\
-    $N_{\mathbb{M}}$      & Number of extinct and extant taxa created by a multiple-birth event \\
-    $N_{\mathbb{B}}$      & Number of extinct and extant taxa created by a single-birth event \\
-    $s$                   & Extent/strength of the multiple birth process \\
-    $L_c$                 & MCMC chain length \\
-    $L_{sc}$              & MCMC sub-chain length \\
-    $\epsilon$            & relative error tolerance in marginal likelihood estimation \\
-    \hline
-  \end{tabular}
-  \caption{
-    Overview of the symbols. 
-  }
-  \label{table:symbols}
-\end{table}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table}
-  \centering 
-  \begin{tabular}{r l}
-    \hline
-                          & Value(s) \\
-    \hline
-    \hline
-    $s$                   & 0.0, 0.25, 0.5, 0.75, 1.0 \\
-    $\mu$                 & 0.0, 0.1, 0.2 [TODO: can be more extreme] \\
-    $L_c$                 & $10^6$ \\
-    $L_{sc}$              & $10^4$ \\
-    $\epsilon$            & $10^{-13}$ \\
-    \hline
-  \end{tabular}
-  \caption{
-    Overview of the simulation parameters. 
-  }
-  \label{table:parameters}
-\end{table}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table}
-  \centering 
-  \begin{tabular}{ l | c | c | c | c   c   c | c }
-    \hline
-idx & $n_{taxa}$ & $f_{taxa}^{MB}$ & $\mu$ & MB regime & $n_{\nu}$ evts & $q$ & $\lambda$ \\
-    \hline
-    \hline
-1 & 50 & 0 & 0 & many modest  & 8 & 1 / 8 & derive \\
-2 & 50 & 0 & 0 & intermediate & 4 & 1 / 4 & derive \\
-3 & 50 & 0 & 0 & few intense & 2 & 1 / 2 & derive \\
-4 & 50 & 0 & 0.1 & many modest  & 8 & 1 / 8 & derive \\
-5 & 50 & 0 & 0.1 & intermediate & 4 & 1 / 4 & derive \\
-6 & 50 & 0 & 0.1 & few intense & 2 & 1 / 2 & derive \\
-7 & 50 & 0 & 0.2 & many modest  & 8 & 1 / 8 & derive \\
-8 & 50 & 0 & 0.2 & intermediate & 4 & 1 / 4 & derive \\
-9 & 50 & 0 & 0.2 & few intense & 2 & 1 / 2 & derive \\
-10 & 50 & 0 & 0.2 & many modest  & 8 & 1 / 8 & derive \\
-11 & 50 & 0.25 & 0 & intermediate & 4 & 1 / 4 & derive \\
-12 & 50 & 0.25 & 0 & few intense & 2 & 1 / 2 & derive \\
-13 & 50 & 0.25 & 0 & many modest  & 8 & 1 / 8 & derive \\
-14 & 50 & 0.25 & 0.1 & intermediate & 4 & 1 / 4 & derive \\
-15 & 50 & 0.25 & 0.1 & few intense & 2 & 1 / 2 & derive \\
-16 & 50 & 0.25 & 0.1 & many modest  & 8 & 1 / 8 & derive \\
-17 & 50 & 0.25 & 0.2 & intermediate & 4 & 1 / 4 & derive \\
-18 & 50 & 0.25 & 0.2 & few intense & 2 & 1 / 2 & derive \\
-19 & 50 & 0.25 & 0.2 & many modest  & 8 & 1 / 8 & derive \\
-20 & 50 & 0.25 & 0.2 & intermediate & 4 & 1 / 4 & derive \\
-21 & 50 & 0.5 & 0 & few intense & 2 & 1 / 2 & derive \\
-22 & 50 & 0.5 & 0 & many modest  & 8 & 1 / 8 & derive \\
-23 & 50 & 0.5 & 0 & intermediate & 4 & 1 / 4 & derive \\
-24 & 50 & 0.5 & 0.1 & few intense & 2 & 1 / 2 & derive \\
-25 & 50 & 0.5 & 0.1 & many modest  & 8 & 1 / 8 & derive \\
-26 & 50 & 0.5 & 0.1 & intermediate & 4 & 1 / 4 & derive \\
-27 & 50 & 0.5 & 0.2 & few intense & 2 & 1 / 2 & derive \\
-28 & 50 & 0.5 & 0.2 & many modest  & 8 & 1 / 8 & derive \\
-29 & 50 & 0.5 & 0.2 & intermediate & 4 & 1 / 4 & derive \\
-30 & 50 & 0.5 & 0.2 & few intense & 2 & 1 / 2 & derive \\
-31 & 50 & 0.75 & 0 & many modest  & 8 & 1 / 8 & derive \\
-32 & 50 & 0.75 & 0 & intermediate & 4 & 1 / 4 & derive \\
-33 & 50 & 0.75 & 0 & few intense & 2 & 1 / 2 & derive \\
-34 & 50 & 0.75 & 0.1 & many modest  & 8 & 1 / 8 & derive \\
-35 & 50 & 0.75 & 0.1 & intermediate & 4 & 1 / 4 & derive \\
-36 & 50 & 0.75 & 0.1 & few intense & 2 & 1 / 2 & derive \\
-37 & 50 & 0.75 & 0.2 & many modest  & 8 & 1 / 8 & derive \\
-38 & 50 & 0.75 & 0.2 & intermediate & 4 & 1 / 4 & derive \\
-39 & 50 & 0.75 & 0.2 & few intense & 2 & 1 / 2 & derive \\
-40 & 50 & 0.75 & 0.2 & many modest  & 8 & 1 / 8 & derive \\
-41 & 50 & 1 & 0 & intermediate & 4 & 1 / 4 & 0 \\
-42 & 50 & 1 & 0 & few intense & 2 & 1 / 2 & 0 \\
-43 & 50 & 1 & 0 & many modest  & 8 & 1 / 8 & 0 \\
-44 & 50 & 1 & 0.1 & intermediate & 4 & 1 / 4 & 0 \\
-45 & 50 & 1 & 0.1 & few intense & 2 & 1 / 2 & 0 \\
-46 & 50 & 1 & 0.1 & many modest  & 8 & 1 / 8 & 0 \\
-47 & 50 & 1 & 0.2 & intermediate & 4 & 1 / 4 & 0 \\
-48 & 50 & 1 & 0.2 & few intense & 2 & 1 / 2 & 0 \\
-49 & 50 & 1 & 0.2 & many modest  & 8 & 1 / 8 & 0 \\
-50 & 50 & 1 & 0.2 & intermediate & 4 & 1 / 4 & 0 \\
-    \hline
-  \end{tabular}
-  \caption{
-    Overview of the MBD parameters 1/3. 
-  }
-  \label{table:mbd_parameters_1}
-\end{table}
-
-\begin{table}
-  \centering 
-  \begin{tabular}{ l | c | c | c | c   c   c | c }
-    \hline
-idx & $n_{taxa}$ & $f_{taxa}^{MB}$ & $\mu$ & MB regime & $n_{\nu}$ evts & $q$ & $\lambda$ \\
-    \hline
-    \hline
-51 & 100 & 0 & 0 & few intense & 2 & 1 / 2 & derive \\
-52 & 100 & 0 & 0 & many modest  & 8 & 1 / 8 & derive \\
-53 & 100 & 0 & 0 & intermediate & 4 & 1 / 4 & derive \\
-54 & 100 & 0 & 0.1 & few intense & 2 & 1 / 2 & derive \\
-55 & 100 & 0 & 0.1 & many modest  & 8 & 1 / 8 & derive \\
-56 & 100 & 0 & 0.1 & intermediate & 4 & 1 / 4 & derive \\
-57 & 100 & 0 & 0.2 & few intense & 2 & 1 / 2 & derive \\
-58 & 100 & 0 & 0.2 & many modest  & 8 & 1 / 8 & derive \\
-59 & 100 & 0 & 0.2 & intermediate & 4 & 1 / 4 & derive \\
-60 & 100 & 0 & 0.2 & few intense & 2 & 1 / 2 & derive \\
-61 & 100 & 0.25 & 0 & many modest  & 8 & 1 / 8 & derive \\
-62 & 100 & 0.25 & 0 & intermediate & 4 & 1 / 4 & derive \\
-63 & 100 & 0.25 & 0 & few intense & 2 & 1 / 2 & derive \\
-64 & 100 & 0.25 & 0.1 & many modest  & 8 & 1 / 8 & derive \\
-65 & 100 & 0.25 & 0.1 & intermediate & 4 & 1 / 4 & derive \\
-66 & 100 & 0.25 & 0.1 & few intense & 2 & 1 / 2 & derive \\
-67 & 100 & 0.25 & 0.2 & many modest  & 8 & 1 / 8 & derive \\
-68 & 100 & 0.25 & 0.2 & intermediate & 4 & 1 / 4 & derive \\
-69 & 100 & 0.25 & 0.2 & few intense & 2 & 1 / 2 & derive \\
-70 & 100 & 0.25 & 0.2 & many modest  & 8 & 1 / 8 & derive \\
-71 & 100 & 0.5 & 0 & intermediate & 4 & 1 / 4 & derive \\
-72 & 100 & 0.5 & 0 & few intense & 2 & 1 / 2 & derive \\
-73 & 100 & 0.5 & 0 & many modest  & 8 & 1 / 8 & derive \\
-74 & 100 & 0.5 & 0.1 & intermediate & 4 & 1 / 4 & derive \\
-75 & 100 & 0.5 & 0.1 & few intense & 2 & 1 / 2 & derive \\
-76 & 100 & 0.5 & 0.1 & many modest  & 8 & 1 / 8 & derive \\
-77 & 100 & 0.5 & 0.2 & intermediate & 4 & 1 / 4 & derive \\
-78 & 100 & 0.5 & 0.2 & few intense & 2 & 1 / 2 & derive \\
-79 & 100 & 0.5 & 0.2 & many modest  & 8 & 1 / 8 & derive \\
-80 & 100 & 0.5 & 0.2 & intermediate & 4 & 1 / 4 & derive \\
-81 & 100 & 0.75 & 0 & few intense & 2 & 1 / 2 & derive \\
-82 & 100 & 0.75 & 0 & many modest  & 8 & 1 / 8 & derive \\
-83 & 100 & 0.75 & 0 & intermediate & 4 & 1 / 4 & derive \\
-84 & 100 & 0.75 & 0.1 & few intense & 2 & 1 / 2 & derive \\
-85 & 100 & 0.75 & 0.1 & many modest  & 8 & 1 / 8 & derive \\
-86 & 100 & 0.75 & 0.1 & intermediate & 4 & 1 / 4 & derive \\
-87 & 100 & 0.75 & 0.2 & few intense & 2 & 1 / 2 & derive \\
-88 & 100 & 0.75 & 0.2 & many modest  & 8 & 1 / 8 & derive \\
-89 & 100 & 0.75 & 0.2 & intermediate & 4 & 1 / 4 & derive \\
-90 & 100 & 0.75 & 0.2 & few intense & 2 & 1 / 2 & derive \\
-91 & 100 & 1 & 0 & many modest  & 8 & 1 / 8 & 0 \\
-92 & 100 & 1 & 0 & intermediate & 4 & 1 / 4 & 0 \\
-93 & 100 & 1 & 0 & few intense & 2 & 1 / 2 & 0 \\
-94 & 100 & 1 & 0.1 & many modest  & 8 & 1 / 8 & 0 \\
-95 & 100 & 1 & 0.1 & intermediate & 4 & 1 / 4 & 0 \\
-96 & 100 & 1 & 0.1 & few intense & 2 & 1 / 2 & 0 \\
-97 & 100 & 1 & 0.2 & many modest  & 8 & 1 / 8 & 0 \\
-98 & 100 & 1 & 0.2 & intermediate & 4 & 1 / 4 & 0 \\
-99 & 100 & 1 & 0.2 & few intense & 2 & 1 / 2 & 0 \\
-100 & 200 & 1 & 0.2 & many modest  & 8 & 1 / 8 & 0 \\
-    \hline
-  \end{tabular}
-  \caption{
-    Overview of the MBD parameters 2/3. 
-  }
-  \label{table:mbd_parameters_2}
-\end{table}
-
-\begin{table}
-  \centering
-  \begin{tabular}{ l | c | c | c | c   c   c | c }
-    \hline
-idx & $n_{taxa}$ & $f_{taxa}^{MB}$ & $\mu$ & MB regime & $n_{\nu}$ evts & $q$ & $\lambda$ \\
-    \hline
-    \hline
-101 & 200 & 0 & 0 & intermediate & 4 & 1 / 4 & derive \\
-102 & 200 & 0 & 0 & few intense & 2 & 1 / 2 & derive \\
-103 & 200 & 0 & 0 & many modest  & 8 & 1 / 8 & derive \\
-104 & 200 & 0 & 0.1 & intermediate & 4 & 1 / 4 & derive \\
-105 & 200 & 0 & 0.1 & few intense & 2 & 1 / 2 & derive \\
-106 & 200 & 0 & 0.1 & many modest  & 8 & 1 / 8 & derive \\
-107 & 200 & 0 & 0.2 & intermediate & 4 & 1 / 4 & derive \\
-108 & 200 & 0 & 0.2 & few intense & 2 & 1 / 2 & derive \\
-109 & 200 & 0 & 0.2 & many modest  & 8 & 1 / 8 & derive \\
-110 & 200 & 0 & 0.2 & intermediate & 4 & 1 / 4 & derive \\
-111 & 200 & 0.25 & 0 & few intense & 2 & 1 / 2 & derive \\
-112 & 200 & 0.25 & 0 & many modest  & 8 & 1 / 8 & derive \\
-113 & 200 & 0.25 & 0 & intermediate & 4 & 1 / 4 & derive \\
-114 & 200 & 0.25 & 0.1 & few intense & 2 & 1 / 2 & derive \\
-115 & 200 & 0.25 & 0.1 & many modest  & 8 & 1 / 8 & derive \\
-116 & 200 & 0.25 & 0.1 & intermediate & 4 & 1 / 4 & derive \\
-117 & 200 & 0.25 & 0.2 & few intense & 2 & 1 / 2 & derive \\
-118 & 200 & 0.25 & 0.2 & many modest  & 8 & 1 / 8 & derive \\
-119 & 200 & 0.25 & 0.2 & intermediate & 4 & 1 / 4 & derive \\
-120 & 200 & 0.25 & 0.2 & few intense & 2 & 1 / 2 & derive \\
-121 & 200 & 0.5 & 0 & many modest  & 8 & 1 / 8 & derive \\
-122 & 200 & 0.5 & 0 & intermediate & 4 & 1 / 4 & derive \\
-123 & 200 & 0.5 & 0 & few intense & 2 & 1 / 2 & derive \\
-124 & 200 & 0.5 & 0.1 & many modest  & 8 & 1 / 8 & derive \\
-125 & 200 & 0.5 & 0.1 & intermediate & 4 & 1 / 4 & derive \\
-126 & 200 & 0.5 & 0.1 & few intense & 2 & 1 / 2 & derive \\
-127 & 200 & 0.5 & 0.2 & many modest  & 8 & 1 / 8 & derive \\
-128 & 200 & 0.5 & 0.2 & intermediate & 4 & 1 / 4 & derive \\
-129 & 200 & 0.5 & 0.2 & few intense & 2 & 1 / 2 & derive \\
-130 & 200 & 0.5 & 0.2 & many modest  & 8 & 1 / 8 & derive \\
-131 & 200 & 0.75 & 0 & intermediate & 4 & 1 / 4 & derive \\
-132 & 200 & 0.75 & 0 & few intense & 2 & 1 / 2 & derive \\
-133 & 200 & 0.75 & 0 & many modest  & 8 & 1 / 8 & derive \\
-134 & 200 & 0.75 & 0.1 & intermediate & 4 & 1 / 4 & derive \\
-135 & 200 & 0.75 & 0.1 & few intense & 2 & 1 / 2 & derive \\
-136 & 200 & 0.75 & 0.1 & many modest  & 8 & 1 / 8 & derive \\
-137 & 200 & 0.75 & 0.2 & intermediate & 4 & 1 / 4 & derive \\
-138 & 200 & 0.75 & 0.2 & few intense & 2 & 1 / 2 & derive \\
-139 & 200 & 0.75 & 0.2 & many modest  & 8 & 1 / 8 & derive \\
-140 & 200 & 0.75 & 0.2 & intermediate & 4 & 1 / 4 & derive \\
-141 & 200 & 1 & 0 & few intense & 2 & 1 / 2 & 0 \\
-142 & 200 & 1 & 0 & many modest  & 8 & 1 / 8 & 0 \\
-143 & 200 & 1 & 0 & intermediate & 4 & 1 / 4 & 0 \\
-144 & 200 & 1 & 0.1 & few intense & 2 & 1 / 2 & 0 \\
-145 & 200 & 1 & 0.1 & many modest  & 8 & 1 / 8 & 0 \\
-146 & 200 & 1 & 0.1 & intermediate & 4 & 1 / 4 & 0 \\
-147 & 200 & 1 & 0.2 & few intense & 2 & 1 / 2 & 0 \\
-148 & 200 & 1 & 0.2 & many modest  & 8 & 1 / 8 & 0 \\
-149 & 200 & 1 & 0.2 & intermediate & 4 & 1 / 4 & 0 \\
-150 & 200 & 1 & 0.2 & few intense & 2 & 1 / 2 & 0 \\
-    \hline
-  \end{tabular}
-  \caption{
-    Overview of the MBD parameters 3/3. 
-  }
-  \label{table:mbd_parameters_3}
-\end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\include{supplement}
 

### 2019-05-07 content.tex
Split up in article.tex (big picture), content.tex (the content) and (supplement.tex) supplementary materials

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

  % From 'How to construct a Nature summary paragraph'

  % A short abstract of fewer than 200 words is acceptable.

  % One or two sentences providing a basic
  % introduction to the field,
  % comprehensible to a scientist in any discipline.
  There exist millions of species on Earth,
  all originating from a common ancestor billions
  of years ago.
  The field of phylogenetics uses heritable material
  to determine which species are closest related and what are
  the mathematics that shape speciation.

  % Two to three sentences of
  % more detailed background, comprehensible to
  % scientists in related disciplines.
  In Bayesian phylogenetics, a DNA/RNA/protein alignment
  is used to infer a distribution of phylogenies and parameter estimates.
  To do so, we use assumptions that may be biologically unrealistic,
  but may give tolerable errors.

  % One sentence clearly stating the general
  % problem being addressed by this particular
  % study.
  Contemporary inference assumes that speciation
  never co-occurs.

  % One sentence summarising the main
  % result (with the words “here we show”
  % their equivalent).
  Here we show the error we make in our inference,
  when nature has varying degrees of co-occuring speciation.

  % Two or three sentences explaining what
  % the main result reveals in direct
  % comparison to what was thought to be the case
  % previously, or how the main result adds to
  % previous knowledge.

  % One or two sentences to put the results into a
  % more general context.

  % Two or three sentences to provide a
  % broader perspective, readily comprehensible
  % to a scientist in any discipline, may be included
  % in the first paragraph if the editor considers that the accessibility of the paper is significantly enhanced
  % by their inclusion. Under these circumstances, the length of the paragraph can be up to 300 words.
  % (The above example is 190 words without the final section, and 250 words with it).

\end{abstract}

{\bf Keywords:} computational biology, evolution, phylogenetics, Bayesian analysis, tree prior
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\begin{itemize}

\item There are many contemporary tools that provide the possibility 
to infer a phylogeny from genetic data (DNA, RNA, proteins). 
A popular Bayesian phylogenetic tool is called BEAST (\cite{beast}) 
and its cousin BEAST2 (\cite{beast2}).

\item BEAST is very flexible, providing the user with the option 
to set up all possible phylogenetic priors (e.g. site/clock/speciation model).

%Current limits in current tools.
\item However, currently available priors can be not suitable 
to analyze some specific datasets. 
With this work we aim to test whether or not 
the implementation of a new prior model 
is beneficial to study a specific kind of diversification process.

\item BEAST2 gives us the possibility to introduce new tree priors 
to infer phylogenies based on different assumptions 
on how the speciation process takes place.

\item One of such speciation processes is the multiple birth hypothesis,
a new model (described below) and thus currently absent in BEAST.

\item The Multiple birth hypothesis can be useful to explain a phenomenon 
that has always puzzled evolutionary biologists: 
what are the drivers of the diversification processes 
for those phylogenies that show an impressive amount of speciation events 
in relatively short times? 
The (constant-rate) birth-death (BD) model embodies the common assumption that 
only a single speciation event can occur at any given time.
The multiple-birth-death (MBD) model 
relaxes this assumption, allowing events in which 
large-scale environmental changes lead to a great number of species 
in relatively short time intervals. 
Such a hypothesis may be a better fit to describe the burst in systems 
like cichlid fish diversification in the 
African Great Lakes: Malawi, Tanganyika and Victoria 
(\cite{janzen2016}, \cite{janzen2017}).

\item However, it may be that current BD tree priors are good enough 
at detecting such events, with a (preferred) lower level of complexity. 
If this is the case one should always be more keen to adopt the simplest model.

\item Here we present our study with the aim of exploring 
when using a more complex MBD tree prior is warranted.
To do so, we simulate phylogenies using the MBD process, with
varying degrees of that process. To be explicit, we define
that degree, $s$, as the number of extinct and extant species created
during a co-occuring speciation event, 
$N_{\mathbb{M}}$, from the total number of extinct and extant species:

\begin{equation}
s = \frac{N_{\mathbb{M}}}{N_{\mathbb{M}} + N_{\mathbb{B}}}
\end{equation}

Here, $N_{\mathbb{B}}$ is the number of extinct and extant species created
during a default single-birth speciation event.

From such phylogenies with different degrees of $s$, we measure the
inference error we make today, would nature follow such a phylogeny.
The inference error we make today is caused by the assumption of a BD process
and by inherent noise in this inference.

We have the hypothesis $\mathcal{H}_1$ that, for a higher $s$, 
the inference error $e$ will increase:

\begin{equation}
e = f(s)
\end{equation}

Where $f$ is a monotonously increasing function of unknown shape.

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{fig_h_1.png}
  \caption{
    Hypothesis 1: we expect the error made be correlated to the number 
    of species created by the multiple-birth process over the total number 
    of species created
  }
  \label{fig:h_1}
\end{figure}

The MBD process has multiple components, that may cause 
different values of $e$ for identical values of $s$.

One MBD component that may cause 
different values of $e$ for identical values of $s$
is the MBD regime: which can be many modest 
multiple-speciation events, or few intense ones. 
As we have no prior expectations, we have the (null) hypothesis, 
$\mathcal{H}_7$, that the MBD regime has no effect on $e$.

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{fig_h_7.png}
  \caption{
    Hypothesis 7: there is no difference in inference errors if the MBD process
    generates many species in many modest MB events or few intense MB events
  }
  \label{fig:h_7}
\end{figure}

Another MBD component that may cause 
different values of $e$ for identical values of $s$
is the effect of extinction. 
As extinctions will hit lineages created by both speciation processes equally,
and we have no additionaly prior expectations,
we have the (null) hypothesis, 
$\mathcal{H}_2$, that extinction has no effect on $e$.

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{fig_h_2.png}
  \caption{
    Hypothesis 2: the effect of extinction rates is neutral
  }
  \label{fig:h_2}
\end{figure}

Another MBD component that may cause 
different values of $e$ for identical values of $s$,
is the timing of a multiple-birth event: be it close to the
crown age or close to the present. 
Compared to a late multiple birth event, an early multiple birth event may have a 
longer-lasting effect (as the next speciation event will be later), but it
will create less new species, as there are still fewer taxa.
As we have no prior expectations,
we have the (null) hypothesis, 
$\mathcal{H}_3$, that the timing of multiple-birth events has no effect on $e$.

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{fig_h_3.png}
  \caption{
    Hypothesis 3: the timing of a multiple birth event has no effect.
  }
  \label{fig:h_3}
\end{figure}

Another MBD component that may cause 
different values of $e$ for identical values of $s$,
is the number of taxa in a phylogeny,
As there is no diversity dependency in any of the processes
and we have no further prior expectations, we have
the (null) hypothesis $\mathcal{H}_5$, that the number of
taxa has no effect on $e$.
As a higher number of taxa increases the information content in a
phylogeny, we have hypothesis $\mathcal{H}_6$ that the
variance in $e$ decreases.

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{fig_h_5.png}
  \caption{
    Hypothesis 5: the number of taxa does not
    have an effect on the error being made
  }
  \label{fig:h_5}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{fig_h_6.png}
  \caption{
    Hypothesis 6: for a higher number of taxa 
    the variance in the error decreases
  }
  \label{fig:h_6}
\end{figure}

\end{itemize}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methods}

\subsection{Model}
\begin{itemize}

\item 
\richel{TODO: move to Introduction}
Current phylogenetic tools assume that only a single speciation event 
can occur at any given time.
While this assumption is useful to construct a wide variety of successful 
models (e.g \cite{Maddison2007biSSE}, \cite{Valente2015}, 
\cite{etienne2012diversity}, \cite{etienne2014estimating}),
they disallow for environmental changes that trigger speciations 
in multiple clades at a same point in time.

\item 
\richel{TODO: move to Introduction}
In the MBD model, parameters $\lambda$ and $\mu$ correspond, respectively, 
to the common per-species speciation and extinction rates present 
also in the standard BD model. 
Additionally, MBD relies on two additional parameters. 
Parameter $\nu$ is the rate at which an environmental change is triggered.
When such event is triggered, 
all species present in the phylogeny at that moment
have a probability $q$ to speciate at that time, which is 
independent on $\lambda$. 
Polytomies are not allowed in such process 
as each species can speciate only once at the time.

\item 
\richel{TODO: move to Introduction}
It is also possible to write down a likelihood function 
for such processes as in \cite{mbd}.

\end{itemize}

\subsection{Simulations}
\begin{itemize}

\item To investigate the effect of $s$ on $e$, we simulate phylogenies
for different values of $s$ spread equally from zero (no multiple-birth
event, thus a BD model) to one (species are created only from multiple-birth
events) and three intermediate values of 0.25, 0.5 and 0.75. Aggregating all
components of $s$, we show the effect of $s$ on $e$ in figure \ref{fig:1}.

\item To see the effect of the number of taxa on the relation 
between $s$ and $e$, we simulate phylogenies
for different number of extant taxa $n$. We use values of $n$
of 50, 100 and 200 extant taxa, as this will result in phylogenies
that are big enough to be useful, yet small of enough to
be computationally feasible. For the different values of $n$, we show
the relation between $s$ and $e$ in figure \ref{fig:5} and its effect
on the variance of $e$ in figure \ref{fig:6}.

\item To see the effect of the extinction rate on the relation 
between $s$ and $e$, we simulate phylogenies
with different extinction rates $\mu$. We use values of $\mu$
of 0.0, 0.1 and 0.2. An extinction rate of zero has two features:
(1) the model falls back to a pure multiple-birth model, 
(2) all multiple-birth events are observed.
For the different values of $\mu$, we show
the relation between $s$ and $e$ in figure \ref{fig:2}.

\item We start simulating $N_{S} = 1000$ MBD trees,
with either 50, 100 and 200 taxa.

\item From each MBD tree, a DNA sequence alignment is simulated. 
For each sequence alignment we then perform a Bayesian analysis 
to recover a posterior distribution of trees, 
each composed of $N_{P}$ phylogenies. 
Such analysis is performed using 
the 'pirouette' package (\cite{pirouette}) to call the BEAST2 tool 
suite from R. 
We let the Bayesian analysis assume a BD prior in both cases, 
to investigate the extent of the error we make under this assumption.

\item For each tree generated under the MBD model 
we aim to generate a "twin" tree under the BD model. 
With the word "twin" 
we denote a tree generated starting from the respective MBD tree, 
in order to perform a fair comparison with it. 
This operation has to be done, 
because we want to compare two trees 
that are generated by different processes. 
To do so we infer the parameters $\lambda_{BD}$ and $\mu_{BD}$ 
from the MBD maximizing the likelihood under a BD model. 
To perform this operation we use the function "\texttt{bd\_ML}" 
from the package "\texttt{DDD}" (\cite{etienne2012diversity}).

\item We then exploit such parameters to generate a BD tree 
using the function "\texttt{tess.sim.taxa.age}" 
from the package "\texttt{TESS}" (\cite{Hoehna2013}). 
We simulate the tree in such a way the new tree 
has the same number of tips and the same crown age as the MBD tree. 
We furthermore require that the BD tree conserve the topology of the MBD tree.
We have hypothesis H4 that, compared to the MBD trees, 
the error will be less in the BD twin tree.
The difference between the errors made in MBD and twin BD trees indicates
the impact the MBD process has on the error we make in inference using a
contemporary BD prior.

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{fig_h_4.png}
  \caption{
    Hypothesis 4: compared to the MBD trees, 
    the error will be less in the BD twin tree
  }
  \label{fig:h_4}
\end{figure}

We want the MBD and twin BD trees to contain the same amount of information, 
i.e. the same number of DNA mutations and the same number of taxa at the present:

\begin{equation}
m_{MBD} = m_{BD} \label{m equivalence}
\end{equation}

The expected number of mutations $m$ of a phylogeny 
with crown age $-T$ (with $T>0$) in fact is given by
\richel{
  So one of use likes '-T', the other likes 'T'. How to resolve this?
}

\begin{equation}
m = L \cdot \rho \cdot \int_{0}^{T} n(t)\ dt \label{m calculation}
\end{equation}

where $L$ is the number of DNA nucleotides, 
$\rho$ is the per-site per-species mutation rate and
$n(t)$ the number of species at each time.

The parameter we'll tune is $\rho$ ... \richel{elaborate here :-)}

Since we cannot know $n_{BD}(t)$ before running simulations
we need to replace it with a proxy. 
For this reason we will use the average number of
species in time according to the BD model. 
It's well known that this is equal to \gio{insert proper citation}

\begin{equation}
    <n_{BD}>(t) = n_{0} \cdot e^{(\mu_{BD} - \lambda_{BD})t} \label{BD average n}
\end{equation}

where $n_{0} = n_{BD}(-T) = n_{MBD}(-T)$ is the initial number of species 
at the crown age.
From \ref{m equivalence}, \ref{m calculation} and \ref{BD average n} follows:

\begin{equation}
m_{MBD} = L \cdot \rho \cdot \int_{0}^{T} <n_{BD}>(t)\ dt \\
= L \cdot \rho \cdot n_{0} \cdot \left[ \frac{e^{(\mu_{BD} - \lambda_{BD})T} - 1}{\mu_{BD} - \lambda_{BD}} \right]
\end{equation}

If we set $\mu_{BD} = \mu_{MBD}$ and reverse this relation 
we can extrapolate the value of $\lambda_{BD}$ to use to generate BD trees.

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{mbd.jpg}
  \caption{
    How to create twin trees and alignments. 
    From a focal MBD tree, a twin tree is produced as 
    such: (1) estimate the $\lambda_{BD}$ to get 
    the same expected number of tips, (2) simulate a BD tree 
    with that amount of tips (discard trees with different number of tips), 
    (3) estimate a mutation rate to get an alignment 
    with the same expected number of mutations, (4) simulate alignments 
    with that amount of mutations (discard those that don't, 
    the picture shows an alignment that should be discarded) 
  }
\end{figure}

\item We explained how we set the parameters for each twin BD tree. 
Using this rules we generate a BD dataset. 
We repeat the analysis, producing alignments for each tree 
and subsequently using BEAST to produce a posterior for each of them.

\subsection{Measuring the inference error}

\item So far we have simulated two datasets of trees under the two models: 
$\{T_{i}^{BD}\}_{i=1}^{N_{S}}$ and $\{T_{i}^{MBD}\}_{i=1}^{N_{S}}$.
We used them to generate a dataset of alignments for each model: $\{X^{BD}_{i}\}_{i=1}^{N_{S}}$ and $\{X^{MBD}_{i}\}_{i=1}^{N_{S}}$. From each dataset we produced a posterior distribution from a BD prior: 
$P_{i}(\theta | X^{BD}_{i}, BD)$ and $P_{i}(\theta | X^{MBD}_{i}, BD)$.
\gio{
  1) We might want to rename the models, e.g. BD = (0) and MBD = (1). 
  These names with capital letters are too big and ugly;
}
\richel{
  I would strongly prefer MBD and BD, as I feel replacing the big ugly 
  capital letters by short pretty numbers hurts readability even more 
}

\item To compare the results for the two models we measure the inference 
error using the nLTT statistic between known/true tree and 
posterior/inferred trees (\cite{nltt}). 
To obtain such statistics the procedure is the following:

- From each tree $T_{i,j}^{M}$ (with $j=1,...,N_{S}$) 
  belonging to the posterior $P_{i}(\theta | X^{M}_{i}, BD)$ 
  and relative to the model $M$, we extrapolate the lineage-through-time (LTT), 
  in other words we measure the number of species as a function of 
  time $n_{i,j}(t)$. To allow a comparison we normalize dividing by the 
  maximum number of species of each tree, i.e. the number of tips at the 
  present $N_{i,j}(t)=\frac{n_{i,j}(t)}{n^{max}_{i,j}}$. We then define the 
  nLTT measure as

$nLTT_{i,j} = \int_{0}^{T} | N_{i,j}(t) - N_{T_{i}} | dt$

\gio{I am running out of letters :(}
\richel{Haha! I suggest to use the same equation and symbols 
  as equation 1 in
  the nLTT article of Janzen, Hoehna and Etienne, 2015:
}

$$
\Delta nLTT = \int_{0}^{1} | nLTT_1(t) - nLTT_2{t} | dt
$$

\subsection{Model selection}

We simulate alignments using the simplest nucleotide substitution model (JC69),
the simplest clock model (strict). It is thus imminent to assume these
models in our Bayesian inference. Nevertheless, the phylogeny the alignment
was based on, could have followed either an MBD or BD tree model, 
where we in both cases assume a BD tree model. This will have 
an unknown effect on our inference: it may theoretically be that an MBD model
generates (a tree that generates) an alignment in which a different site 
and/or clock model is favored.

We investigate this by measuring if the generative model (with the simplest
nucleotide substitution and simplest clock model) is indeed selected 
to be the best fitting model. 
To be precise, we look at the model 
with the highest marginal likelihood 
(also called evidence \cite{mackay2003information}),
$f(D|M)$, which is the probability of the data D given model M.
In the context of this research, D consists of the DNA alignment,
and M is the combination of site, clock and tree models.

To estimate the marginal likelihood, 
we use an algorithm named nested sampling \cite{skilling2006nested}.
Nested sampling is attractive to use
in a phylogentic context, as it gives a good estimation,
requires little tuning \cite{maturana2018}.
Nested sampling is available as a BEAST2 package
and can be used by babette \cite{babette}.

The nested sampling algorithm stops its run 
when the marginal likelihood estimation error 
reaches below a certain tolerance.  
Similar to \cite{maturana2018},
we use a (relative) error tolerance $\epsilon$ of $10^{-13}$,
1 particle to explore the parameter space
and 100 active points. 
To achieve the latter, we use the MCMC chain length $L_c$ of 1M 
(as also used in the parameter estimates),
and a sub-chain length $L_{sc}$ of 10K.

The models we use in our model comparison are the four combinations
of two site models and two clock models. We use the JC69 site model, which
is the (generative and) simplest model and GTR, the site model with most
degrees of freedom. For the clock models, we use the strict clock model,
which is the (generative and) simplest clock model, and the RLN clock model.
\richel{Could also just be all site models and clock models = 8 models}

From these four marginal likelihood estimates, we calculate the weight of
the generative model and plot this in figure 2. We do this for both the 
alignments derived from the MBD tree and the BD twin tree. We expect that
the generative model has the heighest weight in both the MBD and BD alignments.
We expect this weight to be higher in the BD alignments.

\end{itemize}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
\begin{itemize}

\item We expected the effect of $s$ on $e$ to be monotonously increasing
function (figure \ref{fig:h_1}). Figure \ref{fig:1} shows the effect 
measured in this experiment.

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{fig_1.png}
  \caption{
    The effect of s on e
  }
  \label{fig:1}
\end{figure}

\item We expected the effect of the number of taxa on the relation
between $s$ and $e$ to be neutral (figure \ref{fig:h_5}). 
Figure \ref{fig:5} shows the effect 
measured in this experiment.

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{fig_5.png}
  \caption{
    The effect of number of taxa on the relation between s on e
  }
  \label{fig:5}
\end{figure}

\item We expected the effect of the number of taxa on the variance
of $e$ for similar values of $s$ to decrease (figure \ref{fig:h_6}). 
Figure \ref{fig:6} shows the effect 
measured in this experiment.

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{fig_6.png}
  \caption{
    The effect of number of taxa on the variance in e
    for similar values of s
  }
  \label{fig:6}
\end{figure}

\item We expected the effect of extinction on the relation
between $s$ and $e$ to be neutral (figure \ref{fig:h_2}). 
Figure \ref{fig:2} shows the effect 
measured in this experiment.

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{fig_2.png}
  \caption{
    The effect of extinction on the relation between s on e
  }
  \label{fig:2}
\end{figure}

\item

\end{itemize}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Bibliography % MEE style
\bibliographystyle{mee}
\bibliography{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-06-07 article.tex
Updates from Overleaf
Changes @Giappo

@@
 \documentclass{article}
+\usepackage{natbib} % Bibliography
+\bibpunct{(}{)}{;}{a}{}{;} % Bibliography
+\setcitestyle{authoryear,open={},close={}} % Use 'It was found that something is something (Name 1234)' style
 
-% Target journal: Molecular Phylogenetics and Evolution
-%
-% From author guidelines:
-%
-% Short communications of approximately 3000 words are also accepted. 
-% These papers should contain no more than two figures, two tables, and thirty references. 
-% A short abstract of fewer than 200 words is acceptable.
-
-% Annotation/feedback commands
+% Comments
 \newcommand*\rampal[1]{\textcolor{red}{\textbf{[RSE: #1]}}}
 \newcommand*\richel[1]{\textcolor{orange}{\textbf{[RJCB: #1]}}}
-\newcommand*\gio[1]{\textcolor{green}{\textbf{[GL: #1]}}}
+\newcommand*\gio[1]{\textcolor{blue}{\textbf{[GL: #1]}}}
 
-% Bibliography
-\usepackage{natbib}
-\bibpunct{(}{)}{;}{a}{}{;}
-
-\usepackage{amssymb}
-\usepackage[english]{babel}
-
-\usepackage{bbm}
-\usepackage{dsfont}
-
-% Use 'It was found that something is something (Name 1234)' style
-\setcitestyle{authoryear,open={},close={}}
-
-% Affiliations
-\usepackage{authblk}
-\title{The error in Bayesian phylogenetic reconstruction when speciation co-occurs}
-
-\author[1]{Giovanni Laudanno}
-\author[1]{Rich\`el J.C. Bilderbeek}
-\author[1]{Rampal S. Etienne}
-\affil[1]{Groningen Institute for Evolutionary Life Sciences, University of Groningen, Groningen, The Netherlands}
-
-% Use double spacing
-\usepackage{setspace}
-\doublespacing
-
-\usepackage{pgf}
-\usepackage{hyperref}
-\usepackage{verbatim}
-\usepackage{comment}
-
-% Adds numbered lines
-\usepackage{lineno}
-\linenumbers
+% Packages
+\input{razzo-packages.tex}
 
@@
 
+\input{razzo-header.tex}
+
 \begin{document}
@@
 
-\include{content}
+\input{razzo-abstract.tex}
 
-\include{supplement}
+\input{razzo-introduction.tex}
+
+\input{razzo-methods.tex}
+
+\input{razzo-results.tex}
+
+\bigskip\bigskip
+
+\bibliographystyle{mee}
+\bibliography{razzo-bibliography.bib}
+
+\bigskip\bigskip
+
+\input{razzo-supplement.tex}
 

### 2019-06-07 razzo-introduction.tex
Changes @Giappo

--- 2019-05-07 content.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{abstract}
-
-  % From 'How to construct a Nature summary paragraph'
-
-  % A short abstract of fewer than 200 words is acceptable.
-
-  % One or two sentences providing a basic
-  % introduction to the field,
-  % comprehensible to a scientist in any discipline.
-  There exist millions of species on Earth,
-  all originating from a common ancestor billions
-  of years ago.
-  The field of phylogenetics uses heritable material
-  to determine which species are closest related and what are
-  the mathematics that shape speciation.
-
-  % Two to three sentences of
-  % more detailed background, comprehensible to
-  % scientists in related disciplines.
-  In Bayesian phylogenetics, a DNA/RNA/protein alignment
-  is used to infer a distribution of phylogenies and parameter estimates.
-  To do so, we use assumptions that may be biologically unrealistic,
-  but may give tolerable errors.
-
-  % One sentence clearly stating the general
-  % problem being addressed by this particular
-  % study.
-  Contemporary inference assumes that speciation
-  never co-occurs.
-
-  % One sentence summarising the main
-  % result (with the words “here we show”
-  % their equivalent).
-  Here we show the error we make in our inference,
-  when nature has varying degrees of co-occuring speciation.
-
-  % Two or three sentences explaining what
-  % the main result reveals in direct
-  % comparison to what was thought to be the case
-  % previously, or how the main result adds to
-  % previous knowledge.
-
-  % One or two sentences to put the results into a
-  % more general context.
-
-  % Two or three sentences to provide a
-  % broader perspective, readily comprehensible
-  % to a scientist in any discipline, may be included
-  % in the first paragraph if the editor considers that the accessibility of the paper is significantly enhanced
-  % by their inclusion. Under these circumstances, the length of the paragraph can be up to 300 words.
-  % (The above example is 190 words without the final section, and 250 words with it).
-
-\end{abstract}
-
-{\bf Keywords:} computational biology, evolution, phylogenetics, Bayesian analysis, tree prior
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
@@
 
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_h_1.png}
-  \caption{
-    Hypothesis 1: we expect the error made be correlated to the number 
-    of species created by the multiple-birth process over the total number 
-    of species created
-  }
-  \label{fig:h_1}
-\end{figure}
+\input{razzo-fig_h_1.tex}
 
@@
 As we have no prior expectations, we have the (null) hypothesis, 
-$\mathcal{H}_7$, that the MBD regime has no effect on $e$.
+$\mathcal{H}_0$, that the MBD regime has no effect on $e$.
 
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_h_7.png}
-  \caption{
-    Hypothesis 7: there is no difference in inference errors if the MBD process
-    generates many species in many modest MB events or few intense MB events
-  }
-  \label{fig:h_7}
-\end{figure}
+\include{razzo-fig_h_0.tex}
 
@@
 
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_h_2.png}
-  \caption{
-    Hypothesis 2: the effect of extinction rates is neutral
-  }
-  \label{fig:h_2}
-\end{figure}
+\include{razzo-fig_h_2.tex}
 
@@
 
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_h_3.png}
-  \caption{
-    Hypothesis 3: the timing of a multiple birth event has no effect.
-  }
-  \label{fig:h_3}
-\end{figure}
+\include{razzo-fig_h_3.tex}
 
@@
 
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_h_5.png}
-  \caption{
-    Hypothesis 5: the number of taxa does not
-    have an effect on the error being made
-  }
-  \label{fig:h_5}
-\end{figure}
+\include{razzo-fig_h_5.tex}
 
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_h_6.png}
-  \caption{
-    Hypothesis 6: for a higher number of taxa 
-    the variance in the error decreases
-  }
-  \label{fig:h_6}
-\end{figure}
+\include{razzo-fig_h_6.tex}
 
 \end{itemize}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Methods}
-
-\subsection{Model}
-\begin{itemize}
-
-\item 
-\richel{TODO: move to Introduction}
-Current phylogenetic tools assume that only a single speciation event 
-can occur at any given time.
-While this assumption is useful to construct a wide variety of successful 
-models (e.g \cite{Maddison2007biSSE}, \cite{Valente2015}, 
-\cite{etienne2012diversity}, \cite{etienne2014estimating}),
-they disallow for environmental changes that trigger speciations 
-in multiple clades at a same point in time.
-
-\item 
-\richel{TODO: move to Introduction}
-In the MBD model, parameters $\lambda$ and $\mu$ correspond, respectively, 
-to the common per-species speciation and extinction rates present 
-also in the standard BD model. 
-Additionally, MBD relies on two additional parameters. 
-Parameter $\nu$ is the rate at which an environmental change is triggered.
-When such event is triggered, 
-all species present in the phylogeny at that moment
-have a probability $q$ to speciate at that time, which is 
-independent on $\lambda$. 
-Polytomies are not allowed in such process 
-as each species can speciate only once at the time.
-
-\item 
-\richel{TODO: move to Introduction}
-It is also possible to write down a likelihood function 
-for such processes as in \cite{mbd}.
-
-\end{itemize}
-
-\subsection{Simulations}
-\begin{itemize}
-
-\item To investigate the effect of $s$ on $e$, we simulate phylogenies
-for different values of $s$ spread equally from zero (no multiple-birth
-event, thus a BD model) to one (species are created only from multiple-birth
-events) and three intermediate values of 0.25, 0.5 and 0.75. Aggregating all
-components of $s$, we show the effect of $s$ on $e$ in figure \ref{fig:1}.
-
-\item To see the effect of the number of taxa on the relation 
-between $s$ and $e$, we simulate phylogenies
-for different number of extant taxa $n$. We use values of $n$
-of 50, 100 and 200 extant taxa, as this will result in phylogenies
-that are big enough to be useful, yet small of enough to
-be computationally feasible. For the different values of $n$, we show
-the relation between $s$ and $e$ in figure \ref{fig:5} and its effect
-on the variance of $e$ in figure \ref{fig:6}.
-
-\item To see the effect of the extinction rate on the relation 
-between $s$ and $e$, we simulate phylogenies
-with different extinction rates $\mu$. We use values of $\mu$
-of 0.0, 0.1 and 0.2. An extinction rate of zero has two features:
-(1) the model falls back to a pure multiple-birth model, 
-(2) all multiple-birth events are observed.
-For the different values of $\mu$, we show
-the relation between $s$ and $e$ in figure \ref{fig:2}.
-
-\item We start simulating $N_{S} = 1000$ MBD trees,
-with either 50, 100 and 200 taxa.
-
-\item From each MBD tree, a DNA sequence alignment is simulated. 
-For each sequence alignment we then perform a Bayesian analysis 
-to recover a posterior distribution of trees, 
-each composed of $N_{P}$ phylogenies. 
-Such analysis is performed using 
-the 'pirouette' package (\cite{pirouette}) to call the BEAST2 tool 
-suite from R. 
-We let the Bayesian analysis assume a BD prior in both cases, 
-to investigate the extent of the error we make under this assumption.
-
-\item For each tree generated under the MBD model 
-we aim to generate a "twin" tree under the BD model. 
-With the word "twin" 
-we denote a tree generated starting from the respective MBD tree, 
-in order to perform a fair comparison with it. 
-This operation has to be done, 
-because we want to compare two trees 
-that are generated by different processes. 
-To do so we infer the parameters $\lambda_{BD}$ and $\mu_{BD}$ 
-from the MBD maximizing the likelihood under a BD model. 
-To perform this operation we use the function "\texttt{bd\_ML}" 
-from the package "\texttt{DDD}" (\cite{etienne2012diversity}).
-
-\item We then exploit such parameters to generate a BD tree 
-using the function "\texttt{tess.sim.taxa.age}" 
-from the package "\texttt{TESS}" (\cite{Hoehna2013}). 
-We simulate the tree in such a way the new tree 
-has the same number of tips and the same crown age as the MBD tree. 
-We furthermore require that the BD tree conserve the topology of the MBD tree.
-We have hypothesis H4 that, compared to the MBD trees, 
-the error will be less in the BD twin tree.
-The difference between the errors made in MBD and twin BD trees indicates
-the impact the MBD process has on the error we make in inference using a
-contemporary BD prior.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_h_4.png}
-  \caption{
-    Hypothesis 4: compared to the MBD trees, 
-    the error will be less in the BD twin tree
-  }
-  \label{fig:h_4}
-\end{figure}
-
-We want the MBD and twin BD trees to contain the same amount of information, 
-i.e. the same number of DNA mutations and the same number of taxa at the present:
-
-\begin{equation}
-m_{MBD} = m_{BD} \label{m equivalence}
-\end{equation}
-
-The expected number of mutations $m$ of a phylogeny 
-with crown age $-T$ (with $T>0$) in fact is given by
-\richel{
-  So one of use likes '-T', the other likes 'T'. How to resolve this?
-}
-
-\begin{equation}
-m = L \cdot \rho \cdot \int_{0}^{T} n(t)\ dt \label{m calculation}
-\end{equation}
-
-where $L$ is the number of DNA nucleotides, 
-$\rho$ is the per-site per-species mutation rate and
-$n(t)$ the number of species at each time.
-
-The parameter we'll tune is $\rho$ ... \richel{elaborate here :-)}
-
-Since we cannot know $n_{BD}(t)$ before running simulations
-we need to replace it with a proxy. 
-For this reason we will use the average number of
-species in time according to the BD model. 
-It's well known that this is equal to \gio{insert proper citation}
-
-\begin{equation}
-    <n_{BD}>(t) = n_{0} \cdot e^{(\mu_{BD} - \lambda_{BD})t} \label{BD average n}
-\end{equation}
-
-where $n_{0} = n_{BD}(-T) = n_{MBD}(-T)$ is the initial number of species 
-at the crown age.
-From \ref{m equivalence}, \ref{m calculation} and \ref{BD average n} follows:
-
-\begin{equation}
-m_{MBD} = L \cdot \rho \cdot \int_{0}^{T} <n_{BD}>(t)\ dt \\
-= L \cdot \rho \cdot n_{0} \cdot \left[ \frac{e^{(\mu_{BD} - \lambda_{BD})T} - 1}{\mu_{BD} - \lambda_{BD}} \right]
-\end{equation}
-
-If we set $\mu_{BD} = \mu_{MBD}$ and reverse this relation 
-we can extrapolate the value of $\lambda_{BD}$ to use to generate BD trees.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{mbd.jpg}
-  \caption{
-    How to create twin trees and alignments. 
-    From a focal MBD tree, a twin tree is produced as 
-    such: (1) estimate the $\lambda_{BD}$ to get 
-    the same expected number of tips, (2) simulate a BD tree 
-    with that amount of tips (discard trees with different number of tips), 
-    (3) estimate a mutation rate to get an alignment 
-    with the same expected number of mutations, (4) simulate alignments 
-    with that amount of mutations (discard those that don't, 
-    the picture shows an alignment that should be discarded) 
-  }
-\end{figure}
-
-\item We explained how we set the parameters for each twin BD tree. 
-Using this rules we generate a BD dataset. 
-We repeat the analysis, producing alignments for each tree 
-and subsequently using BEAST to produce a posterior for each of them.
-
-\subsection{Measuring the inference error}
-
-\item So far we have simulated two datasets of trees under the two models: 
-$\{T_{i}^{BD}\}_{i=1}^{N_{S}}$ and $\{T_{i}^{MBD}\}_{i=1}^{N_{S}}$.
-We used them to generate a dataset of alignments for each model: $\{X^{BD}_{i}\}_{i=1}^{N_{S}}$ and $\{X^{MBD}_{i}\}_{i=1}^{N_{S}}$.
-From each dataset we produced a posterior distribution from a BD prior:
-$P_{i}(\theta | X^{BD}_{i}, BD)$ and $P_{i}(\theta | X^{MBD}_{i}, BD)$.
-\gio{
-  1) We might want to rename the models, e.g. BD = (0) and MBD = (1). 
-  These names with capital letters are too big and ugly;
-}
-\richel{
-  I would strongly prefer MBD and BD, as I feel replacing the big ugly 
-  capital letters by short pretty numbers hurts readability even more 
-}
-
-\item To compare the results for the two models we measure the inference 
-error using the nLTT statistic between known/true tree and 
-posterior/inferred trees (\cite{nltt}). 
-To obtain such statistics the procedure is the following:
-
-- From each tree $T_{i,j}^{M}$ (with $j=1,...,N_{S}$) 
-  belonging to the posterior $P_{i}(\theta | X^{M}_{i}, BD)$ 
-  and relative to the model $M$, we extrapolate the lineage-through-time (LTT), 
-  in other words we measure the number of species as a function of 
-  time $n_{i,j}(t)$. To allow a comparison we normalize dividing by the 
-  maximum number of species of each tree, i.e. the number of tips at the 
-  present $N_{i,j}(t)=\frac{n_{i,j}(t)}{n^{max}_{i,j}}$. We then define the 
-  nLTT measure as
-
-$nLTT_{i,j} = \int_{0}^{T} | N_{i,j}(t) - N_{T_{i}} | dt$
-
-\gio{I am running out of letters :(}
-\richel{Haha! I suggest to use the same equation and symbols 
-  as equation 1 in
-  the nLTT article of Janzen, Hoehna and Etienne, 2015:
-}
-
-$$
-\Delta nLTT = \int_{0}^{1} | nLTT_1(t) - nLTT_2{t} | dt
-$$
-
-\subsection{Model selection}
-
-We simulate alignments using the simplest nucleotide substitution model (JC69),
-the simplest clock model (strict). It is thus imminent to assume these
-models in our Bayesian inference. Nevertheless, the phylogeny the alignment
-was based on, could have followed either an MBD or BD tree model, 
-where we in both cases assume a BD tree model. This will have 
-an unknown effect on our inference: it may theoretically be that an MBD model
-generates (a tree that generates) an alignment in which a different site 
-and/or clock model is favored.
-
-We investigate this by measuring if the generative model (with the simplest
-nucleotide substitution and simplest clock model) is indeed selected 
-to be the best fitting model. 
-To be precise, we look at the model 
-with the highest marginal likelihood 
-(also called evidence \cite{mackay2003information}),
-$f(D|M)$, which is the probability of the data D given model M.
-In the context of this research, D consists of the DNA alignment,
-and M is the combination of site, clock and tree models.
-
-To estimate the marginal likelihood, 
-we use an algorithm named nested sampling \cite{skilling2006nested}.
-Nested sampling is attractive to use
-in a phylogentic context, as it gives a good estimation,
-requires little tuning \cite{maturana2018}.
-Nested sampling is available as a BEAST2 package
-and can be used by babette \cite{babette}.
-
-The nested sampling algorithm stops its run 
-when the marginal likelihood estimation error 
-reaches below a certain tolerance.  
-Similar to \cite{maturana2018},
-we use a (relative) error tolerance $\epsilon$ of $10^{-13}$,
-1 particle to explore the parameter space
-and 100 active points. 
-To achieve the latter, we use the MCMC chain length $L_c$ of 1M 
-(as also used in the parameter estimates),
-and a sub-chain length $L_{sc}$ of 10K.
-
-The models we use in our model comparison are the four combinations
-of two site models and two clock models. We use the JC69 site model, which
-is the (generative and) simplest model and GTR, the site model with most
-degrees of freedom. For the clock models, we use the strict clock model,
-which is the (generative and) simplest clock model, and the RLN clock model.
-\richel{Could also just be all site models and clock models = 8 models}
-
-From these four marginal likelihood estimates, we calculate the weight of
-the generative model and plot this in figure 2. We do this for both the 
-alignments derived from the MBD tree and the BD twin tree. We expect that
-the generative model has the heighest weight in both the MBD and BD alignments.
-We expect this weight to be higher in the BD alignments.
-
-\end{itemize}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
-\begin{itemize}
-
-\item We expected the effect of $s$ on $e$ to be monotonously increasing
-function (figure \ref{fig:h_1}). Figure \ref{fig:1} shows the effect 
-measured in this experiment.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_1.png}
-  \caption{
-    The effect of s on e
-  }
-  \label{fig:1}
-\end{figure}
-
-\item We expected the effect of the number of taxa on the relation
-between $s$ and $e$ to be neutral (figure \ref{fig:h_5}). 
-Figure \ref{fig:5} shows the effect 
-measured in this experiment.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_5.png}
-  \caption{
-    The effect of number of taxa on the relation between s on e
-  }
-  \label{fig:5}
-\end{figure}
-
-\item We expected the effect of the number of taxa on the variance
-of $e$ for similar values of $s$ to decrease (figure \ref{fig:h_6}). 
-Figure \ref{fig:6} shows the effect 
-measured in this experiment.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_6.png}
-  \caption{
-    The effect of number of taxa on the variance in e
-    for similar values of s
-  }
-  \label{fig:6}
-\end{figure}
-
-\item We expected the effect of extinction on the relation
-between $s$ and $e$ to be neutral (figure \ref{fig:h_2}). 
-Figure \ref{fig:2} shows the effect 
-measured in this experiment.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_2.png}
-  \caption{
-    The effect of extinction on the relation between s on e
-  }
-  \label{fig:2}
-\end{figure}
-
-\item
-
-\end{itemize}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Bibliography % MEE style
-\bibliographystyle{mee}
-\bibliography{article}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-06-07 razzo-methods.tex
Changes @Giappo

--- 2019-05-07 content.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{abstract}
-
-  % From 'How to construct a Nature summary paragraph'
-
-  % A short abstract of fewer than 200 words is acceptable.
-
-  % One or two sentences providing a basic
-  % introduction to the field,
-  % comprehensible to a scientist in any discipline.
-  There exist millions of species on Earth,
-  all originating from a common ancestor billions
-  of years ago.
-  The field of phylogenetics uses heritable material
-  to determine which species are closest related and what are
-  the mathematics that shape speciation.
-
-  % Two to three sentences of
-  % more detailed background, comprehensible to
-  % scientists in related disciplines.
-  In Bayesian phylogenetics, a DNA/RNA/protein alignment
-  is used to infer a distribution of phylogenies and parameter estimates.
-  To do so, we use assumptions that may be biologically unrealistic,
-  but may give tolerable errors.
-
-  % One sentence clearly stating the general
-  % problem being addressed by this particular
-  % study.
-  Contemporary inference assumes that speciation
-  never co-occurs.
-
-  % One sentence summarising the main
-  % result (with the words “here we show”
-  % their equivalent).
-  Here we show the error we make in our inference,
-  when nature has varying degrees of co-occuring speciation.
-
-  % Two or three sentences explaining what
-  % the main result reveals in direct
-  % comparison to what was thought to be the case
-  % previously, or how the main result adds to
-  % previous knowledge.
-
-  % One or two sentences to put the results into a
-  % more general context.
-
-  % Two or three sentences to provide a
-  % broader perspective, readily comprehensible
-  % to a scientist in any discipline, may be included
-  % in the first paragraph if the editor considers that the accessibility of the paper is significantly enhanced
-  % by their inclusion. Under these circumstances, the length of the paragraph can be up to 300 words.
-  % (The above example is 190 words without the final section, and 250 words with it).
-
-\end{abstract}
-
-{\bf Keywords:} computational biology, evolution, phylogenetics, Bayesian analysis, tree prior
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-\begin{itemize}
-
-\item There are many contemporary tools that provide the possibility 
-to infer a phylogeny from genetic data (DNA, RNA, proteins). 
-A popular Bayesian phylogenetic tool is called BEAST (\cite{beast}) 
-and its cousin BEAST2 (\cite{beast2}).
-
-\item BEAST is very flexible, providing the user with the option 
-to set up all possible phylogenetic priors (e.g. site/clock/speciation model).
-
-%Current limits in current tools.
-\item However, currently available priors can be not suitable 
-to analyze some specific datasets. 
-With this work we aim to test whether or not 
-the implementation of a new prior model 
-is beneficial to study a specific kind of diversification process.
-
-\item BEAST2 gives us the possibility to introduce new tree priors 
-to infer phylogenies based on different assumptions 
-on how the speciation process takes place.
-
-\item One of such speciation processes is the multiple birth hypothesis,
-a new model (described below) and thus currently absent in BEAST.
-
-\item The Multiple birth hypothesis can be useful to explain a phenomenon 
-that has always puzzled evolutionary biologists: 
-what are the drivers of the diversification processes 
-for those phylogenies that show an impressive amount of speciation events 
-in relatively short times? 
-The (constant-rate) birth-death (BD) model embodies the common assumption that 
-only a single speciation event can occur at any given time.
-The multiple-birth-death (MBD) model 
-relaxes this assumption, allowing events in which 
-large-scale environmental changes lead to a great number of species 
-in relatively short time intervals. 
-Such a hypothesis may be a better fit to describe the burst in systems 
-like cichlid fish diversification in the 
-African Great Lakes: Malawi, Tanganyika and Victoria 
-(\cite{janzen2016}, \cite{janzen2017}).
-
-\item However, it may be that current BD tree priors are good enough 
-at detecting such events, with a (preferred) lower level of complexity. 
-If this is the case one should always be more keen to adopt the simplest model.
-
-\item Here we present our study with the aim of exploring 
-when using a more complex MBD tree prior is warranted.
-To do so, we simulate phylogenies using the MBD process, with
-varying degrees of that process. To be explicit, we define
-that degree, $s$, as the number of extinct and extant species created
-during a co-occuring speciation event, 
-$N_{\mathbb{M}}$, from the total number of extinct and extant species:
-
-\begin{equation}
-s = \frac{N_{\mathbb{M}}}{N_{\mathbb{M}} + N_{\mathbb{B}}}
-\end{equation}
-
-Here, $N_{\mathbb{B}}$ is the number of extinct and extant species created
-during a default single-birth speciation event.
-
-From such phylogenies with different degrees of $s$, we measure the
-inference error we make today, would nature follow such a phylogeny.
-The inference error we make today is caused by the assumption of a BD process
-and by inherent noise in this inference.
-
-We have the hypothesis $\mathcal{H}_1$ that, for a higher $s$, 
-the inference error $e$ will increase:
-
-\begin{equation}
-e = f(s)
-\end{equation}
-
-Where $f$ is a monotonously increasing function of unknown shape.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_h_1.png}
-  \caption{
-    Hypothesis 1: we expect the error made be correlated to the number 
-    of species created by the multiple-birth process over the total number 
-    of species created
-  }
-  \label{fig:h_1}
-\end{figure}
-
-The MBD process has multiple components, that may cause 
-different values of $e$ for identical values of $s$.
-
-One MBD component that may cause 
-different values of $e$ for identical values of $s$
-is the MBD regime: which can be many modest 
-multiple-speciation events, or few intense ones. 
-As we have no prior expectations, we have the (null) hypothesis, 
-$\mathcal{H}_7$, that the MBD regime has no effect on $e$.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_h_7.png}
-  \caption{
-    Hypothesis 7: there is no difference in inference errors if the MBD process
-    generates many species in many modest MB events or few intense MB events
-  }
-  \label{fig:h_7}
-\end{figure}
-
-Another MBD component that may cause 
-different values of $e$ for identical values of $s$
-is the effect of extinction. 
-As extinctions will hit lineages created by both speciation processes equally,
-and we have no additionaly prior expectations,
-we have the (null) hypothesis, 
-$\mathcal{H}_2$, that extinction has no effect on $e$.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_h_2.png}
-  \caption{
-    Hypothesis 2: the effect of extinction rates is neutral
-  }
-  \label{fig:h_2}
-\end{figure}
-
-Another MBD component that may cause 
-different values of $e$ for identical values of $s$,
-is the timing of a multiple-birth event: be it close to the
-crown age or close to the present. 
-Compared to a late multiple birth event, an early multiple birth event may have a 
-longer-lasting effect (as the next speciation event will be later), but it
-will create less new species, as there are still fewer taxa.
-As we have no prior expectations,
-we have the (null) hypothesis, 
-$\mathcal{H}_3$, that the timing of multiple-birth events has no effect on $e$.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_h_3.png}
-  \caption{
-    Hypothesis 3: the timing of a multiple birth event has no effect.
-  }
-  \label{fig:h_3}
-\end{figure}
-
-Another MBD component that may cause 
-different values of $e$ for identical values of $s$,
-is the number of taxa in a phylogeny,
-As there is no diversity dependency in any of the processes
-and we have no further prior expectations, we have
-the (null) hypothesis $\mathcal{H}_5$, that the number of
-taxa has no effect on $e$.
-As a higher number of taxa increases the information content in a
-phylogeny, we have hypothesis $\mathcal{H}_6$ that the
-variance in $e$ decreases.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_h_5.png}
-  \caption{
-    Hypothesis 5: the number of taxa does not
-    have an effect on the error being made
-  }
-  \label{fig:h_5}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_h_6.png}
-  \caption{
-    Hypothesis 6: for a higher number of taxa 
-    the variance in the error decreases
-  }
-  \label{fig:h_6}
-\end{figure}
-
-\end{itemize}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Methods}
@@
 
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_h_4.png}
-  \caption{
-    Hypothesis 4: compared to the MBD trees, 
-    the error will be less in the BD twin tree
-  }
-  \label{fig:h_4}
-\end{figure}
+\include{razzo-fig_h_4.tex}
 
@@
 \begin{equation}
-m_{MBD} = L \cdot \rho \cdot \int_{0}^{T} <n_{BD}>(t)\ dt \\
-= L \cdot \rho \cdot n_{0} \cdot \left[ \frac{e^{(\mu_{BD} - \lambda_{BD})T} - 1}{\mu_{BD} - \lambda_{BD}} \right]
+    m_{MBD} = 
+    L \cdot \rho \cdot \int_{0}^{T} <n_{BD}>(t)\ dt 
+    \\ =
+    L \cdot \rho \cdot n_{0} \cdot \left[ \frac{e^{(\mu_{BD} - \lambda_{BD})T} - 1}{\mu_{BD} - \lambda_{BD}} \right]
 \end{equation}
@@
 
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{mbd.jpg}
-  \caption{
-    How to create twin trees and alignments. 
-    From a focal MBD tree, a twin tree is produced as 
-    such: (1) estimate the $\lambda_{BD}$ to get 
-    the same expected number of tips, (2) simulate a BD tree 
-    with that amount of tips (discard trees with different number of tips), 
-    (3) estimate a mutation rate to get an alignment 
-    with the same expected number of mutations, (4) simulate alignments 
-    with that amount of mutations (discard those that don't, 
-    the picture shows an alignment that should be discarded) 
-  }
-\end{figure}
+\input{razzo-fig_mbd.tex}
 
@@
 \end{itemize}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
-\begin{itemize}
-
-\item We expected the effect of $s$ on $e$ to be monotonously increasing
-function (figure \ref{fig:h_1}). Figure \ref{fig:1} shows the effect 
-measured in this experiment.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_1.png}
-  \caption{
-    The effect of s on e
-  }
-  \label{fig:1}
-\end{figure}
-
-\item We expected the effect of the number of taxa on the relation
-between $s$ and $e$ to be neutral (figure \ref{fig:h_5}). 
-Figure \ref{fig:5} shows the effect 
-measured in this experiment.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_5.png}
-  \caption{
-    The effect of number of taxa on the relation between s on e
-  }
-  \label{fig:5}
-\end{figure}
-
-\item We expected the effect of the number of taxa on the variance
-of $e$ for similar values of $s$ to decrease (figure \ref{fig:h_6}). 
-Figure \ref{fig:6} shows the effect 
-measured in this experiment.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_6.png}
-  \caption{
-    The effect of number of taxa on the variance in e
-    for similar values of s
-  }
-  \label{fig:6}
-\end{figure}
-
-\item We expected the effect of extinction on the relation
-between $s$ and $e$ to be neutral (figure \ref{fig:h_2}). 
-Figure \ref{fig:2} shows the effect 
-measured in this experiment.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_2.png}
-  \caption{
-    The effect of extinction on the relation between s on e
-  }
-  \label{fig:2}
-\end{figure}
-
-\item
-
-\end{itemize}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Bibliography % MEE style
-\bibliographystyle{mee}
-\bibliography{article}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-06-07 razzo-tables/razzo-table_1.tex
Changes @Giappo

\begin{table}
  \centering 
  \begin{tabular}{r c c c}
    \hline
    Symbol                & Expectation & Expected figure & Measured figure \\
    \hline
    \hline
    $\mathcal{H}_1$       & For a higher $s$, $e$ will increase monotonously & \ref{fig:h_1} & \ref{fig:1} \\
    $\mathcal{H}_2$       & Extinction rates have no effect on $e$ & \ref{fig:h_2} & \ref{fig:2} \\
    $\mathcal{H}_3$       & The timing of multiple-birth events has no effect on $e$ & \ref{fig:h_3} & Not yet \\
    $\mathcal{H}_4$       & An MBD tree will have a higher $e$ than its twin BD counterpart & \ref{fig:h_4} & Not yet \\
    $\mathcal{H}_5$       & The number of taxa has no effect on $e$ & \ref{fig:h_5} & \ref{fig:5} \\
    $\mathcal{H}_6$       & A higher number of taxa decreases the variance in $e$ & \ref{fig:h_6} & \ref{fig:6} \\
    $\mathcal{H}_7$       & The MBD regime has no effect on $e$ & \ref{fig:h_7} & Not yet \\
    \hline
  \end{tabular}
  \caption{
    Overview of the symbols. 
  }
  \label{table:hypotheses}
\end{table}

### 2019-06-07 content.tex

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\iffalse
 \begin{abstract}
@@
 {\bf Keywords:} computational biology, evolution, phylogenetics, Bayesian analysis, tree prior
+\fi
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-\begin{itemize}
 
-\item There are many contemporary tools that provide the possibility 
-to infer a phylogeny from genetic data (DNA, RNA, proteins). 
-A popular Bayesian phylogenetic tool is called BEAST (\cite{beast}) 
-and its cousin BEAST2 (\cite{beast2}).
-
-\item BEAST is very flexible, providing the user with the option 
-to set up all possible phylogenetic priors (e.g. site/clock/speciation model).
-
-%Current limits in current tools.
-\item However, currently available priors can be not suitable 
-to analyze some specific datasets. 
-With this work we aim to test whether or not 
-the implementation of a new prior model 
-is beneficial to study a specific kind of diversification process.
-
-\item BEAST2 gives us the possibility to introduce new tree priors 
-to infer phylogenies based on different assumptions 
-on how the speciation process takes place.
-
-\item One of such speciation processes is the multiple birth hypothesis,
-a new model (described below) and thus currently absent in BEAST.
-
-\item The Multiple birth hypothesis can be useful to explain a phenomenon 
-that has always puzzled evolutionary biologists: 
-what are the drivers of the diversification processes 
-for those phylogenies that show an impressive amount of speciation events 
-in relatively short times? 
-The (constant-rate) birth-death (BD) model embodies the common assumption that 
-only a single speciation event can occur at any given time.
-The multiple-birth-death (MBD) model 
-relaxes this assumption, allowing events in which 
-large-scale environmental changes lead to a great number of species 
-in relatively short time intervals. 
-Such a hypothesis may be a better fit to describe the burst in systems 
-like cichlid fish diversification in the 
-African Great Lakes: Malawi, Tanganyika and Victoria 
-(\cite{janzen2016}, \cite{janzen2017}).
-
-\item However, it may be that current BD tree priors are good enough 
-at detecting such events, with a (preferred) lower level of complexity. 
-If this is the case one should always be more keen to adopt the simplest model.
-
-\item Here we present our study with the aim of exploring 
-when using a more complex MBD tree prior is warranted.
-To do so, we simulate phylogenies using the MBD process, with
-varying degrees of that process. To be explicit, we define
-that degree, $s$, as the number of extinct and extant species created
-during a co-occuring speciation event, 
-$N_{\mathbb{M}}$, from the total number of extinct and extant species:
-
-\begin{equation}
-s = \frac{N_{\mathbb{M}}}{N_{\mathbb{M}} + N_{\mathbb{B}}}
-\end{equation}
-
-Here, $N_{\mathbb{B}}$ is the number of extinct and extant species created
-during a default single-birth speciation event.
-
-From such phylogenies with different degrees of $s$, we measure the
-inference error we make today, would nature follow such a phylogeny.
-The inference error we make today is caused by the assumption of a BD process
-and by inherent noise in this inference.
-
-We have the hypothesis $\mathcal{H}_1$ that, for a higher $s$, 
-the inference error $e$ will increase:
-
-\begin{equation}
-e = f(s)
-\end{equation}
-
-Where $f$ is a monotonously increasing function of unknown shape.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_h_1.png}
-  \caption{
-    Hypothesis 1: we expect the error made be correlated to the number 
-    of species created by the multiple-birth process over the total number 
-    of species created
-  }
-  \label{fig:h_1}
-\end{figure}
-
-The MBD process has multiple components, that may cause 
-different values of $e$ for identical values of $s$.
-
-One MBD component that may cause 
-different values of $e$ for identical values of $s$
-is the MBD regime: which can be many modest 
-multiple-speciation events, or few intense ones. 
-As we have no prior expectations, we have the (null) hypothesis, 
-$\mathcal{H}_7$, that the MBD regime has no effect on $e$.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_h_7.png}
-  \caption{
-    Hypothesis 7: there is no difference in inference errors if the MBD process
-    generates many species in many modest MB events or few intense MB events
-  }
-  \label{fig:h_7}
-\end{figure}
-
-Another MBD component that may cause 
-different values of $e$ for identical values of $s$
-is the effect of extinction. 
-As extinctions will hit lineages created by both speciation processes equally,
-and we have no additionaly prior expectations,
-we have the (null) hypothesis, 
-$\mathcal{H}_2$, that extinction has no effect on $e$.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_h_2.png}
-  \caption{
-    Hypothesis 2: the effect of extinction rates is neutral
-  }
-  \label{fig:h_2}
-\end{figure}
-
-Another MBD component that may cause 
-different values of $e$ for identical values of $s$,
-is the timing of a multiple-birth event: be it close to the
-crown age or close to the present. 
-Compared to a late multiple birth event, an early multiple birth event may have a 
-longer-lasting effect (as the next speciation event will be later), but it
-will create less new species, as there are still fewer taxa.
-As we have no prior expectations,
-we have the (null) hypothesis, 
-$\mathcal{H}_3$, that the timing of multiple-birth events has no effect on $e$.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_h_3.png}
-  \caption{
-    Hypothesis 3: the timing of a multiple birth event has no effect.
-  }
-  \label{fig:h_3}
-\end{figure}
-
-Another MBD component that may cause 
-different values of $e$ for identical values of $s$,
-is the number of taxa in a phylogeny,
-As there is no diversity dependency in any of the processes
-and we have no further prior expectations, we have
-the (null) hypothesis $\mathcal{H}_5$, that the number of
-taxa has no effect on $e$.
-As a higher number of taxa increases the information content in a
-phylogeny, we have hypothesis $\mathcal{H}_6$ that the
-variance in $e$ decreases.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_h_5.png}
-  \caption{
-    Hypothesis 5: the number of taxa does not
-    have an effect on the error being made
-  }
-  \label{fig:h_5}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_h_6.png}
-  \caption{
-    Hypothesis 6: for a higher number of taxa 
-    the variance in the error decreases
-  }
-  \label{fig:h_6}
-\end{figure}
-
-\end{itemize}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Methods}
 
-\subsection{Model}
-\begin{itemize}
-
-\item 
-\richel{TODO: move to Introduction}
-Current phylogenetic tools assume that only a single speciation event 
-can occur at any given time.
-While this assumption is useful to construct a wide variety of successful 
-models (e.g \cite{Maddison2007biSSE}, \cite{Valente2015}, 
-\cite{etienne2012diversity}, \cite{etienne2014estimating}),
-they disallow for environmental changes that trigger speciations 
-in multiple clades at a same point in time.
-
-\item 
-\richel{TODO: move to Introduction}
-In the MBD model, parameters $\lambda$ and $\mu$ correspond, respectively, 
-to the common per-species speciation and extinction rates present 
-also in the standard BD model. 
-Additionally, MBD relies on two additional parameters. 
-Parameter $\nu$ is the rate at which an environmental change is triggered.
-When such event is triggered, 
-all species present in the phylogeny at that moment
-have a probability $q$ to speciate at that time, which is 
-independent on $\lambda$. 
-Polytomies are not allowed in such process 
-as each species can speciate only once at the time.
-
-\item 
-\richel{TODO: move to Introduction}
-It is also possible to write down a likelihood function 
-for such processes as in \cite{mbd}.
-
-\end{itemize}
-
-\subsection{Simulations}
-\begin{itemize}
-
-\item To investigate the effect of $s$ on $e$, we simulate phylogenies
-for different values of $s$ spread equally from zero (no multiple-birth
-event, thus a BD model) to one (species are created only from multiple-birth
-events) and three intermediate values of 0.25, 0.5 and 0.75. Aggregating all
-components of $s$, we show the effect of $s$ on $e$ in figure \ref{fig:1}.
-
-\item To see the effect of the number of taxa on the relation 
-between $s$ and $e$, we simulate phylogenies
-for different number of extant taxa $n$. We use values of $n$
-of 50, 100 and 200 extant taxa, as this will result in phylogenies
-that are big enough to be useful, yet small of enough to
-be computationally feasible. For the different values of $n$, we show
-the relation between $s$ and $e$ in figure \ref{fig:5} and its effect
-on the variance of $e$ in figure \ref{fig:6}.
-
-\item To see the effect of the extinction rate on the relation 
-between $s$ and $e$, we simulate phylogenies
-with different extinction rates $\mu$. We use values of $\mu$
-of 0.0, 0.1 and 0.2. An extinction rate of zero has two features:
-(1) the model falls back to a pure multiple-birth model, 
-(2) all multiple-birth events are observed.
-For the different values of $\mu$, we show
-the relation between $s$ and $e$ in figure \ref{fig:2}.
-
-\item We start simulating $N_{S} = 1000$ MBD trees,
-with either 50, 100 and 200 taxa.
-
-\item From each MBD tree, a DNA sequence alignment is simulated. 
-For each sequence alignment we then perform a Bayesian analysis 
-to recover a posterior distribution of trees, 
-each composed of $N_{P}$ phylogenies. 
-Such analysis is performed using 
-the 'pirouette' package (\cite{pirouette}) to call the BEAST2 tool 
-suite from R. 
-We let the Bayesian analysis assume a BD prior in both cases, 
-to investigate the extent of the error we make under this assumption.
-
-\item For each tree generated under the MBD model 
-we aim to generate a "twin" tree under the BD model. 
-With the word "twin" 
-we denote a tree generated starting from the respective MBD tree, 
-in order to perform a fair comparison with it. 
-This operation has to be done, 
-because we want to compare two trees 
-that are generated by different processes. 
-To do so we infer the parameters $\lambda_{BD}$ and $\mu_{BD}$ 
-from the MBD maximizing the likelihood under a BD model. 
-To perform this operation we use the function "\texttt{bd\_ML}" 
-from the package "\texttt{DDD}" (\cite{etienne2012diversity}).
-
-\item We then exploit such parameters to generate a BD tree 
-using the function "\texttt{tess.sim.taxa.age}" 
-from the package "\texttt{TESS}" (\cite{Hoehna2013}). 
-We simulate the tree in such a way the new tree 
-has the same number of tips and the same crown age as the MBD tree. 
-We furthermore require that the BD tree conserve the topology of the MBD tree.
-We have hypothesis H4 that, compared to the MBD trees, 
-the error will be less in the BD twin tree.
-The difference between the errors made in MBD and twin BD trees indicates
-the impact the MBD process has on the error we make in inference using a
-contemporary BD prior.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_h_4.png}
-  \caption{
-    Hypothesis 4: compared to the MBD trees, 
-    the error will be less in the BD twin tree
-  }
-  \label{fig:h_4}
-\end{figure}
-
-We want the MBD and twin BD trees to contain the same amount of information, 
-i.e. the same number of DNA mutations and the same number of taxa at the present:
-
-\begin{equation}
-m_{MBD} = m_{BD} \label{m equivalence}
-\end{equation}
-
-The expected number of mutations $m$ of a phylogeny 
-with crown age $-T$ (with $T>0$) in fact is given by
-\richel{
-  So one of use likes '-T', the other likes 'T'. How to resolve this?
-}
-
-\begin{equation}
-m = L \cdot \rho \cdot \int_{0}^{T} n(t)\ dt \label{m calculation}
-\end{equation}
-
-where $L$ is the number of DNA nucleotides, 
-$\rho$ is the per-site per-species mutation rate and
-$n(t)$ the number of species at each time.
-
-The parameter we'll tune is $\rho$ ... \richel{elaborate here :-)}
-
-Since we cannot know $n_{BD}(t)$ before running simulations
-we need to replace it with a proxy. 
-For this reason we will use the average number of
-species in time according to the BD model. 
-It's well known that this is equal to \gio{insert proper citation}
-
-\begin{equation}
-    <n_{BD}>(t) = n_{0} \cdot e^{(\mu_{BD} - \lambda_{BD})t} \label{BD average n}
-\end{equation}
-
-where $n_{0} = n_{BD}(-T) = n_{MBD}(-T)$ is the initial number of species 
-at the crown age.
-From \ref{m equivalence}, \ref{m calculation} and \ref{BD average n} follows:
-
-\begin{equation}
-m_{MBD} = L \cdot \rho \cdot \int_{0}^{T} <n_{BD}>(t)\ dt \\
-= L \cdot \rho \cdot n_{0} \cdot \left[ \frac{e^{(\mu_{BD} - \lambda_{BD})T} - 1}{\mu_{BD} - \lambda_{BD}} \right]
-\end{equation}
-
-If we set $\mu_{BD} = \mu_{MBD}$ and reverse this relation 
-we can extrapolate the value of $\lambda_{BD}$ to use to generate BD trees.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{mbd.jpg}
-  \caption{
-    How to create twin trees and alignments. 
-    From a focal MBD tree, a twin tree is produced as 
-    such: (1) estimate the $\lambda_{BD}$ to get 
-    the same expected number of tips, (2) simulate a BD tree 
-    with that amount of tips (discard trees with different number of tips), 
-    (3) estimate a mutation rate to get an alignment 
-    with the same expected number of mutations, (4) simulate alignments 
-    with that amount of mutations (discard those that don't, 
-    the picture shows an alignment that should be discarded) 
-  }
-\end{figure}
-
-\item We explained how we set the parameters for each twin BD tree. 
-Using this rules we generate a BD dataset. 
-We repeat the analysis, producing alignments for each tree 
-and subsequently using BEAST to produce a posterior for each of them.
-
-\subsection{Measuring the inference error}
-
-\item So far we have simulated two datasets of trees under the two models: 
-$\{T_{i}^{BD}\}_{i=1}^{N_{S}}$ and $\{T_{i}^{MBD}\}_{i=1}^{N_{S}}$.
-We used them to generate a dataset of alignments for each model: $\{X^{BD}_{i}\}_{i=1}^{N_{S}}$ and $\{X^{MBD}_{i}\}_{i=1}^{N_{S}}$.
-From each dataset we produced a posterior distribution from a BD prior:
-$P_{i}(\theta | X^{BD}_{i}, BD)$ and $P_{i}(\theta | X^{MBD}_{i}, BD)$.
-\gio{
-  1) We might want to rename the models, e.g. BD = (0) and MBD = (1). 
-  These names with capital letters are too big and ugly;
-}
-\richel{
-  I would strongly prefer MBD and BD, as I feel replacing the big ugly 
-  capital letters by short pretty numbers hurts readability even more 
-}
-
-\item To compare the results for the two models we measure the inference 
-error using the nLTT statistic between known/true tree and 
-posterior/inferred trees (\cite{nltt}). 
-To obtain such statistics the procedure is the following:
-
-- From each tree $T_{i,j}^{M}$ (with $j=1,...,N_{S}$) 
-  belonging to the posterior $P_{i}(\theta | X^{M}_{i}, BD)$ 
-  and relative to the model $M$, we extrapolate the lineage-through-time (LTT), 
-  in other words we measure the number of species as a function of 
-  time $n_{i,j}(t)$. To allow a comparison we normalize dividing by the 
-  maximum number of species of each tree, i.e. the number of tips at the 
-  present $N_{i,j}(t)=\frac{n_{i,j}(t)}{n^{max}_{i,j}}$. We then define the 
-  nLTT measure as
-
-$nLTT_{i,j} = \int_{0}^{T} | N_{i,j}(t) - N_{T_{i}} | dt$
-
-\gio{I am running out of letters :(}
-\richel{Haha! I suggest to use the same equation and symbols 
-  as equation 1 in
-  the nLTT article of Janzen, Hoehna and Etienne, 2015:
-}
-
-$$
-\Delta nLTT = \int_{0}^{1} | nLTT_1(t) - nLTT_2{t} | dt
-$$
-
-\subsection{Model selection}
-
-We simulate alignments using the simplest nucleotide substitution model (JC69),
-the simplest clock model (strict). It is thus imminent to assume these
-models in our Bayesian inference. Nevertheless, the phylogeny the alignment
-was based on, could have followed either an MBD or BD tree model, 
-where we in both cases assume a BD tree model. This will have 
-an unknown effect on our inference: it may theoretically be that an MBD model
-generates (a tree that generates) an alignment in which a different site 
-and/or clock model is favored.
-
-We investigate this by measuring if the generative model (with the simplest
-nucleotide substitution and simplest clock model) is indeed selected 
-to be the best fitting model. 
-To be precise, we look at the model 
-with the highest marginal likelihood 
-(also called evidence \cite{mackay2003information}),
-$f(D|M)$, which is the probability of the data D given model M.
-In the context of this research, D consists of the DNA alignment,
-and M is the combination of site, clock and tree models.
-
-To estimate the marginal likelihood, 
-we use an algorithm named nested sampling \cite{skilling2006nested}.
-Nested sampling is attractive to use
-in a phylogentic context, as it gives a good estimation,
-requires little tuning \cite{maturana2018}.
-Nested sampling is available as a BEAST2 package
-and can be used by babette \cite{babette}.
-
-The nested sampling algorithm stops its run 
-when the marginal likelihood estimation error 
-reaches below a certain tolerance.  
-Similar to \cite{maturana2018},
-we use a (relative) error tolerance $\epsilon$ of $10^{-13}$,
-1 particle to explore the parameter space
-and 100 active points. 
-To achieve the latter, we use the MCMC chain length $L_c$ of 1M 
-(as also used in the parameter estimates),
-and a sub-chain length $L_{sc}$ of 10K.
-
-The models we use in our model comparison are the four combinations
-of two site models and two clock models. We use the JC69 site model, which
-is the (generative and) simplest model and GTR, the site model with most
-degrees of freedom. For the clock models, we use the strict clock model,
-which is the (generative and) simplest clock model, and the RLN clock model.
-\richel{Could also just be all site models and clock models = 8 models}
-
-From these four marginal likelihood estimates, we calculate the weight of
-the generative model and plot this in figure 2. We do this for both the 
-alignments derived from the MBD tree and the BD twin tree. We expect that
-the generative model has the heighest weight in both the MBD and BD alignments.
-We expect this weight to be higher in the BD alignments.
-
-\end{itemize}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
-\begin{itemize}
 
-\item We expected the effect of $s$ on $e$ to be monotonously increasing
-function (figure \ref{fig:h_1}). Figure \ref{fig:1} shows the effect 
-measured in this experiment.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_1.png}
-  \caption{
-    The effect of s on e
-  }
-  \label{fig:1}
-\end{figure}
-
-\item We expected the effect of the number of taxa on the relation
-between $s$ and $e$ to be neutral (figure \ref{fig:h_5}). 
-Figure \ref{fig:5} shows the effect 
-measured in this experiment.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_5.png}
-  \caption{
-    The effect of number of taxa on the relation between s on e
-  }
-  \label{fig:5}
-\end{figure}
-
-\item We expected the effect of the number of taxa on the variance
-of $e$ for similar values of $s$ to decrease (figure \ref{fig:h_6}). 
-Figure \ref{fig:6} shows the effect 
-measured in this experiment.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_6.png}
-  \caption{
-    The effect of number of taxa on the variance in e
-    for similar values of s
-  }
-  \label{fig:6}
-\end{figure}
-
-\item We expected the effect of extinction on the relation
-between $s$ and $e$ to be neutral (figure \ref{fig:h_2}). 
-Figure \ref{fig:2} shows the effect 
-measured in this experiment.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fig_2.png}
-  \caption{
-    The effect of extinction on the relation between s on e
-  }
-  \label{fig:2}
-\end{figure}
-
-\item
-
-\end{itemize}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % Bibliography % MEE style
-\bibliographystyle{mee}
-\bibliography{article}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-06-07 razzo-fig_mbd.tex

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{razzo-figures/fig_mbd.jpg}
  \caption{
    How to create twin trees and alignments. 
    From a focal MBD tree, a twin tree is produced as 
    such: (1) estimate the $\lambda_{BD}$ to get 
    the same expected number of tips, (2) simulate a BD tree 
    with that amount of tips (discard trees with different number of tips), 
    (3) estimate a mutation rate to get an alignment 
    with the same expected number of mutations, (4) simulate alignments 
    with that amount of mutations (discard those that don't, 
    the picture shows an alignment that should be discarded) 
  }
\end{figure}

### 2019-06-07 razzo-results.tex

\section{Results}
\begin{itemize}

\item We expected the effect of $s$ on $e$ to be monotonously increasing
function (figure \ref{fig:h_1}). Figure \ref{fig:1} shows the effect 
measured in this experiment.

\input{razzo-fig_1.tex}

\item We expected the effect of the number of taxa on the relation
between $s$ and $e$ to be neutral (figure \ref{fig:h_5}). 
Figure \ref{fig:5} shows the effect 
measured in this experiment.

\input{razzo-fig_5.tex}

\item We expected the effect of the number of taxa on the variance
of $e$ for similar values of $s$ to decrease (figure \ref{fig:h_6}). 
Figure \ref{fig:6} shows the effect 
measured in this experiment.

\input{razzo-fig_6.tex}

\item We expected the effect of extinction on the relation
between $s$ and $e$ to be neutral (figure \ref{fig:h_2}). 
Figure \ref{fig:2} shows the effect 
measured in this experiment.

\input{razzo-fig_2.tex}

\item

\end{itemize}

### 2019-06-19 razzo-methods.tex
Updates from Overleaf

@@
 \subsection{Model}
-\begin{itemize}
 
-\item 
-\richel{TODO: move to Introduction}
-Current phylogenetic tools assume that only a single speciation event 
-can occur at any given time.
-While this assumption is useful to construct a wide variety of successful 
-models (e.g \cite{Maddison2007biSSE}, \cite{Valente2015}, 
-\cite{etienne2012diversity}, \cite{etienne2014estimating}),
-they disallow for environmental changes that trigger speciations 
-in multiple clades at a same point in time.
-
-\item 
-\richel{TODO: move to Introduction}
 In the MBD model, parameters $\lambda$ and $\mu$ correspond, respectively, 
@@
 Parameter $\nu$ is the rate at which an environmental change is triggered.
-When such event is triggered, 
-all species present in the phylogeny at that moment
-have a probability $q$ to speciate at that time, which is 
-independent on $\lambda$. 
-Polytomies are not allowed in such process 
-as each species can speciate only once at the time.
+When such event is triggered, each species present in the phylogeny at that moment has a probability $q$ to speciate at that time.
+This kind of speciation is of a different nature respect to the one triggered by $\lambda$.
+In fact, whereas parameter $\lambda$ can be seen as describing a sympatric process, $\nu$ induces the rise of geographical barriers interrupting the gene flow and leading to an allopatric speciation.
+Even though multiple speciations can co-occur, polytomies are not allowed in such process as each species can speciate only once at the time.
+A likelihood expression for the process is provided in \cite{mbd}.
 
-\item 
-\richel{TODO: move to Introduction}
-It is also possible to write down a likelihood function 
-for such processes as in \cite{mbd}.
+\subsection{Tree simulations}
 
-\end{itemize}
+We can easily simulate such processes in continuous time using the Doob-Gillespie algorithm. Simulations are performed using the function "mbd\_sim" from the \verb;mbd; package (\citep{mbd}).
+We let parameters vary using all possible combinations of value included in Table~\ref{tab:simulation_parameters} to simulate $1000$ independent trees for each parameter setting.
 
-\subsection{Simulations}
-\begin{itemize}
+\input{razzo-tables/tab_simparameters.tex}
 
-\item To investigate the effect of $s$ on $e$, we simulate phylogenies
-for different values of $s$ spread equally from zero (no multiple-birth
-event, thus a BD model) to one (species are created only from multiple-birth
-events) and three intermediate values of 0.25, 0.5 and 0.75. Aggregating all
-components of $s$, we show the effect of $s$ on $e$ in figure \ref{fig:1}.
+\subsection{Pirouette}
 
-\item To see the effect of the number of taxa on the relation 
-between $s$ and $e$, we simulate phylogenies
-for different number of extant taxa $n$. We use values of $n$
-of 50, 100 and 200 extant taxa, as this will result in phylogenies
-that are big enough to be useful, yet small of enough to
-be computationally feasible. For the different values of $n$, we show
-the relation between $s$ and $e$ in figure \ref{fig:5} and its effect
-on the variance of $e$ in figure \ref{fig:6}.
+Once the MBD dataset has been simulated we exploit the \verb;pirouette; package (\citep{pirouette}) to assess the error made by BEAST2, executing  the inference using a BD tree prior.
 
-\item To see the effect of the extinction rate on the relation 
-between $s$ and $e$, we simulate phylogenies
-with different extinction rates $\mu$. We use values of $\mu$
-of 0.0, 0.1 and 0.2. An extinction rate of zero has two features:
-(1) the model falls back to a pure multiple-birth model, 
-(2) all multiple-birth events are observed.
-For the different values of $\mu$, we show
-the relation between $s$ and $e$ in figure \ref{fig:2}.
+We briefly summarize here how the \verb;pirouette; routine works.
+From each MBD tree a DNA sequence alignment is simulated. 
+For each sequence alignment we then perform a Bayesian inference analysis (using BEAST2) to recover a posterior distribution of trees. 
+For each parameter setting, this process leads to an inference error distribution.
+To evaluate the extent of this error we also run a \verb;pirouette; "twin" pipeline, through which we mimic the original pipeline in every aspect but starting instead from "twin" trees.
+From each MBD simulated tree we can produce its BD twin, inferring the most likely BD parameters (i.e.
+$\lambda$ and $\mu$) through maximum likelihood, and use them to simulate a standard BD tree (see \cite{pirouette} for more information), keeping the topology of the original tree.
+We let the Bayesian analysis assume a BD prior in both cases, to investigate the extent of the error we make under this assumption.
+The twin pipeline serves as an estimation of the baseline error, as its error it is not due by the mismatch of the generative prior with the inference prior.
 
-\item We start simulating $N_{S} = 1000$ MBD trees,
-with either 50, 100 and 200 taxa.
-
-\item From each MBD tree, a DNA sequence alignment is simulated. 
-For each sequence alignment we then perform a Bayesian analysis 
-to recover a posterior distribution of trees, 
-each composed of $N_{P}$ phylogenies. 
-Such analysis is performed using 
-the 'pirouette' package (\cite{pirouette}) to call the BEAST2 tool 
-suite from R. 
-We let the Bayesian analysis assume a BD prior in both cases, 
-to investigate the extent of the error we make under this assumption.
-
-\item For each tree generated under the MBD model 
-we aim to generate a "twin" tree under the BD model. 
-With the word "twin" 
-we denote a tree generated starting from the respective MBD tree, 
-in order to perform a fair comparison with it. 
-This operation has to be done, 
-because we want to compare two trees 
-that are generated by different processes. 
-To do so we infer the parameters $\lambda_{BD}$ and $\mu_{BD}$ 
-from the MBD maximizing the likelihood under a BD model. 
-To perform this operation we use the function "\texttt{bd\_ML}" 
-from the package "\texttt{DDD}" (\cite{etienne2012diversity}).
-
-\item We then exploit such parameters to generate a BD tree 
-using the function "\texttt{tess.sim.taxa.age}" 
-from the package "\texttt{TESS}" (\cite{Hoehna2013}). 
-We simulate the tree in such a way the new tree 
-has the same number of tips and the same crown age as the MBD tree. 
-We furthermore require that the BD tree conserve the topology of the MBD tree.
-We have hypothesis H4 that, compared to the MBD trees, 
-the error will be less in the BD twin tree.
-The difference between the errors made in MBD and twin BD trees indicates
-the impact the MBD process has on the error we make in inference using a
-contemporary BD prior.
-
-\include{razzo-fig_h_4.tex}
-
-We want the MBD and twin BD trees to contain the same amount of information, 
-i.e. the same number of DNA mutations and the same number of taxa at the present:
-
-\begin{equation}
-m_{MBD} = m_{BD} \label{m equivalence}
-\end{equation}
-
-The expected number of mutations $m$ of a phylogeny 
-with crown age $-T$ (with $T>0$) in fact is given by
-\richel{
-  So one of use likes '-T', the other likes 'T'. How to resolve this?
-}
-
-\begin{equation}
-m = L \cdot \rho \cdot \int_{0}^{T} n(t)\ dt \label{m calculation}
-\end{equation}
-
-where $L$ is the number of DNA nucleotides, 
-$\rho$ is the per-site per-species mutation rate and
-$n(t)$ the number of species at each time.
-
-The parameter we'll tune is $\rho$ ... \richel{elaborate here :-)}
-
-Since we cannot know $n_{BD}(t)$ before running simulations
-we need to replace it with a proxy. 
-For this reason we will use the average number of
-species in time according to the BD model. 
-It's well known that this is equal to \gio{insert proper citation}
-
-\begin{equation}
-    <n_{BD}>(t) = n_{0} \cdot e^{(\mu_{BD} - \lambda_{BD})t} \label{BD average n}
-\end{equation}
-
-where $n_{0} = n_{BD}(-T) = n_{MBD}(-T)$ is the initial number of species 
-at the crown age.
-From \ref{m equivalence}, \ref{m calculation} and \ref{BD average n} follows:
-
-\begin{equation}
-    m_{MBD} = 
-    L \cdot \rho \cdot \int_{0}^{T} <n_{BD}>(t)\ dt 
-    \\ =
-    L \cdot \rho \cdot n_{0} \cdot \left[ \frac{e^{(\mu_{BD} - \lambda_{BD})T} - 1}{\mu_{BD} - \lambda_{BD}} \right]
-\end{equation}
-
-If we set $\mu_{BD} = \mu_{MBD}$ and reverse this relation 
-we can extrapolate the value of $\lambda_{BD}$ to use to generate BD trees.
-
-\input{razzo-fig_mbd.tex}
-
-\item We explained how we set the parameters for each twin BD tree. 
-Using this rules we generate a BD dataset. 
-We repeat the analysis, producing alignments for each tree 
-and subsequently using BEAST to produce a posterior for each of them.
-
-\subsection{Measuring the inference error}
-
-\item So far we have simulated two datasets of trees under the two models: 
-$\{T_{i}^{BD}\}_{i=1}^{N_{S}}$ and $\{T_{i}^{MBD}\}_{i=1}^{N_{S}}$.
-We used them to generate a dataset of alignments for each model: $\{X^{BD}_{i}\}_{i=1}^{N_{S}}$ and $\{X^{MBD}_{i}\}_{i=1}^{N_{S}}$.
-From each dataset we produced a posterior distribution from a BD prior:
-$P_{i}(\theta | X^{BD}_{i}, BD)$ and $P_{i}(\theta | X^{MBD}_{i}, BD)$.
-\gio{
-  1) We might want to rename the models, e.g. BD = (0) and MBD = (1). 
-  These names with capital letters are too big and ugly;
-}
-\richel{
-  I would strongly prefer MBD and BD, as I feel replacing the big ugly 
-  capital letters by short pretty numbers hurts readability even more 
-}
-
-\item To compare the results for the two models we measure the inference 
-error using the nLTT statistic between known/true tree and 
-posterior/inferred trees (\cite{nltt}). 
-To obtain such statistics the procedure is the following:
-
-- From each tree $T_{i,j}^{M}$ (with $j=1,...,N_{S}$) 
-  belonging to the posterior $P_{i}(\theta | X^{M}_{i}, BD)$ 
-  and relative to the model $M$, we extrapolate the lineage-through-time (LTT), 
-  in other words we measure the number of species as a function of 
-  time $n_{i,j}(t)$. To allow a comparison we normalize dividing by the 
-  maximum number of species of each tree, i.e. the number of tips at the 
-  present $N_{i,j}(t)=\frac{n_{i,j}(t)}{n^{max}_{i,j}}$. We then define the 
-  nLTT measure as
-
-$nLTT_{i,j} = \int_{0}^{T} | N_{i,j}(t) - N_{T_{i}} | dt$
-
-\gio{I am running out of letters :(}
-\richel{Haha! I suggest to use the same equation and symbols 
-  as equation 1 in
-  the nLTT article of Janzen, Hoehna and Etienne, 2015:
-}
-
-$$
-\Delta nLTT = \int_{0}^{1} | nLTT_1(t) - nLTT_2{t} | dt
-$$
-
+\iffalse
 \subsection{Model selection}
@@
 We expect this weight to be higher in the BD alignments.
-
-\end{itemize}
+\fi

### 2019-06-19 razzo-introduction.tex

@@
 \section{Introduction}
-\begin{itemize}
 
-\item There are many contemporary tools that provide the possibility 
-to infer a phylogeny from genetic data (DNA, RNA, proteins). 
-A popular Bayesian phylogenetic tool is called BEAST (\cite{beast}) 
-and its cousin BEAST2 (\cite{beast2}).
+Modern computational techniques allow to infer phylogenies from genetic data such as DNA, RNA or proteins.
+BEAST (\cite{beast}) and its descendant BEAST2 (\cite{beast2}) are widely used tools to perform this task, which they can achieve by running a Bayesian analysis given data and tree priors.
 
-\item BEAST is very flexible, providing the user with the option 
-to set up all possible phylogenetic priors (e.g. site/clock/speciation model).
+BEAST2 gives to the user the option to set up several possible phylogenetic priors (e.g.
+substitution/clock/speciation models).
+However, currently available priors can be not suitable to analyze some specific datasets.
 
-%Current limits in current tools.
-\item However, currently available priors can be not suitable 
-to analyze some specific datasets. 
-With this work we aim to test whether or not 
-the implementation of a new prior model 
-is beneficial to study a specific kind of diversification process.
+For this reason BEAST2 provides users with the possibility to introduce new tree priors, to infer phylogenies based on different assumptions on how the speciation process takes place.
 
-\item BEAST2 gives us the possibility to introduce new tree priors 
-to infer phylogenies based on different assumptions 
-on how the speciation process takes place.
+Current phylogenetic tools assume that only a single speciation event can occur at any given time.
+While this assumption has been proved to be useful to construct a wide variety of successful models (e.g \cite{Maddison2007biSSE}, \cite{Valente2015}, \cite{etienne2012diversity}, \cite{etienne2014estimating}), they disallow for environmental changes that trigger speciations in multiple clades at a same point in time.
 
-\item One of such speciation processes is the multiple birth hypothesis,
-a new model (described below) and thus currently absent in BEAST.
+We explore such case introducing the multiple birth model, currently absent in BEAST.
+The multiple birth hypothesis aims to include species pump mechanisms to investigate drivers and modes of such diversification processes whose phylogenies show an impressive amount of speciation events in relatively short times.
 
-\item The Multiple birth hypothesis can be useful to explain a phenomenon 
-that has always puzzled evolutionary biologists: 
-what are the drivers of the diversification processes 
-for those phylogenies that show an impressive amount of speciation events 
-in relatively short times? 
-The (constant-rate) birth-death (BD) model embodies the common assumption that 
-only a single speciation event can occur at any given time.
-The multiple-birth-death (MBD) model 
-relaxes this assumption, allowing events in which 
-large-scale environmental changes lead to a great number of species 
-in relatively short time intervals. 
-Such a hypothesis may be a better fit to describe the burst in systems 
-like cichlid fish diversification in the 
-African Great Lakes: Malawi, Tanganyika and Victoria 
-(\cite{janzen2016}, \cite{janzen2017}).
+The (constant-rate) birth-death (BD) model embodies the common assumption that only a single speciation event can occur at any given time.
+The multiple-birth-death (MBD) model relaxes this assumption allowing, in addition to standard BD events, also events in which large-scale environmental changes lead to speciation bursts.
+Such hypothesis can be useful to better understand the history of systems of particular interests for evolutionary biologists, such as the  diversification of cichlid fish in the African Great Lakes (Malawi, Tanganyika and Victoria), where water level fluctuations are thought to play an important role in promoting diversification (\citep{verheyen1996mitochondrial}, \citep{sturmbauer2001lake}, \citep{janzen2016}, \citep{janzen2017}).
 
-\item However, it may be that current BD tree priors are good enough 
-at detecting such events, with a (preferred) lower level of complexity. 
+However, the introduction of new tree priors is not always desirable (\citep{bilderbeek2019pirouette}).
+Current BD tree priors might, in principle, prove to be good enough at detecting such events despite the lower level of complexity.
 If this is the case one should always be more keen to adopt the simplest model.
 
-\item Here we present our study with the aim of exploring 
-when using a more complex MBD tree prior is warranted.
-To do so, we simulate phylogenies using the MBD process, with
-varying degrees of that process. To be explicit, we define
-that degree, $s$, as the number of extinct and extant species created
-during a co-occuring speciation event, 
-$N_{\mathbb{M}}$, from the total number of extinct and extant species:
+We used the R package \verb;pirouette; (\citep{pirouette}) to perform such test, starting on phylogenies simulated under the MBD regime using the \verb;mbd; package (\citep{mbd}).
+From such phylogenies we measure the inference error made adopting a standard BD tree prior in the inference process.
 
-\begin{equation}
-s = \frac{N_{\mathbb{M}}}{N_{\mathbb{M}} + N_{\mathbb{B}}}
-\end{equation}
-
-Here, $N_{\mathbb{B}}$ is the number of extinct and extant species created
-during a default single-birth speciation event.
-
-From such phylogenies with different degrees of $s$, we measure the
-inference error we make today, would nature follow such a phylogeny.
-The inference error we make today is caused by the assumption of a BD process
-and by inherent noise in this inference.
-
-We have the hypothesis $\mathcal{H}_1$ that, for a higher $s$, 
-the inference error $e$ will increase:
-
-\begin{equation}
-e = f(s)
-\end{equation}
-
-Where $f$ is a monotonously increasing function of unknown shape.
-
-\input{razzo-fig_h_1.tex}
-
-The MBD process has multiple components, that may cause 
-different values of $e$ for identical values of $s$.
-
-One MBD component that may cause 
-different values of $e$ for identical values of $s$
-is the MBD regime: which can be many modest 
-multiple-speciation events, or few intense ones. 
-As we have no prior expectations, we have the (null) hypothesis, 
-$\mathcal{H}_0$, that the MBD regime has no effect on $e$.
-
-\include{razzo-fig_h_0.tex}
-
-Another MBD component that may cause 
-different values of $e$ for identical values of $s$
-is the effect of extinction. 
-As extinctions will hit lineages created by both speciation processes equally,
-and we have no additionaly prior expectations,
-we have the (null) hypothesis, 
-$\mathcal{H}_2$, that extinction has no effect on $e$.
-
-\include{razzo-fig_h_2.tex}
-
-Another MBD component that may cause 
-different values of $e$ for identical values of $s$,
-is the timing of a multiple-birth event: be it close to the
-crown age or close to the present. 
-Compared to a late multiple birth event, an early multiple birth event may have a 
-longer-lasting effect (as the next speciation event will be later), but it
-will create less new species, as there are still fewer taxa.
-As we have no prior expectations,
-we have the (null) hypothesis, 
-$\mathcal{H}_3$, that the timing of multiple-birth events has no effect on $e$.
-
-\include{razzo-fig_h_3.tex}
-
-Another MBD component that may cause 
-different values of $e$ for identical values of $s$,
-is the number of taxa in a phylogeny,
-As there is no diversity dependency in any of the processes
-and we have no further prior expectations, we have
-the (null) hypothesis $\mathcal{H}_5$, that the number of
-taxa has no effect on $e$.
-As a higher number of taxa increases the information content in a
-phylogeny, we have hypothesis $\mathcal{H}_6$ that the
-variance in $e$ decreases.
-
-\include{razzo-fig_h_5.tex}
-
-\include{razzo-fig_h_6.tex}
-
-\end{itemize}
+With this work we aim, using such inference error distributions, to test whether or not it is advantageous to implement a new prior model that can allow the construction of trees where multiple speciations can co-occur at the same time.

### 2019-06-19 razzo-abstract.tex

--- 2019-06-07 content.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\iffalse
 \begin{abstract}
-
 % From 'How to construct a Nature summary paragraph'
@@
 Here we show the error we make in our inference,
-  when nature has varying degrees of co-occuring speciation.
+when nature has varying degrees of co-occurring speciation.
 
@@
 % by their inclusion. Under these circumstances, the length of the paragraph can be up to 300 words.
-  % (The above example is 190 words without the final section, and 250 words with it).
-
 \end{abstract}
-
 {\bf Keywords:} computational biology, evolution, phylogenetics, Bayesian analysis, tree prior
-\fi
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Bibliography % MEE style
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-06-19 razzo-results.tex

@@
 \section{Results}
-\begin{itemize}
+Here we present the results as show in Fig.~\ref{fig:results}...
 
-\item We expected the effect of $s$ on $e$ to be monotonously increasing
-function (figure \ref{fig:h_1}). Figure \ref{fig:1} shows the effect 
-measured in this experiment.
-
-\input{razzo-fig_1.tex}
-
-\item We expected the effect of the number of taxa on the relation
-between $s$ and $e$ to be neutral (figure \ref{fig:h_5}). 
-Figure \ref{fig:5} shows the effect 
-measured in this experiment.
-
-\input{razzo-fig_5.tex}
-
-\item We expected the effect of the number of taxa on the variance
-of $e$ for similar values of $s$ to decrease (figure \ref{fig:h_6}). 
-Figure \ref{fig:6} shows the effect 
-measured in this experiment.
-
-\input{razzo-fig_6.tex}
-
-\item We expected the effect of extinction on the relation
-between $s$ and $e$ to be neutral (figure \ref{fig:h_2}). 
-Figure \ref{fig:2} shows the effect 
-measured in this experiment.
-
-\input{razzo-fig_2.tex}
-
-\item
-
-\end{itemize}
+\input{razzo-figures/fig_results.tex}

### 2019-06-19 razzo-abstract.tex
Revised abstract

@@
 of years ago.
-The field of phylogenetics uses heritable material
-to determine which species are closest related and what are
-the mathematics that shape speciation.
+The field of phylogenetics uses heritable material (e.g. DNA)
+to determine the evolutionary history of species.
 
@@
 % scientists in related disciplines.
-In Bayesian phylogenetics, a DNA/RNA/protein alignment
-is used to infer a distribution of phylogenies and parameter estimates.
-To do so, we use assumptions that may be biologically unrealistic,
-but may give tolerable errors.
+Starting from heritable material and explicit assumptions,
+Bayesian phylogenetics allows to infer a jointly-estimated 
+phylogeny and parameter estimates distribution.
+One of these assumptions in the speciation model, which
+mathematically describes the branching process of a phylogeny in time.
+The most used speciation model assumes that speciation events are
+independent, where we know that certain events can trigger speciation 
+events in multiple species.
 
@@
 % study.
-Contemporary inference assumes that speciation
-never co-occurs.
+This research answers the question what the impact is of using a
+species tree model that assumes speciation is independent, when it
+is used on phylogenies created by a tree model in which speciation can
+co-occur.
 
@@
 % their equivalent).
-Here we show the error we make in our inference,
-when nature has varying degrees of co-occurring speciation.
+Here we show the inference error made,
+when nature has varying degrees of co-occurring speciation
+over a wide range of parameter settings.
 
@@
 % previous knowledge.
+We show that the inference error correlates
+with the amount of co-occuring speciation events,
+which valudates
 
@@
 % more general context.
+These results allow phylogeneticist to judge under which
+circumstances the commonly used speciation model can be safely 
+used.
 
@@
 % by their inclusion. Under these circumstances, the length of the paragraph can be up to 300 words.
+In a bigger picture, these results showcase the use of a general and 
+flexible method we used to assess the impact of using an oversimplistic tree 
+prior, helping phylogeneticists to find the line between 'too simple'
+and 'too complex' speciation models.
+
 \end{abstract}
-{\bf Keywords:} computational biology, evolution, phylogenetics, Bayesian analysis, tree prior
+
+{\bf Keywords:} computational biology, evolution, phylogenetics, Bayesian analysis, tree prior,
+pirouette, BEAST2, babette

### 2019-06-19 razzo-methods.tex
Folded TeX

@@
 Parameter $\nu$ is the rate at which an environmental change is triggered.
-When such event is triggered, each species present in the phylogeny at that moment has a probability $q$ to speciate at that time.
-This kind of speciation is of a different nature respect to the one triggered by $\lambda$.
-In fact, whereas parameter $\lambda$ can be seen as describing a sympatric process, $\nu$ induces the rise of geographical barriers interrupting the gene flow and leading to an allopatric speciation.
-Even though multiple speciations can co-occur, polytomies are not allowed in such process as each species can speciate only once at the time.
+When such event is triggered, each species present in the phylogeny at that mome
+nt has a probability $q$ to speciate at that time.
+This kind of speciation is of a different nature respect to the one triggered by
+ $\lambda$. In fact, whereas parameter $\lambda$ can be seen as describing a sym
+patric process, $\nu$ induces the rise of geographical barriers interrupting the
+ gene flow and leading to an allopatric speciation.
+Even though multiple speciations can co-occur, polytomies are not allowed in suc
+h process as each species can speciate only once at the time.
 A likelihood expression for the process is provided in \cite{mbd}.
@@
 
-We can easily simulate such processes in continuous time using the Doob-Gillespie algorithm. Simulations are performed using the function "mbd\_sim" from the \verb;mbd; package (\citep{mbd}).
-We let parameters vary using all possible combinations of value included in Table~\ref{tab:simulation_parameters} to simulate $1000$ independent trees for each parameter setting.
+We can easily simulate such processes in continuous time using the Doob-Gillespi
+e algorithm. Simulations are performed using the function "mbd\_sim" from the \v
+erb;mbd; package (\citep{mbd}).
+We let parameters vary using all possible combinations of value included in Tabl
+e~\ref{tab:simulation_parameters} to simulate $1000$ independent trees for each 
+parameter setting.
 
@@
 
-Once the MBD dataset has been simulated we exploit the \verb;pirouette; package (\citep{pirouette}) to assess the error made by BEAST2, executing  the inference using a BD tree prior.
+Once the MBD dataset has been simulated we exploit the \verb;pirouette; package 
+(\citep{pirouette}) to assess the error made by BEAST2, executing  the inference
+ using a BD tree prior.
 
@@
 From each MBD tree a DNA sequence alignment is simulated. 
-For each sequence alignment we then perform a Bayesian inference analysis (using BEAST2) to recover a posterior distribution of trees. 
-For each parameter setting, this process leads to an inference error distribution.
-To evaluate the extent of this error we also run a \verb;pirouette; "twin" pipeline, through which we mimic the original pipeline in every aspect but starting instead from "twin" trees.
-From each MBD simulated tree we can produce its BD twin, inferring the most likely BD parameters (i.e.
-$\lambda$ and $\mu$) through maximum likelihood, and use them to simulate a standard BD tree (see \cite{pirouette} for more information), keeping the topology of the original tree.
-We let the Bayesian analysis assume a BD prior in both cases, to investigate the extent of the error we make under this assumption.
-The twin pipeline serves as an estimation of the baseline error, as its error it is not due by the mismatch of the generative prior with the inference prior.
+For each sequence alignment we then perform a Bayesian inference analysis (using
+ BEAST2) to recover a posterior distribution of trees. 
+For each parameter setting, this process leads to an inference error distributio
+n. To evaluate the extent of this error we also run a \verb;pirouette; "twin" pi
+peline, through which we mimic the original pipeline in every aspect but startin
+g instead from "twin" trees.
+From each MBD simulated tree we can produce its BD twin, inferring the most like
+ly BD parameters (i.e. $\lambda$ and $\mu$) through maximum likelihood, and use 
+them to simulate a standard BD tree (see \cite{pirouette} for more information),
+ keeping the topology of the original tree.
+We let the Bayesian analysis assume a BD prior in both cases, to investigate the
+ extent of the error we make under this assumption. The twin pipeline serves as 
+an estimation of the baseline error, as its error it is not due by the mismatch 
+of the generative prior with the inference prior.
 

### 2019-06-19 razzo-introduction.tex
Folded TeX

@@
 
-Modern computational techniques allow to infer phylogenies from genetic data such as DNA, RNA or proteins.
-BEAST (\cite{beast}) and its descendant BEAST2 (\cite{beast2}) are widely used tools to perform this task, which they can achieve by running a Bayesian analysis given data and tree priors.
+Modern computational techniques allow to infer phylogenies from genetic data suc
+h as DNA, RNA or proteins. BEAST (\cite{beast}) and its descendant BEAST2 (\cite
+{beast2}) are widely used tools to perform this task, which they can achieve by 
+running a Bayesian analysis given data and tree priors.
 
-BEAST2 gives to the user the option to set up several possible phylogenetic priors (e.g.
-substitution/clock/speciation models).
-However, currently available priors can be not suitable to analyze some specific datasets.
+BEAST2 gives to the user the option to set up several possible phylogenetic prio
+rs (e.g. substitution/clock/speciation models). However, currently available pri
+ors can be not suitable to analyze some specific datasets.
 
-For this reason BEAST2 provides users with the possibility to introduce new tree priors, to infer phylogenies based on different assumptions on how the speciation process takes place.
+For this reason BEAST2 provides users with the possibility to introduce new tree
+ priors, to infer phylogenies based on different assumptions on how the speciati
+on process takes place.
 
-Current phylogenetic tools assume that only a single speciation event can occur at any given time.
-While this assumption has been proved to be useful to construct a wide variety of successful models (e.g \cite{Maddison2007biSSE}, \cite{Valente2015}, \cite{etienne2012diversity}, \cite{etienne2014estimating}), they disallow for environmental changes that trigger speciations in multiple clades at a same point in time.
+Current phylogenetic tools assume that only a single speciation event can occur 
+at any given time.
+While this assumption has been proved to be useful to construct a wide variety o
+f successful models (e.g \cite{Maddison2007biSSE}, \cite{Valente2015}, \cite{eti
+enne2012diversity}, \cite{etienne2014estimating}), they disallow for environment
+al changes that trigger speciations in multiple clades at a same point in time.
 
-We explore such case introducing the multiple birth model, currently absent in BEAST.
-The multiple birth hypothesis aims to include species pump mechanisms to investigate drivers and modes of such diversification processes whose phylogenies show an impressive amount of speciation events in relatively short times.
+We explore such case introducing the multiple birth model, currently absent in B
+EAST. The multiple birth hypothesis aims to include species pump mechanisms to i
+nvestigate drivers and modes of such diversification processes whose phylogenies
+ show an impressive amount of speciation events in relatively short times.
 
-The (constant-rate) birth-death (BD) model embodies the common assumption that only a single speciation event can occur at any given time.
-The multiple-birth-death (MBD) model relaxes this assumption allowing, in addition to standard BD events, also events in which large-scale environmental changes lead to speciation bursts.
-Such hypothesis can be useful to better understand the history of systems of particular interests for evolutionary biologists, such as the  diversification of cichlid fish in the African Great Lakes (Malawi, Tanganyika and Victoria), where water level fluctuations are thought to play an important role in promoting diversification (\citep{verheyen1996mitochondrial}, \citep{sturmbauer2001lake}, \citep{janzen2016}, \citep{janzen2017}).
+The (constant-rate) birth-death (BD) model embodies the common assumption that o
+nly a single speciation event can occur at any given time.
+The multiple-birth-death (MBD) model relaxes this assumption allowing, in additi
+on to standard BD events, also events in which large-scale environmental changes
+ lead to speciation bursts.
+Such hypothesis can be useful to better understand the history of systems of par
+ticular interests for evolutionary biologists, such as the  diversification of c
+ichlid fish in the African Great Lakes (Malawi, Tanganyika and Victoria), where 
+water level fluctuations are thought to play an important role in promoting dive
+rsification (\citep{verheyen1996mitochondrial}, \citep{sturmbauer2001lake}, \cit
+ep{janzen2016}, \citep{janzen2017}).
 
-However, the introduction of new tree priors is not always desirable (\citep{bilderbeek2019pirouette}).
-Current BD tree priors might, in principle, prove to be good enough at detecting such events despite the lower level of complexity.
-If this is the case one should always be more keen to adopt the simplest model.
+However, the introduction of new tree priors is not always desirable (\citep{bil
+derbeek2019pirouette}). Current BD tree priors might, in principle, prove to be 
+good enough at detecting such events despite the lower level of complexity. If t
+his is the case one should always be more keen to adopt the simplest model.
 
-We used the R package \verb;pirouette; (\citep{pirouette}) to perform such test, starting on phylogenies simulated under the MBD regime using the \verb;mbd; package (\citep{mbd}).
-From such phylogenies we measure the inference error made adopting a standard BD tree prior in the inference process.
+We used the R package \verb;pirouette; (\citep{pirouette}) to perform such test,
+ starting on phylogenies simulated under the MBD regime using the \verb;mbd; pac
+kage (\citep{mbd}).
+From such phylogenies we measure the inference error made adopting a standard BD
+ tree prior in the inference process.
 
-With this work we aim, using such inference error distributions, to test whether or not it is advantageous to implement a new prior model that can allow the construction of trees where multiple speciations can co-occur at the same time.
+With this work we aim, using such inference error distributions, to test whether
+ or not it is advantageous to implement a new prior model that can allow the con
+struction of trees where multiple speciations can co-occur at the same time.

### 2019-06-19 razzo-introduction.tex
Folded in a readable way

--- 2019-06-19 razzo-introduction.tex
@@
 
-Modern computational techniques allow to infer phylogenies from genetic data such as DNA, RNA or proteins.
-BEAST (\cite{beast}) and its descendant BEAST2 (\cite{beast2}) are widely used tools to perform this task, which they can achieve by running a Bayesian analysis given data and tree priors.
+Modern computational techniques allow to infer phylogenies from genetic data 
+such as DNA, RNA or proteins. BEAST (\cite{beast}) and its descendant 
+BEAST2 (\cite{beast2}) are widely used tools to perform this task, which they 
+can achieve by running a Bayesian analysis given data and tree priors.
 
-BEAST2 gives to the user the option to set up several possible phylogenetic priors (e.g.
-substitution/clock/speciation models).
-However, currently available priors can be not suitable to analyze some specific datasets.
+BEAST2 gives to the user the option to set up several possible phylogenetic 
+priors (e.g. substitution/clock/speciation models). However, currently 
+available priors can be not suitable to analyze some specific datasets.
 
-For this reason BEAST2 provides users with the possibility to introduce new tree priors, to infer phylogenies based on different assumptions on how the speciation process takes place.
+For this reason BEAST2 provides users with the possibility to introduce new 
+tree priors, to infer phylogenies based on different assumptions on how the 
+speciation process takes place.
 
-Current phylogenetic tools assume that only a single speciation event can occur at any given time.
-While this assumption has been proved to be useful to construct a wide variety of successful models (e.g \cite{Maddison2007biSSE}, \cite{Valente2015}, \cite{etienne2012diversity}, \cite{etienne2014estimating}), they disallow for environmental changes that trigger speciations in multiple clades at a same point in time.
+Current phylogenetic tools assume that only a single speciation event can occur 
+at any given time.
+While this assumption has been proved to be useful to construct a wide 
+variety of successful models (e.g \cite{Maddison2007biSSE}, 
+\cite{Valente2015}, \cite{etienne2012diversity}, \cite{etienne2014estimating}), 
+they disallow for environment
+al changes that trigger speciations in multiple clades at a same point in time.
 
-We explore such case introducing the multiple birth model, currently absent in BEAST.
-The multiple birth hypothesis aims to include species pump mechanisms to investigate drivers and modes of such diversification processes whose phylogenies show an impressive amount of speciation events in relatively short times.
+We explore such case introducing the multiple birth model, currently absent 
+in BEAST. The multiple birth hypothesis aims to include species pump 
+mechanisms to investigate drivers and modes of such diversification processes 
+whose phylogenies show an impressive amount of speciation events 
+in relatively short times.
 
-The (constant-rate) birth-death (BD) model embodies the common assumption that only a single speciation event can occur at any given time.
-The multiple-birth-death (MBD) model relaxes this assumption allowing, in addition to standard BD events, also events in which large-scale environmental changes lead to speciation bursts.
-Such hypothesis can be useful to better understand the history of systems of particular interests for evolutionary biologists, such as the  diversification of cichlid fish in the African Great Lakes (Malawi, Tanganyika and Victoria), where water level fluctuations are thought to play an important role in promoting diversification (\citep{verheyen1996mitochondrial}, \citep{sturmbauer2001lake}, \citep{janzen2016}, \citep{janzen2017}).
+The (constant-rate) birth-death (BD) model embodies the common assumption that 
+only a single speciation event can occur at any given time. The 
+multiple-birth-death (MBD) model relaxes this assumption allowing, in additi
+on to standard BD events, also events in which large-scale environmental 
+changes lead to speciation bursts.
+Such hypothesis can be useful to better understand the history of systems of 
+particular interests for evolutionary biologists, such as the 
+diversification of cichlid fish in the African Great Lakes (Malawi, 
+Tanganyika and Victoria), where 
+water level fluctuations are thought to play an important role in promoting 
+diversification (\citep{verheyen1996mitochondrial}, \citep{sturmbauer2001lake}, 
+\citep{janzen2016}, \citep{janzen2017}).
 
-However, the introduction of new tree priors is not always desirable (\citep{bilderbeek2019pirouette}).
-Current BD tree priors might, in principle, prove to be good enough at detecting such events despite the lower level of complexity.
+However, the introduction of new tree priors is not always 
+desirable (\citep{bilderbeek2019pirouette}). Current BD tree priors might, 
+in principle, prove to be 
+good enough at detecting such events despite the lower level of complexity. 
 If this is the case one should always be more keen to adopt the simplest model.
 
-We used the R package \verb;pirouette; (\citep{pirouette}) to perform such test, starting on phylogenies simulated under the MBD regime using the \verb;mbd; package (\citep{mbd}).
-From such phylogenies we measure the inference error made adopting a standard BD tree prior in the inference process.
+We used the R package \verb;pirouette; (\citep{pirouette}) to perform such 
+test, starting on phylogenies simulated under the MBD regime using 
+the \verb;mbd; package (\citep{mbd}).
+From such phylogenies we measure the inference error made adopting a 
+standard BD tree prior in the inference process.
 
-With this work we aim, using such inference error distributions, to test whether or not it is advantageous to implement a new prior model that can allow the construction of trees where multiple speciations can co-occur at the same time.
+With this work we aim, using such inference error distributions, to test 
+whether or not it is advantageous to implement a new prior model that can 
+allow the construction of trees where multiple speciations can co-occur at 
+the same time.

### 2019-06-19 razzo-discussion.tex
Start of discussion

\section{Discussion}

From the four MBD parameters $\lambda$, $\mu$, $\nu$ and $q$,
we investigated 1, 2, 4 and 3 different values respectively.
We chose to use only one $\lambda$, as the proportion of species
created in a co-occurent speciation event is dependent on the ratio
between $\lambda$ and a combination of $\nu$ and $q$.

We selected our parameters in such a way that the simulated trees 
had usually less than 200 taxa. One could argue that starting from
trees with more taxa would result in a clearer inference, which we
agree upon. We chose to use more replicates over more taxa, 
as we could easily add more replicates in a scheduled way.

### 2019-06-19 razzo-results.tex
Write down expected results

@@
 \section{Results}
-Here we present the results as show in Fig.~\ref{fig:results}...
+
+The inference error made for each of the parameter combinations
+is shown in Fig.~\ref{fig:results}. For both extinction rates,
+we find, as expected, that the error increases for increased $\nu$
+or $q$. Also in line with our predictions, we find no difference
+between the two extinction rates.
 

### 2019-06-19 razzo-methods.tex
Folded in a readable way
Improve methods

--- 2019-06-19 razzo-methods.tex
@@
 Parameter $\nu$ is the rate at which an environmental change is triggered.
-When such event is triggered, each species present in the phylogeny at that moment has a probability $q$ to speciate at that time.
-This kind of speciation is of a different nature respect to the one triggered by $\lambda$.
-In fact, whereas parameter $\lambda$ can be seen as describing a sympatric process, $\nu$ induces the rise of geographical barriers interrupting the gene flow and leading to an allopatric speciation.
-Even though multiple speciations can co-occur, polytomies are not allowed in such process as each species can speciate only once at the time.
+When such event is triggered, each species present in the phylogeny at that 
+moment has a probability $q$ to speciate at that time.
+This kind of speciation is of a different nature respect to the one triggered 
+by $\lambda$. In fact, whereas parameter $\lambda$ can be seen as describing 
+a sympatric process, $\nu$ induces the rise of geographical barriers 
+interrupting the gene flow and leading to an allopatric speciation.
+Even though multiple speciations can co-occur, polytomies are not allowed 
+in such process as each species can speciate only once at the time.
 A likelihood expression for the process is provided in \cite{mbd}.
@@
 
-We can easily simulate such processes in continuous time using the Doob-Gillespie algorithm. Simulations are performed using the function "mbd\_sim" from the \verb;mbd; package (\citep{mbd}).
-We let parameters vary using all possible combinations of value included in Table~\ref{tab:simulation_parameters} to simulate $1000$ independent trees for each parameter setting.
+We can easily simulate such processes in continuous time using the 
+Doob-Gillespie algorithm. Simulations are performed using the 
+function "mbd\_sim" from the \verb;mbd; package (\citep{mbd}).
+We let parameters vary using all possible combinations of values
+as shown in Table~\ref{tab:simulation_parameters}.
+For each parameter setting, we simulate $1000$ independent trees.
+
+We have picked the parameters in such a way that in the most speciose
+setting, the simulated trees have usually less than 200 taxa.
 
@@
 
-Once the MBD dataset has been simulated we exploit the \verb;pirouette; package (\citep{pirouette}) to assess the error made by BEAST2, executing  the inference using a BD tree prior.
+Once the MBD dataset has been simulated we exploit the \verb;pirouette; 
+package (\citep{pirouette}) to assess the error made by BEAST2, executing 
+the inference using a BD tree prior.
 
@@
 From each MBD tree a DNA sequence alignment is simulated. 
-For each sequence alignment we then perform a Bayesian inference analysis (using BEAST2) to recover a posterior distribution of trees. 
-For each parameter setting, this process leads to an inference error distribution.
-To evaluate the extent of this error we also run a \verb;pirouette; "twin" pipeline, through which we mimic the original pipeline in every aspect but starting instead from "twin" trees.
-From each MBD simulated tree we can produce its BD twin, inferring the most likely BD parameters (i.e.
-$\lambda$ and $\mu$) through maximum likelihood, and use them to simulate a standard BD tree (see \cite{pirouette} for more information), keeping the topology of the original tree.
-We let the Bayesian analysis assume a BD prior in both cases, to investigate the extent of the error we make under this assumption.
-The twin pipeline serves as an estimation of the baseline error, as its error it is not due by the mismatch of the generative prior with the inference prior.
+For each sequence alignment we then perform a Bayesian inference 
+analysis (using BEAST2) to recover a posterior distribution of trees. 
+For each parameter setting, this process leads to an inference error 
+distribution. To evaluate the extent of this error we also run 
+a \verb;pirouette; "twin" pipeline, through which we mimic the original 
+pipeline in every aspect but starting instead from "twin" trees.
+From each MBD simulated tree we can produce its BD twin, inferring the most 
+likely BD parameters (i.e. $\lambda$ and $\mu$) through maximum likelihood, 
+and use them to simulate a standard BD tree (see \cite{pirouette} for more 
+information), keeping the topology of the original tree.
+We let the Bayesian analysis assume a BD prior in both cases, to investigate 
+the extent of the error we make under this assumption. The twin pipeline 
+serves as an estimation of the baseline error, as its error it is not due by 
+the mismatch of the generative prior with the inference prior.
 

### 2019-06-19 razzo-methods.tex
Add figs

@@
 \fi
+
+\subsection{Quantifying the impact of the speciation model}
+\richel{this is a proposal}
+
+To quantify the impact of the speciation model, we propose to
+use impact = complete error - baseline error / complete error.
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{razzo-figures/pir_ex_3_errors.png}
+  \caption{
+    Uses Yule tree, impact of speciation model should be low
+  }
+  \label{fig:results}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{razzo-figures/pir_ex_6_errors.png}
+  \caption{
+    Uses non-standard tree, impact of speciation model should be high
+  }
+  \label{fig:results}
+\end{figure}

### 2019-06-20 razzo-methods.tex
Deleted impact definition

--- 2019-06-19 razzo-methods.tex
@@
 also in the standard BD model. 
-Additionally, MBD relies on two additional parameters. 
-Parameter $\nu$ is the rate at which an environmental change is triggered.
+Additionally, MBD relies on two additional parameters, $\nu$ and $q$. 
+$\nu$ is the rate at which an environmental change is triggered.
 When such event is triggered, each species present in the phylogeny at that 
@@
 
-We can easily simulate such processes in continuous time using the 
-Doob-Gillespie algorithm. Simulations are performed using the 
-function "mbd\_sim" from the \verb;mbd; package (\citep{mbd}).
+We simulate the speciation process in continuous time using the 
+Doob-Gillespie algorithm, using 
+the \verb;mbd_sim; function in the \verb;mbd; package (\citep{mbd}).
 We let parameters vary using all possible combinations of values
 as shown in Table~\ref{tab:simulation_parameters}.
-For each parameter setting, we simulate $1000$ independent trees.
-
+For each parameter setting, we generate $1000$ independent 
+reconstructed trees of the same crown age.
 We have picked the parameters in such a way that in the most speciose
@@
 
-\subsection{Pirouette}
+\subsection{Inference error estimation}
 
-Once the MBD dataset has been simulated we exploit the \verb;pirouette; 
-package (\citep{pirouette}) to assess the error made by BEAST2, executing 
-the inference using a BD tree prior.
+From each MBD tree, we measure the impact 
+of using the simpler BD tree prior, 
+using the \verb;pirouette; R package, as described
+in detail in \citep{pirouette}.
 
-We briefly summarize here how the \verb;pirouette; routine works.
-From each MBD tree a DNA sequence alignment is simulated. 
-For each sequence alignment we then perform a Bayesian inference 
-analysis (using BEAST2) to recover a posterior distribution of trees. 
-For each parameter setting, this process leads to an inference error 
-distribution. To evaluate the extent of this error we also run 
-a \verb;pirouette; "twin" pipeline, through which we mimic the original 
-pipeline in every aspect but starting instead from "twin" trees.
-From each MBD simulated tree we can produce its BD twin, inferring the most 
-likely BD parameters (i.e. $\lambda$ and $\mu$) through maximum likelihood, 
-and use them to simulate a standard BD tree (see \cite{pirouette} for more 
-information), keeping the topology of the original tree.
-We let the Bayesian analysis assume a BD prior in both cases, to investigate 
-the extent of the error we make under this assumption. The twin pipeline 
-serves as an estimation of the baseline error, as its error it is not due by 
-the mismatch of the generative prior with the inference prior.
+In brief: \verb;pirouette; starts from a 'true' (but 
+unobservable in nature) starting phylogeny,
+from which a DNA sequence alignment (which is observable 
+in nature) is simulated. 
+From each sequence alignment, a Bayesian inference is run, 
+to obtain a posterior distribution of jointly-estimated 
+trees and model parameter estimates.
+By comparing the true tree and the posterior trees,
+an inference error distribution is generated. 
+We use the twinning option available in \verb;pirouette;
+to measure a minimum and full error, in which
+the minimum error is caused
+by stochasticity in the full pipeline, where the full error
+is the added error from using an invalid but standard speciation model.
 
-\iffalse
-\subsection{Model selection}
+In our context, the alignments are 1000 nucleotides in length,
+with a known root sequence of four 250 mono-nucloetide blocks,
+following the simplest nucleotide substitution model (JC69)
+and clock model (strict), with a mutation rate of $\frac{1}{2}\cdot c$, 
+in which $c$ is the crown age. With this mutation rate, each nuclotide is
+expected to mutate (both silently and non-silently) 
+in half of the histories from the root sequence in the past 
+to the sequence in the tips in the present.
 
-We simulate alignments using the simplest nucleotide substitution model (JC69),
-the simplest clock model (strict). It is thus imminent to assume these
-models in our Bayesian inference. Nevertheless, the phylogeny the alignment
-was based on, could have followed either an MBD or BD tree model, 
-where we in both cases assume a BD tree model. This will have 
-an unknown effect on our inference: it may theoretically be that an MBD model
-generates (a tree that generates) an alignment in which a different site 
-and/or clock model is favored.
+For the Bayesian inference, we use a JC69 site model, a strict
+clock model and a BD tree prior. Additionally, we use 
+an MRCA prior equal to the crown age with a normal distribution 
+of width $\sigma = 0.01$. We pick an MCMC setup of 10M states,
+which is sampled each 1k moves.
 
-We investigate this by measuring if the generative model (with the simplest
-nucleotide substitution and simplest clock model) is indeed selected 
-to be the best fitting model. 
-To be precise, we look at the model 
-with the highest marginal likelihood 
-(also called evidence \cite{mackay2003information}),
-$f(D|M)$, which is the probability of the data D given model M.
-In the context of this research, D consists of the DNA alignment,
-and M is the combination of site, clock and tree models.
+For the error measurement, we use the nLTT statistic and
+a burn-in of 0.1.
 
-To estimate the marginal likelihood, 
-we use an algorithm named nested sampling \cite{skilling2006nested}.
-Nested sampling is attractive to use
-in a phylogentic context, as it gives a good estimation,
-requires little tuning \cite{maturana2018}.
-Nested sampling is available as a BEAST2 package
-and can be used by babette \cite{babette}.
+For the twinning, we let the twin trees follow a BD model.
 
-The nested sampling algorithm stops its run 
-when the marginal likelihood estimation error 
-reaches below a certain tolerance.  
-Similar to \cite{maturana2018},
-we use a (relative) error tolerance $\epsilon$ of $10^{-13}$,
-1 particle to explore the parameter space
-and 100 active points. 
-To achieve the latter, we use the MCMC chain length $L_c$ of 1M 
-(as also used in the parameter estimates),
-and a sub-chain length $L_{sc}$ of 10K.
-
-The models we use in our model comparison are the four combinations
-of two site models and two clock models. We use the JC69 site model, which
-is the (generative and) simplest model and GTR, the site model with most
-degrees of freedom. For the clock models, we use the strict clock model,
-which is the (generative and) simplest clock model, and the RLN clock model.
-\richel{Could also just be all site models and clock models = 8 models}
-
-From these four marginal likelihood estimates, we calculate the weight of
-the generative model and plot this in figure 2. We do this for both the 
-alignments derived from the MBD tree and the BD twin tree. We expect that
-the generative model has the heighest weight in both the MBD and BD alignments.
-We expect this weight to be higher in the BD alignments.
-\fi
+For the experiments, we use both a hand-picked generative model
+and a set of candidate models. We set a BD tree prior, JC69 site
+model and strict clock model as the generative model. We used
+all other combinations of four tree priors, two clock models and
+five speciation models, resulting in a set of 39 candidate models.

### 2019-07-01 razzo-abstract.tex
Updates from Overleaf

--- 2019-06-19 razzo-abstract.tex
@@
 % scientists in related disciplines.
-In Bayesian phylogenetics, a DNA/RNA/protein alignment
-is used to infer a distribution of phylogenies and parameter estimates.
-To do so, we use assumptions that may be biologically unrealistic,
-but may give tolerable errors.
+Starting from heritable material and explicit assumptions,
+Bayesian phylogenetics allows to infer a jointly-estimated 
+phylogeny and parameter estimates distribution.
+One of these assumptions is the speciation model, which
+mathematically describes the branching process of a phylogeny in time.
+Speciation models commonly assume that speciation events are independent and disjointed.
+Yet, such assumption may overlook the complexity of certain processes, where environmental changes promote speciation events in multiple lineages.
+% Purpose of the paper
+This new layer of complexity can be captured developing a novel ad hoc speciation model.
+However, the introduction of a new model could be not necessary if current models are capable of describing the process, under simpler assumptions.
+Here we investigate the extent of the discrepancy produced by current BEAST2's tree priors, when trying to infer back trees where speciation can co-occur.
 
-% One sentence clearly stating the general
-% problem being addressed by this particular
-% study.
-Contemporary inference assumes that speciation
-never co-occurs.
+\iffalse
+This research answers the question what the impact is of using a
+species tree model that assumes speciation is independent, when it
+is used on phylogenies created by a tree model in which speciation can
+co-occur.
+\fi
 
-% One sentence summarising the main
-% result (with the words “here we show”
-% their equivalent).
-Here we show the error we make in our inference,
-when nature has varying degrees of co-occurring speciation.
+% Summarising the main result
+We let the impact of co-occurring speciation vary on datasets of simulated trees and show the corresponding error produced during the inference process.
 
@@
 % previous knowledge.
+We show that the extent of the inference error grows with the amount of co-occuring speciation events to establish the limits of BEAST2's standard tree priors.
 
@@
 % more general context.
+These results allow phylogeneticists to judge under which circumstances the commonly used speciation model can be safely used.
 
-% Two or three sentences to provide a
-% broader perspective, readily comprehensible
-% to a scientist in any discipline, may be included
-% in the first paragraph if the editor considers that the accessibility of the paper is significantly enhanced
-% by their inclusion. Under these circumstances, the length of the paragraph can be up to 300 words.
 \end{abstract}
-{\bf Keywords:} computational biology, evolution, phylogenetics, Bayesian analysis, tree prior
+
+{\bf Keywords:} computational biology, evolution, phylogenetics, Bayesian analysis, tree prior,
+pirouette, BEAST2, babette

### 2019-07-01 razzo-methods.tex
Put 'tab_simparameters.tex' in root folder for thesis
Updates from Overleaf

@@
 \section{Methods}
+
+We proceed in the following way: we build simulated datasets generated under the multiple birth model.
+Then we run a \verb;pirouette; analysis, which will lead to error distributions between the inference posterior and the original simulated trees.
+Importantly, the \verb;pirouette; analysis includes also a 'twin' parallel pipeline, which will provide a measure of the baseline error due to pure stochasticity, unavoidably occurring in the process.
 
@@
 
-In the MBD model, parameters $\lambda$ and $\mu$ correspond, respectively, 
-to the common per-species speciation and extinction rates present 
-also in the standard BD model. 
+In the MBD model, parameters $\lambda$ and $\mu$ correspond, respectively, to the common per-species speciation and extinction rates already present in the standard BD model. 
 Additionally, MBD relies on two additional parameters, $\nu$ and $q$. 
-$\nu$ is the rate at which an environmental change is triggered.
-When such event is triggered, each species present in the phylogeny at that 
-moment has a probability $q$ to speciate at that time.
-This kind of speciation is of a different nature respect to the one triggered 
-by $\lambda$. In fact, whereas parameter $\lambda$ can be seen as describing 
-a sympatric process, $\nu$ induces the rise of geographical barriers 
-interrupting the gene flow and leading to an allopatric speciation.
-Even though multiple speciations can co-occur, polytomies are not allowed 
-in such process as each species can speciate only once at the time.
+The first, $\nu$, is the rate at which an environmental change is triggered.
+When such event is triggered, each species present in the phylogeny at that moment has a probability $q$ to speciate at that time.
+This kind of speciation is of a different nature respect to the one triggered by $\lambda$.
+In fact, whereas parameter $\lambda$ can be seen as describing a sympatric process, $\nu$ induces the rise of geographical barriers interrupting the gene flow \giovanni{@richel: please check if biology here is accurate} and leading to an allopatric speciation.
+Even though multiple speciations can co-occur, polytomies are not allowed in such process as each species can speciate only once at the time.
 A likelihood expression for the process is provided in \cite{mbd}.
@@
 
-We simulate the speciation process in continuous time using the 
-Doob-Gillespie algorithm, using 
-the \verb;mbd_sim; function in the \verb;mbd; package (\citep{mbd}).
-We let parameters vary using all possible combinations of values
-as shown in Table~\ref{tab:simulation_parameters}.
-For each parameter setting, we generate $1000$ independent 
-reconstructed trees of the same crown age.
-We have picked the parameters in such a way that in the most speciose
-setting, the simulated trees have usually less than 200 taxa.
+We simulate the speciation process in continuous time using the Doob-Gillespie algorithm, using the \verb;mbd_sim; function from the \verb;mbd; package (\citep{mbd}).
+We let parameters vary using all possible combinations of values as shown in Table~\ref{tab:simulation_parameters}.
+For each parameter setting, we generate $1000$ independent reconstructed trees of the same crown age.
+\giovanni{Do we need to add more information on the Doob-Gillespie algorithm or it is overkilling?}
 
@@
 
-From each MBD tree, we measure the impact 
-of using the simpler BD tree prior, 
-using the \verb;pirouette; R package, as described
-in detail in \citep{pirouette}.
+From each MBD tree, we measure the impact of using the simpler BD tree prior on the inference, using the \verb;pirouette; R package, as described in detail in \citep{pirouette}.
 
-In brief: \verb;pirouette; starts from a 'true' (but 
-unobservable in nature) starting phylogeny,
-from which a DNA sequence alignment (which is observable 
-in nature) is simulated. 
-From each sequence alignment, a Bayesian inference is run, 
-to obtain a posterior distribution of jointly-estimated 
-trees and model parameter estimates.
-By comparing the true tree and the posterior trees,
-an inference error distribution is generated. 
-We use the twinning option available in \verb;pirouette;
-to measure a minimum and full error, in which
-the minimum error is caused
-by stochasticity in the full pipeline, where the full error
-is the added error from using an invalid but standard speciation model.
+In brief: \verb;pirouette; starts from a 'true' (but unobservable in nature) starting phylogeny, from which a DNA sequence alignment (which is observable in nature) is simulated. 
+From each sequence alignment, a Bayesian inference is run, to obtain a posterior distribution of jointly-estimated trees and model parameter estimates.
+By comparing the true tree and the posterior trees, an inference error distribution is generated. 
+We use the twinning option available in \verb;pirouette; to measure a minimum and full error, in which the minimum error is caused by stochasticity in the full pipeline, where the full error is the added error from using an invalid but standard speciation model \giovanni{I am not sure I get this.
+What do you mean?}.
 
-In our context, the alignments are 1000 nucleotides in length,
-with a known root sequence of four 250 mono-nucloetide blocks,
-following the simplest nucleotide substitution model (JC69)
-and clock model (strict), with a mutation rate of $\frac{1}{2}\cdot c$, 
-in which $c$ is the crown age. With this mutation rate, each nuclotide is
-expected to mutate (both silently and non-silently) 
-in half of the histories from the root sequence in the past 
-to the sequence in the tips in the present.
+In our context, the alignments are 1000 nucleotides in length, with a known root sequence of four 250 mono-nucleotide blocks, following the simplest nucleotide substitution model (JC69) and clock model (strict), with a mutation rate of $\frac{1}{2}\cdot c$,  in which $c$ is the crown age.
+With this mutation rate, each nucleotide is expected to mutate (both silently and non-silently)
+in half of the histories from the root sequence in the past to the sequence in the tips in the present.
 
@@
 
-For the error measurement, we use the nLTT statistic and
-a burn-in of 0.1.
+For the error measurement, we use the nLTT statistic \citet{janzen2015} and
+a burn-in fraction of $10\%$.
 

### 2019-07-01 razzo-results.tex
Updates from Overleaf

@@
 is shown in Fig.~\ref{fig:results}. For both extinction rates,
-we find, as expected, that the error increases for increased $\nu$
-or $q$. Also in line with our predictions, we find no difference
-between the two extinction rates.
+we find that the error increases as $\nu$
+and $q$ increase. Also in line with our predictions, we find no difference
+between the two extinction rates \giovanni{I don't actually have predictions on that.}.
 

### 2019-07-01 razzo-introduction.tex

--- 2019-06-19 razzo-introduction.tex
@@
 
-Modern computational techniques allow to infer phylogenies from genetic data such as DNA, RNA or proteins.
-BEAST (\cite{beast}) and its descendant BEAST2 (\cite{beast2}) are widely used tools to perform this task, which they can achieve by running a Bayesian analysis given data and tree priors.
+\giovanni{Introduce "BD" and "MBD" acronyms at some point.}
+
+Modern computational techniques allow to infer phylogenies from genetic data 
+such as DNA, RNA or proteins. BEAST (\cite{beast}) and its descendant 
+BEAST2 (\cite{beast2}) are widely used tools to perform such task. They 
+return posterior distributions of phylogenies and estimated parameters by running a Bayesian analysis, given genetic data and tree priors.
+A tree prior is a mathematical description of the a priori characteristics that we expect to be reflected in posterior phylogenies.
+The choice of a specific prior is, by definition, arbitrary but the consequences of such choice can be vet analyzing the so-obtained posteriors.
 
@@
 substitution/clock/speciation models).
-However, currently available priors can be not suitable to analyze some specific datasets.
-
+However, currently available priors might be not suitable to analyze some specific datasets.
 For this reason BEAST2 provides users with the possibility to introduce new tree priors, to infer phylogenies based on different assumptions on how the speciation process takes place.
@@
 Current phylogenetic tools assume that only a single speciation event can occur at any given time.
-While this assumption has been proved to be useful to construct a wide variety of successful models (e.g \cite{Maddison2007biSSE}, \cite{Valente2015}, \cite{etienne2012diversity}, \cite{etienne2014estimating}), they disallow for environmental changes that trigger speciations in multiple clades at a same point in time.
+While this assumption has been proved to be useful to construct a wide variety of successful models (e.g \cite{Maddison2007biSSE}, \cite{Valente2015}, \cite{etienne2012diversity}, \cite{etienne2014estimating}), they do not take into account the possibility for environmentally-driven contemporary speciations on multiple lineages.
 
-We explore such case introducing the multiple birth model, currently absent in BEAST.
-The multiple birth hypothesis aims to include species pump mechanisms to investigate drivers and modes of such diversification processes whose phylogenies show an impressive amount of speciation events in relatively short times.
+We explore such case introducing the multiple-birth death model (MBD), currently absent in BEAST2.
+The multiple birth hypothesis aims to describe how large-scale environmental changes (also known as species pumps \giovanni{add some ref?}) can lead phylogenies to sport an impressive amount of speciation events in relatively short times.
 
+\iffalse
+\giovanni{This might be good for the abstract}
 The (constant-rate) birth-death (BD) model embodies the common assumption that only a single speciation event can occur at any given time.
 The multiple-birth-death (MBD) model relaxes this assumption allowing, in addition to standard BD events, also events in which large-scale environmental changes lead to speciation bursts.
+\fi
+
 Such hypothesis can be useful to better understand the history of systems of particular interests for evolutionary biologists, such as the diversification of cichlid fish in the African Great Lakes (Malawi, Tanganyika and Victoria), where water level fluctuations are thought to play an important role in promoting diversification (\citep{verheyen1996mitochondrial}, \citep{sturmbauer2001lake}, \citep{janzen2016}, \citep{janzen2017}).
@@
 However, the introduction of new tree priors is not always desirable (\citep{bilderbeek2019pirouette}).
-Current BD tree priors might, in principle, prove to be good enough at detecting such events despite the lower level of complexity.
-If this is the case one should always be more keen to adopt the simplest model.
+Current standard birth death (BD) tree priors might, in principle, prove to be good enough at detecting such events despite their lower level of complexity.
+If this is the case one should always choose to adopt the simplest model.
 

### 2019-07-01 razzo-discussion.tex

@@
 between $\lambda$ and a combination of $\nu$ and $q$.
-
-We selected our parameters in such a way that the simulated trees 
-had usually less than 200 taxa. One could argue that starting from
-trees with more taxa would result in a clearer inference, which we
-agree upon. We chose to use more replicates over more taxa, 
-as we could easily add more replicates in a scheduled way.

### 2019-07-01 fig_results.tex
Move figures and their TeX files to the same folder as the other TeX, as using a subfolder does not scale (that is, renders the TeX unusable to be used in a thesis)

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{razzo_results.png}
  \caption{
    The inference error distribution 
    for the different biological
    parameter settings. In all cases, $\lambda = 0.2$ and 
    crown age equals 6. 
  }
  \label{fig:results}
\end{figure}

### 2019-08-22 razzo-methods.tex
Do not use '\input' in file that is already added by an '\input' as it does not scale (that is, this TeX cannot be used as part of a thesis)
Improve wording, added to describe the nested sampling

@@
 
-In the MBD model, parameters $\lambda$ and $\mu$ correspond, respectively, to the common per-species speciation and extinction rates already present in the standard BD model. 
+In the MBD model, parameters $\lambda$ and $\mu$ correspond, respectively, 
+to the common per-species speciation and extinction rates 
+already present in the standard BD model. 
 Additionally, MBD relies on two additional parameters, $\nu$ and $q$. 
 The first, $\nu$, is the rate at which an environmental change is triggered.
-When such event is triggered, each species present in the phylogeny at that moment has a probability $q$ to speciate at that time.
-This kind of speciation is of a different nature respect to the one triggered by $\lambda$.
-In fact, whereas parameter $\lambda$ can be seen as describing a sympatric process, $\nu$ induces the rise of geographical barriers interrupting the gene flow \giovanni{@richel: please check if biology here is accurate} and leading to an allopatric speciation.
-Even though multiple speciations can co-occur, polytomies are not allowed in such process as each species can speciate only once at the time.
+When such event is triggered, each species present in the phylogeny 
+at that moment has a probability $q$ to speciate at that time.
+This kind of speciation is of a different nature with respect to 
+the one triggered by $\lambda$. 
+In fact, whereas parameter $\lambda$ can be seen as 
+describing a sympatric process, $\nu$ induces the rise of 
+geographical barriers that interrupt the gene flow,
+resulting in allopatric speciation. 
+Even though multiple speciations can co-occur, 
+polytomies are not allowed in such process as each species can speciate only 
+once during a co-occuring speciation event.
 A likelihood expression for the process is provided in \cite{mbd}.
@@
 
-We simulate the speciation process in continuous time using the Doob-Gillespie algorithm, using the \verb;mbd_sim; function from the \verb;mbd; package (\citep{mbd}).
-We let parameters vary using all possible combinations of values as shown in Table~\ref{tab:simulation_parameters}.
-For each parameter setting, we generate $1000$ independent reconstructed trees of the same crown age.
-\giovanni{Do we need to add more information on the Doob-Gillespie algorithm or it is overkilling?}
-
-\input{razzo-tables/tab_simparameters.tex}
+We simulate the speciation process in continuous time 
+using the \verb;mbd_sim; function from the \verb;mbd; R package (\citep{mbd}).
+We let parameters vary using all possible combinations of values 
+as shown in Table~\ref{tab:simulation_parameters}.
+For each parameter setting, 
+we generate $2$ \richel{Will need to calculate how many replicates we can do} 
+independent reconstructed trees of the same crown age.
 
@@
 
-From each MBD tree, we measure the impact of using the simpler BD tree prior on the inference, using the \verb;pirouette; R package, as described in detail in \citep{pirouette}.
+From each simulated 'true' MBD tree, we measure the impact of
+ignoring the more complex and non-standard MBD tree prior in
+Bayesian phylogentic inference.
+We do so by assuming a simpler and standard BD tree prior,
+and compare the 'true' tree with the inferred trees.
+To do so, we use the \verb;pirouette; R package \citep{pirouette}.
 
-In brief: \verb;pirouette; starts from a 'true' (but unobservable in nature) starting phylogeny, from which a DNA sequence alignment (which is observable in nature) is simulated. 
-From each sequence alignment, a Bayesian inference is run, to obtain a posterior distribution of jointly-estimated trees and model parameter estimates.
-By comparing the true tree and the posterior trees, an inference error distribution is generated. 
-We use the twinning option available in \verb;pirouette; to measure a minimum and full error, in which the minimum error is caused by stochasticity in the full pipeline, where the full error is the added error from using an invalid but standard speciation model \giovanni{I am not sure I get this.
-What do you mean?}.
+In brief: \verb;pirouette; starts from 
+a 'true' starting phylogeny (in our case: the simulated MBD tree), 
+from which a DNA sequence alignment is simulated. 
+From each sequence alignment, a Bayesian 
+inference is run (in our case: with assuming a BD tree prior), 
+to obtain a posterior distribution of jointly-estimated trees and model parameter estimates.
+By comparing the true tree and the posterior trees, 
+an inference error distribution is generated.
+We use the twinning option available in \verb;pirouette;
+that allows to quantify the impact of assuming a wrong
+tree prior in a Bayesian inference.
 
-In our context, the alignments are 1000 nucleotides in length, with a known root sequence of four 250 mono-nucleotide blocks, following the simplest nucleotide substitution model (JC69) and clock model (strict), with a mutation rate of $\frac{1}{2}\cdot c$,  in which $c$ is the crown age.
-With this mutation rate, each nucleotide is expected to mutate (both silently and non-silently)
-in half of the histories from the root sequence in the past to the sequence in the tips in the present.
+In our experiment, the alignments are 1000 nucleotides in length, 
+with a known root sequence of four 250 mono-nucleotide blocks, 
+generated using the simplest nucleotide substitution model (JC69) 
+and clock model (strict), with a mutation rate of $\frac{1}{2}\cdot c$,
+in which $c$ is the crown age. 
+With this mutation rate, each nucleotide has a 50\% chance
+to mutate (both silently and non-silently) from the ancestral root sequence 
+to any of the contemporary species' sequences at the tips.
 
-For the Bayesian inference, we use a JC69 site model, a strict
-clock model and a BD tree prior. Additionally, we use 
-an MRCA prior equal to the crown age with a normal distribution 
-of width $\sigma = 0.01$. We pick an MCMC setup of 10M states,
-which is sampled each 1k moves.
+For the Bayesian inference, we assume a generative model of 
+a JC69 site model, a strict clock model and a BD tree prior. 
+Additionally, we use an MRCA prior equal to the crown age with a normal distribution 
+of width $\sigma = 0.01$. We pick an MCMC setup of $10^6$ states
+with a sampling interval of once per $10^3$ states. Of the resulting
+$10^3$ states, we discard a burn-in of $10\%$.
 
-For the error measurement, we use the nLTT statistic \citet{janzen2015} and
-a burn-in fraction of $10\%$.
+To verify that the generative model (with a BD tree prior) is
+indeed the most appropriate standard model, we compare it to
+39 other candidate inference models. These candidate models 
+are all other combinations of four tree priors, 
+two clock models and five speciation models, 
+excluding the one combination used as a generative model.
+We let the best candidate model also generate an error distribution,
+which is expected to have a similar or high median.
+For models comparison, \verb;pirouette; uses a nested sampling approach
+to estimate marginal likelihoods,
+as described in [\cite{maturana2018model}], 
+using the \verb;NS; BEAST2 package. We setup the MCMC for the nested sampling
+using its default values, which are a relative error $\epsilon$ of $10^-12$, 
+1 particle, a sub-chain sampling interval of $5.10^3$ states with
+a maximum chain length of $10^6$ states.
 
-For the twinning, we let the twin trees follow a BD model.
+To compare the true tree with any of the posterior trees,
+we use the absolute nLTT statistic \citet{janzen2015},
+which will result in an error distribution with values
+ranging from zero (when the inferred tree is identical 
+to the true tree) to a maximum of one (trees are completely different).
 
-For the experiments, we use both a hand-picked generative model
-and a set of candidate models. We set a BD tree prior, JC69 site
-model and strict clock model as the generative model. We used
-all other combinations of four tree priors, two clock models and
-five speciation models, resulting in a set of 39 candidate models.
+For the twinning, we let a birth-death process generate one random tree,
+so that the twin inference can be done with the correct phylogenetic
+model of a JC69 site model, strict clock model and a BD tree prior.

### 2019-08-22 razzo-introduction.tex
Do not refer to species pump

@@
 
-We explore such case introducing the multiple-birth death model (MBD), currently absent in BEAST2.
-The multiple birth hypothesis aims to describe how large-scale environmental changes (also known as species pumps \giovanni{add some ref?}) can lead phylogenies to sport an impressive amount of speciation events in relatively short times.
+We explore such case introducing the multiple-birth death model (MBD), 
+currently absent in BEAST2. 
+The multiple birth hypothesis aims to describe how 
+large-scale environmental changes can lead phylogenies 
+with a high abundance of speciation events in relatively short times.
 
@@
 \giovanni{This might be good for the abstract}
-The (constant-rate) birth-death (BD) model embodies the common assumption that only a single speciation event can occur at any given time.
-The multiple-birth-death (MBD) model relaxes this assumption allowing, in addition to standard BD events, also events in which large-scale environmental changes lead to speciation bursts.
+The (constant-rate) birth-death (BD) model embodies the common assumption 
+that only a single speciation event can occur at any given time. 
+The multiple-birth-death (MBD) model relaxes this assumption allowing, 
+in addition to standard BD events, 
+also events in which large-scale environmental changes lead 
+to speciation bursts.
 \fi

### 2019-09-02 razzo-abstract.tex
Updates from Overleaf

@@
 This new layer of complexity can be captured developing a novel ad hoc speciation model.
-However, the introduction of a new model could be not necessary if current models are capable of describing the process, under simpler assumptions.
+However, the introduction of a new model could be not necessary if current models are already capable of describing the process to a satisfactory extent, under simpler assumptions.
 Here we investigate the extent of the discrepancy produced by current BEAST2's tree priors, when trying to infer back trees where speciation can co-occur.
-
-\iffalse
-This research answers the question what the impact is of using a
-species tree model that assumes speciation is independent, when it
-is used on phylogenies created by a tree model in which speciation can
-co-occur.
-\fi
 

### 2019-09-02 razzo-methods.tex
Updates from Overleaf

--- 2019-07-01 razzo-methods.tex
@@
 The first, $\nu$, is the rate at which an environmental change is triggered.
-When such event is triggered, each species present in the phylogeny at that moment has a probability $q$ to speciate at that time.
+When such event is triggered, each species present in the phylogeny at that time has a probability $q$ to speciate.
 This kind of speciation is of a different nature respect to the one triggered by $\lambda$.
@@
 
-We simulate the speciation process in continuous time using the Doob-Gillespie algorithm, using the \verb;mbd_sim; function from the \verb;mbd; package (\citep{mbd}).
+We simulate the speciation process in continuous time exploiting the Doob-Gillespie algorithm, using the \verb;mbd_sim; function from the \verb;mbd; package (\citep{mbd}).
 We let parameters vary using all possible combinations of values as shown in Table~\ref{tab:simulation_parameters}.
-For each parameter setting, we generate $1000$ independent reconstructed trees of the same crown age.
+For each parameter setting, we generate $1000$ independent reconstructed trees with the same crown age.
 \giovanni{Do we need to add more information on the Doob-Gillespie algorithm or it is overkilling?}
-
-\input{razzo-tables/tab_simparameters.tex}
 
@@
 
-From each MBD tree, we measure the impact of using the simpler BD tree prior on the inference, using the \verb;pirouette; R package, as described in detail in \citep{pirouette}.
+From each MBD tree, we measure the impact of using the simpler BD tree prior on the inference, using the \verb;pirouette; R package, as described in detail in \citet{bilderbeek2019pirouette}.
 
-In brief: \verb;pirouette; starts from a 'true' (but unobservable in nature) starting phylogeny, from which a DNA sequence alignment (which is observable in nature) is simulated. 
+In brief: \verb;pirouette; starts from a 'true' (but unobservable in nature) starting phylogeny, from which a DNA sequence alignment (which is instead observable) is simulated. 
 From each sequence alignment, a Bayesian inference is run, to obtain a posterior distribution of jointly-estimated trees and model parameter estimates.
@@
 model and strict clock model as the generative model. We used
-all other combinations of four tree priors, two clock models and
-five speciation models, resulting in a set of 39 candidate models.
+all other combinations of five tree priors, four site models and two clock models, resulting in a set of 39 candidate models.

### 2019-09-03 razzo-discussion.tex
Constrain
Improve discussion

@@
 
+We \richel{(at least I. If you do not expect this, let me know the 
+reasoning why)} expected to find a bigger inference error 
+when increasing the amount of co-occurring speciation events.
+The reason is simple: the tree prior used assumes speciation does not
+co-occur. The stronger the amount of co-occurring speciation, the stronger
+the deviation from the species tree prior.
+
+\subsection{Research}
+
+The MBD model is a BD model that allows for co-occurring speciation,
+with the same drawbacks as the BD model: it allows for an unlimited 
+number of species and does not take into account that speciation takes time.
+That does make the MBD model a very focussed model, only taking into 
+account the essence of co-occurring speciation. The MBD model can probably
+be extended to have a diversity-dependent speciation rate (as in the 
+DDD model) and/or making speciation take time by adding 
+an incipient species state (as in the PBD model).
+The idea of this experiment, however, is to measure the impact a novel
+tree prior has on the inference error. To do so, we could have picked
+any other non-standard (as in, not included within BEAST2) tree prior.
+
+\subsection{Methodological choices}
+
+When simulating a DNA alignment, we chose to use the simplest (JC69) site
+model and the simplest (strict) clock model. One could argue these models
+are overly simplistic and biologically irrelevant. We argue that this
+actually improves the clarity of our results for two reasons. The first
+reason is a practical one: because the computational load was lowered,
+we were able to perform more replicates. For a process with this
+amount of stochaticity, we think this is highly favored. The second reason
+is .. less noise,
+more replicates.
+
+ * Choice of generative (JC69 and strict) DNA alignment 
+ * Choice of generative tree prior (BD)
+ * Using Nested Sampling to do model comparison
+ * Using the absolute nLTT statistic as an error function
+
+\subsection{Choices to constrain time}
+
+ * MBD parameters
 From the four MBD parameters $\lambda$, $\mu$, $\nu$ and $q$,
@@
 between $\lambda$ and a combination of $\nu$ and $q$.
+
+ * Crown age, and thus the number of taxa in the true species tree
+ * Number of replicates
+ * DNA sequence length
+ * Effective sample size
+ * Choice of candidate models
+
+We had to constrain ourselves on the number of replicates.
+
+We had to constrain ourselves on the number of taxa in
+the true species trees.
+
+\subsection{Future research}
+
+This research measures the impact an MBD tree prior 
+has on the inference error.
+
+Because the MBD tree prior has only recently been 
+developed, it's biological relevance is unknown.
+Would the MBD tree prior be employed on biological data of a system
+in which co-occuring speciation is known to have taken a strong role,
+it will be hard to conclude its performance is superior to a
+rate-shift model or an early-burst model. One way to do so, is
+to add the MBD other speciation models as a species tree prior
+in a phylogentic tool and do a model comparision (e.g. using Nested Sampling).
+Another way to do so, is measure the impact a rate-shift or early-burst model
+has on the inference error in the same way as this research.
+Doing so would also allow to put the quantitative results into perspective,
+helping us in how to judge the error distributions.
+
+Expressing the inference error of an MBD process as a distribution
+is a good first step in judging the impact an MBD process has on
+our predictions, when compared to standard models. It is only a first
+step as one still needs to subjectively judge these error distributions,
+which was exactly what \verb;pirouette; was hoping to avoid. 
+One way to improve our judgement, is by testing it on known species tree
+priors, for example, by simulating true trees with a BD model, when assuming
+a Yule tree prior.

### 2019-09-03 razzo-discussion.tex
Depend on master branch

@@
 reason is a practical one: because the computational load was lowered,
-we were able to perform more replicates. For a process with this
+we were able to perform more replicates. For a process with this high
 amount of stochaticity, we think this is highly favored. The second reason
-is .. less noise,
-more replicates.
+is that we think this reduces the noise of our results, without 
+hurting the experiment: in our setup, it is essential
+to assume the same generative site and clock model as the one that is actually 
+used. Sure, we could have generated the alignment with a 
+complex (GTR) site model and a complex (relaxed log-normal) clock model, then
+correctly assume these were the generative site and clock model.
+But that would be besides the point of the experiment: to measure the impact
+of the tree prior.
+
+For the generative tree prior, we picked the BD tree prior.
 
  * Choice of generative (JC69 and strict) DNA alignment
+
  * Choice of generative tree prior (BD)

### 2019-09-03 razzo-discussion.tex
Improve discussion

@@
 
-For the generative tree prior, we picked the BD tree prior.
+When simulating our phylogenies, we used an MBD model with and without
+extinction. In the case in which extinction is present,
+picking the BD tree prior as the generative tree prior seems justified:
+it is the best-fitting standard tree prior. For the MBD model without
+extinction, \richel{@Giappo: I think this is an interesting insight:} 
+we could have picked the Yule tree prior. We \richel{at least I do} predicted
+that the Yule tree prior will be selected as the best candidate model
+more often in this scenario, which shows the sensitity of the method we used.
 
- * Choice of generative (JC69 and strict) DNA alignment
+To compare multiple inference models, we picked the Nested Sampling
+approach to directly compare marginal likelihoods. We could have used
+alternative methods like \richel{forgot the term for now: where you
+do an MCMC that can switch between models}. The reason we used Nested
+Sampling was mostly pragmatic and we think other methods are equivalent.
 
- * Choice of generative tree prior (BD)
- * Using Nested Sampling to do model comparison
- * Using the absolute nLTT statistic as an error function
+We used the absolute nLTT statistic to measure the difference between
+two phylogenies. Because the crown age of all phylogenies are equal,
+normalization on the time has no effect at all. Also because of these
+same crown ages, we could have picked many tree difference metrics, like
+\richel{must check:} Pagel's delta or Pigot's delta-r. The reason we used
+the nLTT statistic was mostly pragmatic. \verb;pirouette; does allow to 
+use a different tree difference metric.
 
@@
 
- * MBD parameters
 From the four MBD parameters $\lambda$, $\mu$, $\nu$ and $q$,
@@
 
- * Crown age, and thus the number of taxa in the true species tree
- * Number of replicates
- * DNA sequence length
- * Effective sample size
- * Choice of candidate models
+The crown age we selected for our experiments was based on trial and error
+of pilot experiments. Because the number of species in an MBD process is
+expected to increase exponentially, increasing the crown age has a
+profound effect on the number of taxa, with the consequence of having
+runs that took days to calculate. We decided to prefer a lower number of
+taxa at the gain of having more replicates. As a consequence, 
+this made our research findings 
+\richel{that is, I predict, a bigger error in settings with
+more co-occurring speciation} more likely to happen: we know that the
+choice of a Bayesian prior is most important in weaker data.
 
-We had to constrain ourselves on the number of replicates.
+The DNA sequence length we used was 1000 base pairs. This is a DNA
+sequence length that we \richel{at least I} feel is informative enough
+for the amount of taxa we simulate. Additionally (although sequencing
+techniques advance at a great pace), this is -to our knowledge- 
+of the same magnitude as the DNA length sequenced in the field.
 
-We had to constrain ourselves on the number of taxa in
-the true species trees.
+Ideally, our ESSes would all be above the recommended value of 
+200 \richel{I predict not all will be above 200, simply due to 
+stochasticity}. Also here, we preferred to have more replicates
+and a minimum median of 200 over less replicates.
+
+We compare our generative to model to \richel{I predict we will} 
+a limited set of candidate models. Also this was for practical
+reasons.
 
@@
 a Yule tree prior.
+
+The effect of using the absolute nLTT tree statistic to quantify the error
+is unknown. A pilot experiment \richel{pirouette example 14} shows that
+using the logarithmically transformed nLTT statistic gives a messier and
+flatter error distribution, but this is not to be taken for granted.
+
+We chose to use few taxa (as we preferred more replicates), but we do not
+know the results would the data be more informative (that is, have more taxa).
+A cool feature of \verb;pirouette; is that it can quantify when this happen;
+when the prior is uninformative due to strong data.

### 2019-09-05 razzo-discussion.tex
Diminishing returns in ESSes, both figure and text (40x longer MCMC for a decreases of 'ESS < 200' of 50%)

@@
 200 \richel{I predict not all will be above 200, simply due to 
-stochasticity}. Also here, we preferred to have more replicates
-and a minimum median of 200 over less replicates.
+stochasticity}. We are aware ESSes can be increased easily by
+making the MCMC chain longer. The shape of the ESS distribution
+(see figure \ref{fig:esses_cumulative}), however, is disadvantageous
+to do so: to reduce the percentage of $ESS < 200$ by fifty percent,
+we would have to increase the MCMC chain lengths by a factor of forty.
+We invested this run-time in doing more replicates.
 

### 2019-10-07 razzo-results.tex
Move figures and their TeX files to the same folder as the other TeX, as using a subfolder does not scale (that is, renders the TeX unusable to be used in a thesis)
Do not '\input' in a file added by '\input', as it scales badly (that is, it renders the TeX unusable for a thesis)
Depend on master branch

--- 2019-06-19 razzo-results.tex
@@
 \section{Results}
-Here we present the results as show in Fig.~\ref{fig:results}...
 
-\input{razzo-figures/fig_results.tex}
+The inference error made for each of the parameter combinations
+is shown in Figures ~\ref{fig:errors_yule} and ~\ref{fig:errors_bd}. 
+For both extinction rates,
+we find that the error increases as $\nu$
+and $q$ increase. 
+Also in line with RJCB's predictions, 
+we find no difference between the two extinction rates.

### 2019-11-25 razzo-methods.tex
Describe how we picked the parameter settings

--- 2019-08-22 razzo-methods.tex
@@
 The first, $\nu$, is the rate at which an environmental change is triggered.
+
 When such event is triggered, each species present in the phylogeny 
@@
 once during a co-occuring speciation event.
+
 A likelihood expression for the process is provided in \cite{mbd}.
+
+\subsection{Parameter settings}
+
+We ran multiple pilot experiments with 1 replicate to arrive at our final
+parameter settings. We devised a set of rules to make a verdict about the
+settings \richel{(script at \url{https://github.com/richelbilderbeek/razzo_project/blob/master/scripts/90_collect_run_times.R},
+resulting verdicts at \url{https://github.com/richelbilderbeek/razzo_project/blob/master/verdict.md})}:
+\itemize{
+  \item quality: 95\% of all individual runs should have an ESS of at least 200.
+  \item feasibility: 95\% of all individual runs should finish within 10 days.
+  \item reproducibility: the mean run-time of all finished runs should be less than 24 hours 
+    \richel{I suggest}.
+  \item relevance: the percentage of taxa created by the MB process should be
+    as high as possible
+}
+
+We searched through parameter space until these criteria were met.
+All the parameter settings used in the pilot experiments can be found at 
+\url{https://github.com/richelbilderbeek/razzo_project/blob/master/overview.md}.
+Due to the low number of replicates, we were unable (nor tried)
+to draw conclusions based on the results.
 

### 2019-11-25 razzo-hypothesis.tex
Add hypothesis section

\section{Hypothesis}

RJCB expects that the inference error is lowest for
parameter settings without multiple-births, for these settings, the
MBD model falls back to being a BD model, which is the standard speciation
model we assume to have generated the tree.

RJCB expects that the inference error is highest for
parameter settings with a higher percentage of multiple-births.
For these settings, the mismatch between the the speciation model that
generated the true tree (which is profoundly-MB MBD) is biggest with
the tree prior assumed to be generative (which is BD).

RJCB expects that an increased extinction rate
has a neutral effect on the error made, for runs with an equal percentage of MBness,
because extinction affects all species equally.

RJCB expects that an increased number of taxa
has a negative effect on the variance of the errors made, as there
is more information available to base inference on.

RJCB expects that an increased extinction rate
increases the variance of the errors made,
due to the decrease of the number of taxa, reducing the amount of information
to base inference on.

RJCB expects that the distribution of nLTT statistics between
true trees and expected twin trees has the same shape
as the difference between the error between the true and twin phylogeny \richel{(see
Figure \ref{fig:nltt_as_proxy})}.
Due to this, the nLTT statistic can act as good shortcut
to arrive at the same conclusion.

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{20191125_nltt_as_proxy.png}
  \caption{
    The nLTT statistic between true and twin tree is strongly related
    to the error between true and twin phylogeny.
  }
  \label{fig:nltt_as_proxy}
\end{figure}

### 2019-11-27 razzo-hypothesis.tex
Improved hypotheses

@@
 
-RJCB expects that the inference error is lowest for
-parameter settings without multiple-births, for these settings, the
-MBD model falls back to being a BD model, which is the standard speciation
-model we assume to have generated the tree.
+RJCB expects that the inference error is lowest when the true
+tree is generated under MBD parameter settings without multiple-births, i.e
+$nu = 0 \vee q = 0$. For such parameters settings, the
+true is in practice generated by a BD model, which
+matches the tree prior used in inference. Without a mismatch between
+true tree prior and assumed tree prior, this source of errors will be absent,
+where the other two sources of errors (stochasticity in the
+simulation of the alignment and the MCMC algorithm) 
+will remain the same.
 
-RJCB expects that the inference error is highest for
-parameter settings with a higher percentage of multiple-births.
-For these settings, the mismatch between the the speciation model that
+RJCB expects that, on average, the inference error is highest for
+parameter settings with a higher percentage of multiple-births,
+i.e. $nu > 0 \wedge q > 0$, as for these settings, 
+the mismatch between the the speciation model that
 generated the true tree (which is profoundly-MB MBD) is biggest with
@@
 
+RJCB expects that the inference error is higher for true trees
+with a higher percentage of species generated in an MB event,
+regardless of the MBD parameters.
+These multiple-born species are one of the three (and the most 
+interesting) sources of error, as a BD model -as a feature- will never
+infer a co-occuring speciation event.
+
 RJCB expects that an increased extinction rate
-has a neutral effect on the error made, for runs with an equal percentage of MBness,
+has a neutral effect on the error made, 
+for runs with an equal percentage of MBness,
 because extinction affects all species equally.
@@
 
-RJCB expects that the distribution of nLTT statistics between
-true trees and expected twin trees has the same shape
-as the difference between the error between the true and twin phylogeny \richel{(see
-Figure \ref{fig:nltt_as_proxy})}.
-Due to this, the nLTT statistic can act as good shortcut
-to arrive at the same conclusion.
+RJCB expects that the nLTT statistic between a true and twin
+tree at the start of the pipeline,
+will correlate strongly with the difference
+between the higest posterior density of the 
+errors of the true and twin tree 
+generated at the end of the pipeline. 
+This hypothesis
+stems from the idea of assuming a setup without noise; that is, with
+a DNA alignment of infinite length. From such a DNA alignment 
+with infinite information, the MCMC
+is able to infer the close-to-correct phylogeny.
+\richel{if this is true, we can use this shortcut (instead of running
+multiple pirouette replicates) to draw stronger conclusions}
 
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{20191125_nltt_as_proxy.png}
+  \includegraphics[width=\textwidth]{20191126_nltt_as_proxy.png}
   \caption{
-    The nLTT statistic between true and twin tree is strongly related
-    to the error between true and twin phylogeny.
+    The nLTT statistic between true and twin tree correlates strongly
+    with the difference between (the highest posterior density
+    of) the errors of the true and twin tree.
   }

### 2019-11-27 razzo-introduction.tex
Updates from Overleaf
Improved hypotheses

@@
 \section{Introduction}
-
-\giovanni{Introduce "BD" and "MBD" acronyms at some point.}
 
@@
 BEAST2 (\cite{beast2}) are widely used tools to perform such task. They 
-return posterior distributions of phylogenies and estimated parameters by running a Bayesian analysis, given genetic data and tree priors.
-A tree prior is a mathematical description of the a priori characteristics that we expect to be reflected in posterior phylogenies.
-The choice of a specific prior is, by definition, arbitrary but the consequences of such choice can be vet analyzing the so-obtained posteriors.
+return posterior distributions of phylogenies 
+and estimated parameters by running a Bayesian analysis, 
+given genetic data and tree priors.
 
-BEAST2 gives to the user the option to set up several possible phylogenetic priors (e.g.
-substitution/clock/speciation models).
-However, currently available priors might be not suitable to analyze some specific datasets.
-For this reason BEAST2 provides users with the possibility to introduce new tree priors, to infer phylogenies based on different assumptions on how the speciation process takes place.
+A tree prior is a mathematical description 
+of the a priori characteristics 
+that we expect to be reflected in posterior phylogenies. 
+The choice of a specific prior is, by definition, 
+arbitrary but the consequences of such choice 
+can be vet analyzing the so-obtained posteriors.
 
-Current phylogenetic tools assume that only a single speciation event can occur at any given time.
-While this assumption has been proved to be useful to construct a wide variety of successful models (e.g \cite{Maddison2007biSSE}, \cite{Valente2015}, \cite{etienne2012diversity}, \cite{etienne2014estimating}), they do not take into account the possibility for environmentally-driven contemporary speciations on multiple lineages.
+BEAST2 gives to the user the option to set up 
+several possible phylogenetic 
+priors (e.g. substitution/clock/speciation models). 
+However, currently available priors 
+might be not suitable to analyze some specific datasets.
+For this reason BEAST2 provides users with the possibility 
+to introduce new tree priors, 
+to infer phylogenies based on different assumptions 
+on how the speciation process takes place.
+
+Current phylogenetic tools assume that 
+only a single speciation event can occur at any given time.
+While this assumption has been proved to be useful 
+to construct a wide variety of successful 
+models (e.g \cite{Maddison2007biSSE}, \cite{Valente2015}, 
+\cite{etienne2012diversity}, \cite{etienne2014estimating}), 
+they do not take into account the possibility 
+for environmentally-driven contemporary speciations on multiple lineages.
 

### 2019-11-27 fig_results.tex
Moved old pics to pics, use new pretty one-mu-per-figure figures
Rename duplicate figure labels
Add 10 replicate figures
Diminishing returns in ESSes, both figure and text (40x longer MCMC for a decreases of 'ESS < 200' of 50%)
Add figures

@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{razzo_results.png}
+  \includegraphics[width=\textwidth]{figure_1a.png}
   \caption{
-    The inference error distribution 
-    for the different biological
-    parameter settings. In all cases, $\lambda = 0.2$ and 
-    crown age equals 6. 
+    The error distribution for a extinction rate $\mu = 0.0$.
   }
-  \label{fig:results}
+  \label{fig:errors_yule}
 \end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{figure_1b.png}
+  \caption{
+    The error distribution for extinction rate $\mu = 0.15$.
+  }
+  \label{fig:errors_bd}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{figure_2.png}
+  \caption{
+    The relation between percentage of observable MBness and
+    the inference error. Each point corresponds to the error from
+    one posterior tree.
+  }
+  \label{fig:mbness_and_error}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{figure_3.png}
+  \caption{
+    The relation between the number of taxa and the
+    variance of the inference error. 
+    Each point corresponds to the variance of one posterior.
+    \richel{move to supplementary materials?}
+  }
+  \label{fig:n_taxa_and_error_variance}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{figure_4.png}
+  \caption{
+    The relation between the extinction rate and the
+    variance of the inference error. 
+    \richel{move to supplementary materials?}
+  }
+  \label{fig:ext_rate_and_error_variance}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{figure_5.png}
+  \caption{
+    The relation between the nLTT statistic and the difference
+    between the HPD of the true and twin trees' errors. 
+    Each point corresponds to the difference of 
+    the HPD error between true and twin posterior.
+    The grey line is the identity line.
+  }
+  \label{fig:nltt_and_error_difference}
+\end{figure}

### 2019-11-27 razzo-hypotheses.tex
Renamed section to 'hypotheses'
Add figures

--- 2019-11-27 razzo-hypothesis.tex
@@
-\section{Hypothesis}
+\section{Hypotheses}
 
-RJCB expects that the inference error is lowest when the true
+$\mathcal{H}_1$: RJCB expects that the inference error is lowest when the true
 tree is generated under MBD parameter settings without multiple-births, i.e
@@
 simulation of the alignment and the MCMC algorithm) 
-will remain the same.
+will remain the same. \richel{figure 1a and b}
 
-RJCB expects that, on average, the inference error is highest for
+$\mathcal{H}_2$: RJCB expects that, on average, the inference error is highest for
 parameter settings with a higher percentage of multiple-births,
@@
 generated the true tree (which is profoundly-MB MBD) is biggest with
-the tree prior assumed to be generative (which is BD).
+the tree prior assumed to be generative (which is BD). \richel{figure 1a and b}
 
-RJCB expects that the inference error is higher for true trees
-with a higher percentage of species generated in an MB event,
+$\mathcal{H}_3$: RJCB expects that the inference error is higher for true trees
+with a higher percentage of species observably generated in an MB event,
 regardless of the MBD parameters.
@@
 interesting) sources of error, as a BD model -as a feature- will never
-infer a co-occuring speciation event.
+infer a co-occuring speciation event. \richel{figure 2}
 
-RJCB expects that an increased extinction rate
-has a neutral effect on the error made, 
-for runs with an equal percentage of MBness,
-because extinction affects all species equally.
+$\mathcal{H}_4$: RJCB expects that, on average, the inference error is equal for
+parameter settings with an equal $\lambda$, $\nu$ and $q$, i.e. extinction
+has no effect on the error made, because extinction affects all species equally. 
+\richel{figure 1a and b, or a new figure}
 
-RJCB expects that an increased number of taxa
+$\mathcal{H}_5$: RJCB expects that an increased number of taxa
 has a negative effect on the variance of the errors made, as there
 is more information available to base inference on.
+\richel{figure 3}
 
-RJCB expects that an increased extinction rate
+$\mathcal{H}_6$: RJCB expects that an increased extinction rate
 increases the variance of the errors made,
@@
 to base inference on.
+\richel{figure 4}
 
-RJCB expects that the nLTT statistic between a true and twin
+$\mathcal{H}_7$: RJCB expects that the nLTT statistic between a true and twin
 tree at the start of the pipeline,
 will correlate strongly with the difference
-between the higest posterior density of the 
+between the higest posterior density (HPD) of the 
 errors of the true and twin tree 
@@
 multiple pirouette replicates) to draw stronger conclusions}
+\richel{figure 5}
 
@@
 \end{figure}
+
+$\mathcal{H}_8$: RJCB expects that the JC69 and strict clock 
+candidate models, the close

### 2019-11-27 razzo-methods.tex
Add figures

@@
 
-We proceed in the following way: we build simulated datasets generated under the multiple birth model.
-Then we run a \verb;pirouette; analysis, which will lead to error distributions between the inference posterior and the original simulated trees.
-Importantly, the \verb;pirouette; analysis includes also a 'twin' parallel pipeline, which will provide a measure of the baseline error due to pure stochasticity, unavoidably occurring in the process.
+We proceed in the following way:
+under a range a parameter setting, 
+we build simulated datasets generated under the multiple birth model. 
+Then we run a \verb;pirouette; analysis, 
+which will lead to error distributions
+between the inference posterior and the original simulated trees. 
+Importantly, the \verb;pirouette; analysis includes 
+also a 'twin' parallel pipeline, 
+which will provide a measure of the baseline error 
+due to pure stochasticity, unavoidably occurring in the process.
+We then analyse the complete and baseline errors.
 
@@
 For each parameter setting, 
-we generate $2$ \richel{Will need to calculate how many replicates we can do} 
+we generate [as much as time permits] 
+\richel{Will need to calculate how many replicates we can do} 
 independent reconstructed trees of the same crown age.

### 2019-11-27 razzo-results.tex
Add figures

@@
 is shown in Figures ~\ref{fig:errors_yule} and ~\ref{fig:errors_bd}.
-For both extinction rates,
+
+\richel{if true:}The lowest errors are at settings where $nu = 0 \vee q = 0$,
+confirming $\mathcal{H}_1$.
+
+\richel{if true:}For both extinction rates,
 we find that the error increases as $\nu$
-and $q$ increase. 
-Also in line with RJCB's predictions, 
-we find no difference between the two extinction rates.
+and $q$ increase, confirming $\mathcal{H}_2$
+
+\richel{if true:}There is no difference between the
+errors for the two different extinction rates, which
+confirms $\mathcal{H}_4$
+
+[other hypotheses being rejected/accepted here]

### 2019-11-28 razzo-hypotheses.tex
Hypothesis (of no biological relevance) regarding candidate models

@@
 
-$\mathcal{H}_8$: RJCB expects that the JC69 and strict clock 
-candidate models, the close
+$\mathcal{H}_8$: RJCB expects that for settings without extinctions,
+the candidate model with JC69, strict and Yule will come up as the best
+model most, as that model is the generative model, where the HKY, strict, BD
+will be seleted least, as both its site and tree model are needlessly complex.
+For settings with extinctions, RJCB predict HKY, strict and Yule will come up 
+as the worst model mostly, as it uses both an incorrect site model and incorrect
+tree model. \richel{I would enjoy a hypothesis of actual biological relevance for the
+candidate models. If we cannot devise one, I suggest to remove the
+use of candidate models}

### 2019-12-09 razzo-hypotheses.tex
Updates from Overleaf

@@
 \richel{figure 1a and b, or a new figure}
+\giovanni{A higher extinction rate increases variance and tend to reduce to number of taxa in the tree.
+According to your other hypotheses (e.g.
+H5) these effects do affect the inference.
+Therefore you need to adjust for these variables to prove your hypothesis.}
 
@@
 will correlate strongly with the difference
-between the higest posterior density (HPD) of the 
+between the highest posterior density (HPD) of the 
 errors of the true and twin tree 
@@
 multiple pirouette replicates) to draw stronger conclusions}
+\giovanni{I can agree with the principle.
+However there are also other factors that you might overlook with such approximation.
+It might be useful, though, for a fast comparison to have a preliminary idea, if the hypothesys will be confirmed by the experiment.}
 \richel{figure 5}
@@
 use of candidate models}
+
+$\mathcal{H}_9$: GL expects a positive correlation between $\nu \, q(1-q)$ and the error made by the inference process.
+This is due to the fact that this factor regulates the variance of the underlying binomial distribution that regulates multiple births events.

### 2020-01-23 razzo-hypotheses.tex

@@
 \richel{figure 1a and b, or a new figure}
-\giovanni{A higher extinction rate increases variance and tend to reduce to number of taxa in the tree.
+\giovanni{A higher extinction rate increases variance and tends to reduce to number of taxa in the tree.
 According to your other hypotheses (e.g.
@@
 $\mathcal{H}_9$: GL expects a positive correlation between $\nu \, q(1-q)$ and the error made by the inference process.
-This is due to the fact that this factor regulates the variance of the underlying binomial distribution that regulates multiple births events.
+This is due to the fact that this factor expresses the variance of the underlying binomial distribution that regulates multiple births events.

### 2020-02-06 article.tex
Do not use '\input' in file that is already added by an '\input' as it does not scale (that is, this TeX cannot be used as part of a thesis)
Do not '\input' in a file added by '\input', as it scales badly (that is, it renders the TeX unusable for a thesis)
'speciation completion' are two words
Add hypothesis section
Renamed section to 'hypotheses'
Changes Giappo

@@
 \newcommand*\richel[1]{\textcolor{orange}{\textbf{[RJCB: #1]}}}
-\newcommand*\gio[1]{\textcolor{blue}{\textbf{[GL: #1]}}}
+\newcommand*\giovanni[1]{\textcolor{blue}{\textbf{[GL: #1]}}}
 
@@
   cre-ate
-  spe-ci-a-tion-com-ple-tion
+  spe-ci-a-tion
+  com-ple-tion
   pro-trac-ted
@@
 \input{razzo-introduction.tex}
-
+\iffalse % probably not present in the final version
+\input{razzo-hypotheses.tex}
+\fi
 \input{razzo-methods.tex}
@@
 \input{razzo-results.tex}
+
+\input{fig_results.tex}
+
+\input{razzo-discussion.tex}
 

### 2020-02-06 razzo-abstract.tex
Changes Giappo

@@
 % comprehensible to a scientist in any discipline.
-There exist millions of species on Earth,
-all originating from a common ancestor billions
+Millions of species exist on Earth, all originating from a common ancestor billions
 of years ago.
-The field of phylogenetics uses heritable material (e.g. DNA)
-to determine the evolutionary history of species.
+The field of phylogenetics uses heritable material (e.g. DNA) to determine the evolutionary history of species.
 
@@
 % scientists in related disciplines.
-Starting from heritable material and explicit assumptions,
-Bayesian phylogenetics allows to infer a jointly-estimated 
-phylogeny and parameter estimates distribution.
-One of these assumptions is the speciation model, which
-mathematically describes the branching process of a phylogeny in time.
+Starting from heritable material and explicit assumptions, Bayesian phylogenetics allows to infer a jointly-estimated phylogeny and parameter estimates distribution.
+One of these assumptions is the speciation model, which mathematically describes how species are added or removed from a phylogeny in time.
 Speciation models commonly assume that speciation events are independent and disjointed.
@@
 % previous knowledge.
-We show that the extent of the inference error grows with the amount of co-occuring speciation events to establish the limits of BEAST2's standard tree priors.
+We show that the extent of the inference error increases with the amount of co-occuring speciation events to establish the limits of BEAST2's standard tree priors in this circumstance.
 
@@
 % more general context.
-These results allow phylogeneticists to judge under which circumstances the commonly used speciation model can be safely used.
+These results allow phylogeneticists to judge under which circumstances the commonly used speciation model can be safely used. \giovanni{Not entirely sure if this is the best sentence to put here.}
 

### 2020-02-06 razzo-discussion.tex
Changes Giappo

@@
 We \richel{(at least I. If you do not expect this, let me know the 
-reasoning why)} expected to find a bigger inference error 
+reasoning why)} \giovanni{I expect the same thing.} expected to find a bigger inference error 
 when increasing the amount of co-occurring speciation events.
@@
 
+\giovanni{Further conclusions are to be drawn after results are produced.}
+
+\iffalse % A model can always be made more accurate.
+There are two principles to follow when building one: accuracy, on one hand, is surely important but, at the same time, simplicity is important too as overcomplicated models miss the point of a research question.
+This is, for example, summarized in the request of a low AIC, which weights positively the accuracy of the model (high likelihood) and negatively its complexity (for which the number of parameters is a proxy).
 \subsection{Research}
@@
 \subsection{Methodological choices}
+\fi
 
+\iffalse % I choose to remove this too, at least for now, as it's too soon to write any conclusion when no result is available.
 When simulating a DNA alignment, we chose to use the simplest (JC69) site
@@
 we were able to perform more replicates. For a process with this high
-amount of stochaticity, we think this is highly favored. The second reason
+amount of stochasticity, we think this is highly favored. The second reason
 is that we think this reduces the noise of our results, without 
@@
 that the Yule tree prior will be selected as the best candidate model
-more often in this scenario, which shows the sensitity of the method we used.
+more often in this scenario, which shows the sensitivity of the method we used.
 
@@
 when the prior is uninformative due to strong data.
+\fi

### 2020-02-06 razzo-introduction.tex
Changes Giappo

@@
 currently absent in BEAST2. 
-The multiple birth hypothesis aims to describe how 
-large-scale environmental changes can lead phylogenies 
-with a high abundance of speciation events in relatively short times.
+The multiple birth (MB) hypothesis aims to describe how large-scale environmental changes can lead phylogenies with a high abundance of speciation events in relatively short times.
 

### 2020-02-06 razzo-methods.tex
Complete supplementary materials
Changes Giappo

@@
 
-We proceed in the following way:
-under a range a parameter setting, 
-we build simulated datasets generated under the multiple birth model. 
-Then we run a \verb;pirouette; analysis, 
-which will lead to error distributions
-between the inference posterior and the original simulated trees. 
-Importantly, the \verb;pirouette; analysis includes 
-also a 'twin' parallel pipeline, 
-which will provide a measure of the baseline error 
-due to pure stochasticity, unavoidably occurring in the process.
+We proceed in the following way: under a range of parameter setting, we build simulated datasets generated under the multiple birth model. 
+Then we run a \verb;pirouette; analysis, which will lead to error distributions between the inferred posterior and the original simulated trees, for each setting. 
+Importantly, the \verb;pirouette; analysis includes also a 'twin' parallel pipeline, which will provide a measure of the baseline error due to pure stochasticity, unavoidably occurring in the process.
 We then analyse the complete and baseline errors.
+\subsection{The Framework}
 
-\subsection{Model}
+\subsubsection{Simulation Model}
 
 In the MBD model, parameters $\lambda$ and $\mu$ correspond, respectively, 
-to the common per-species speciation and extinction rates 
+to the traditional per-species speciation and extinction rates 
 already present in the standard BD model. 
@@
 In fact, whereas parameter $\lambda$ can be seen as 
-describing a sympatric process, $\nu$ induces the rise of 
-geographical barriers that interrupt the gene flow,
-resulting in allopatric speciation. 
+describing a sympatric process, $\nu$ induces instead the rise of 
+geographical barriers able to interrupt the gene flow in the population,
+resulting in a possible allopatric speciation for each of the species.  
 Even though multiple speciations can co-occur, 
@@
 
+\subsubsection{Estimating the inference error}
+
+From each simulated 'true' MBD tree, we measure the impact of
+ignoring the more complex and non-standard MBD tree prior in
+Bayesian phylogenetic inference.
+We do so by assuming a simpler and standard BD tree prior,
+and compare the 'true' tree with the inferred trees.
+To do so, we use the \verb;pirouette; R package [\citep{pirouette}].
+
+Here we summarize briefly how \verb;pirouette; works: it starts from a 'true' starting phylogeny (in our case: the simulated MBD tree), from which a DNA sequence alignment is simulated. 
+From each sequence alignment, a Bayesian inference is run (in our case: with assuming a BD tree prior), to obtain a posterior distribution of jointly-estimated trees and model parameter estimates.
+By comparing the true tree and the posterior trees, an inference error distribution is generated.
+We use the twinning option available in \verb;pirouette; that allows to quantify the impact of assuming a wrong tree prior in a Bayesian inference compared to a reference background error.
+
 \subsection{Parameter settings}
 
+\iffalse
 We ran multiple pilot experiments with 1 replicate to arrive at our final
@@
   \item quality: 95\% of all individual runs should have an ESS of at least 200.
+    \richel{figure \ref{fig:esses}}
   \item feasibility: 95\% of all individual runs should finish within 10 days.
+    \richel{figure \ref{fig:runtimes}}
   \item reproducibility: the mean run-time of all finished runs should be less than 24 hours 
     \richel{I suggest}.
-  \item relevance: the percentage of taxa created by the MB process should be
+    \richel{figure \ref{fig:runtimes}}
+  \item relevance 1: the percentage of taxa created by the MB process should be
+    as high as possible
+  \item relevance 2: the percentage of taxa should be
     as high as possible
 }
-
 We searched through parameter space until these criteria were met.
@@
 to draw conclusions based on the results.
+\fi
 
-\subsection{Tree simulations}
+When working on simulated datasets the choice of the generating parameters is always an arbitrary choice.
+In principle one would choose a parameter range as wide as possible to characterize the most of the available parameter space.
+However, there are technical aspects that must be taken in consideration to be sure that both the simulation and the inference processes work out in an optimal way.
+We required the following criteria to be satisfied \giovanni{Does this apply only to the mbd parameters or also to the pirouette ones?}:
+\itemize{
+  \item Quality: At least 95\% of all individual runs should have an Estimated Sample Size (ESS) of at least 200 (see \cite{beast} for reference);
+    \richel{figure \ref{fig:esses}}
+  \item Feasibility: As time is a limited resource we impose that at least 95\% of all individual runs should finish within 10 days;
+    \richel{figure \ref{fig:runtimes}}
+  \item Reproducibility: the mean run-time of all finished runs should be less than 24 hours \richel{figure \ref{fig:runtimes}}
 
-We simulate the speciation process in continuous time 
-using the \verb;mbd_sim; function from the \verb;mbd; R package (\citep{mbd}).
-We let parameters vary using all possible combinations of values 
-as shown in Table~\ref{tab:simulation_parameters}.
-For each parameter setting, 
-we generate [as much as time permits] 
+    \richel{I suggest}. \giovanni{What's the difference with the previous point?}
+  \item Relevance 1: In order to have a meaningful inference, we demand that the MB signal is detectable to a satisfactory extent.
+For this reason the percentage of taxa created by the MB process should be as high as possible;
+  \item Relevance 2: Small trees might not contain a significant amount of information to be analyzed.
+For this reason we demand that the number of taxa should be as high as possible, as long as all the other requirements are met;
+}
+
+Additional information about the choice of a suitable parameter setting can be found in the Supplementary Material.
+\iffalse
+\giovanni{This could be put in the appendix, if we manage to put in a handy format.}
+We searched through parameter space until these criteria were met.
+All the parameter settings used in the pilot experiments can be found at 
+\url{https://github.com/richelbilderbeek/razzo_project/blob/master/overview.md}.
+Due to the low number of replicates, we were unable (nor tried)
+to draw conclusions based on the results.
+\fi
+
+\subsubsection{Simulations setting}
+
+We simulate the speciation process in continuous time using the \verb;mbd_sim; function from the \verb;mbd; R package [\citep{mbd}].
+We let parameters vary using all possible combinations of values as shown in Table~\ref{tab:simulation_parameters}.
+For each parameter setting, we generate [as much as time permits] 
 \richel{Will need to calculate how many replicates we can do} 
-independent reconstructed trees of the same crown age.
+independent trees of the same crown age.
+Note that trees that are actually passed as \verb;pirouette; input are not full trees, but reconstructed trees instead.
+A reconstructed tree is a tree where all the species that went extinct before the present time are removed, as this information is usually unavailable on real data phylogenies.
 
-\subsection{Inference error estimation}
+\input{tab_simparameters.tex}
 
-From each simulated 'true' MBD tree, we measure the impact of
-ignoring the more complex and non-standard MBD tree prior in
-Bayesian phylogentic inference.
-We do so by assuming a simpler and standard BD tree prior,
-and compare the 'true' tree with the inferred trees.
-To do so, we use the \verb;pirouette; R package \citep{pirouette}.
+\subsubsection{pirouette setting}
 
-In brief: \verb;pirouette; starts from 
-a 'true' starting phylogeny (in our case: the simulated MBD tree), 
-from which a DNA sequence alignment is simulated. 
-From each sequence alignment, a Bayesian 
-inference is run (in our case: with assuming a BD tree prior), 
-to obtain a posterior distribution of jointly-estimated trees and model parameter estimates.
-By comparing the true tree and the posterior trees, 
-an inference error distribution is generated.
-We use the twinning option available in \verb;pirouette;
-that allows to quantify the impact of assuming a wrong
-tree prior in a Bayesian inference.
+In our experiment, the alignments are 1000 nucleotides in length, with a known root sequence of four 250 mono-nucleotide blocks, generated using the simplest nucleotide substitution model (Jukes Cantor, JC69) and clock model (strict), with a mutation rate of $\frac{t_c}{2}$, where $t_c$ is the crown age.
+With this mutation rate, each nucleotide has a 50\% chance to mutate (both silently and non-silently) from the ancestral root sequence to any of the contemporary species' sequences at the tips.
 
-In our experiment, the alignments are 1000 nucleotides in length, 
-with a known root sequence of four 250 mono-nucleotide blocks, 
-generated using the simplest nucleotide substitution model (JC69) 
-and clock model (strict), with a mutation rate of $\frac{1}{2}\cdot c$,
-in which $c$ is the crown age. 
-With this mutation rate, each nucleotide has a 50\% chance
-to mutate (both silently and non-silently) from the ancestral root sequence 
-to any of the contemporary species' sequences at the tips.
+For the Bayesian inference, we assume a generative model of a JC69 site model, a strict clock model and a BD tree prior.
+Additionally, we use a Most Recent Common Ancestor (MRCA) prior equal to the crown age with a normal distribution of width $\sigma = 0.01$.
+We pick a Markov Chain Monte Carlo (MCMC) setup of $10^6$ states with a sampling interval of once per $10^3$ states.
+Of the resulting $10^3$ states, we discard a burn-in of $10\%$.
 
-For the Bayesian inference, we assume a generative model of 
-a JC69 site model, a strict clock model and a BD tree prior. 
-Additionally, we use an MRCA prior equal to the crown age with a normal distribution 
-of width $\sigma = 0.01$. We pick an MCMC setup of $10^6$ states
-with a sampling interval of once per $10^3$ states. Of the resulting
-$10^3$ states, we discard a burn-in of $10\%$.
+To verify that the generative model (with a BD tree prior) is indeed the most appropriate standard model, we compare it to 39 other candidate inference models.
+These candidate models are all other combinations of four tree priors, two clock models and five speciation models, excluding the one combination used as a generative model.
+We let the best candidate model also generate an error distribution, which is expected to have a similar or high median.
+For models comparison, \verb;pirouette; uses a nested sampling approach to estimate marginal likelihoods, as described in \citet{russel2019model}, using the "\verb;NS;" BEAST2 package.
+We setup the MCMC for the nested sampling using its default values, which are a relative error $\epsilon$ of $10^{-12}$,
+1 particle, a sub-chain sampling interval of $5 \cdot 10^3$ states with a maximum chain length of $10^6$ states.
 
-To verify that the generative model (with a BD tree prior) is
-indeed the most appropriate standard model, we compare it to
-39 other candidate inference models. These candidate models 
-are all other combinations of four tree priors, 
-two clock models and five speciation models, 
-excluding the one combination used as a generative model.
-We let the best candidate model also generate an error distribution,
-which is expected to have a similar or high median.
-For models comparison, \verb;pirouette; uses a nested sampling approach
-to estimate marginal likelihoods,
-as described in [\cite{maturana2018model}], 
-using the \verb;NS; BEAST2 package. We setup the MCMC for the nested sampling
-using its default values, which are a relative error $\epsilon$ of $10^-12$, 
-1 particle, a sub-chain sampling interval of $5.10^3$ states with
-a maximum chain length of $10^6$ states.
-
-To compare the true tree with any of the posterior trees,
-we use the absolute nLTT statistic \citet{janzen2015},
-which will result in an error distribution with values
-ranging from zero (when the inferred tree is identical 
-to the true tree) to a maximum of one (trees are completely different).
+To compare the true tree with any of the posterior trees, we use the absolute nLTT statistic by \citet{janzen2015}, which will result in an error distribution with values ranging from zero (when the inferred tree is identical to the true tree) to a maximum of one (trees are completely different).
 
@@
 model of a JC69 site model, strict clock model and a BD tree prior.
+
+\giovanni{Add mention to Jensen-Shannon divergence metric, if used.}

### 2020-03-24 razzo-abstract.tex
Feedback rsetienne

@@
 % comprehensible to a scientist in any discipline.
-Millions of species exist on Earth, all originating from a common ancestor billions
-of years ago.
-The field of phylogenetics uses heritable material (e.g. DNA) to determine the evolutionary history of species.
-
+The field of phylogenetics uses heritable material such as DNA to determine the (shared) evolutionary history of a set of species which are summarized in a phylogenetic tree.
 % Two to three sentences of
@@
 % scientists in related disciplines.
-Starting from heritable material and explicit assumptions, Bayesian phylogenetics allows to infer a jointly-estimated phylogeny and parameter estimates distribution.
-One of these assumptions is the speciation model, which mathematically describes how species are added or removed from a phylogeny in time.
-Speciation models commonly assume that speciation events are independent and disjointed.
-Yet, such assumption may overlook the complexity of certain processes, where environmental changes promote speciation events in multiple lineages.
+Bayesian phylogenetic methods allow us to jointly infer probability distributions for the phylogenetic tree and parameters of the various models underlying the methods.
+One of these models is the diversification model, which mathematically describes the dynamics of speciation addition and removal through time.
+In Bayesian analyses these are called the (species) tree priors.
+Tree priors commonly assume that speciation events occur independently.
+The Bayesian tools heavily rely on this assumption.
+However, under species pump dynamics, this assumption is violated.
+By species pump dynamics we mean the (repeated) simultaneous formation of multiple species due to environment-driven isolation which results in temporally aligned (or clustered) divergence times.
 % Purpose of the paper
-This new layer of complexity can be captured developing a novel ad hoc speciation model.
-However, the introduction of a new model could be not necessary if current models are already capable of describing the process to a satisfactory extent, under simpler assumptions.
-Here we investigate the extent of the discrepancy produced by current BEAST2's tree priors, when trying to infer back trees where speciation can co-occur.
+Current Bayesian phylogenetic tools do not contain such species pump diversification models as tree priors.
+This may not be a problem if currently implemented tree priors are already capable of inferring a phylogenetic tree to a satisfactory extent.
+Here we investigate the extent of the error made by one such Bayesian phylogenetic tool (BEAST2) when inferring a phylogenetic tree generated by a known species pump diversification model with a standard tree prior.
 
 % Summarising the main result
-We let the impact of co-occurring speciation vary on datasets of simulated trees and show the corresponding error produced during the inference process.
+To this end we simulate our species pump model, which we call the multiple birth model because it produces multiple simultaneous speciation events, under various parameter settings, and evaluate the corresponding error produced during the inference process.
+We compare this error with the error made when the generating model is the same as the tree prior used in inference.
 
@@
 % previous knowledge.
-We show that the extent of the inference error increases with the amount of co-occuring speciation events to establish the limits of BEAST2's standard tree priors in this circumstance.
+We show that the extent of the inference error increases with the number of multiple birth events, as expected.
+\rampal{We need a bit more reults here}
 
@@
 % more general context.
-These results allow phylogeneticists to judge under which circumstances the commonly used speciation model can be safely used. \giovanni{Not entirely sure if this is the best sentence to put here.}
+These results allow phylogeneticists to judge under which circumstances the standard tree priors can be safely used, and under what circumstances phylogenetic inferences are unreliable.
+\giovanni{Not entirely sure if this is the best sentence to put here.}
+\rampal{Would be nice if we gave some guidelines here}
 

### 2020-03-24 razzo-introduction.tex
Feedback rsetienne

--- 2019-11-27 razzo-introduction.tex
@@
 
-Modern computational techniques allow to infer phylogenies from genetic data 
-such as DNA, RNA or proteins. BEAST (\cite{beast}) and its descendant 
-BEAST2 (\cite{beast2}) are widely used tools to perform such task. They 
+Modern computational techniques, such as BEAST (\cite{beast}) and its descendant 
+BEAST2 (\cite{beast2}), allow to infer phylogenetic trees from genetic data 
+such as DNA, RNA or proteins. \rampal{mention MrBayes and revBayes as well}. They 
 return posterior distributions of phylogenies 
 and estimated parameters by running a Bayesian analysis, 
-given genetic data and tree priors.
-
-A tree prior is a mathematical description 
-of the a priori characteristics 
-that we expect to be reflected in posterior phylogenies. 
-The choice of a specific prior is, by definition, 
-arbitrary but the consequences of such choice 
-can be vet analyzing the so-obtained posteriors.
-
-BEAST2 gives to the user the option to set up 
+given aligned sequence data and a set of models. 
+One of these models is the diversification model, for which a prior distribution must be provided. This is called a tree prior; it is a mathematical description 
+of the probability distribution of possible branching patterns before looking at the data.
+Together with the signal from the data, this tree prior will determine the posterior distribution of phylogenies, i.e.
+after considering the data.
+Other models include the nucleotide substitution model (i.e.
+a model of relative transition rates between different nucleotides through time) and the clock model (a model determining the absolute rate of changes for each lineage).
+For each of these models choices must be made and prior distributions must be specified for their parameters.
+BEAST2 gives the user the option to set up
 several possible phylogenetic 
-priors (e.g. substitution/clock/speciation models). 
+priors (e.g. substitution/clock/diversification models). 
 However, currently available priors 
@@
 For this reason BEAST2 provides users with the possibility 
-to introduce new tree priors, 
-to infer phylogenies based on different assumptions 
-on how the speciation process takes place.
+to introduce new models and corresponding priors. Particularly, one can specify the tree prior for a new model of diversification.
 
-Current phylogenetic tools assume that 
+Current phylogenetic tools such as BEAST2 assume that 
 only a single speciation event can occur at any given time.
-While this assumption has been proved to be useful 
-to construct a wide variety of successful 
-models (e.g \cite{Maddison2007biSSE}, \cite{Valente2015}, 
-\cite{etienne2012diversity}, \cite{etienne2014estimating}), 
-they do not take into account the possibility 
-for environmentally-driven contemporary speciations on multiple lineages.
+This assumption is consistent with many different diversification models (e.g \cite{Maddison2007biSSE}, \cite{Valente2015}, 
+\cite{etienne2012diversity}, \cite{etienne2014estimating}).
+However, multiple speciation events can take place simultaneously and repeatedly when populations are intermittently disconnected and connected, for example due to climatic fluctuations.
+This has been called the species pump hypothesis.\rampal{Cite: Haffer, J.
+(1969).
+Speciation in Amazonian forest birds.
+Science, 165,
+131–137.
+https://doi.org/10.1126/science.165.3889.131.} and has been invoked particularly in mountainous areas that underwent glaciation.
+\rampal{ Origins of global mountain plant biodiversity: Testing the 'mountain-geobiodiversity hypothesis'
+By: Muellner-Riehl, Alexandra N.; Schnitzler, Jan; Kissling, W. Daniel; et al.
+JOURNAL OF BIOGEOGRAPHY}.
+Our own interest in the species pump hypotheses arose from its potential explanation of the radiation of cichlid fish in the African rift lakes (Malawi, Tanganyika and Victoria), where water level drops created multiple smaller lakes providing the opportunity for allopatric speciation in multiple species.
+(\citep{verheyen1996mitochondrial}, \citep{sturmbauer2001lake}, \citep{janzen2016}, \citep{janzen2017}).
 
-We explore such case introducing the multiple-birth death model (MBD), 
-currently absent in BEAST2. 
-The multiple birth hypothesis aims to describe how 
-large-scale environmental changes can lead phylogenies 
-with a high abundance of speciation events in relatively short times.
-
-\iffalse
-\giovanni{This might be good for the abstract}
-The (constant-rate) birth-death (BD) model embodies the common assumption 
-that only a single speciation event can occur at any given time. 
-The multiple-birth-death (MBD) model relaxes this assumption allowing, 
-in addition to standard BD events, 
-also events in which large-scale environmental changes lead 
-to speciation bursts.
-\fi
-
-Such hypothesis can be useful to better understand the history of systems of particular interests for evolutionary biologists, such as the diversification of cichlid fish in the African Great Lakes (Malawi, Tanganyika and Victoria), where water level fluctuations are thought to play an important role in promoting diversification (\citep{verheyen1996mitochondrial}, \citep{sturmbauer2001lake}, \citep{janzen2016}, \citep{janzen2017}).
-
-However, the introduction of new tree priors is not always desirable (\citep{bilderbeek2019pirouette}).
-Current standard birth death (BD) tree priors might, in principle, prove to be good enough at detecting such events despite their lower level of complexity.
-If this is the case one should always choose to adopt the simplest model.
-
-We used the R package \verb;pirouette; (\citep{pirouette}) to perform such test, starting on phylogenies simulated under the MBD regime using the \verb;mbd; package (\citep{mbd}).
-From such phylogenies we measure the inference error made adopting a standard BD tree prior in the inference process.
-
-With this work we aim, using such inference error distributions, to test whether or not it is advantageous to implement a new prior model that can allow the construction of trees where multiple speciations can co-occur at the same time.
+One could study whether the species pump hypothesis is a viable explanation in empirical ystems by comparing divergence times of sister taxa \rampal{Oaks et al - doi:10.1111/evo.13754}.
+A more inclusive approach would involve using a model allowing multiple simultaneous speciation events as a new species tree prior in phylogenetic reconstruction.
+However, introducing a new tree prior may be computationally prohibitive (\citep{bilderbeek2019pirouette}), and may also not be necessary, as current standard birth death (BD) tree priors might prove to be good enough at inferring the correct tree.
+Here we use the R package \verb;pirouette; (\citep{pirouette}) to check whether this is the case by simulating phylogenies under a species pump model, i.e.
+the multiple-birth-death model (MBD), with the \verb;mbd; package (\citep{mbd}), then simulating sequence alignments for each of these trees and finally inferring a phylogenetic tree using BEAST2 from these alignment.
+By comparing the inferred phylogeny with the true (simulated) one, we measure the inference error made by adopting a standard BD tree prior.

### 2020-03-24 razzo-methods.tex
Feedback rsetienne

@@
 
-We proceed in the following way: under a range of parameter setting, we build simulated datasets generated under the multiple birth model. 
-Then we run a \verb;pirouette; analysis, which will lead to error distributions between the inferred posterior and the original simulated trees, for each setting. 
-Importantly, the \verb;pirouette; analysis includes also a 'twin' parallel pipeline, which will provide a measure of the baseline error due to pure stochasticity, unavoidably occurring in the process.
-We then analyse the complete and baseline errors.
-\subsection{The Framework}
+\subsection{Simulation model}
 
-\subsubsection{Simulation Model}
-
-In the MBD model, parameters $\lambda$ and $\mu$ correspond, respectively, 
-to the traditional per-species speciation and extinction rates 
-already present in the standard BD model. 
-Additionally, MBD relies on two additional parameters, $\nu$ and $q$. 
-The first, $\nu$, is the rate at which an environmental change is triggered.
-
-When such event is triggered, each species present in the phylogeny 
-at that moment has a probability $q$ to speciate at that time.
-This kind of speciation is of a different nature with respect to 
-the one triggered by $\lambda$. 
-In fact, whereas parameter $\lambda$ can be seen as 
-describing a sympatric process, $\nu$ induces instead the rise of 
+The multiple-birth-death (MBD) model inherits the parameters $\lambda$ and $\mu$ from the BD model; they correspond, respectively, 
+to the traditional per-species speciation and extinction rates. 
+Additionally, the MBD model assumes that external events, occuring at rate $\nu$ triggers a speciation initiation event in each lineage which leads to a full new species with probability $q$. 
+Whereas parameter $\lambda$ can be interpreted as the rate of sympatric speciation, $\nu$ is the rate of appearance of 
 geographical barriers able to interrupt the gene flow in the population,
 resulting in a possible allopatric speciation for each of the species.  
-Even though multiple speciations can co-occur, 
-polytomies are not allowed in such process as each species can speciate only 
-once during a co-occuring speciation event.
+Even though multiple speciation events can occur simultaneouly, it does not lead to  
+polytomies, because each species can only split once after a trigger event.
+This model can be easily simulated with a Doob-Gillespie algorithm.
+A probability distribution for the phylogeny under the MBD model can also be formulated using the integration approach developed for diversity-dependent diversification models \citep{etienne2012diversity}, which has been explored in \cite{mbd}.\rampal{Is this the mbd chapter?}.
+While this probability distribution could in principle be used as a tree prior in Bayesian phylogenetic inference, it is computationally very demanding, particularly for large trees.
+With the MBD model we generate simulated datasets for various parameter settings, using the \verb;mbd_sim; function from the \verb;mbd; R package [\citep{mbd}].
 
-A likelihood expression for the process is provided in \cite{mbd}.
-
-\subsubsection{Estimating the inference error}
+\subsection{Estimating the inference error}
 
@@
 ignoring the more complex and non-standard MBD tree prior in
-Bayesian phylogenetic inference.
-We do so by assuming a simpler and standard BD tree prior,
-and compare the 'true' tree with the inferred trees.
-To do so, we use the \verb;pirouette; R package [\citep{pirouette}].
+Bayesian phylogenetic inference with the R package \verb;pirouette; [\citep{pirouette}].
+ \verb;pirouette; starts from a 'true'  phylogeny (in our case: the simulated MBD tree), and simulates a DNA sequence alignment on it using a known nucleotide substitution model and a clocl model. 
+From each sequence alignment, a Bayesian inference is run with a particular choice of tree prior and substitution and clock models.
+One can choose the same substitution and clock models as used in generating the tree, or pick the ones that fit the data best.
+For the tree pior we assume the BD model (as the effect of this choice is our focus).
+We obtain a posterior distribution of jointly-estimated trees and model parameter estimates.
+By comparing the true tree and the posterior trees, an inference error distribution is generated.
+For this comparison we used the absolute nLTT statistic by \citet{janzen2015}, which results in an error distribution with values ranging from zero (when the inferred tree is identical to the true tree) to a maximum of one (trees are completely different).
 
-Here we summarize briefly how \verb;pirouette; works: it starts from a 'true' starting phylogeny (in our case: the simulated MBD tree), from which a DNA sequence alignment is simulated. 
-From each sequence alignment, a Bayesian inference is run (in our case: with assuming a BD tree prior), to obtain a posterior distribution of jointly-estimated trees and model parameter estimates.
-By comparing the true tree and the posterior trees, an inference error distribution is generated.
-We use the twinning option available in \verb;pirouette; that allows to quantify the impact of assuming a wrong tree prior in a Bayesian inference compared to a reference background error.
+We used the twinning option available in \verb;pirouette; that allows to quantify the impact of assuming a wrong tree prior in a Bayesian inference compared to a reference background error that would arise even if the models used in inference were identical to those used in generating the tree (i.e.
+the twin tree was generated with a BD model).
+
+\giovanni{Add mention to Jensen-Shannon divergence metric, if used.}
 
@@
 
-When working on simulated datasets the choice of the generating parameters is always an arbitrary choice.
-In principle one would choose a parameter range as wide as possible to characterize the most of the available parameter space.
-However, there are technical aspects that must be taken in consideration to be sure that both the simulation and the inference processes work out in an optimal way.
-We required the following criteria to be satisfied \giovanni{Does this apply only to the mbd parameters or also to the pirouette ones?}:
+When working with simulated datasets the choice of the generating parameters is always somewhat arbitrary.
+In principle one would choose a parameter range as wide as possible to characterize most of parameter space.
+However, both simulation and inference need to be possible computationally.
+To ensure this
+we required the following criteria to be satisfied for a parameter set to be chosen: \giovanni{Does this apply only to the mbd parameters or also to the pirouette ones?}:
 \itemize{
-  \item Quality: At least 95\% of all individual runs should have an Estimated Sample Size (ESS) of at least 200 (see \cite{beast} for reference);
+  \item Quality: At least 95\% of all inferences should have an Estimated Sample Size (ESS) of at least 200 (see \cite{beast} for reference);
     \richel{figure \ref{fig:esses}}
-  \item Feasibility: As time is a limited resource we impose that at least 95\% of all individual runs should finish within 10 days;
+  \item Feasibility: At least 95\% of all individual runs should finish within 10 days;
     \richel{figure \ref{fig:runtimes}}
@@
 
-Additional information about the choice of a suitable parameter setting can be found in the Supplementary Material.
+This resulted in the simulation parameters in Table~\ref{tab:simulation_parameters}.
+
+\input{tab_simparameters.tex}
+
+We assumed that the alignments are 1000 nucleotides in length, with a known root sequence of four 250 mono-nucleotide blocks, generated using the simplest nucleotide substitution model (Jukes Cantor, JC69) and clock model (strict), with a mutation rate of $\frac{t_c}{2}$, where $t_c$ is the crown age.
+With this mutation rate, each nucleotide has a 50\% chance to mutate (both silently and non-silently) from the ancestral root sequence to any of the contemporary species' sequences at the tips.
+
+For the Bayesian inference, we assumed a generative model of a JC69 site model, a strict clock model and a BD tree prior.
+Additionally, we used a Most Recent Common Ancestor (MRCA) prior equal to the crown age with a normal distribution of width $\sigma = 0.01$.
+We used a Markov Chain Monte Carlo (MCMC) setup of $10^6$ states with a sampling interval of once per $10^3$ states.
+Of the resulting $10^3$ states, we discarded a burn-in of $10\%$.
+
+To verify that the generative model (with a BD tree prior) is indeed the most appropriate standard model, we compared it to 39 other candidate inference models.
+These candidate models are all other combinations of four tree priors, two clock models and five speciation models, excluding the one combination used as a generative model.
+We let the best candidate model also generate an error distribution, which is expected to have a similar or high median. \rampal{What do you mean by 'high median'?}
+For model comparison, \verb;pirouette; uses a nested sampling approach to estimate marginal likelihoods, as described in \citet{russel2019model}, using the "\verb;NS;" BEAST2 package.
+We setup the MCMC for the nested sampling using its default values, which are a relative error $\epsilon$ of $10^{-12}$,
+1 particle, a sub-chain sampling interval of $5 \cdot 10^3$ states with a maximum chain length of $10^6$ states.
+
+Additional information about the choice of suitable parameter settings can be found in the Supplementary Material.\rampal{What kind of info is this? Explain. Why not put it here?}
 \iffalse
@@
 \fi
-
-\subsubsection{Simulations setting}
-
-We simulate the speciation process in continuous time using the \verb;mbd_sim; function from the \verb;mbd; R package [\citep{mbd}].
-We let parameters vary using all possible combinations of values as shown in Table~\ref{tab:simulation_parameters}.
-For each parameter setting, we generate [as much as time permits] 
-\richel{Will need to calculate how many replicates we can do} 
-independent trees of the same crown age.
-Note that trees that are actually passed as \verb;pirouette; input are not full trees, but reconstructed trees instead.
-A reconstructed tree is a tree where all the species that went extinct before the present time are removed, as this information is usually unavailable on real data phylogenies.
-
-\input{tab_simparameters.tex}
-
-\subsubsection{pirouette setting}
-
-In our experiment, the alignments are 1000 nucleotides in length, with a known root sequence of four 250 mono-nucleotide blocks, generated using the simplest nucleotide substitution model (Jukes Cantor, JC69) and clock model (strict), with a mutation rate of $\frac{t_c}{2}$, where $t_c$ is the crown age.
-With this mutation rate, each nucleotide has a 50\% chance to mutate (both silently and non-silently) from the ancestral root sequence to any of the contemporary species' sequences at the tips.
-
-For the Bayesian inference, we assume a generative model of a JC69 site model, a strict clock model and a BD tree prior.
-Additionally, we use a Most Recent Common Ancestor (MRCA) prior equal to the crown age with a normal distribution of width $\sigma = 0.01$.
-We pick a Markov Chain Monte Carlo (MCMC) setup of $10^6$ states with a sampling interval of once per $10^3$ states.
-Of the resulting $10^3$ states, we discard a burn-in of $10\%$.
-
-To verify that the generative model (with a BD tree prior) is indeed the most appropriate standard model, we compare it to 39 other candidate inference models.
-These candidate models are all other combinations of four tree priors, two clock models and five speciation models, excluding the one combination used as a generative model.
-We let the best candidate model also generate an error distribution, which is expected to have a similar or high median.
-For models comparison, \verb;pirouette; uses a nested sampling approach to estimate marginal likelihoods, as described in \citet{russel2019model}, using the "\verb;NS;" BEAST2 package.
-We setup the MCMC for the nested sampling using its default values, which are a relative error $\epsilon$ of $10^{-12}$,
-1 particle, a sub-chain sampling interval of $5 \cdot 10^3$ states with a maximum chain length of $10^6$ states.
-
-To compare the true tree with any of the posterior trees, we use the absolute nLTT statistic by \citet{janzen2015}, which will result in an error distribution with values ranging from zero (when the inferred tree is identical to the true tree) to a maximum of one (trees are completely different).
-
-For the twinning, we let a birth-death process generate one random tree,
-so that the twin inference can be done with the correct phylogenetic
-model of a JC69 site model, strict clock model and a BD tree prior.
-
-\giovanni{Add mention to Jensen-Shannon divergence metric, if used.}

### 2020-03-26 razzo-introduction.tex
Changes @Giappo

@@
 
-Modern computational techniques, such as BEAST (\cite{beast}) and its descendant 
-BEAST2 (\cite{beast2}), allow to infer phylogenetic trees from genetic data 
-such as DNA, RNA or proteins. \rampal{mention MrBayes and revBayes as well}. They 
-return posterior distributions of phylogenies 
+Modern computational techniques, such as BEAST (\citep{beast,beast2}), RevBayes (\citep{hohna2016revbayes}) and MrBayes (\citep{huelsenbeck2001mrbayes,ronquist2003mrbayes}), allow to infer phylogenetic trees from genetic data
+such as DNA, RNA or proteins.
+They return posterior distributions of phylogenies 
 and estimated parameters by running a Bayesian analysis, 
 given aligned sequence data and a set of models. 
-One of these models is the diversification model, for which a prior distribution must be provided. This is called a tree prior; it is a mathematical description 
+One of these models is the diversification model, for which a prior distribution must be provided. Within the Bayesian framework this is called a tree prior; it is a mathematical description 
 of the probability distribution of possible branching patterns before looking at the data.
@@
 However, multiple speciation events can take place simultaneously and repeatedly when populations are intermittently disconnected and connected, for example due to climatic fluctuations.
-This has been called the species pump hypothesis.\rampal{Cite: Haffer, J.
-(1969).
-Speciation in Amazonian forest birds.
-Science, 165,
-131–137.
-https://doi.org/10.1126/science.165.3889.131.} and has been invoked particularly in mountainous areas that underwent glaciation.
-\rampal{ Origins of global mountain plant biodiversity: Testing the 'mountain-geobiodiversity hypothesis'
-By: Muellner-Riehl, Alexandra N.; Schnitzler, Jan; Kissling, W. Daniel; et al.
-JOURNAL OF BIOGEOGRAPHY}.
+This has been called the species pump hypothesis (\citep{haffer1969speciation}) and has been invoked particularly in mountainous areas that underwent glaciation (\citep{muellner2019origins}).
 Our own interest in the species pump hypotheses arose from its potential explanation of the radiation of cichlid fish in the African rift lakes (Malawi, Tanganyika and Victoria), where water level drops created multiple smaller lakes providing the opportunity for allopatric speciation in multiple species.
@@
 
-One could study whether the species pump hypothesis is a viable explanation in empirical ystems by comparing divergence times of sister taxa \rampal{Oaks et al - doi:10.1111/evo.13754}.
+One could study whether the species pump hypothesis is a viable explanation in empirical ystems by comparing divergence times of sister taxa (\citep{oaks2019comparative}).
 A more inclusive approach would involve using a model allowing multiple simultaneous speciation events as a new species tree prior in phylogenetic reconstruction.

### 2020-03-26 razzo-methods.tex
Changes @Giappo

@@
 Bayesian phylogenetic inference with the R package \verb;pirouette; [\citep{pirouette}].
- \verb;pirouette; starts from a 'true'  phylogeny (in our case: the simulated MBD tree), and simulates a DNA sequence alignment on it using a known nucleotide substitution model and a clocl model. 
+ \verb;pirouette; starts from a 'true'  phylogeny (in our case: the simulated MBD tree), and simulates a DNA sequence alignment on it using a known nucleotide substitution model and a clock model. 
 From each sequence alignment, a Bayesian inference is run with a particular choice of tree prior and substitution and clock models.
 One can choose the same substitution and clock models as used in generating the tree, or pick the ones that fit the data best.
-For the tree pior we assume the BD model (as the effect of this choice is our focus).
+For the tree prior we assume the BD model (as the effect of this choice is our focus).
 We obtain a posterior distribution of jointly-estimated trees and model parameter estimates.

### 2020-03-28 razzo-methods.tex
Correct itemization

@@
 resulting verdicts at \url{https://github.com/richelbilderbeek/razzo_project/blob/master/verdict.md})}:
-\itemize{
+\begin{itemize}
   \item quality: 95\% of all individual runs should have an ESS of at least 200.
@@
     as high as possible
-}
+\end{itemize}
 We searched through parameter space until these criteria were met.
@@
 we required the following criteria to be satisfied for a parameter set to be chosen: \giovanni{Does this apply only to the mbd parameters or also to the pirouette ones?}:
-\itemize{
+\begin{itemize}
   \item Quality: At least 95\% of all inferences should have an Estimated Sample Size (ESS) of at least 200 (see \cite{beast} for reference);
@@
 For this reason we demand that the number of taxa should be as high as possible, as long as all the other requirements are met;
-}
+\end{itemize}
 

### 2020-03-31 scripts/merge_files.sh
Add 'merge_files.sh' script, needed for thesis

@@ -1,0 +1,11 @@
+#!/bin/bash
+#
+# Merge all .tex files to one file 'razzo_article.tex'.
+# 'razzo_article.tex' is ignored by git
+#
+# Usage:
+#
+#   ./scripts/merge_files.sh
+#
+#
+

### 2020-03-31 scripts/merge_files.sh
Can merge files

@@ -8,4 +8,14 @@
 #   ./scripts/merge_files.sh
 #
 #
+# 
+# Follow same structure as article.tex
+
+cat razzo-abstract.tex \
+  razzo-introduction.tex \
+  razzo-hypotheses.tex \
+  razzo-methods.tex \
+  razzo-results.tex \
+  fig_results.tex \
+  razzo-discussion.tex > razzo_article.tex
 

### 2020-03-31 scripts/merge_files.sh
Improve merge file script for thesis

@@ -19,3 +19,13 @@ cat razzo-abstract.tex \
   fig_results.tex \
   razzo-discussion.tex > razzo_article.tex
 
+
+echo " " >> razzo_article.tex
+echo "%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%" >> razzo_article.tex
+echo "% Bibliography" >> razzo_article.tex
+echo "%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%" >> razzo_article.tex
+echo "% MEE style" >> razzo_article.tex
+echo "\bibliographystyle{razzo_mee}" >> razzo_article.tex
+echo "\bibliography{razzo_article}" >> razzo_article.tex
+echo "%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%" >> razzo_article.tex
+

### 2020-03-31 razzo-discussion.tex
Add awesome 20200204 results, with 40 replicates

--- 2019-09-05 razzo-discussion.tex
@@
 
-We \richel{(at least I. If you do not expect this, let me know the 
-reasoning why)} expected to find a bigger inference error 
+We expected to find a bigger inference error 
 when increasing the amount of co-occurring speciation events.
@@
 it is the best-fitting standard tree prior. For the MBD model without
-extinction, \richel{@Giappo: I think this is an interesting insight:} 
-we could have picked the Yule tree prior. We \richel{at least I do} predicted
-that the Yule tree prior will be selected as the best candidate model
-more often in this scenario, which shows the sensitity of the method we used.
+extinction, we could have picked the Yule tree prior. 
+We predicted that the Yule tree prior will be selected as the best candidate model
+more often in this scenario, which shows the sensitivity of the method we used.
 
@@
 alternative methods like \richel{forgot the term for now: where you
-do an MCMC that can switch between models}. The reason we used Nested
-Sampling was mostly pragmatic and we think other methods are equivalent.
+do an MCMC that can switch between models}. The reason we used nested
+sampling was mostly pragmatic and we think other methods are equivalent.
 

### 2020-03-31 scripts/create_article.sh
Update scripts, show awesome figures

@@ -1,4 +1,6 @@
 #!/bin/bash
+./scripts/merge_files.sh
+
 pdflatex article.tex --shell-escape
 bibtex article.aux
 pdflatex article.tex --shell-escape

### 2020-03-31 scripts/merge_files.sh

@@ -28,4 +28,7 @@ echo "% MEE style" >> razzo_article.tex
 echo "\bibliographystyle{razzo_mee}" >> razzo_article.tex
 echo "\bibliography{razzo_article}" >> razzo_article.tex
 echo "%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%" >> razzo_article.tex
+echo " " >> razzo_article.tex
+
+cat razzo-supplement.tex >> razzo_article.tex
 

### 2020-03-31 razzo-abstract.tex
Abstract

@@
 % previous knowledge.
-We show that the extent of the inference error increases with the number of multiple birth events, as expected.
-\rampal{We need a bit more reults here}
+We show that the extent of the inference error 
+does not notably increase with the number of multiple birth events.
+Instead, the phylogenetic inference fails to converge more
+often under these settings.
+This reduced convergence is profound and easily detectable,
+yet its cause is unknown.
 
@@
 % more general context.
-These results allow phylogeneticists to judge under which circumstances the standard tree priors can be safely used, and under what circumstances phylogenetic inferences are unreliable.
-\giovanni{Not entirely sure if this is the best sentence to put here.}
-\rampal{Would be nice if we gave some guidelines here}
+These results show that using standard tree priors for biological systems
+following a species pump model is warranted, as long as convergence can 
+be attained. For settings that do not converge, the addition of 
+a new species tree prior to the current phylogenetic software may resolve this.
 

### 2020-03-31 razzo-methods.tex
Methods is fine

@@
 This model can be easily simulated with a Doob-Gillespie algorithm.
-A probability distribution for the phylogeny under the MBD model can also be formulated using the integration approach developed for diversity-dependent diversification models \citep{etienne2012diversity}, which has been explored in \cite{mbd}.\rampal{Is this the mbd chapter?}.
+A probability distribution for the phylogeny under the MBD model can also be formulated using the integration approach developed for diversity-dependent diversification models \citep{etienne2012diversity}.
 While this probability distribution could in principle be used as a tree prior in Bayesian phylogenetic inference, it is computationally very demanding, particularly for large trees.
@@
 For this comparison we used the absolute nLTT statistic by \citet{janzen2015}, which results in an error distribution with values ranging from zero (when the inferred tree is identical to the true tree) to a maximum of one (trees are completely different).
+Another advantage of using the nLTT statistic is that its behavior is best
+explored in \citet{bilderbeek2019pirouette}.
 
@@
 the twin tree was generated with a BD model).
-
-\giovanni{Add mention to Jensen-Shannon divergence metric, if used.}
+If the error distribution of the true tree matches the error distribution
+of the twin tree, the effect of using an incorrect tree prior is negligible \citet{bilderbeek2019pirouette}.
 
@@
 
-\iffalse
-We ran multiple pilot experiments with 1 replicate to arrive at our final
-parameter settings. We devised a set of rules to make a verdict about the
-settings \richel{(script at \url{https://github.com/richelbilderbeek/razzo_project/blob/master/scripts/90_collect_run_times.R},
-resulting verdicts at \url{https://github.com/richelbilderbeek/razzo_project/blob/master/verdict.md})}:
+We ran multiple pilot experiments with increasingly more replicate to arrive 
+at our final parameter settings. We devised a set of rules to make a 
+verdict about the settings:
+
 \begin{itemize}
   \item quality: 95\% of all individual runs should have an ESS of at least 200.
-    \richel{figure \ref{fig:esses}}
   \item feasibility: 95\% of all individual runs should finish within 10 days.
-    \richel{figure \ref{fig:runtimes}}
   \item reproducibility: the mean run-time of all finished runs should be less than 24 hours 
-    \richel{I suggest}.
-    \richel{figure \ref{fig:runtimes}}
   \item relevance 1: the percentage of taxa created by the MB process should be
@@
 \end{itemize}
+
 We searched through parameter space until these criteria were met.
@@
 \url{https://github.com/richelbilderbeek/razzo_project/blob/master/overview.md}.
-Due to the low number of replicates, we were unable (nor tried)
-to draw conclusions based on the results.
-\fi
-
-When working with simulated datasets the choice of the generating parameters is always somewhat arbitrary.
-In principle one would choose a parameter range as wide as possible to characterize most of parameter space.
-However, both simulation and inference need to be possible computationally.
-To ensure this
-we required the following criteria to be satisfied for a parameter set to be chosen: \giovanni{Does this apply only to the mbd parameters or also to the pirouette ones?}:
-\begin{itemize}
-  \item Quality: At least 95\% of all inferences should have an Estimated Sample Size (ESS) of at least 200 (see \cite{beast} for reference);
-    \richel{figure \ref{fig:esses}}
-  \item Feasibility: At least 95\% of all individual runs should finish within 10 days;
-    \richel{figure \ref{fig:runtimes}}
-  \item Reproducibility: the mean run-time of all finished runs should be less than 24 hours \richel{figure \ref{fig:runtimes}}
-
-    \richel{I suggest}. \giovanni{What's the difference with the previous point?}
-  \item Relevance 1: In order to have a meaningful inference, we demand that the MB signal is detectable to a satisfactory extent.
-For this reason the percentage of taxa created by the MB process should be as high as possible;
-  \item Relevance 2: Small trees might not contain a significant amount of information to be analyzed.
-For this reason we demand that the number of taxa should be as high as possible, as long as all the other requirements are met;
-\end{itemize}
-
+Due to the low number of replicates used in these pilot experiments, 
+we were unable (nor tried) to draw conclusions based on these pilot results.
 This resulted in the simulation parameters in Table~\ref{tab:simulation_parameters}.
@@
 
-We assumed that the alignments are 1000 nucleotides in length, with a known root sequence of four 250 mono-nucleotide blocks, generated using the simplest nucleotide substitution model (Jukes Cantor, JC69) and clock model (strict), with a mutation rate of $\frac{t_c}{2}$, where $t_c$ is the crown age.
+We assumed that the alignments are 1000 nucleotides in length, with a known root sequence of four 250 mono-nucleotide blocks, generated using the simplest nucleotide substitution model (Jukes Cantor, JC69) and clock model (strict), with a mutation rate of $\frac{0.5}{t_c}$, where $t_c$ is the crown age.
 With this mutation rate, each nucleotide has a 50\% chance to mutate (both silently and non-silently) from the ancestral root sequence to any of the contemporary species' sequences at the tips.
 
-For the Bayesian inference, we assumed a generative model of a JC69 site model, a strict clock model and a BD tree prior.
+For the Bayesian inference, we assumed a generative model of a site model with 
+a JC69 nucleotide substitution model, a strict clock model and a BD tree prior.
 Additionally, we used a Most Recent Common Ancestor (MRCA) prior equal to the crown age with a normal distribution of width $\sigma = 0.01$.
-We used a Markov Chain Monte Carlo (MCMC) setup of $10^6$ states with a sampling interval of once per $10^3$ states.
+We used a Markov Chain Monte Carlo (MCMC) setup of $10^7$ states with a sampling interval of once per $10^4$ states.
 Of the resulting $10^3$ states, we discarded a burn-in of $10\%$.
 
-To verify that the generative model (with a BD tree prior) is indeed the most appropriate standard model, we compared it to 39 other candidate inference models.
-These candidate models are all other combinations of four tree priors, two clock models and five speciation models, excluding the one combination used as a generative model.
-We let the best candidate model also generate an error distribution, which is expected to have a similar or high median. \rampal{What do you mean by 'high median'?}
-For model comparison, \verb;pirouette; uses a nested sampling approach to estimate marginal likelihoods, as described in \citet{russel2019model}, using the "\verb;NS;" BEAST2 package.
-We setup the MCMC for the nested sampling using its default values, which are a relative error $\epsilon$ of $10^{-12}$,
-1 particle, a sub-chain sampling interval of $5 \cdot 10^3$ states with a maximum chain length of $10^6$ states.
-
-Additional information about the choice of suitable parameter settings can be found in the Supplementary Material.\rampal{What kind of info is this? Explain. Why not put it here?}
-\iffalse
-\giovanni{This could be put in the appendix, if we manage to put in a handy format.}
-We searched through parameter space until these criteria were met.
 All the parameter settings used in the pilot experiments can be found at 
 \url{https://github.com/richelbilderbeek/razzo_project/blob/master/overview.md}.
-Due to the low number of replicates, we were unable (nor tried)
-to draw conclusions based on the results.
-\fi
+For each of the parameter settings, we simulated 40 MBD trees. We ran all
+960 \verb;pirouette; experiments on the computer cluster of the University
+of Groningen, which took 10 days. We used additionally R scripts to visualize
+the results.

### 2020-03-31 scripts/merge_files.sh
No hypotheses for now

@@ -13,7 +13,6 @@
 
 cat razzo-abstract.tex \
   razzo-introduction.tex \
-  razzo-hypotheses.tex \
   razzo-methods.tex \
   razzo-results.tex \
   fig_results.tex \

### 2020-03-31 fig_results.tex
Remove obsolete figures
Use nicer figures
Update scripts, show awesome figures
Moved all figures to main text

--- 2019-07-01 fig_results.tex
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{razzo_results.png}
+  \includegraphics[width=\textwidth]{20200204_fig_esses.png}
   \caption{
-    The inference error distribution 
-    for the different biological
-    parameter settings. In all cases, $\lambda = 0.2$ and 
-    crown age equals 6. 
+    Effective sample sizes of the likelihood of all the experiments that
+    finished within 10 days. 
+    Each point is one parameter setting.
   }
-  \label{fig:results}
+  \label{fig:esses}
 \end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{20200204_fig_n_taxa.png}
+  \caption{
+    Histogram of the number of taxa in the full experiment.
+  }
+  \label{fig:n_taxa}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{20200204_figure_1a.png}
+  \caption{
+    The error distribution for a extinction rate $\mu = 0.0$.
+  }
+  \label{fig:errors_yule}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{20200204_figure_1b.png}
+  \caption{
+    The error distribution for extinction rate $\mu = 0.1$.
+  }
+  \label{fig:errors_bd}
+\end{figure}

### 2020-03-31 razzo-methods.tex
Moved all figures to main text

@@
 Bayesian phylogenetic inference with the R package \verb;pirouette; [\citep{pirouette}].
+
 \verb;pirouette; starts from a 'true'  phylogeny (in our case: the simulated MBD tree), and simulates a DNA sequence alignment on it using a known nucleotide substitution model and a clock model. 
@@
 
-We ran multiple pilot experiments with increasingly more replicate to arrive 
+We ran multiple pilot experiments with increasingly more replicates to arrive 
 at our final parameter settings. We devised a set of rules to make a 
-verdict about the settings:
+verdict about the settings.
 
 \begin{itemize}
-  \item quality: 95\% of all individual runs should have an ESS of at least 200.
+  \item quality: 95\% of all individual runs should have an effective
+    sample size (ESS) of at least 200, as is recommended by \cite{beastbook}.
   \item feasibility: 95\% of all individual runs should finish within 10 days.
@@
 We searched through parameter space until these criteria were met.
-All the parameter settings used in the pilot experiments can be found at 
-\url{https://github.com/richelbilderbeek/razzo_project/blob/master/overview.md}.
-Due to the low number of replicates used in these pilot experiments, 
-we were unable (nor tried) to draw conclusions based on these pilot results.
 This resulted in the simulation parameters in Table~\ref{tab:simulation_parameters}.
@@
 
-We assumed that the alignments are 1000 nucleotides in length, with a known root sequence of four 250 mono-nucleotide blocks, generated using the simplest nucleotide substitution model (Jukes Cantor, JC69) and clock model (strict), with a mutation rate of $\frac{0.5}{t_c}$, where $t_c$ is the crown age.
+We generated alignments that are 1000 nucleotides in length, with a known root sequence of four 250 mono-nucleotide blocks, generated using the simplest nucleotide substitution model (Jukes Cantor, JC69) and clock model (strict), with a mutation rate of $\frac{0.5}{t_c}$, where $t_c$ is the crown age.
 With this mutation rate, each nucleotide has a 50\% chance to mutate (both silently and non-silently) from the ancestral root sequence to any of the contemporary species' sequences at the tips.
 
-For the Bayesian inference, we assumed a generative model of a site model with 
-a JC69 nucleotide substitution model, a strict clock model and a BD tree prior.
+For the Bayesian inference, we assumed a generative model of a site model that
+follows a JC69 nucleotide substitution model, a strict clock model and a BD tree prior.
 Additionally, we used a Most Recent Common Ancestor (MRCA) prior equal to the crown age with a normal distribution of width $\sigma = 0.01$.
@@
 
-All the parameter settings used in the pilot experiments can be found at 
-\url{https://github.com/richelbilderbeek/razzo_project/blob/master/overview.md}.
-For each of the parameter settings, we simulated 40 MBD trees. We ran all
-960 \verb;pirouette; experiments on the computer cluster of the University
-of Groningen, which took 10 days. We used additionally R scripts to visualize
-the results.
+For each of the MBD parameter settings, we simulated 40 different trees
+that we put through the same \verb;pirouette; pipeline as described above.
+Of each of these 40 replicates, we aggregated the error distributions.

### 2020-03-31 razzo-results.tex
Changes Giappo
Moved all figures to main text

--- 2019-10-07 razzo-results.tex
@@
 is shown in Figures ~\ref{fig:errors_yule} and ~\ref{fig:errors_bd}.
-For both extinction rates,
-we find that the error increases as $\nu$
-and $q$ increase. 
-Also in line with RJCB's predictions, 
-we find no difference between the two extinction rates.
+The two figures, that only differ in their extinction rate, 
+show the same error distributions. There is no clear difference between
+the true and twin error distribution. For $\nu \leqslant 0.5$, the
+range of errors, however, differ. For $\nu \geq 1.0$, the error
+distributions become bimodal.
+
+All intermediate data can be downloaded from \url{https://richelbilderbeek.nl/razzo_project_20200204.zip}.

### 2020-03-31 scripts/merge_files.sh

@@ -29,5 +29,3 @@ echo "\bibliography{razzo_article}" >> razzo_article.tex
 echo "%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%" >> razzo_article.tex
 echo " " >> razzo_article.tex
 
-cat razzo-supplement.tex >> razzo_article.tex
-

### 2020-03-31 razzo-results.tex
Discuss all figures in the results

@@
 \section{Results}
+
+98% of all runs finished within 10 days. 
+Of these runs, figure \ref{fig:esses} shows the distribution 
+of their effective samples sizes, with 88\% above the recommended 
+value of 200.
+
+Figure \ref{fig:n_taxa} shows the number of taxa for all runs. 
+Most runs have less than 100 taxa, of which the median and mean both lie
+beloe 30 taxa.
 
@@
 
-All intermediate data can be downloaded from \url{https://richelbilderbeek.nl/razzo_project_20200204.zip}.
+All simulated data can be downloaded from \url{https://richelbilderbeek.nl/razzo_project_20200204.zip}.

### 2020-03-31 razzo-discussion.tex
Good Enough

@@
 co-occur. The stronger the amount of co-occurring speciation, the stronger
-the deviation from the species tree prior.
+the deviation from the species tree prior. That we did not find this, 
+is unexpected.
 
-\subsection{Research}
+The first reason why this may be the case, is because of the noise
+generated by the runs that did not converge. We expect that the bimodality
+in the error distribution is caused by most runs converging (at the lower
+errors) and runs with low ESSes (at the higher errors). Filtering away
+the runs with a low ESS may help to get a better idea if the expected
+inference error still cannot be detected.
+
+The second reason why this we did not detect an inference error may
+be due to the choice of our error measuring method: by using the nLTT
+statistic. Another method may be better at detecting an inference error.
+
+A third reason might be that the MBD parameter settings chosen, resulted
+in too few multiple-birth events. We chose low values of $\nu$ to
+prevent doing inference on huge trees. Being able to work on bigger
+trees might resolve this problem.
 
@@
 
-\subsection{Methodological choices}
-
-When simulating a DNA alignment, we chose to use the simplest (JC69) site
-model and the simplest (strict) clock model. One could argue these models
+For our experiment, we used the default \verb;pirouette; setup:
+when simulating a DNA alignment, the simplest (JC69) nucleotide
+substitution model and the simplest (strict) clock model are used. 
+One could argue these models
 are overly simplistic and biologically irrelevant. We argue that this
@@
 to assume the same generative site and clock model as the one that is actually 
-used. Sure, we could have generated the alignment with a 
-complex (GTR) site model and a complex (relaxed log-normal) clock model, then
-correctly assume these were the generative site and clock model.
-But that would be besides the point of the experiment: to measure the impact
-of the tree prior.
+used.
 
@@
 it is the best-fitting standard tree prior. For the MBD model without
-extinction, we could have picked the Yule tree prior. 
-We predicted that the Yule tree prior will be selected as the best candidate model
-more often in this scenario, which shows the sensitivity of the method we used.
-
-To compare multiple inference models, we picked the Nested Sampling
-approach to directly compare marginal likelihoods. We could have used
-alternative methods like \richel{forgot the term for now: where you
-do an MCMC that can switch between models}. The reason we used nested
-sampling was mostly pragmatic and we think other methods are equivalent.
-
-We used the absolute nLTT statistic to measure the difference between
-two phylogenies. Because the crown age of all phylogenies are equal,
-normalization on the time has no effect at all. Also because of these
-same crown ages, we could have picked many tree difference metrics, like
-\richel{must check:} Pagel's delta or Pigot's delta-r. The reason we used
-the nLTT statistic was mostly pragmatic. \verb;pirouette; does allow to 
-use a different tree difference metric.
+extinction, we could have picked the Yule tree prior. We expect, however,
+that this is of minor relevance.
 
@@
 runs that took days to calculate. We decided to prefer a lower number of
-taxa at the gain of having more replicates. As a consequence, 
-this made our research findings 
-\richel{that is, I predict, a bigger error in settings with
-more co-occurring speciation} more likely to happen: we know that the
-choice of a Bayesian prior is most important in weaker data.
-
-The DNA sequence length we used was 1000 base pairs. This is a DNA
-sequence length that we \richel{at least I} feel is informative enough
-for the amount of taxa we simulate. Additionally (although sequencing
-techniques advance at a great pace), this is -to our knowledge- 
-of the same magnitude as the DNA length sequenced in the field.
+taxa at the gain of having more replicates.
 
 Ideally, our ESSes would all be above the recommended value of 
-200 \richel{I predict not all will be above 200, simply due to 
-stochasticity}. We are aware ESSes can be increased easily by
-making the MCMC chain longer. The shape of the ESS distribution
-(see figure \ref{fig:esses_cumulative}), however, is disadvantageous
-to do so: to reduce the percentage of $ESS < 200$ by fifty percent,
+200. We are aware ESSes can be increased easily by
+making the MCMC chain longer. However, the bimodal
+distribution of ESSes is disadvantageous: to reduce the percentage 
+of $ESS < 200$ by fifty percent,
 we would have to increase the MCMC chain lengths by a factor of forty.
 We invested this run-time in doing more replicates.
-
-We compare our generative to model to \richel{I predict we will} 
-a limited set of candidate models. Also this was for practical
-reasons.
 
@@
 to add the MBD other speciation models as a species tree prior
-in a phylogentic tool and do a model comparision (e.g. using Nested Sampling).
+in a phylogentic tool and do a model comparision (e.g. using nested sampling).
 Another way to do so, is measure the impact a rate-shift or early-burst model
@@
 helping us in how to judge the error distributions.
-
-Expressing the inference error of an MBD process as a distribution
-is a good first step in judging the impact an MBD process has on
-our predictions, when compared to standard models. It is only a first
-step as one still needs to subjectively judge these error distributions,
-which was exactly what \verb;pirouette; was hoping to avoid. 
-One way to improve our judgement, is by testing it on known species tree
-priors, for example, by simulating true trees with a BD model, when assuming
-a Yule tree prior.
-
-The effect of using the absolute nLTT tree statistic to quantify the error
-is unknown. A pilot experiment \richel{pirouette example 14} shows that
-using the logarithmically transformed nLTT statistic gives a messier and
-flatter error distribution, but this is not to be taken for granted.
-
-We chose to use few taxa (as we preferred more replicates), but we do not
-know the results would the data be more informative (that is, have more taxa).
-A cool feature of \verb;pirouette; is that it can quantify when this happen;
-when the prior is uninformative due to strong data.

### 2020-04-03 fig_results.tex
Discuss all figures in the results
Changes @rsetienne

--- 2019-07-01 fig_results.tex
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{razzo_results.png}
+  \includegraphics[width=1.0\textwidth]{20200204_fig_esses.png}
   \caption{
-    The inference error distribution 
-    for the different biological
-    parameter settings. In all cases, $\lambda = 0.2$ and 
-    crown age equals 6. 
+    Frequency distribution of effective sample sizes of all experiments that
+    finished within 10 days. 
+    Each point represents one parameter setting.
   }
-  \label{fig:results}
+  \label{fig:esses}
 \end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=1.0\textwidth]{20200204_fig_n_taxa.png}
+  \caption{
+    Frequency distribution of the number of taxa across all experiments.
+  }
+  \label{fig:n_taxa}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{20200204_figure_1a.png}
+  \caption{
+    The inference error distribution for a generating extinction rate of $\mu = 0$.
+  }
+  \label{fig:errors_yule}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{20200204_figure_1b.png}
+  \caption{
+    The inference error distribution for a generating extinction rate of $\mu = 0.1$.
+  }
+  \label{fig:errors_bd}
+\end{figure}

### 2020-04-03 razzo-discussion.tex
Changes @rsetienne

@@
 
-We expected to find a bigger inference error 
+We expected to find a larger inference error 
 when increasing the amount of co-occurring speciation events.
-The reason is simple: the tree prior used assumes speciation does not
-co-occur. The stronger the amount of co-occurring speciation, the stronger
-the deviation from the species tree prior. That we did not find this, 
-is unexpected.
+The reason is simple: the tree prior used assumes that speciation events do not occur simultaneously.
+The higher the number of co-occurring speciation events, the stronger
+the deviation from the species tree prior. However, we did not find evidence for this prediction..
 
 The first reason why this may be the case, is because of the noise
-generated by the runs that did not converge. We expect that the bimodality
-in the error distribution is caused by most runs converging (at the lower
-errors) and runs with low ESSes (at the higher errors). Filtering away
-the runs with a low ESS may help to get a better idea if the expected
-inference error still cannot be detected.
+generated by the runs that did not converge. We think that the bimodality
+in the error distribution is caused by the converging runs giving a low error, and runs with low ESSes giving higher errors. This can be studied further by Filtering away
+the runs with a low ESS. However, this will only remove the second mode, but will not explain the similarity of the two error distributions for true and twin trees.
 
-The second reason why this we did not detect an inference error may
-be due to the choice of our error measuring method: by using the nLTT
-statistic. Another method may be better at detecting an inference error.
+A second reason might be that our choice of MBD parameter settings resulted
+in too few multiple-birth events. We chose low values of $\nu$ to
+prevent doing inference on huge trees.
+For larger trees, one might perhaps find a difference in the error distributions.
+However, we do remark the frequency of multiple-birth events was often larger than 50\%.
 
-A third reason might be that the MBD parameter settings chosen, resulted
-in too few multiple-birth events. We chose low values of $\nu$ to
-prevent doing inference on huge trees. Being able to work on bigger
-trees might resolve this problem.
+The third reason why this we did not detect a difference in inference errors between true and twin trees may
+be due to the choice of measuring our error. We used the nLTT
+statistic, which may not be sensitive enough to what we consider to be different trees: trees with aligned speciation events and trees where these are not aligned.
+It is, however, not straightforward to define a metric that is more sensitive to this difference than nLTT.
+One could perhaps try a shotgun approach by simply applying many commonly used metrics, but we believe that this is unlikely to yield a different outcome.
+The nLTT statistic at least picks up the sudden increase in number of lineages present in MBD true trees, but not in BD twin trees, but this difference is apparently small relative to other stochastically arising differences between true/twin and the inferred posterior distribution of trees.
 
-The MBD model is a BD model that allows for co-occurring speciation,
-with the same drawbacks as the BD model: it allows for an unlimited 
-number of species and does not take into account that speciation takes time.
-That does make the MBD model a very focussed model, only taking into 
-account the essence of co-occurring speciation. The MBD model can probably
+The MBD model is a BD model that allows for speciation events taking place simultaneously, but has
+the same drawbacks as the BD model: the expected number of species increases quickly with time (when net formation of species exceeds extinction) and the model does not take into account that speciation takes time.
+The MBD model can, at least in the simulations, easily
 be extended to have a diversity-dependent speciation rate (as in the 
@@
 an incipient species state (as in the PBD model).
-The idea of this experiment, however, is to measure the impact a novel
-tree prior has on the inference error. To do so, we could have picked
-any other non-standard (as in, not included within BEAST2) tree prior.
+The idea of this experiment, however, was to measure the impact of species pump dynamics on phylogenetic inference that does not assume this, and the comparison of a BD and MBD model seems therefore appropriate, particularly because for diversity-dependent or protracted single-birth-death models there is no standard species tree prior available which would have complicated our analyses.
+Furthermore, we do not expect that addition of diversity-dependence or protracted speciation alters the inference on whether speciation events take place simultaneously or not.
 
@@
 we were able to perform more replicates. For a process with this high
-amount of stochasticity, we think this is highly favored. The second reason
-is that we think this reduces the noise of our results, without 
+amount of stochasticity, we think this is highly preferable. The second reason
+is that we think that this reduces the noise of our results, without 
 hurting the experiment: in our setup, it is essential
-to assume the same generative site and clock model as the one that is actually 
-used.
+to assume the same site and clock model in inference as the one that is actually used in generating the data.
 
@@
 it is the best-fitting standard tree prior. For the MBD model without
-extinction, we could have picked the Yule tree prior. We expect, however,
-that this is of minor relevance.
+extinction, we could have picked the Yule tree prior. We believe, however,
+that this is of minor importance.
 
-\subsection{Choices to constrain time}
-
-From the four MBD parameters $\lambda$, $\mu$, $\nu$ and $q$,
+For the four MBD parameters $\lambda$, $\mu$, $\nu$ and $q$,
 we investigated 1, 2, 4 and 3 different values respectively.
-We chose to use only one $\lambda$, as the proportion of species
-created in a co-occurent speciation event is dependent on the ratio
+We chose to use only one value for $\lambda$, because the proportion of multiple-birth events depends on the ratio
 between $\lambda$ and a combination of $\nu$ and $q$.  
-
 The crown age we selected for our experiments was based on trial and error
@@
 profound effect on the number of taxa, with the consequence of having
-runs that took days to calculate. We decided to prefer a lower number of
-taxa at the gain of having more replicates.
+runs that took days to calculate. We decided to priortize the number of replicates at the cost of a lower number of taxa.
 
-Ideally, our ESSes would all be above the recommended value of 
+Ideally, we would have liked all our ESSes to be above the recommended value of 
 200. We are aware ESSes can be increased easily by
@@
 of $ESS < 200$ by fifty percent,
-we would have to increase the MCMC chain lengths by a factor of forty.
-We invested this run-time in doing more replicates.
+we would have had to increase the MCMC chain lengths by a factor of forty.
+We preferred to invest this run-time in doing more replicates.
 
-\subsection{Future research}
-
-This research measures the impact an MBD tree prior 
-has on the inference error.
-
-Because the MBD tree prior has only recently been 
-developed, it's biological relevance is unknown.
-Would the MBD tree prior be employed on biological data of a system
-in which co-occuring speciation is known to have taken a strong role,
-it will be hard to conclude its performance is superior to a
-rate-shift model or an early-burst model. One way to do so, is
-to add the MBD other speciation models as a species tree prior
-in a phylogentic tool and do a model comparision (e.g. using nested sampling).
-Another way to do so, is measure the impact a rate-shift or early-burst model
-has on the inference error in the same way as this research.
-Doing so would also allow to put the quantitative results into perspective,
-helping us in how to judge the error distributions.
+This research measures the impact that the use of a non-MBD tree prior 
+has on the inference error in phylogenetic construction for species that are subject to species pump dynamics.
+We found that the trees constructed with the standard BD species tree prior are similar (as measured by nLTT) to the true tree.
+This may be considered to be good news, as the implementation of an MBD species tree prior does not seem necessary.
+However, our results also suggest that it will be difficult to determine the parameter of the MBD process when one wants to fit the model to data of a system that is known to be subject to species pump dynamics.
+One way to test this would be to try  birth-death skyline species tree priors and see if they pick up a signal of elevated speciation rates during co-occurring speciation events.
+These models may be prone to overparametrization because they have to assume a speciation rate for each interval in which the multiple speciation events take place, whereas the MBD model provides a dynamic explanation for these elevated speciation rates.
+But if they do pick up a signal, then it may still be worth implementing a species tree prior for the MBD model in Bayesian phylogenetic reconstruction tools.

### 2020-04-03 razzo-results.tex
:heart: sloppypar, but don't tell my mom
Changes @rsetienne

--- 2020-03-31 razzo-results.tex
@@
 
-The inference error made for each of the parameter combinations
-is shown in Figures ~\ref{fig:errors_yule} and ~\ref{fig:errors_bd}.
-The two figures, that only differ in their extinction rate, 
-show the same error distributions. There is no clear difference between
-the true and twin error distribution. For $\nu \leqslant 0.5$, the
-range of errors, however, differ. For $\nu \geq 1.0$, the error
-distributions become bimodal.
+Of all runs 98\% finished within 10 days.
+Of these runs, 88\% had effective sample sizes above the recommended value of 200, and most well above this (Fig.
+\ref{fig:esses}).
+Most runs had less than 100 taxa, of which the median and mean both lie below 30 taxa (Fig.
+\ref{fig:n_taxa}).
 
-All intermediate data can be downloaded from \url{https://richelbilderbeek.nl/razzo_project_20200204.zip}.
+We find no clear difference between the error distributions for the true and twin trees, regardless of whether there is extinction or not in the generating process (Fig.~\ref{fig:errors_yule} and Fig.
+~\ref{fig:errors_bd}.
+We do observe that for generating values of $\nu \leqslant 0.5$, there is only one mode close to 0, whereas for $\nu \geq 1.0$, the error
+distributions become bimodal, with a second mode around 0.25. 
+All simulated data can be downloaded from 
+\begin{sloppypar}
+\url{https://richelbilderbeek.nl/razzo_project_20200204.zip}.
+\end{sloppypar}
